### 2017-07-29 README.Rmd
many changes -- added confint methods, added islptw, various tests, print methods, etc...

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# R/`drtmle`

[![Travis-CI Build Status](https://travis-ci.org/benkeser/drtmle.svg?branch=master)](https://travis-ci.org/benkeser/drtmle)
[![AppVeyor Build  Status](https://ci.appveyor.com/api/projects/status/github/benkeser/drtmle?branch=master&svg=true)](https://ci.appveyor.com/project/benkeser/drtmle)
[![Coverage Status](https://img.shields.io/codecov/c/github/benkeser/drtmle/master.svg)](https://codecov.io/github/benkeser/drtmle?branch=master)
[![CRAN](http://www.r-pkg.org/badges/version/drtmle)](http://www.r-pkg.org/pkg/drtmle)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)

> Doubly Robust, Asymptotically Linear Targeted Minimum Loss-Based Estimation

__Author:__ [David Benkeser](https://www.benkeserstatistics.com/)

---

## Description

`drtmle` is an R package that computes marginal effect estimators for binary 
treatments on continuous and binary outcomes. The TMLE estimators are doubly
robust, not only with respect to consistency, but also with respect to asymptotic 
normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017,(_accepted Biometrika_, [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).

---

## Installation

- Install the most recent _stable release_:
  `devtools::install_github("benkeser/drtmle")`

- To contribute, install the _development version_:
  `devtools::install_github("benkeser/drtmle", ref = "develop")`

---

## Use

This package can be used to estimate covariate-adjusted marginal means under multiple discrete levels of a treatment. It may be used in situations where data consist of a vector of baseline covariates (`W`), a multi-level treatment assignment (`A`), and a continuous or binary-valued outcome (`Y`). The function `drtmle` may be used to estimate $E[E(Y | A=a_0, W)]$ for user-selected values of $a_0$ (via option `a_0`). The resulting targeted minimum loss-based estimates are doubly robust with respect to both consistency and asymptotic normality. The function computes doubly robust variance estimators that can be used to construct doubly robust confidence intervals for covariate-adjusted marginal means and contrasts between these means (via the `drconfint` function) for different levels of $a_0$. Doubly robust hypothesis tests are also available (via `drtest` function). A simple example on simulated data is shown below.

```{r}
# simulate simple data structure
set.seed(1234)
n <- 500
W <- data.frame(W1 = runif(n,-2,2), W2 = rbinom(n,1,0.5))
A <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W2))
Y <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W2 + A))

# estimate the covariate-adjusted marginal mean for A = 1 and A = 0
# here, we do not properly estimate the propensity score
fit <- drtmle(W = W, A = A, Y = Y, # input data
              a_0 = c(0,1), # return estimates for A = 0 and A = 1
              SL_Q = "SL.npreg", # use kernel regression for E(Y | A=a, W)
              glm_g = "W1 + W2", # use misspecified main terms glm for E(A | W)
              SL_Qr = "SL.npreg", # use kernel regression to guard against
              					  # misspecification of outcome regression
              SL_gr = "SL.npreg" # use kernel regression to guard against 
              					 # misspecification of propensity score
              )
# print the output
fit

# get confidence intervals for means
ci_fit <- drconfint(fit)
# print the output
ci_fit

# get hypothesis test for means 
test_fit <- drtest(fit)
# print the output
test_fit
```
---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/benkeser/survtmle/issues).

---

## Citation

After using the `drtmle` R package, please cite the following:

        @article{benkeser2017improved,
          year  = {2017},
          author = {Benkeser, David C and Carone, Marco and van der Laan, Mark J and Gilbert, Peter B},
          title = {Doubly-robust nonparametric inference on the average treatment effect},
          journal = {Biometrika}
        }

---

## License

&copy; 2016-2017 [David C. Benkeser](http://www.benkeserstatistics.com)

The contents of this repository are distributed under the MIT license. See
below for details:
```
The MIT License (MIT)

Copyright (c) 2016-2017 David C. Benkeser

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
```

### 2017-07-29 R/confint.R

@@ -1,0 +1,285 @@
+#' Confidence intervals for drtmle objects
+#' 
+#' Estimate doubly-robust confidence intervals for
+#' an object of class \code{"drtmle"}
+#' 
+#' @param object An object of class \code{"drtmle"}
+#' @param parm A vector indicating for which estimators to return a 
+#' confidence interval. Possible estimators include the TMLE with doubly robust
+#' inference (\code{"drtmle"}, recommended), the AIPTW with additional correction 
+#' for misspecification (\code{"aiptw_c"}, not recommended), the standard TMLE 
+#' (\code{"tmle"}, recommended only for comparison to "drtmle"), the standard 
+#' AIPTW (\code{"aiptw"}, recommended only for comparison to "drtmle").  
+#' @param level The level of the confidence interval
+#' @param contrast This option specifies what parameter to return confidence intervals for.
+#' If \code{contrast=NULL}, then compute confidence intervals for the covariate-adjusted
+#' marginal means. \code{contrast} can also be input as a numeric vector of ones, negative ones,
+#' and zeros to define linear combinations of the various means (e.g., to estimate an average
+#' treatment effect, see examples). \code{contrast} can also be a list with named functions
+#' \code{f}, \code{f_inv}, \code{h}, and \code{h_grad}. The first two functions should take
+#' as input argument \code{eff}. Respectively, these specify which transformation 
+#' of the effect measure to compute the confidence interval for and the inverse
+#' transformation to put the confidence interval back on the original scale. The function \code{h}
+#' defines the contrast to be estimated and should take as input \code{parm}, a vector
+#' of the same length as \code{object$a_0}, and output the desired contrast. The function
+#' \code{h_grad} is the gradient of the function \code{h}. See examples. 
+#' @param ... Other options (not currently used)
+#' @importFrom stats qnorm
+#' @export
+#' @method confint drtmle
+#' @return An object of class \code{"confint.drtmle"} with point estimates and
[…]

### 2017-07-29 R/drtmle.R

@@ -1,37 +1,36 @@
-#' drtmle 
+#' TMLE estimate of the average treatment effect with doubly-robust inference
 #' 
-#' Main function to estimate TMLE with doubly-robust inference
 #' 
 #' @param W A \code{data.frame} of named covariates
 #' @param A A vector of binary treatment assignment (assumed to be equal to 0 or 1)
 #' @param Y A numeric of continuous or binary outcomes. 
-#' @param a0 A vector of treatment levels at which to compute the adjusted mean outcome. 
+#' @param a_0 A vector of treatment levels at which to compute the adjusted mean outcome. 
 #' @param family A \code{family} object equal to either \code{binomial()} or \code{gaussian()}, 
 #' to be passed to the \code{SuperLearner} or \code{glm} function.
 #' @param stratify A \code{boolean} indicating whether to estimate the outcome regression separately
 #' for observations with \code{A} equal to 0/1 (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
-#' @param libraryQ A vector of characters or a list describing the Super Learner library to be used 
+#' @param SL_Q A vector of characters or a list describing the Super Learner library to be used 
 #' for the outcome regression. See \code{link{SuperLearner::SuperLearner}} for details.
-#' @param libraryg A vector of characters or a list describing the Super Learner library to be used 
+#' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
 #' for the propensity score. See \code{link{SuperLearner::SuperLearner}} for details.
-#' @param libraryQr A vector of characters or a list describing the Super Learner library to be used 
+#' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
 #' for the first reduced-dimension regression. 
-#' @param librarygr A vector of characters or a list describing the Super Learner library to be used 
+#' @param SL_gr A vector of characters or a list describing the Super Learner library to be used 
 #' for the second reduced-dimension regression.
-#' @param glmQ A character describing a formula to be used in the call to \code{glm} for the outcome regression. Ignored
-#' if \code{libraryQ!=NULL}.
[…]

### 2017-07-29 R/estimate.R

@@ -5,72 +5,88 @@
 #' @importFrom plyr alply
 #' @param A A vector of binary treatment assignment (assumed to be equal to 0 or 1)
 #' @param W A \code{data.frame} of named covariates
-#' @param libraryg A vector of characters or a list describing the Super Learner library to be used 
+#' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
 #' for the propensity score
 #' @param tolg A numeric indicating the minimum value for estimates of the propensity score.
 #' @param verbose A boolean indicating whether to print status updates.
 #' @param returnModels A boolean indicating whether to return model fits for the outcome regression, propensity score,
 #' and reduced-dimension regressions.
-#' @param glmg A character describing a formula to be used in the call to \code{glm} for the propensity score
-#' @param a0 A list of fixed treatment values 
+#' @param glm_g A character describing a formula to be used in the call to \code{glm} for the propensity score
+#' @param a_0 A list of fixed treatment values 
 #' 
 #' 
 #' @importFrom SuperLearner SuperLearner trimLogit
 #' @importFrom stats predict glm as.formula
 #' 
 
-estimateG <- function(A, W,libraryg,glmg,a0,tolg,verbose=FALSE, returnModels=FALSE){
-  if(is.null(libraryg) & is.null(glmg)) stop("Specify Super Learner library or GLM formula for g")
-  if(!is.null(libraryg) & !is.null(glmg)){
+estimateG <- function(A, W,SL_g,glm_g,a_0,tolg,verbose=FALSE, returnModels=FALSE){
+  if(is.null(SL_g) & is.null(glm_g)) stop("Specify Super Learner library or GLM formula for g")
+  if(!is.null(SL_g) & !is.null(glm_g)){
     warning("Super Learner library and GLM formula specified. Proceeding with Super Learner only.")
-    glmg <- NULL
+    glm_g <- NULL
   }
   # Super Learner
-  if(!is.null(libraryg)){
-    if(length(libraryg)>1 | is.list(libraryg)){
-      if(length(a0)==length(unique(A)) & length(unique(A))==2){
-        fm <- SuperLearner::SuperLearner(Y=as.numeric(A==a0[1]), X=W, 
+  if(!is.null(SL_g)){
+    if(length(SL_g)>1 | is.list(SL_g)){
[…]

### 2017-07-29 R/fluctuate.R

@@ -8,14 +8,14 @@
 #' @param W The covariates
 #' @param Qn A list of outcome regression estimates evaluated on observed data
 #' @param gn A list of propensity regression estimates evaluated on observed data
-#' @param a0 A list of fixed treatment values 
+#' @param a_0 A list of fixed treatment values 
 #' 
 #' @importFrom SuperLearner trimLogit
 #' @importFrom stats predict glm
 #' 
 
-fluctuateQ1 <- function(Y,A,W, Qn, gn, a0){
-  QnStar <- mapply(a=a0,Q=Qn,g=gn,FUN=function(x, a, Q, g){
+fluctuateQ1 <- function(Y,A,W, Qn, gn, a_0){
+  QnStar <- mapply(a=a_0,Q=Qn,g=gn,FUN=function(x, a, Q, g){
       l <- min(Y); u <- max(Y)
       Yscale <- (Y-l)/(u-l)
       off <- SuperLearner::trimLogit((Q-l)/(u-l))
@@ -40,20 +40,19 @@ fluctuateQ1 <- function(Y,A,W, Qn, gn, a0){
 #' @param Y The outcome
 #' @param A The treatment
 #' @param W The covariates
-#' @param Qn A list of outcome regression estimates evaluated on observed data
 #' @param gn A list of propensity regression estimates evaluated on observed data
 #' @param Qrn A list of reduced-dimension regression estimates evaluated on observed data
 #' @param coefTol A tolerance level on the magnitude of the coefficient that flags the
 #' result as potentially the result of numeric instability.
 #' @param tolg The lower bound on propensity score estimates
-#' @param a0 A list of fixed treatment values 
+#' @param a_0 A list of fixed treatment values 
 #' 
 #' @importFrom SuperLearner trimLogit
 #' @importFrom stats predict glm
 #' 
 
-fluctuateG <- function(Y, A, W, a0, Qn, gn, Qrn, tolg, coefTol=1e5){
-  gnStar <- mapply(a=a0, Q=Qn, g=gn, Qr=Qrn, FUN=function(x, a, Q, g, Qr){
+fluctuateG <- function(Y, A, W, a_0, gn, Qrn, tolg, coefTol=1e5){
+  gnStar <- mapply(a=a_0, g=gn, Qr=Qrn, FUN=function(x, a, g, Qr){
     H1 <- Qr/g
[…]

### 2017-07-29 R/islptw.R

@@ -1,0 +1,180 @@
+#' Compute asymptotically linear IPTW estimators with super learning 
+#' for the propensity score
+#' 
+#' @param W A \code{data.frame} of named covariates
+#' @param A A vector of binary treatment assignment (assumed to be equal to 0 or 1)
+#' @param Y A numeric of continuous or binary outcomes. 
+#' @param a_0 A vector of treatment levels at which to compute the adjusted mean outcome. 
+#' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
+#' for the propensity score. See \code{link{SuperLearner::SuperLearner}} for details.
+#' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
+#' for the first reduced-dimension regression. 
+#' @param glm_g A character describing a formula to be used in the call to \code{glm} for the propensity score. Ignored
+#' if \code{SL_g!=NULL}.
+#' @param glm_Qr A character describing a formula to be used in the call to \code{glm} for the first reduced-dimension regression. Ignored
+#' if \code{SL_Qr!=NULL}. The formula should use the variable name \code{'gn'}.
+#' @param maxIter A numeric that sets the maximum number of iterations the TMLE can perform in its fluctuation step.
+#' @param tolIC A numeric that defines the stopping criteria based on the empirical mean
+#' of the scores of the fluctuation submodels submodels. Setting to \code{"default"}
+#' @param tolg A numeric indicating the minimum value for estimates of the propensity score.
+#' @param verbose A boolean indicating whether to print status updates.
+#' @param returnModels A boolean indicating whether to return model fits for the propensity score
+#' and reduced-dimension regressions.
+#' @return An object of class \code{"islptw"}.
+#' \describe{
+#'  \item{\code{islptw_tmle}}{A \code{list} of point estimates and 
+#'        covariance matrix for the IPTW estimator based on a targeted propensity 
+#' 		  score. }
[…]

### 2017-07-29 R/sl_npreg.R

@@ -1,4 +1,3 @@
-
 #' Super learner wrapper for kernel regression
 #' 
 #' Kernel regression based on the \href{https://cran.r-project.org/web/packages/np/}{np}
@@ -22,11 +21,16 @@
 SL.npreg <- function (Y, X, newX, family, obsWeights, 
                       rangeThresh=1e-7, ...) 
 {
+  # turn off annoying messages
+  options(np.messages=FALSE)
+  # check range threshold
   if(abs(diff(range(Y))) <= rangeThresh){
     thisMod <- glm(Y ~ 1, data=X)
   }else{
-  bw <- np::npregbw(stats::as.formula(paste("Y ~", paste(names(X),collapse="+"))), data=X,
-                ftol=0.01, tol=0.01, remin=FALSE)
+    # bandwidth selection
+    bw <- np::npregbw(stats::as.formula(
+            paste("Y ~", paste(names(X),collapse="+"))), data=X,
+          ftol=0.01, tol=0.01, remin=FALSE)
   
   # fit the kernel regression
   thisMod <- np::npreg(bw)

### 2017-07-29 R/utils.R

@@ -1,0 +1,65 @@
+#--------------------------------------------------
+# Print methods
+#--------------------------------------------------
+#' Print the output of a \code{"drtmle"} object.
+#' 
+#' @param x A \code{"drtmle"} object
+#' @param ... Other arguments (not used)
+#' @export
+#' @method print drtmle
+
+print.drtmle <- function(x,...){
+  tmp <- list(est = cbind(x$drtmle$est),
+  	          cov = x$drtmle$cov)
+  row.names(tmp$est) <- x$a_0
+  row.names(tmp$cov) <- colnames(tmp$cov) <- x$a_0
+  if(length(x$a_0) <= 4){
+  	print(tmp)
+  }else{
+  	tmp$cov <- diag(tmp$cov)
+  }
+}
+
+#' Print the output of a \code{"islptw"} object.
+#' 
+#' @param x A \code{"islptw"} object.
+#' @param ... Other arguments (not used)
+#' @export
+#' @method print islptw
+
+print.islptw <- function(x,...){
+  tmp <- list(est = cbind(x$islptw_tmle$est),
+  	          cov = x$islptw_tmle$cov)
+  row.names(tmp$est) <- x$a_0
+  row.names(tmp$cov) <- colnames(tmp$cov) <- x$a_0
+  if(length(x$a_0) <= 4){
+  	print(tmp)
+  }else{
+  	tmp$cov <- diag(tmp$cov)
+  }
+}
+
+
+#' Print the output of confint.drtmle
+#' @export
+#' @param x An object of class drconfint
+#' @param digits Number of digits to round to 
+#' @param ... Other options (not currently used)
+#' @method print confint.drtmle	
+
+print.confint.drtmle <- function(x,digits = 3,...){
+	tmp <- lapply(x, round, digits = digits)
+	print(tmp)
+}
+
+#' Print the output of confint.islptw
+#' @export
+#' @param x An object of class drconfint
+#' @param digits Number of digits to round to 
+#' @param ... Other options (not currently used)
+#' @method print confint.islptw 
+
+print.confint.islptw <- function(x,digits = 3,...){
+	tmp <- lapply(x, round, digits = digits)
+	print(tmp)
+}

### 2017-07-29 tests/testthat/test-confint.R

@@ -1,0 +1,124 @@
+library(drtmle)
+library(SuperLearner)
+
+context("Testing confint.drtmle method ")
+test_that("confint.drtmle works as expected",{
+	# simulate data
+	set.seed(123456)
+	n <- 200
+	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+	A <- rbinom(n,1,plogis(W$W1 - W$W2))
+	Y <- rbinom(n, 1, plogis(W$W1*W$W2*A))
+	# fit a drtmle
+	fit1 <- drtmle(W = W, A = A, Y = Y, a_0 = c(1,0),
+					  family=binomial(),
+	               stratify=FALSE,
+	               SL_Q="SL.glm",
+	               SL_g="SL.glm",
+	               SL_Qr="SL.glm",
+	               SL_gr="SL.glm")
+
+	# get confidence intervals for each mean for only drtmle
+	tmp <- confint(fit1)
+	# correct class
+	expect_true(class(tmp)=="confint.drtmle")
+	# no NAs
+	expect_true(sum(is.na(unlist(tmp)))==0)
+
+	# get confidence intervals for each mean for drtmle and tmle
+	tmp <- confint(fit1, parm = c("drtmle","tmle","aiptw","aiptw_c","gcomp"))
+
+	# correct class
+	expect_true(class(tmp)=="confint.drtmle")
+	# no NAs
+	expect_true(sum(is.na(unlist(tmp)))==0)
+	expect_true(length(tmp) == 5)
+	expect_true(all(names(tmp) == c("drtmle","tmle","aiptw","aiptw_c","gcomp")))
+	expect_true(nrow(tmp$drtmle) == 2)
+	expect_true(nrow(tmp$tmle) == 2)
+
+	# get confidence intervals for ATE
+	tmp <- confint(fit1, contrast = c(1,-1))
+	# correct class
+	expect_true(class(tmp)=="confint.drtmle")
+	# correct row name
+	expect_true(row.names(tmp$drtmle) == "E[Y(1)]-E[Y(0)]")
+	# no NAs
+	expect_true(sum(is.na(unlist(tmp)))==0)
+
+	# throws error if crazy contrast is put in 
+	expect_error(confint(fit1, contrast = c(10214,NA)))
+
+	# get confidence intervals for risk ratio 
+	# by inputting own contrast function
+	# this computes CI on log scale and back transforms
+	myContrast <- list(f = function(eff){ log(eff) },
[…]

### 2017-07-29 tests/testthat/test-drtmle.R

@@ -1,15 +1,9 @@
 # install.packages("~/Dropbox/R/drtmle",repos=NULL,type="source")
 library(drtmle)
-
+library(SuperLearner)
 context("Testing drtmle function")
 
 test_that("drtmle executes as expected with stratify = TRUE", {
-	set.seed(123456)
-	n <- 200
-	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
-	A <- rbinom(n,1,plogis(W$W1 - W$W2))
-	Y <- rnorm(n, W$W1*W$W2*A, 2)
-
 	set.seed(123456)
 	n <- 200
 	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
@@ -21,132 +15,132 @@ test_that("drtmle executes as expected with stratify = TRUE", {
 	fit1 <- drtmle(W = W, A = A, Y = Y, 
                    family=gaussian(),
                       stratify=TRUE,
-                      glmQ="W1 + W2",
-                      glmg="W1 + W2",
-                      glmQr="gn",
-                      glmgr="Qn",
+                      glm_Q="W1 + W2",
+                      glm_g="W1 + W2",
+                      glm_Qr="gn",
+                      glm_gr="Qn",
                       guard=c("Q","g"),
                       reduction="univariate")
 
-	expect_true(is.numeric(fit1$naive$est))
+	expect_true(is.numeric(fit1$gcomp$est))
 	expect_true(is.numeric(fit1$tmle$est))
 	expect_true(is.numeric(fit1$tmle$est))
 	expect_true(is.numeric(fit1$tmle$cov))
-	expect_true(is.numeric(fit1$tmle.dral$est))
-	expect_true(is.numeric(fit1$tmle.dral$cov))
-	expect_true(is.numeric(fit1$os$est))
-	expect_true(is.numeric(fit1$os$cov))
-	expect_true(is.numeric(fit1$os.dral$est))
-	expect_true(is.numeric(fit1$os.dral$cov))
+	expect_true(is.numeric(fit1$drtmle$est))
+	expect_true(is.numeric(fit1$drtmle$cov))
+	expect_true(is.numeric(fit1$aiptw$est))
+	expect_true(is.numeric(fit1$aiptw$cov))
+	expect_true(is.numeric(fit1$aiptw_c$est))
+	expect_true(is.numeric(fit1$aiptw_c$cov))
 
 	# bivariate reduction with 
 	# all GLMs + stratify 
[…]

### 2017-07-29 tests/testthat/test-islptw.R

@@ -1,0 +1,26 @@
+library(drtmle)
+library(SuperLearner)
+library(np)
+
+# TO DO: Add tests for multiple treatment levels
+context("Testing islptw works")
+
+test_that("islptw works as expected",{
+	set.seed(123456)
+	n <- 200
+	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+	A <- rbinom(n,1,plogis(W$W1 - W$W2))
+	Y <- rnorm(n, W$W1*W$W2*A, 2)
+
+	fit1 <- islptw(W = W, A = A, Y = Y, 
+	               a_0 = c(0,1),
+                  glm_g="W1 + W2",
+                  glm_Qr="gn",verbose = TRUE)
+	expect_true(all(!is.na(fit1$islptw_tmle$est)))
+	expect_true(all(!is.na(fit1$islptw_tmle$cov)))
+	expect_true(all(!is.na(fit1$islptw_os$est)))
+	expect_true(all(!is.na(fit1$islptw_os$cov)))
+	expect_true(all(!is.na(fit1$iptw$est)))
+	expect_true(all(!is.na(fit1$iptw$cov)))
+	expect_true(class(fit1)=="islptw")
+})	

### 2017-07-29 tests/testthat/test-slWrappers.R

@@ -1,242 +1,0 @@
-library(drtmle)
-library(caret)
-library(SuperLearner)
-library(randomForest)
-library(rpart)
-library(gam)
-library(gbm)
-library(kernlab)
-library(glmnet)
-library(np)
-library(randomGLM)
-library(e1071)
-library(nnet)
-
-context("Testing super learner wrappers")
-
-test_that("SL.npreg works as expected",{
-	set.seed(1234)
-	n <- 100
-	X <- data.frame(X1 = rnorm(n))
-	Y <- rnorm(n,X$X1,1)
-	fit <- SL.npreg(Y=Y,X=X,newX=X,
-	                obsWeights = rep(1,n))
-	expect_true(is.numeric(fit$pred))
-	expect_true(all(!is.na(fit$pred)))
-	expect_true(class(fit$fit)=="SL.npreg")
-	# test predict function
-	newX <- data.frame(X1=seq(-1,1,length=10))
-	pred <- predict(fit$fit, newdata=newX)
-	expect_true(is.numeric(pred))
-	expect_true(all(!is.na(pred)))
-})
-test_that("SL.svmLinear.caretMod works as expected",{
-	set.seed(1234)
-	n <- 100
-	X <- data.frame(X1 = rnorm(n))
-	Y <- rnorm(n,X$X1,1)
-	fit <- SL.svmLinear.caretMod(
-     Y=Y,X=X,newX=X,tuneLength = 2, 
-     trControl = caret::trainControl(method = "cv", number = 2),
-     obsWeights = rep(1,n)
-    )
-	expect_true(is.numeric(fit$pred))
-	expect_true(all(!is.na(fit$pred)))
-	expect_true(class(fit$fit)=="SL.caret")
-	# test predict function
-	newX <- data.frame(X1=seq(-1,1,length=10))
-	pred <- predict(fit$fit, newdata=newX)
-	expect_true(is.numeric(pred))
-	expect_true(all(!is.na(pred)))
-})
-test_that("SL.rpart.caretMod works as expected",{
-	set.seed(1234)
-	n <- 100
-	X <- data.frame(X1 = rnorm(n))
-	Y <- rnorm(n,X$X1,1)
-	fit <- SL.rpart.caretMod(
-     Y=Y,X=X,newX=X,tuneLength = 2, 
-     trControl = caret::trainControl(method = "cv", number = 2),
-     obsWeights = rep(1,n)
-    )
[…]

### 2017-07-29 README.Rmd
added nloptr to depends. minor readme updates

@@
 
-> Doubly Robust, Asymptotically Linear Targeted Minimum Loss-Based Estimation
+> Nonparametric estimators of the average treatment effect with doubly-robust confidence intervals and hypothesis tests
 
@@
 `drtmle` is an R package that computes marginal effect estimators for binary 
-treatments on continuous and binary outcomes. The TMLE estimators are doubly
-robust, not only with respect to consistency, but also with respect to asymptotic 
-normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017,(_accepted Biometrika_, [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).
+treatments on continuous and binary outcomes.
+The targeted minimum loss-based (TMLE) estimators are doubly robust, not only with respect to consistency, but also with respect to asymptotic normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017,(_accepted Biometrika_, [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).
+Therefore, it is possible to construct doubly-robust confidence intervals and hypothesis tests.
+Also included are methods for computing valid confidence intervals for an inverse probability of treatment weighted (IPTW) estimator of the average treatment effect when the propensity score is estimated via super learning, as discussed in [van der Laan, 2014](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
 

### 2017-07-29 R/estimate.R

@@ -135,7 +135,7 @@ estimateQ <- function(Y,A,W,SL_Q,glm_Q,a_0,stratify,family,verbose=FALSE,returnM
                                          verbose=verbose,family=family,
                                          SL.library=SL_Q,
                                          method =ifelse(family$family=="binomial",
-                                                           "method.CC_LS","method.CC_LS"),
+                                                        "method.CC_LS","method.CC_LS"),
                                          ...)
       
         Qn <- alply(a_0,1,function(x){

### 2017-07-29 README.Rmd
update md

@@
 ```{r}
+# load package
+library(drtmle)
+
 # simulate simple data structure
 set.seed(1234)
-n <- 500
+n <- 200
 W <- data.frame(W1 = runif(n,-2,2), W2 = rbinom(n,1,0.5))
@@
 # get confidence intervals for means
-ci_fit <- drconfint(fit)
+ci_fit <- confint(fit)
 # print the output
@@
 
-# get hypothesis test for means 
-test_fit <- drtest(fit)
-# print the output
-test_fit
 ```

### 2017-07-29 README.Rmd
remove warnings about fluctuation failures

@@
 treatments on continuous and binary outcomes.
-The targeted minimum loss-based (TMLE) estimators are doubly robust, not only with respect to consistency, but also with respect to asymptotic normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017,(_accepted Biometrika_, [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).
+The targeted minimum loss-based (TMLE) estimators are doubly robust, not only with respect to consistency, but also with respect to asymptotic normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017 (_accepted Biometrika_; [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).
 Therefore, it is possible to construct doubly-robust confidence intervals and hypothesis tests.

### 2017-07-29 R/fluctuate.R

@@ -66,7 +66,7 @@ fluctuateG <- function(Y, A, W, a_0, gn, Qrn, tolg, coefTol=1e5){
                 data=data.frame(thisA=thisA, off=off, H1=H1), family="binomial")
       )
       if(!fm$converged | abs(fm$coef) > coefTol){
-        warning("No sane fluctuation found for G this iteration. Check mean of IC.") 
+        # warning("No sane fluctuation found for G this iteration. Check mean of IC.") 
       }
     }
     pred <- stats::predict(fm, type="response")
@@ -117,7 +117,7 @@ fluctuateQ2 <- function(Y,A,W,Qn,gn,grn,a_0,reduction,coefTol=1e5){
                   data=data.frame(Y=Y, off=off, H2=H2), family="binomial")
       )
       if(!fm$converged | abs(max(fm$coef)) > coefTol){
-        warning("No sane fluctuation found. Proceeding using current estimates.")
+        # warning("No sane fluctuation found. Proceeding using current estimates.")
         if(reduction=="univariate"){
           return(list(est=Q,eps=rep(0,2)))
         }else if(reduction=="bivariate"){
@@ -178,7 +178,7 @@ fluctuateQ <- function(Y,A,W,Qn,gn,grn,a_0,reduction,coefTol=1e5){
                 data=data.frame(Y=Y, off=off, H1=H1, H2=H2), family="binomial")
       )
       if(!fm$converged | abs(max(fm$coef)) > coefTol){
-        warning("No sane fluctuation found. Proceeding using current estimates.")
+        # warning("No sane fluctuation found. Proceeding using current estimates.")
         if(reduction=="univariate"){
           return(list(est=Q,eps=rep(0,2)))
         }else if(reduction=="bivariate"){

### 2017-07-29 README.Rmd
add to README. add tests for multiple level treatment

@@
 
+### Doubly-robust inference for average treatment effect
+
 This package can be used to estimate covariate-adjusted marginal means under multiple discrete levels of a treatment.
@@
 The function computes doubly robust variance estimators that can be used to construct doubly robust confidence intervals for covariate-adjusted marginal means and contrasts between these means (via the `drconfint` function) for different levels of $a_0$.
-Doubly robust hypothesis tests are also available (via `drtest` function).
 A simple example on simulated data is shown below.
@@
 ```{r}
-# load package
+# load packages
 library(drtmle)
+library(SuperLearner)
 
@@
 # here, we do not properly estimate the propensity score
-fit <- drtmle(W = W, A = A, Y = Y, # input data
+fit1 <- drtmle(W = W, A = A, Y = Y, # input data
               a_0 = c(0,1), # return estimates for A = 0 and A = 1
@@
 # print the output
-fit
+fit1
 
-# get confidence intervals for means
-ci_fit <- confint(fit)
+# get confidence intervals for marginal means
+ci_fit1 <- confint(fit1)
 # print the output
-ci_fit
+ci_fit1
 
+# get confidence intervals for ate
+ci_ate1 <- confint(fit1,contrast = c(-1,1))
+# print the output
+ci_ate1
+```
+
+### Inference for super learner-based IPTW
+
+The package additionally includes a function for computing valid confidence intervals about an inverse probability of treatment weight (IPTW) estimator when super learning is used to estimate the propensity score.
+We call this estimator the inverse super learning probability of treatment estimator (ISLPTW).
+
+```{r}
+# fit iptw 
+fit2 <- islptw(Y = Y, A = A, W = W, a_0 = c(0,1),
+               SL_g = c("SL.glm","SL.mean","SL.step.interaction"),
+               SL_Qr = "SL.npreg")
+# print the output
+fit2
+
+# compute a confidence interval for margin means
+ci_fit2 <- confint(fit2)
+# print the output
+ci_fit2
+
+# compute a confidence interval for the ate
+ci_ate2 <- confint(fit2, contrast = c(-1,1))
+# print the output
+ci_ate2
 ```

### 2017-07-29 R/confint.R

@@ -88,8 +88,8 @@ confint.drtmle <- function(object, parm = c("drtmle"), level = 0.95,
 		if(length(contrast) != length(object$a_0)){
 			stop("length of contrast vector not equal to length of a_0")
 		}
-		if(!all(contrast %in% c(-1,1))){
-			stop("contrast should only be -1 or 1")
+		if(!all(contrast %in% c(-1,1,0))){
+			stop("contrast should only be -1, 1, or 0")
 		}
 		for(i in seq_along(parm)){
 			out[[i]] <- matrix(NA, nrow = 1, ncol = 3)
@@ -230,8 +230,8 @@ confint.islptw <- function(object, parm = c("islptw_tmle"), level = 0.95,
 		if(length(contrast) != length(object$a_0)){
 			stop("length of contrast vector not equal to length of a_0")
 		}
-		if(!all(contrast %in% c(-1,1))){
-			stop("contrast should only be -1 or 1")
+		if(!all(contrast %in% c(-1,1,0))){
+			stop("contrast should only be -1, 1, or 0")
 		}
 		for(i in seq_along(parm)){
 			out[[i]] <- matrix(NA, nrow = 1, ncol = 3)

### 2017-07-29 R/estimate.R

@@ -415,7 +415,7 @@ estimategrn <- function(Y, A, W, Qn, gn, SL_gr, tolg, glm_gr, a_0, reduction,ret
         glm_gr <- "1"
       }
       fm2 <- stats::glm(stats::as.formula(paste0("A~",glm_gr)), family="binomial",
-           data=data.frame(A=A, Qn=Q))
+           data=data.frame(A=as.numeric(A==a), Qn=Q))
       grn2 <- stats::predict(fm2, type="response")
       grn2[grn2 < tolg] <- tolg
       if(reduction == "univariate"){

### 2017-07-29 tests/testthat/test-multileveltrt.R

@@ -1,0 +1,128 @@
+library(drtmle)
+library(SuperLearner)
+context("Testing drtmle function with > 2 treatment levels")
+
+test_that("drtmle executes as expected with multiple treatment levels", {
+	set.seed(123456)
+	n <- 200
+	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+	A <- rbinom(n,1,plogis(W$W1 - W$W2)) + rbinom(n,1,plogis(W$W2))
+	Y <- rnorm(n, W$W1*W$W2*A, 2)
+
+	# univariate reduction with
+	# all GLMs + stratify
+	fit1 <- drtmle(W = W, A = A, Y = Y, a_0=c(0,1,2),
+                   family=gaussian(),
+                      stratify=TRUE,
+                      glm_Q="W1 + W2",
+                      glm_g="W1 + W2",
+                      glm_Qr="gn",
+                      glm_gr="Qn",
+                      guard=c("Q","g"),
+                      reduction="univariate")
+
+	expect_true(is.numeric(fit1$gcomp$est))
+	expect_true(is.numeric(fit1$tmle$est))
+	expect_true(is.numeric(fit1$tmle$est))
+	expect_true(is.numeric(fit1$tmle$cov))
+	expect_true(is.numeric(fit1$drtmle$est))
+	expect_true(is.numeric(fit1$drtmle$cov))
+	expect_true(is.numeric(fit1$aiptw$est))
+	expect_true(is.numeric(fit1$aiptw$cov))
+	expect_true(is.numeric(fit1$aiptw_c$est))
+	expect_true(is.numeric(fit1$aiptw_c$cov))
+
+	# check confint works
+	ci <- confint(fit1)
+	expect_true(length(fit1$drtmle$est)==3)
+	# check contrasts work
+	ci2 <- confint(fit1, contrast = c(1,-1,0))
+	expect_true(row.names(ci2$drtmle) == "E[Y(0)]-E[Y(1)]")
+	ci3 <- confint(fit1, contrast = c(1,0,-1))
+	expect_true(row.names(ci3$drtmle) == "E[Y(0)]-E[Y(2)]")
+
+
+	# same thing but with super learner for g
+	fit1 <- drtmle(W = W, A = A, Y = Y, a_0=c(0,1,2),
+                   family=gaussian(),
+                      stratify=TRUE,
[…]

### 2017-07-31 README.Rmd
added testing functions, changed confint to ci

@@
 The resulting targeted minimum loss-based estimates are doubly robust with respect to both consistency and asymptotic normality.
-The function computes doubly robust variance estimators that can be used to construct doubly robust confidence intervals for covariate-adjusted marginal means and contrasts between these means (via the `drconfint` function) for different levels of $a_0$.
+The function computes doubly robust variance estimators that can be used to construct doubly robust confidence intervals for covariate-adjusted marginal means and contrasts between these means (via the `ci` method) for different levels of $a_0$.
 A simple example on simulated data is shown below.
@@
 # get confidence intervals for marginal means
-ci_fit1 <- confint(fit1)
+ci_fit1 <- ci(fit1)
 # print the output
@@
 # get confidence intervals for ate
-ci_ate1 <- confint(fit1,contrast = c(-1,1))
+ci_ate1 <- ci(fit1,contrast = c(-1,1))
 # print the output
@@
 # compute a confidence interval for margin means
-ci_fit2 <- confint(fit2)
+ci_fit2 <- ci(fit2)
 # print the output
@@
 # compute a confidence interval for the ate
-ci_ate2 <- confint(fit2, contrast = c(-1,1))
+ci_ate2 <- ci(fit2, contrast = c(-1,1))
 # print the output

### 2017-07-31 R/sl_npreg.R

@@ -32,8 +32,8 @@ SL.npreg <- function (Y, X, newX, family, obsWeights,
             paste("Y ~", paste(names(X),collapse="+"))), data=X,
           ftol=0.01, tol=0.01, remin=FALSE)
   
-  # fit the kernel regression
-  thisMod <- np::npreg(bw)
+    # fit the kernel regression
+    thisMod <- np::npreg(bw)
   }
   
   pred <- stats::predict(thisMod, newdata=newX)

### 2017-07-31 R/test.R

@@ -1,11 +1,11 @@
-#' Compute confidence intervals for drtmle and islptw@
+#' Wald tests for drtmle and islptw objects
 #' @param ... Arguments to be passed to method
 #' @export
-ci <- function(...){
-  UseMethod("ci")
+wald_test <- function(...){
+	UseMethod("wald_test")
 }
 
-#' Confidence intervals for drtmle objects
+#' Wald tests for drtmle objects
 #' 
 #' @param object An object of class \code{"drtmle"}
 #' @param est A vector indicating for which estimators to return a 
@@ -14,24 +14,29 @@ ci <- function(...){
 #' for misspecification (\code{"aiptw_c"}, not recommended), the standard TMLE 
 #' (\code{"tmle"}, recommended only for comparison to "drtmle"), the standard 
 #' AIPTW (\code{"aiptw"}, recommended only for comparison to "drtmle"), and 
-#' G-computation (\code{"gcomp"}, not recommended).  
-#' @param level The level of the confidence interval
+#' G-computation (\code{"gcomp"}, not recommended).
+#' @param null The null hypothesis value. 
 #' @param contrast This option specifies what parameter to return confidence intervals for.
-#' If \code{contrast=NULL}, then compute confidence intervals for the covariate-adjusted
-#' marginal means. \code{contrast} can also be input as a numeric vector of ones, negative ones,
+#' If \code{contrast=NULL}, then test the null hypothesis that the covariate-adjusted
+#' marginal means equal the value(s) specified in \code{null}. 
+#' \code{contrast} can also be a numeric vector of ones, negative ones,
 #' and zeros to define linear combinations of the various means (e.g., to estimate an average
-#' treatment effect, see examples). \code{contrast} can also be a list with named functions
-#' \code{f}, \code{f_inv}, \code{h}, and \code{h_grad}. The first two functions should take
-#' as input argument \code{eff}. Respectively, these specify which transformation 
-#' of the effect measure to compute the confidence interval for and the inverse
-#' transformation to put the confidence interval back on the original scale. The function \code{h}
-#' defines the contrast to be estimated and should take as input \code{est}, a vector
[…]

### 2017-07-31 R/utils.R

@@ -40,26 +40,50 @@ print.islptw <- function(x,...){
 }
 
 
-#' Print the output of confint.drtmle
+#' Print the output of ci.drtmle
 #' @export
-#' @param x An object of class drconfint
+#' @param x An object of class ci.drtmle
 #' @param digits Number of digits to round to 
 #' @param ... Other options (not currently used)
-#' @method print confint.drtmle	
+#' @method print ci.drtmle	
 
-print.confint.drtmle <- function(x,digits = 3,...){
+print.ci.drtmle <- function(x,digits = 3,...){
 	tmp <- lapply(x, round, digits = digits)
 	print(tmp)
 }
 
-#' Print the output of confint.islptw
+#' Print the output of ci.islptw
 #' @export
-#' @param x An object of class drconfint
+#' @param x An object of class ci.islptw
 #' @param digits Number of digits to round to 
 #' @param ... Other options (not currently used)
-#' @method print confint.islptw 
+#' @method print ci.islptw 
 
-print.confint.islptw <- function(x,digits = 3,...){
+print.ci.islptw <- function(x,digits = 3,...){
 	tmp <- lapply(x, round, digits = digits)
 	print(tmp)
+}
+
+#' Print the output of wald_test.drtmle
+#' @export
+#' @param x An object of class wald_test.drtmle
+#' @param digits Number of digits to round to 
+#' @param ... Other options (not currently used)
+#' @method print wald_test.drtmle
+
+print.wald_test.drtmle <- function(x,digits = 3,...){
+  tmp <- lapply(x, round, digits = digits)
+  print(tmp)
+}
+
+#' Print the output of wald_test.islptw
+#' @export
+#' @param x An object of class wald_test.islptw
+#' @param digits Number of digits to round to 
+#' @param ... Other options (not currently used)
+#' @method print wald_test.islptw 
+
+print.wald_test.islptw <- function(x,digits = 3,...){
+  tmp <- lapply(x, round, digits = digits)
+  print(tmp)
 }

### 2017-07-31 tests/testthat/test-multileveltrt.R

@@ -35,13 +35,13 @@ test_that("drtmle executes as expected with multiple treatment levels", {
 	expect_true(is.numeric(fit1$aiptw_c$est))
 	expect_true(is.numeric(fit1$aiptw_c$cov))
 
-	# check confint works
-	ci <- confint(fit1)
+	# check ci works
+	ci <- ci(fit1)
 	expect_true(length(fit1$drtmle$est)==3)
 	# check contrasts work
-	ci2 <- confint(fit1, contrast = c(1,-1,0))
+	ci2 <- ci(fit1, contrast = c(1,-1,0))
 	expect_true(row.names(ci2$drtmle) == "E[Y(0)]-E[Y(1)]")
-	ci3 <- confint(fit1, contrast = c(1,0,-1))
+	ci3 <- ci(fit1, contrast = c(1,0,-1))
 	expect_true(row.names(ci3$drtmle) == "E[Y(0)]-E[Y(2)]")
 
 
@@ -68,13 +68,13 @@ test_that("drtmle executes as expected with multiple treatment levels", {
 	expect_true(is.numeric(fit1$aiptw_c$est))
 	expect_true(is.numeric(fit1$aiptw_c$cov))
 
-	# check confint works
-	ci <- confint(fit1)
+	# check ci works
+	ci <- ci(fit1)
 	expect_true(length(fit1$drtmle$est)==3)
 	# check contrasts work
-	ci2 <- confint(fit1, contrast = c(1,-1,0))
+	ci2 <- ci(fit1, contrast = c(1,-1,0))
 	expect_true(row.names(ci2$drtmle) == "E[Y(0)]-E[Y(1)]")
-	ci3 <- confint(fit1, contrast = c(1,0,-1))
+	ci3 <- ci(fit1, contrast = c(1,0,-1))
 	expect_true(row.names(ci3$drtmle) == "E[Y(0)]-E[Y(2)]")
 })
 
@@ -98,13 +98,13 @@ test_that("islptw executes as expected with multiple treatment levels", {
 	expect_true(is.numeric(fit1$islptw_os$est))
 	expect_true(is.numeric(fit1$islptw_os$cov))
 	expect_true(is.numeric(fit1$islptw_tmle$cov))
-	# check confint works
-	ci <- confint(fit1)
+	# check ci works
+	ci <- ci(fit1)
 	expect_true(length(fit1$islptw_tmle$est)==3)
[…]

### 2017-07-31 tests/testthat/test-tests.R

@@ -1,8 +1,8 @@
 library(drtmle)
 library(SuperLearner)
 
-context("Testing ci.drtmle method ")
-test_that("ci.drtmle works as expected",{
+context("Testing wald_test.drtmle method ")
+test_that("wald_test.drtmle works as expected",{
 	# simulate data
 	set.seed(123456)
 	n <- 200
@@ -18,18 +18,18 @@ test_that("ci.drtmle works as expected",{
 	               SL_Qr="SL.glm",
 	               SL_gr="SL.glm")
 
-	# get confidence intervals for each mean for only drtmle
-	tmp <- ci(fit1)
+	# get test for each mean for only drtmle
+	tmp <- wald_test(fit1)
 	# correct class
-	expect_true(class(tmp)=="ci.drtmle")
+	expect_true(class(tmp)=="wald_test.drtmle")
 	# no NAs
 	expect_true(sum(is.na(unlist(tmp)))==0)
 
-	# get confidence intervals for each mean for drtmle and tmle
-	tmp <- ci(fit1, est = c("drtmle","tmle","aiptw","aiptw_c","gcomp"))
+	# get test for each mean for drtmle and tmle
+	tmp <- wald_test(fit1, est = c("drtmle","tmle","aiptw","aiptw_c","gcomp"))
 
 	# correct class
-	expect_true(class(tmp)=="ci.drtmle")
+	expect_true(class(tmp)=="wald_test.drtmle")
 	# no NAs
 	expect_true(sum(is.na(unlist(tmp)))==0)
 	expect_true(length(tmp) == 5)
@@ -37,59 +37,57 @@ test_that("ci.drtmle works as expected",{
 	expect_true(nrow(tmp$drtmle) == 2)
 	expect_true(nrow(tmp$tmle) == 2)
 
-	# get confidence intervals for ATE
-	tmp <- ci(fit1, contrast = c(1,-1))
+	# get test for ATE
+	tmp <- wald_test(fit1, contrast = c(1,-1))
 	# correct class
-	expect_true(class(tmp)=="ci.drtmle")
+	expect_true(class(tmp)=="wald_test.drtmle")
 	# correct row name
-	expect_true(row.names(tmp$drtmle) == "E[Y(1)]-E[Y(0)]")
+	expect_true(row.names(tmp$drtmle) == "H0:E[Y(1)]-E[Y(0)]=0")
 	# no NAs
 	expect_true(sum(is.na(unlist(tmp)))==0)
 
 	# throws error if crazy contrast is put in 
[…]

### 2017-08-15 README.Rmd
update README, family argument for estimateQr in adaptive_iptw, some file renames, add vignette

@@
 ---
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
 
@@
 
-`drtmle` is an R package that computes marginal effect estimators for binary 
-treatments on continuous and binary outcomes.
-The targeted minimum loss-based (TMLE) estimators are doubly robust, not only with respect to consistency, but also with respect to asymptotic normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017 (_accepted Biometrika_; [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).
-Therefore, it is possible to construct doubly-robust confidence intervals and hypothesis tests.
-Also included are methods for computing valid confidence intervals for an inverse probability of treatment weighted (IPTW) estimator of the average treatment effect when the propensity score is estimated via super learning, as discussed in [van der Laan, 2014](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
+`drtmle` is an R package that computes marginal means of an outcome under fixed levels of a treatment.
+The package computes targeted minimum loss-based (TMLE) estimators that are doubly robust, not only with respect to consistency, but also with respect to asymptotic normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017 (_accepted Biometrika_; [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).
+This property facilitates construction of doubly-robust confidence intervals and hypothesis tests.
+
+The package additionally includes methods for computing valid confidence intervals for an inverse probability of treatment weighted (IPTW) estimator of the average treatment effect when the propensity score is estimated via super learning, as discussed in [van der Laan, 2014](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
 
@@
 
-This package can be used to estimate covariate-adjusted marginal means under multiple discrete levels of a treatment.
-It may be used in situations where data consist of a vector of baseline covariates (`W`), a multi-level treatment assignment (`A`), and a continuous or binary-valued outcome (`Y`).
+Suppose the data consist of a vector of baseline covariates (`W`), a multi-level treatment assignment (`A`), and a continuous or binary-valued outcome (`Y`).
 The function `drtmle` may be used to estimate $E[E(Y | A=a_0, W)]$ for user-selected values of $a_0$ (via option `a_0`).
 The resulting targeted minimum loss-based estimates are doubly robust with respect to both consistency and asymptotic normality.
-The function computes doubly robust variance estimators that can be used to construct doubly robust confidence intervals for covariate-adjusted marginal means and contrasts between these means (via the `ci` method) for different levels of $a_0$.
+The function computes doubly robust covariance estimates that can be used to construct doubly robust confidence intervals for marginal means and contrasts between means.
 A simple example on simulated data is shown below.
+We refer users to the vignette for more information and further examples.
 
@@
 # load packages
-library(drtmle)
-library(SuperLearner)
+library(drtmle, quietly = TRUE)
+library(SuperLearner, quietly = TRUE)
 
@@
 The package additionally includes a function for computing valid confidence intervals about an inverse probability of treatment weight (IPTW) estimator when super learning is used to estimate the propensity score.
-We call this estimator the inverse super learning probability of treatment estimator (ISLPTW).
 
@@
 # fit iptw 
-fit2 <- islptw(Y = Y, A = A, W = W, a_0 = c(0,1),
+fit2 <- adaptive_iptw(Y = Y, A = A, W = W, a_0 = c(0,1),
                SL_g = c("SL.glm","SL.mean","SL.step.interaction"),
@@
 If you encounter any bugs or have any specific feature requests, please [file an
-issue](https://github.com/benkeser/survtmle/issues).
+issue](https://github.com/benkeser/drtmle/issues).
 

### 2017-08-15 R/adaptive_iptw.R

@@ -11,6 +11,8 @@ globalVariables(c("v", "%dopar%"))
 #' @param a_0 A vector of treatment levels at which to compute the adjusted mean outcome. 
 #' @param stratify A \code{boolean} indicating whether to estimate the missing outcome regression separately
 #' for observations with different levels of \code{A} (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
+#' @param family A \code{family} object equal to either \code{binomial()} or \code{gaussian()}, 
+#' to be passed to the \code{SuperLearner} or \code{glm} function.
 #' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
 #' for the propensity score. See \code{link{SuperLearner::SuperLearner}} for details.
 #' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
@@ -32,17 +34,17 @@ globalVariables(c("v", "%dopar%"))
 #' @param parallel A boolean indicating whether to use \code{foreach}
 #' to estimate nuisance parameters in parallel. Only useful if there is a registered parallel
 #' backend (see examples) and \code{cvFolds > 1}.
-#' @return An object of class \code{"islptw"}.
+#' @return An object of class \code{"adaptive_iptw"}.
 #' \describe{
-#'  \item{\code{islptw_tmle}}{A \code{list} of point estimates and 
+#'  \item{\code{iptw_tmle}}{A \code{list} of point estimates and 
 #'        covariance matrix for the IPTW estimator based on a targeted propensity 
 #' 		  score. }
-#'  \item{\code{islptw_tmle_nuisance}}{A \code{list} of the final TMLE estimates of the
+#'  \item{\code{iptw_tmle_nuisance}}{A \code{list} of the final TMLE estimates of the
 #'        propensity score (\code{$gnStar}) and reduced-dimension regression (\code{$QrnStar}) 
 #' 		  evaluated at the observed data values.}
-#'  \item{\code{islptw_os}}{A \code{list} of point estimates and covariance matrix for the
+#'  \item{\code{iptw_os}}{A \code{list} of point estimates and covariance matrix for the
[…]

### 2017-08-15 R/confint.R

@@ -1,4 +1,4 @@
-#' Compute confidence intervals for drtmle and islptw@
+#' Compute confidence intervals for drtmle and adaptive_iptw@
 #' @param ... Arguments to be passed to method
 #' @export
 ci <- function(...){
@@ -147,15 +147,15 @@ ci.drtmle <- function(object, est = c("drtmle"), level = 0.95,
 	return(out)
 }
 
-#' Confidence intervals for islptw objects
+#' Confidence intervals for adaptive_iptw objects
 #' 
-#' Estimate confidence intervals for objects of class \code{"islptw"}
+#' Estimate confidence intervals for objects of class \code{"adaptive_iptw"}
 #' 
-#' @param object An object of class \code{"islptw"}
+#' @param object An object of class \code{"adaptive_iptw"}
 #' @param est A vector indicating for which estimators to return a 
 #' confidence interval. Possible estimators include the TMLE IPTW 
-#' (\code{"islptw_tmle"}, recommended), the one-step IPTW 
-#' (\code{"islptw_os"}, not recommended), the standard IPTW 
+#' (\code{"iptw_tmle"}, recommended), the one-step IPTW 
+#' (\code{"iptw_os"}, not recommended), the standard IPTW 
 #' (\code{"iptw"}, recommended only for comparison to the other two estimators).
 #' @param level The level of the confidence interval
 #' @param contrast This option specifies what parameter to return confidence intervals for.
@@ -172,21 +172,21 @@ ci.drtmle <- function(object, est = c("drtmle"), level = 0.95,
 #' \code{fh_grad} is the gradient of the function \code{h}. See examples. 
 #' @param ... Other options (not currently used)
 #' @export
-#' @method ci islptw
-#' @return An object of class \code{"ci.islptw"} with point estimates and
+#' @method ci adaptive_iptw
+#' @return An object of class \code{"ci.adaptive_iptw"} with point estimates and
 #' confidence intervals of the specified level. 
 #' 
 #' @examples
 #' # load super learner
 #' library(SuperLearner)
-#' # fit islptw
+#' # fit adaptive_iptw
 #' set.seed(123456)
 #' n <- 200
 #' W <- data.frame(W1 = runif(n), W2 = rnorm(n))
[…]

### 2017-08-15 R/drtmle.R

@@ -226,12 +226,12 @@ drtmle <- function(Y, A, W,
     if(!parallel){
       QrnOut <- lapply(X = validRows, FUN = estimateQrn, 
                        Y=Y, A=A, W=W, DeltaA = DeltaA, DeltaY = DeltaY, 
-                       Qn=Qn, gn=gn, glm_Qr=glm_Qr, 
+                       Qn=Qn, gn=gn, glm_Qr=glm_Qr, family = stats::gaussian(),
                        SL_Qr=SL_Qr, a_0=a_0,returnModels = returnModels)
     }else{
       QrnOut <- foreach::foreach(v = 1:cvFolds, .packages = "SuperLearner") %dopar% {
         estimateQrn(Y=Y, A=A, W=W, DeltaA = DeltaA, DeltaY = DeltaY, 
-                    Qn=Qn, gn=gn, glm_Qr=glm_Qr, 
+                    Qn=Qn, gn=gn, glm_Qr=glm_Qr, family = stats::gaussian(),
                     SL_Qr=SL_Qr, a_0=a_0,returnModels = returnModels,
                     validRows = validRows[[v]])
       }
@@ -364,7 +364,7 @@ drtmle <- function(Y, A, W,
       }else if(Qsteps==2){
         # do the extra targeting
         QnStarOut2 <- fluctuateQ2(Y=Y, A=A, W=W, DeltaA = DeltaA, DeltaY = DeltaY, 
-                                  a_0=a_0, Qn=QnStar, 
+                                  a_0=a_0, Qn=QnStar,   
                                   gn=gnStar, grn=grnStar, reduction=reduction)
         QnStar <- plyr::llply(QnStarOut2, function(x){unlist(x[[1]])})
         
@@ -387,12 +387,12 @@ drtmle <- function(Y, A, W,
       if(!parallel){
         QrnStarOut <- lapply(X = validRows, FUN = estimateQrn, 
                          Y=Y, A=A, W=W, DeltaA = DeltaA, DeltaY = DeltaY, 
-                         Qn=QnStar, gn=gnStar, glm_Qr=glm_Qr, 
+                         Qn=QnStar, gn=gnStar, glm_Qr=glm_Qr, family = stats::gaussian(), 
                          SL_Qr=SL_Qr, a_0=a_0,returnModels = returnModels)
       }else{
         QrnStarOut <- foreach::foreach(v = 1:cvFolds, .packages = "SuperLearner") %dopar% {
[…]

### 2017-08-15 R/estimate.R

@@ -100,9 +100,12 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
       gn_DeltaA <- stats::predict(fm_DeltaA, newdata = data.frame(DeltaA = validDeltaA, validW),
                                   type = "response")
     }
+    # name for returned models
+    name_DeltaA <- "DeltaA ~ W"
   }else{ 
     # if all DeltaA==1 then put NULL model and 1 predictions
     fm_DeltaA <- NULL
+    name_DeltaA <- ""
     gn_DeltaA <- rep(1, length(validDeltaA))
   }
 
@@ -124,12 +127,15 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
                                            method = "method.CC_nloglik")
         gn_A <- vector(mode="list",length=2)
         gn_A[[1]] <- fm_A$SL.predict; gn_A[[2]] <- 1-gn_A[[1]]
+        # name for this model 
+        name_A <- paste0("I(A = ",a_0[1],") ~ W | DeltaA == 1")
       # if there are more than two unique values of A, then we need more
       # than one call to super learner
       }else{
         a_ct <- 0
         gn_A <- vector(mode = "list", length = length(a_0))
         fm_A <- vector(mode = "list", length = length(a_0) - 1)
+        name_A <- rep(NA, length(a_0) - 1)
         for(a in a_0[1:(length(a_0)-1)]){
           # determine who to include in the regression for this outcome
           if(a_ct == 0){  
@@ -155,6 +161,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
             gn_A[[a_ct + 1]] <- tmp_pred
           }
           fm_A[[a_ct + 1]] <- tmp_fm
+          name_A[a_ct + 1] <- paste0("I(A = ",a,") ~ W | DeltaA == 1")
           a_ct <- a_ct + 1
         }
         # add in final predictions
@@ -170,10 +177,12 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
                       family=stats::binomial()
                       ))
         gn_A[[1]] <- fm_A$pred; gn_A[[2]] <- 1 - fm_A$pred
+        name_A <- paste0("I(A = ",a_0[1],") ~ W | DeltaA == 1")
       }else{
         a_ct <- 0
         gn_A <- vector(mode = "list", length = length(a_0))
[…]

### 2017-08-15 R/test.R

@@ -1,4 +1,4 @@
-#' Wald tests for drtmle and islptw objects
+#' Wald tests for drtmle and adaptive_iptw objects
 #' @param ... Arguments to be passed to method
 #' @export
 wald_test <- function(...){
@@ -166,13 +166,13 @@ wald_test.drtmle <- function(object, est = c("drtmle"), null = 0,
 	return(out)
 }
 
-#' Wald tests for islptw objects
+#' Wald tests for adaptive_iptw objects
 #' 
-#' @param object An object of class \code{"islptw"}
+#' @param object An object of class \code{"adaptive_iptw"}
 #' @param est A vector indicating for which estimators to return a 
 #' confidence interval. Possible estimators include the TMLE IPTW 
-#' (\code{"islptw_tmle"}, recommended), the one-step IPTW 
-#' (\code{"islptw_os"}, not recommended), the standard IPTW 
+#' (\code{"iptw_tmle"}, recommended), the one-step IPTW 
+#' (\code{"iptw_os"}, not recommended), the standard IPTW 
 #' (\code{"iptw"}, recommended only for comparison to the other two estimators).
 #' @param null The null hypothesis value(s). 
 #' @param contrast This option specifies what parameter to return confidence intervals for.
@@ -195,21 +195,21 @@ wald_test.drtmle <- function(object, est = c("drtmle"), null = 0,
 #' 
 #' @importFrom stats pnorm
 #' @export
-#' @method wald_test islptw
-#' @return An object of class \code{"ci.islptw"} with point estimates and
+#' @method wald_test adaptive_iptw
+#' @return An object of class \code{"ci.adaptive_iptw"} with point estimates and
 #' confidence intervals of the specified level. 
 #' 
 #' @examples
 #' # load super learner
 #' library(SuperLearner)
-#' # fit islptw
+#' # fit adaptive_iptw
 #' set.seed(123456)
 #' n <- 200
 #' W <- data.frame(W1 = runif(n), W2 = rnorm(n))
 #' A <- rbinom(n,1,plogis(W$W1 - W$W2))
 #' Y <- rbinom(n, 1, plogis(W$W1*W$W2*A))
 #'
-#' fit1 <- islptw(W = W, A = A, Y = Y, a_0 = c(1,0),
+#' fit1 <- adaptive_iptw(W = W, A = A, Y = Y, a_0 = c(1,0),
 #'                SL_g=c("SL.glm","SL.mean","SL.step"),
[…]

### 2017-08-15 R/utils.R

@@ -22,16 +22,16 @@ print.drtmle <- function(x,...){
   invisible(tmp)
 }
 
-#' Print the output of a \code{"islptw"} object.
+#' Print the output of a \code{"adaptive_iptw"} object.
 #' 
-#' @param x A \code{"islptw"} object.
+#' @param x A \code{"adaptive_iptw"} object.
 #' @param ... Other arguments (not used)
 #' @export
-#' @method print islptw
+#' @method print adaptive_iptw
 
-print.islptw <- function(x,...){
-  tmp <- list(est = cbind(x$islptw_tmle$est),
-  	          cov = x$islptw_tmle$cov)
+print.adaptive_iptw <- function(x,...){
+  tmp <- list(est = cbind(x$iptw_tmle$est),
+  	          cov = x$iptw_tmle$cov)
   row.names(tmp$est) <- x$a_0
   colnames(tmp$est) <- ""
   row.names(tmp$cov) <- colnames(tmp$cov) <- x$a_0
@@ -57,14 +57,14 @@ print.ci.drtmle <- function(x,digits = 3,...){
   invisible(tmp)
 }
 
-#' Print the output of ci.islptw
+#' Print the output of ci.adaptive_iptw
 #' @export
-#' @param x An object of class ci.islptw
+#' @param x An object of class ci.adaptive_iptw
 #' @param digits Number of digits to round to 
 #' @param ... Other options (not currently used)
-#' @method print ci.islptw 
+#' @method print ci.adaptive_iptw 
 
-print.ci.islptw <- function(x,digits = 3,...){
+print.ci.adaptive_iptw <- function(x,digits = 3,...){
 	tmp <- lapply(x, round, digits = digits)
 	print(tmp)
   invisible(tmp)
@@ -83,14 +83,14 @@ print.wald_test.drtmle <- function(x,digits = 3,...){
   invisible(tmp)
 }
 
-#' Print the output of wald_test.islptw
+#' Print the output of wald_test.adaptive_iptw
 #' @export
-#' @param x An object of class wald_test.islptw
+#' @param x An object of class wald_test.adaptive_iptw
 #' @param digits Number of digits to round to 
 #' @param ... Other options (not currently used)
-#' @method print wald_test.islptw 
+#' @method print wald_test.adaptive_iptw 
 
-print.wald_test.islptw <- function(x,digits = 3,...){
+print.wald_test.adaptive_iptw <- function(x,digits = 3,...){
[…]

### 2017-08-15 tests/testthat/test-adaptive_iptw-missingAY.R

@@ -4,8 +4,8 @@ library(np)
 library(testthat)
 
 # TO DO: Add tests for multiple treatment levels
-context("Testing islptw works")
-test_that("islptw works as expected with missing data",{
+context("Testing adaptive_iptw works")
+test_that("adaptive_iptw works as expected with missing data",{
 	set.seed(123456)
 	n <- 200
 	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
@@ -16,14 +16,14 @@ test_that("islptw works as expected with missing data",{
 	Y[DeltaY == 0] <- NA
 	A[DeltaA == 0] <- NA
 
-	fit1 <- islptw(W = W, A = A, Y = Y, 
+	fit1 <- adaptive_iptw(W = W, A = A, Y = Y, 
 	               a_0 = c(0,1),
                   glm_g="W1 + W2",
                   glm_Qr="gn")
-	expect_true(all(!is.na(fit1$islptw_tmle$est)))
-	expect_true(all(!is.na(fit1$islptw_tmle$cov)))
-	expect_true(all(!is.na(fit1$islptw_os$est)))
-	expect_true(all(!is.na(fit1$islptw_os$cov)))
+	expect_true(all(!is.na(fit1$iptw_tmle$est)))
+	expect_true(all(!is.na(fit1$iptw_tmle$cov)))
+	expect_true(all(!is.na(fit1$iptw_os$est)))
+	expect_true(all(!is.na(fit1$iptw_os$cov)))
 	expect_true(all(!is.na(fit1$iptw$est)))
-	expect_true(class(fit1)=="islptw")
+	expect_true(class(fit1)=="adaptive_iptw")
 })

### 2017-08-15 tests/testthat/test-adaptive_iptw.R

@@ -1,0 +1,66 @@
+library(drtmle)
+library(SuperLearner)
+library(np)
+library(foreach)
+
+context("Testing adaptive_iptw works")
+test_that("adaptive_iptw works as expected with parallel = TRUE",{
+	set.seed(123456)
+	n <- 200
+	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+	A <- rbinom(n,1,plogis(W$W1 - W$W2))
+	Y <- rnorm(n, W$W1*W$W2*A, 2)
+
+	fit1 <- adaptive_iptw(W = W, A = A, Y = Y, 
+	               a_0 = c(0,1),
+                  glm_g="W1 + W2",
+                  glm_Qr="gn",parallel = TRUE)
+	expect_true(all(!is.na(fit1$iptw_tmle$est)))
+	expect_true(all(!is.na(fit1$iptw_tmle$cov)))
+	expect_true(all(!is.na(fit1$iptw_os$est)))
+	expect_true(all(!is.na(fit1$iptw_os$cov)))
+	expect_true(all(!is.na(fit1$iptw$est)))
+	expect_true(all(!is.na(fit1$iptw$cov)))
+	expect_true(class(fit1)=="adaptive_iptw")
+})
+test_that("adaptive_iptw works as expected",{
+	set.seed(123456)
+	n <- 200
+	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+	A <- rbinom(n,1,plogis(W$W1 - W$W2))
+	Y <- rnorm(n, W$W1*W$W2*A, 2)
+
+	fit1 <- adaptive_iptw(W = W, A = A, Y = Y, 
+	               a_0 = c(0,1),
+                  glm_g="W1 + W2",
+                  glm_Qr="gn")
+	expect_true(all(!is.na(fit1$iptw_tmle$est)))
+	expect_true(all(!is.na(fit1$iptw_tmle$cov)))
+	expect_true(all(!is.na(fit1$iptw_os$est)))
+	expect_true(all(!is.na(fit1$iptw_os$cov)))
+	expect_true(all(!is.na(fit1$iptw$est)))
+	expect_true(all(!is.na(fit1$iptw$cov)))
+	expect_true(class(fit1)=="adaptive_iptw")
+})	
+
+test_that("adaptive_iptw works as expected with multi-level treatment",{
+	set.seed(123456)
[…]

### 2017-08-15 tests/testthat/test-confint.R

@@ -63,10 +63,10 @@ test_that("ci.drtmle works as expected",{
 	expect_true(sum(is.na(unlist(tmp)))==0)
 })
 
-##### ADD TEST FOR ci.islptw
+##### ADD TEST FOR ci.adaptive_iptw
 
-context("Testing ci.islptw method ")
-test_that("ci.islptw works as expected",{
+context("Testing ci.adaptive_iptw method ")
+test_that("ci.adaptive_iptw works as expected",{
 	# simulate data
 	set.seed(123456)
 	n <- 200
@@ -74,35 +74,35 @@ test_that("ci.islptw works as expected",{
 	A <- rbinom(n,1,plogis(W$W1 - W$W2))
 	Y <- rbinom(n, 1, plogis(W$W1*W$W2*A))
 	# fit a drtmle
-	fit1 <- islptw(W = W, A = A, Y = Y, a_0 = c(1,0),
+	fit1 <- adaptive_iptw(W = W, A = A, Y = Y, a_0 = c(1,0),
 	               SL_g=c("SL.glm","SL.mean","SL.step"),
 	               SL_Qr="SL.glm")
 
 	# get confidence intervals for each 
 	tmp <- ci(fit1)
 	# correct class
-	expect_true(class(tmp)=="ci.islptw")
+	expect_true(class(tmp)=="ci.adaptive_iptw")
 	# no NAs
 	expect_true(sum(is.na(unlist(tmp)))==0)
 
 	# get confidence intervals for each mean 
-	tmp <- ci(fit1, est = c("islptw_tmle","islptw_os"))
+	tmp <- ci(fit1, est = c("iptw_tmle","iptw_os"))
 
 	# correct class
-	expect_true(class(tmp)=="ci.islptw")
+	expect_true(class(tmp)=="ci.adaptive_iptw")
 	# no NAs
 	expect_true(sum(is.na(unlist(tmp)))==0)
 	expect_true(length(tmp) == 2)
-	expect_true(all(names(tmp) == c("islptw_tmle","islptw_os")))
-	expect_true(nrow(tmp$islptw_tmle) == 2)
-	expect_true(nrow(tmp$islptw_os) == 2)
+	expect_true(all(names(tmp) == c("iptw_tmle","iptw_os")))
+	expect_true(nrow(tmp$iptw_tmle) == 2)
+	expect_true(nrow(tmp$iptw_os) == 2)
 
 	# get confidence intervals for ATE
[…]

### 2017-08-15 tests/testthat/test-islptw.R

@@ -1,66 +1,0 @@
-library(drtmle)
-library(SuperLearner)
-library(np)
-
-# TO DO: Add tests for multiple treatment levels
-context("Testing islptw works")
-test_that("islptw works as expected with parallel = TRUE",{
-	set.seed(123456)
-	n <- 200
-	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
-	A <- rbinom(n,1,plogis(W$W1 - W$W2))
-	Y <- rnorm(n, W$W1*W$W2*A, 2)
-
-	fit1 <- islptw(W = W, A = A, Y = Y, 
-	               a_0 = c(0,1),
-                  glm_g="W1 + W2",
-                  glm_Qr="gn",parallel = TRUE)
-	expect_true(all(!is.na(fit1$islptw_tmle$est)))
-	expect_true(all(!is.na(fit1$islptw_tmle$cov)))
-	expect_true(all(!is.na(fit1$islptw_os$est)))
-	expect_true(all(!is.na(fit1$islptw_os$cov)))
-	expect_true(all(!is.na(fit1$iptw$est)))
-	expect_true(all(!is.na(fit1$iptw$cov)))
-	expect_true(class(fit1)=="islptw")
-})
-test_that("islptw works as expected",{
-	set.seed(123456)
-	n <- 200
-	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
-	A <- rbinom(n,1,plogis(W$W1 - W$W2))
-	Y <- rnorm(n, W$W1*W$W2*A, 2)
-
-	fit1 <- islptw(W = W, A = A, Y = Y, 
-	               a_0 = c(0,1),
-                  glm_g="W1 + W2",
-                  glm_Qr="gn")
-	expect_true(all(!is.na(fit1$islptw_tmle$est)))
-	expect_true(all(!is.na(fit1$islptw_tmle$cov)))
-	expect_true(all(!is.na(fit1$islptw_os$est)))
-	expect_true(all(!is.na(fit1$islptw_os$cov)))
-	expect_true(all(!is.na(fit1$iptw$est)))
-	expect_true(all(!is.na(fit1$iptw$cov)))
-	expect_true(class(fit1)=="islptw")
-})	
-
-test_that("islptw works as expected with multi-level treatment",{
-	set.seed(123456)
-	n <- 200
[…]

### 2017-08-15 tests/testthat/test-multileveltrt.R

@@ -79,7 +79,7 @@ test_that("drtmle executes as expected with multiple treatment levels", {
 })
 
 
-test_that("islptw executes as expected with multiple treatment levels", {
+test_that("adaptive_iptw executes as expected with multiple treatment levels", {
 	set.seed(123456)
 	n <- 200
 	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
@@ -88,47 +88,47 @@ test_that("islptw executes as expected with multiple treatment levels", {
 
 	# univariate reduction with
 	# all GLMs + stratify
-	fit1 <- islptw(W = W, A = A, Y = Y, a_0=c(0,1,2),
+	fit1 <- adaptive_iptw(W = W, A = A, Y = Y, a_0=c(0,1,2),
                       glm_g="W1 + W2",
                       returnModels = TRUE,
                       glm_Qr="gn")
 
 	expect_true(is.numeric(fit1$iptw$est))
-	expect_true(is.numeric(fit1$islptw_tmle$est))
-	expect_true(is.numeric(fit1$islptw_os$est))
-	expect_true(is.numeric(fit1$islptw_os$cov))
-	expect_true(is.numeric(fit1$islptw_tmle$cov))
+	expect_true(is.numeric(fit1$iptw_tmle$est))
+	expect_true(is.numeric(fit1$iptw_os$est))
+	expect_true(is.numeric(fit1$iptw_os$cov))
+	expect_true(is.numeric(fit1$iptw_tmle$cov))
 	# check ci works
 	ci <- ci(fit1)
-	expect_true(length(fit1$islptw_tmle$est)==3)
+	expect_true(length(fit1$iptw_tmle$est)==3)
 	# check contrasts work
 	ci2 <- ci(fit1, contrast = c(1,-1,0))
-	expect_true(row.names(ci2$islptw_tmle) == "E[Y(0)]-E[Y(1)]")
+	expect_true(row.names(ci2$iptw_tmle) == "E[Y(0)]-E[Y(1)]")
 	ci3 <- ci(fit1, contrast = c(1,0,-1))
-	expect_true(row.names(ci3$islptw_tmle) == "E[Y(0)]-E[Y(2)]")
+	expect_true(row.names(ci3$iptw_tmle) == "E[Y(0)]-E[Y(2)]")
 
 
 	# same thing but with super learner for g
[…]

### 2017-08-15 tests/testthat/test-tests.R

@@ -63,44 +63,44 @@ test_that("wald_test.drtmle works as expected",{
 	expect_true(sum(is.na(unlist(tmp)))==0)
 })
 
-context("Testing ci.islptw method ")
-test_that("wald_test.islptw works as expected",{
+context("Testing ci.adaptive_iptw method ")
+test_that("wald_test.adaptive_iptw works as expected",{
 	# simulate data
 	set.seed(123456)
 	n <- 200
 	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
 	A <- rbinom(n,1,plogis(W$W1 - W$W2))
 	Y <- rbinom(n, 1, plogis(W$W1*W$W2*A))
-	# fit a islptw
-	fit1 <- islptw(W = W, A = A, Y = Y, a_0 = c(1,0),
+	# fit a adaptive_iptw
+	fit1 <- adaptive_iptw(W = W, A = A, Y = Y, a_0 = c(1,0),
 	               SL_g=c("SL.glm","SL.mean","SL.step"),
 	               SL_Qr="SL.glm")
 
 	# get test for each mean for only drtmle
 	tmp <- wald_test(fit1)
 	# correct class
-	expect_true(class(tmp)=="wald_test.islptw")
+	expect_true(class(tmp)=="wald_test.adaptive_iptw")
 	# no NAs
 	expect_true(sum(is.na(unlist(tmp)))==0)
 
 	# get test for each mean for drtmle and tmle
-	tmp <- wald_test(fit1, est = c("islptw_tmle","islptw_os"))
+	tmp <- wald_test(fit1, est = c("iptw_tmle","iptw_os"))
 
 	# correct class
-	expect_true(class(tmp)=="wald_test.islptw")
+	expect_true(class(tmp)=="wald_test.adaptive_iptw")
 	# no NAs
 	expect_true(sum(is.na(unlist(tmp)))==0)
 	expect_true(length(tmp) == 2)
-	expect_true(all(names(tmp) == c("islptw_tmle","islptw_os")))
-	expect_true(nrow(tmp$islptw_tmle) == 2)
-	expect_true(nrow(tmp$islptw_os) == 2)
+	expect_true(all(names(tmp) == c("iptw_tmle","iptw_os")))
+	expect_true(nrow(tmp$iptw_tmle) == 2)
+	expect_true(nrow(tmp$iptw_os) == 2)
 
 	# get test for ATE
[…]

### 2017-08-15 tests/testthat/testthat-adaptive_iptw.R

@@ -3,24 +3,24 @@ library(SuperLearner)
 library(np)
 library(testthat)
 # TO DO: Add tests for multiple treatment levels
-context("Testing islptw works with cv")
+context("Testing adaptive_iptw works with cv")
 
-test_that("islptw works as expected with cv",{
+test_that("adaptive_iptw works as expected with cv",{
 	set.seed(123456)
 	n <- 200
 	W <- data.frame(W1 = runif(n), W2 = rnorm(n))
 	A <- rbinom(n,1,plogis(W$W1 - W$W2))
 	Y <- rnorm(n, W$W1*W$W2*A, 2)
 
-	fit1 <- islptw(W = W, A = A, Y = Y,
+	fit1 <- adaptive_iptw(W = W, A = A, Y = Y,
 					cvFolds = 2,  
 	               a_0 = c(0,1),
                   glm_g="W1 + W2",
                   glm_Qr="gn")
-	expect_true(all(!is.na(fit1$islptw_tmle$est)))
-	expect_true(all(!is.na(fit1$islptw_tmle$cov)))
-	expect_true(all(!is.na(fit1$islptw_os$est)))
-	expect_true(all(!is.na(fit1$islptw_os$cov)))
+	expect_true(all(!is.na(fit1$iptw_tmle$est)))
+	expect_true(all(!is.na(fit1$iptw_tmle$cov)))
+	expect_true(all(!is.na(fit1$iptw_os$est)))
+	expect_true(all(!is.na(fit1$iptw_os$cov)))
 	expect_true(all(!is.na(fit1$iptw$est)))
-	expect_true(class(fit1)=="islptw")
+	expect_true(class(fit1)=="adaptive_iptw")
 })	

### 2017-08-15 R/adaptive_iptw.R
cleaning up for CRAN submission

@@ -16,10 +16,10 @@ globalVariables(c("v", "%dopar%"))
 #' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
 #' for the propensity score. See \code{link{SuperLearner::SuperLearner}} for details.
 #' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
-#' for the first reduced-dimension regression. 
+#' for the reduced-dimension outcome regression. 
 #' @param glm_g A character describing a formula to be used in the call to \code{glm} for the propensity score. Ignored
 #' if \code{SL_g!=NULL}.
-#' @param glm_Qr A character describing a formula to be used in the call to \code{glm} for the first reduced-dimension regression. Ignored
+#' @param glm_Qr A character describing a formula to be used in the call to \code{glm} for reduced-dimension outcome regression. Ignored
 #' if \code{SL_Qr!=NULL}. The formula should use the variable name \code{'gn'}.
 #' @param maxIter A numeric that sets the maximum number of iterations the TMLE can perform in its fluctuation step.
 #' @param tolIC A numeric that defines the stopping criteria based on the empirical mean
@@ -30,10 +30,10 @@ globalVariables(c("v", "%dopar%"))
 #' and reduced-dimension regressions.
 #' @param cvFolds A numeric equal to the number of folds to be used in cross-validated fitting of 
 #' nuisance parameters. If \code{cvFolds = 1}, no cross-validation is used.
-#' @param ... Other options (not currently used).
 #' @param parallel A boolean indicating whether to use \code{foreach}
 #' to estimate nuisance parameters in parallel. Only useful if there is a registered parallel
-#' backend (see examples) and \code{cvFolds > 1}.
+#' backend and \code{cvFolds > 1}.
+#' @param ... Other options (not currently used).
 #' @return An object of class \code{"adaptive_iptw"}.
 #' \describe{
 #'  \item{\code{iptw_tmle}}{A \code{list} of point estimates and 

### 2017-08-15 R/confint.R

@@ -15,20 +15,21 @@ ci <- function(...){
 #' (\code{"tmle"}, recommended only for comparison to "drtmle"), the standard 
 #' AIPTW (\code{"aiptw"}, recommended only for comparison to "drtmle"), and 
 #' G-computation (\code{"gcomp"}, not recommended).  
-#' @param level The level of the confidence interval
-#' @param contrast This option specifies what parameter to return confidence intervals for.
-#' If \code{contrast=NULL}, then compute confidence intervals for the covariate-adjusted
-#' marginal means. \code{contrast} can also be input as a numeric vector of ones, negative ones,
+#' @param level The nominal coverage probability of the desired confidence interval (should be
+#' between 0 and 1). Default computes 95% confidence intervals. 
+#' @param contrast Specifies the parameter for which to return confidence intervals.
+#' If \code{contrast=NULL}, then confidence intervals for the
+#' marginal means are computed. If instead, \code{contrast} is a numeric vector of ones, negative ones,
 #' and zeros to define linear combinations of the various means (e.g., to estimate an average
-#' treatment effect, see examples). \code{contrast} can also be a list with named functions
+#' treatment effect, see example). Finally, \code{contrast} can be a list with named functions
 #' \code{f}, \code{f_inv}, \code{h}, and \code{fh_grad}. The first two functions should take
 #' as input argument \code{eff}. Respectively, these specify which transformation 
 #' of the effect measure to compute the confidence interval for and the inverse
 #' transformation to put the confidence interval back on the original scale. The function \code{h}
 #' defines the contrast to be estimated and should take as input \code{est}, a vector
 #' of the same length as \code{object$a_0}, and output the desired contrast. The function
-#' \code{fh_grad} is the gradient of the function \code{h}. See examples. 
-#' @param ... Other options (not currently used)
+#' \code{fh_grad} is the gradient of the function \code{h}. See examples and vignette for more information. 
+#' @param ... Other options (not currently used).
 #' @importFrom stats qnorm
 #' @export
 #' @method ci drtmle
@@ -157,19 +158,20 @@ ci.drtmle <- function(object, est = c("drtmle"), level = 0.95,
 #' (\code{"iptw_tmle"}, recommended), the one-step IPTW 
 #' (\code{"iptw_os"}, not recommended), the standard IPTW 
 #' (\code{"iptw"}, recommended only for comparison to the other two estimators).
-#' @param level The level of the confidence interval
-#' @param contrast This option specifies what parameter to return confidence intervals for.
[…]

### 2017-08-15 R/drtmle.R

@@ -2,56 +2,56 @@ globalVariables(c("v", "%dopar%"))
 
 #' TMLE estimate of the average treatment effect with doubly-robust inference
 #' 
-#' @param W A \code{data.frame} of named covariates
-#' @param A A vector of binary treatment assignment (assumed to be equal to 0 or 1)
-#' @param Y A numeric of continuous or binary outcomes. 
-#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if missing 1 if observed)
-#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if missing 1 if observed)
-#' @param a_0 A vector of treatment levels at which to compute the adjusted mean outcome. 
+#' @param W A \code{data.frame} of named covariates.
+#' @param A A vector of discrete-valued treatment assignment (assumed to be equal to 0 or 1).
+#' @param Y numeric of continuous or binary outcomes.
+#' @param DeltaY A vector of missing outcome indicator (assumed to be equal to 0 if missing 1 if observed).
+#' @param DeltaA A vector of missing treatment indicator (assumed to be equal to 0 if missing 1 if observed).
+#' @param a_0 A vector of treatment levels at which to compute the adjusted mean outcome (should correspond to values of \code{A}).
 #' @param family A \code{family} object equal to either \code{binomial()} or \code{gaussian()}, 
 #' to be passed to the \code{SuperLearner} or \code{glm} function.
 #' @param stratify A \code{boolean} indicating whether to estimate the outcome regression separately
-#' for observations with \code{A} equal to 0/1 (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
+#' for different values of \code{A} (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
 #' @param SL_Q A vector of characters or a list describing the Super Learner library to be used 
 #' for the outcome regression. See \code{link{SuperLearner::SuperLearner}} for details.
 #' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
 #' for the propensity score. See \code{link{SuperLearner::SuperLearner}} for details.
 #' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
-#' for the first reduced-dimension regression. 
+#' for the reduced-dimension outcome regression. 
 #' @param SL_gr A vector of characters or a list describing the Super Learner library to be used 
-#' for the second reduced-dimension regression.
+#' for the second reduced-dimension propensity score. 
[…]

### 2017-08-15 R/sl_npreg.R

@@ -4,14 +4,14 @@
 #' package. Uses leave-one-out cross-validation to fit a kernel regression. 
 #' See \code{?npreg} for more details.
 #' 
-#' @param Y Training outcomes
-#' @param X Training predictors 
-#' @param family Not used by the function, but needed for compatibility with \code{SuperLearner}
-#' @param newX Test set predictors 
-#' @param obsWeights Weights for the observations
-#' @param rangeThresh For stability: if the the range of the outcomes is smaller than this number, the method
-#' returns the empirical average of the outcomes. 
-#' @param ... Other arguments (not currently used)
+#' @param Y A vector of outcomes.
+#' @param X A matrix or data.frame of training data predictors. 
+#' @param family Not used by the function directly, but ensures compatibility with \code{SuperLearner}.
+#' @param newX A test set of predictors.
+#' @param obsWeights Not used by the function directly, but ensures compatibility with \code{SuperLearner}.
+#' @param rangeThresh If the the range of the outcomes is smaller than this number, the method
+#' returns the empirical average of the outcomes. Used for computational expediency and stability.  
+#' @param ... Other arguments (not currently used).
 #' 
 #' @importFrom np npregbw npreg
 #' @importFrom stats predict as.formula
@@ -47,9 +47,9 @@ SL.npreg <- function (Y, X, newX, family, obsWeights,
 #' 
 #' Method for predicting SL.npreg objects. 
 #' 
-#' @param object An object of class \code{"SL.npreg"}
-#' @param newdata The new data used to obtain predictions 
-#' @param ... Other arguments passed to predict
+#' @param object An object of class \code{"SL.npreg"}.
+#' @param newdata The new data used to obtain predictions.
+#' @param ... Other arguments passed to predict.
 #' 
 #' @importFrom stats predict
 #' @method predict SL.npreg

### 2017-08-15 R/test.R

@@ -32,7 +32,7 @@ wald_test <- function(...){
 #' \code{fh_grad} is the gradient of the function \code{h(f())}. The function
 #' computes a test of the null hypothesis that \code{h(f(object$est)) = null}. 
 #' See examples. 
-#' @param ... Other options (not currently used)
+#' @param ... Other options (not currently used).
 #' 
 #' @importFrom stats pnorm
 #' @export
@@ -191,7 +191,7 @@ wald_test.drtmle <- function(object, est = c("drtmle"), null = 0,
 #' \code{fh_grad} is the gradient of the function \code{h(f())}. The function
 #' computes a test of the null hypothesis that \code{h(f(object$est)) = null}. 
 #' See examples. 
-#' @param ... Other options (not currently used)
+#' @param ... Other options (not currently used).
 #' 
 #' @importFrom stats pnorm
 #' @export

### 2017-08-15 R/confint.R
escape % causing build error

@@ -16,7 +16,7 @@ ci <- function(...){
 #' AIPTW (\code{"aiptw"}, recommended only for comparison to "drtmle"), and 
 #' G-computation (\code{"gcomp"}, not recommended).  
 #' @param level The nominal coverage probability of the desired confidence interval (should be
-#' between 0 and 1). Default computes 95% confidence intervals. 
+#' between 0 and 1). Default computes 95\% confidence intervals. 
 #' @param contrast Specifies the parameter for which to return confidence intervals.
 #' If \code{contrast=NULL}, then confidence intervals for the
 #' marginal means are computed. If instead, \code{contrast} is a numeric vector of ones, negative ones,
@@ -159,7 +159,7 @@ ci.drtmle <- function(object, est = c("drtmle"), level = 0.95,
 #' (\code{"iptw_os"}, not recommended), the standard IPTW 
 #' (\code{"iptw"}, recommended only for comparison to the other two estimators).
 #' @param level The nominal coverage probability of the desired confidence interval (should be
-#' between 0 and 1). Default computes 95% confidence intervals. 
+#' between 0 and 1). Default computes 95\% confidence intervals. 
 #' @param contrast Specifies the parameter for which to return confidence intervals.
 #' If \code{contrast=NULL}, then confidence intervals for the
 #' marginal means are computed. If instead, \code{contrast} is a numeric vector of ones, negative ones,
@@ -172,7 +172,7 @@ ci.drtmle <- function(object, est = c("drtmle"), level = 0.95,
 #' defines the contrast to be estimated and should take as input \code{est}, a vector
 #' of the same length as \code{object$a_0}, and output the desired contrast. The function
 #' \code{fh_grad} is the gradient of the function \code{h}. See examples and vignette for more information. 
-#' @param ... Other options (not currently used)
+#' @param ... Other options (not currently used).
 #' @export
 #' @method ci adaptive_iptw
 #' @return An object of class \code{"ci.adaptive_iptw"} with point estimates and

### 2017-08-15 R/sl_npreg.R
small changes to pass R CMD BUILD

@@ -1,6 +1,6 @@
 #' Super learner wrapper for kernel regression
 #' 
-#' Kernel regression based on the \href{https://cran.r-project.org/web/packages/np/}{np}
+#' Kernel regression based on the \href{https://CRAN.R-project.org/package=np}{np}
 #' package. Uses leave-one-out cross-validation to fit a kernel regression. 
 #' See \code{?npreg} for more details.
 #' 

### 2017-08-16 R/adaptive_iptw.R
cran documentation clean up, remove some TO DOs, add stopCluster to vignette

@@ -13,12 +13,17 @@ globalVariables(c("v", "%dopar%"))
 #' for observations with different levels of \code{A} (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
 #' @param family A \code{family} object equal to either \code{binomial()} or \code{gaussian()}, 
 #' to be passed to the \code{SuperLearner} or \code{glm} function.
-#' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
-#' for the propensity score. See \code{link{SuperLearner::SuperLearner}} for details.
+#' @param SL_g A vector of characters describing the super learner library to be used
+#' for each of the propensity score regressions (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
+#' library for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
+#' a single library may be input. See \code{link{SuperLearner::SuperLearner}} for details on how super 
+#' learner libraries can be specified.
 #' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
 #' for the reduced-dimension outcome regression. 
-#' @param glm_g A character describing a formula to be used in the call to \code{glm} for the propensity score. Ignored
-#' if \code{SL_g!=NULL}.
+#' @param glm_g A list of characters describing the formulas to be used
+#' for each of the propensity score regressions (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
+#' formula for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
+#' a single character formula may be input.
 #' @param glm_Qr A character describing a formula to be used in the call to \code{glm} for reduced-dimension outcome regression. Ignored
 #' if \code{SL_Qr!=NULL}. The formula should use the variable name \code{'gn'}.
 #' @param maxIter A numeric that sets the maximum number of iterations the TMLE can perform in its fluctuation step.
@@ -139,7 +144,6 @@ adaptive_iptw <- function(W, A, Y,
   }
   # obtain list of propensity score fits
   gnMod <- gnValid[seq(2, length(gnValid), 2)]
-  # TO DO: Add reasonable names to gnMod?
 
  	# compute iptw estimator
 	psi_n <- mapply(a = split(a_0, 1:length(a_0)),g=gn, function(a,g){
[…]

### 2017-08-16 R/drtmle.R

@@ -14,16 +14,21 @@ globalVariables(c("v", "%dopar%"))
 #' for different values of \code{A} (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
 #' @param SL_Q A vector of characters or a list describing the Super Learner library to be used 
 #' for the outcome regression. See \code{link{SuperLearner::SuperLearner}} for details.
-#' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
-#' for the propensity score. See \code{link{SuperLearner::SuperLearner}} for details.
+#' @param SL_g A vector of characters describing the super learner library to be used
+#' for each of the propensity score regressions (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
+#' library for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
+#' a single library may be input. See \code{link{SuperLearner::SuperLearner}} for details on how super 
+#' learner libraries can be specified.
 #' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
 #' for the reduced-dimension outcome regression. 
 #' @param SL_gr A vector of characters or a list describing the Super Learner library to be used 
 #' for the second reduced-dimension propensity score. 
 #' @param glm_Q A character describing a formula to be used in the call to \code{glm} for the outcome regression. Ignored
 #' if \code{SL_Q!=NULL}.
-#' @param glm_g A character describing a formula to be used in the call to \code{glm} for the propensity score. Ignored
-#' if \code{SL_g!=NULL}.
+#' @param glm_g A list of characters describing the formulas to be used
+#' for each of the propensity score regressions (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
+#' formula for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
+#' a single character formula may be input.
 #' @param glm_Qr A character describing a formula to be used in the call to \code{glm} for reduced-dimension outcome regression. Ignored
 #' if \code{SL_Qr!=NULL}. The formula should use the variable name \code{'gn'}.
[…]

### 2017-08-16 R/estimate.R

@@ -9,8 +9,10 @@
 #' @param W A \code{data.frame} of named covariates
 #' @param stratify A \code{boolean} indicating whether to estimate the missing outcome regression separately
 #' for observations with \code{A} equal to 0/1 (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
-#' @param SL_g A vector of characters or a list describing the Super Learner library to be used 
-#' for the propensity score !!!!! TO DO: Add new description.
+#' @param SL_g A vector of characters describing the super learner library to be used
+#' for each of the regression (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
+#' regression for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
+#' a single library may be input. 
 #' @param tolg A numeric indicating the minimum value for estimates of the propensity score.
 #' @param verbose A boolean indicating whether to print status updates.
 #' @param returnModels A boolean indicating whether to return model fits for the outcome regression, propensity score,

### 2017-08-16 R/adaptive_iptw.R
update site, format code

@@ -4,71 +4,79 @@ globalVariables(c("v", "%dopar%"))
 #' for the propensity score
 #' 
 #' @param W A \code{data.frame} of named covariates
-#' @param A A vector of binary treatment assignment (assumed to be equal to 0 or 1)
+#' @param A A vector of binary treatment assignment (assumed to be equal to 0 or
+#'  1)
 #' @param Y A numeric of continuous or binary outcomes.
-#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if missing 1 if observed)
-#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if missing 1 if observed) 
-#' @param a_0 A vector of treatment levels at which to compute the adjusted mean outcome. 
-#' @param stratify A \code{boolean} indicating whether to estimate the missing outcome regression separately
-#' for observations with different levels of \code{A} (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
-#' @param family A \code{family} object equal to either \code{binomial()} or \code{gaussian()}, 
-#' to be passed to the \code{SuperLearner} or \code{glm} function.
-#' @param SL_g A vector of characters describing the super learner library to be used
-#' for each of the propensity score regressions (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
-#' library for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
-#' a single library may be input. See \code{link{SuperLearner::SuperLearner}} for details on how super 
-#' learner libraries can be specified.
-#' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
-#' for the reduced-dimension outcome regression. 
+#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if 
+#' missing 1 if observed)
+#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if 
+#' missing 1 if observed) 
+#' @param a_0 A vector of fixed treatment values at which to return marginal 
+#' mean estimates.
+#' @param stratify A \code{boolean} indicating whether to estimate the missing 
+#' outcome regression separately
+#' for observations with different levels of \code{A} (if \code{TRUE}) or to 
+#' pool across \code{A} (if \code{FALSE}).
[…]

### 2017-08-16 R/drtmle.R

@@ -3,99 +3,121 @@ globalVariables(c("v", "%dopar%"))
 #' TMLE estimate of the average treatment effect with doubly-robust inference
 #' 
 #' @param W A \code{data.frame} of named covariates.
-#' @param A A vector of discrete-valued treatment assignment (assumed to be equal to 0 or 1).
+#' @param A A vector of discrete-valued treatment assignment (assumed to be 
+#' equal to 0 or 1).
 #' @param Y numeric of continuous or binary outcomes.
-#' @param DeltaY A vector of missing outcome indicator (assumed to be equal to 0 if missing 1 if observed).
-#' @param DeltaA A vector of missing treatment indicator (assumed to be equal to 0 if missing 1 if observed).
-#' @param a_0 A vector of treatment levels at which to compute the adjusted mean outcome (should correspond to values of \code{A}).
-#' @param family A \code{family} object equal to either \code{binomial()} or \code{gaussian()}, 
-#' to be passed to the \code{SuperLearner} or \code{glm} function.
-#' @param stratify A \code{boolean} indicating whether to estimate the outcome regression separately
-#' for different values of \code{A} (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
-#' @param SL_Q A vector of characters or a list describing the Super Learner library to be used 
-#' for the outcome regression. See \code{link{SuperLearner::SuperLearner}} for details.
-#' @param SL_g A vector of characters describing the super learner library to be used
-#' for each of the propensity score regressions (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
-#' library for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
-#' a single library may be input. See \code{link{SuperLearner::SuperLearner}} for details on how super 
-#' learner libraries can be specified.
-#' @param SL_Qr A vector of characters or a list describing the Super Learner library to be used 
-#' for the reduced-dimension outcome regression. 
-#' @param SL_gr A vector of characters or a list describing the Super Learner library to be used 
-#' for the second reduced-dimension propensity score. 
-#' @param glm_Q A character describing a formula to be used in the call to \code{glm} for the outcome regression. Ignored
-#' if \code{SL_Q!=NULL}.
+#' @param DeltaY A vector of missing outcome indicator (assumed to be equal to 0 
+#' if missing 1 if observed).
+#' @param DeltaA A vector of missing treatment indicator (assumed to be equal to 
[…]

### 2017-08-16 R/estimate.R

@@ -3,461 +3,458 @@
 #' Function to estimate propensity score
 #' 
 #' @importFrom plyr alply
-#' @param A A vector of binary treatment assignment (assumed to be equal to 0 or 1)
-#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if missing 1 if observed)
-#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if missing 1 if observed)
+#' @param A A vector of binary treatment assignment (assumed to be equal to 0 or
+#' 1)
+#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if 
+#' missing 1 if observed)
+#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if 
+#' missing 1 if observed)
 #' @param W A \code{data.frame} of named covariates
-#' @param stratify A \code{boolean} indicating whether to estimate the missing outcome regression separately
-#' for observations with \code{A} equal to 0/1 (if \code{TRUE}) or to pool across \code{A} (if \code{FALSE}).
-#' @param SL_g A vector of characters describing the super learner library to be used
-#' for each of the regression (\code{DeltaA}, \code{A}, and \code{DeltaY}). To use the same
-#' regression for each of the regressions (or if there is no missing data in \code{A} nor \code{Y}), 
-#' a single library may be input. 
-#' @param tolg A numeric indicating the minimum value for estimates of the propensity score.
+#' @param stratify A \code{boolean} indicating whether to estimate the missing 
+#' outcome regression separately
+#' for observations with \code{A} equal to 0/1 (if \code{TRUE}) or to pool 
+#' across \code{A} (if \code{FALSE}).
+#' @param SL_g A vector of characters describing the super learner library to be 
+#' used for each of the regression (\code{DeltaA}, \code{A}, and \code{DeltaY}). 
+#' To use the same regression for each of the regressions (or if there is no 
+#' missing data in \code{A} nor \code{Y}), a single library may be input. 
+#' @param tolg A numeric indicating the minimum value for estimates of the 
+#' propensity score.
 #' @param verbose A boolean indicating whether to print status updates.
-#' @param returnModels A boolean indicating whether to return model fits for the outcome regression, propensity score,
-#' and reduced-dimension regressions.
[…]

### 2017-08-17 README.Rmd
update README

@@
 
-- Install the most recent _stable release_:
-  `devtools::install_github("benkeser/drtmle")`
+Install the current stable release from
+[CRAN](https://cran.r-project.org/) via
 
-- To contribute, install the _development version_:
-  `devtools::install_github("benkeser/drtmle", ref = "develop")`
+```{r cran-installation, eval = FALSE}
+install.packages("drtmle")
+```
+
+A developmental release may be installed from GitHub via
+[`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with:
+
+```{r gh-installation, eval = FALSE}
+devtools::install_github("benkeser/drtmle")
+```
 

### 2017-08-17 README.Rmd
readme badges

@@
 [![CRAN](http://www.r-pkg.org/badges/version/drtmle)](http://www.r-pkg.org/pkg/drtmle)
+[![CRAN downloads](https://cranlogs.r-pkg.org/badges/drtmle)](https://CRAN.R-project.org/package=drtmle)
 [![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
 [![MIT license](http://img.shields.io/badge/license-MIT-brightgreen.svg)](http://opensource.org/licenses/MIT)
+[![DOI](https://zenodo.org/badge/75324341.svg)](https://zenodo.org/badge/latestdoi/75324341)
 

### 2017-09-27 R/adaptive_iptw.R
initial add of future functionality

@@ -1,91 +1,96 @@
-globalVariables(c("v", "%dopar%"))
+globalVariables(c("v"))
 
-#' Compute asymptotically linear IPTW estimators with super learning 
+#' Compute asymptotically linear IPTW estimators with super learning
 #' for the propensity score
-#' 
+#'
 #' @param W A \code{data.frame} of named covariates
 #' @param A A vector of binary treatment assignment (assumed to be equal to 0 or
 #'  1)
 #' @param Y A numeric of continuous or binary outcomes.
-#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if 
+#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if
 #' missing 1 if observed)
-#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if 
-#' missing 1 if observed) 
-#' @param a_0 A vector of fixed treatment values at which to return marginal 
+#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if
+#' missing 1 if observed)
+#' @param a_0 A vector of fixed treatment values at which to return marginal
 #' mean estimates.
-#' @param stratify A \code{boolean} indicating whether to estimate the missing 
+#' @param stratify A \code{boolean} indicating whether to estimate the missing
 #' outcome regression separately
-#' for observations with different levels of \code{A} (if \code{TRUE}) or to 
+#' for observations with different levels of \code{A} (if \code{TRUE}) or to
 #' pool across \code{A} (if \code{FALSE}).
-#' @param family A \code{family} object equal to either \code{binomial()} or 
-#' \code{gaussian()}, to be passed to the \code{SuperLearner} or \code{glm} 
+#' @param family A \code{family} object equal to either \code{binomial()} or
+#' \code{gaussian()}, to be passed to the \code{SuperLearner} or \code{glm}
 #' function.
-#' @param SL_g A vector of characters describing the super learner library to be 
-#' used for each of the propensity score regressions (\code{DeltaA}, \code{A}, 
-#' and \code{DeltaY}). To use the same library for each of the regressions (or 
-#' if there is no missing data in \code{A} nor \code{Y}), a single library may 
-#' be input. See \code{link{SuperLearner::SuperLearner}} for details on how 
+#' @param SL_g A vector of characters describing the super learner library to be
+#' used for each of the propensity score regressions (\code{DeltaA}, \code{A},
[…]

### 2017-09-27 R/drtmle.R

@@ -1,133 +1,135 @@
-globalVariables(c("v", "%dopar%"))
+globalVariables(c("v"))
 
 #' TMLE estimate of the average treatment effect with doubly-robust inference
-#' 
+#'
 #' @param W A \code{data.frame} of named covariates.
-#' @param A A vector of discrete-valued treatment assignment (assumed to be 
+#' @param A A vector of discrete-valued treatment assignment (assumed to be
 #' equal to 0 or 1).
 #' @param Y numeric of continuous or binary outcomes.
-#' @param DeltaY A vector of missing outcome indicator (assumed to be equal to 0 
+#' @param DeltaY A vector of missing outcome indicator (assumed to be equal to 0
 #' if missing 1 if observed).
-#' @param DeltaA A vector of missing treatment indicator (assumed to be equal to 
+#' @param DeltaA A vector of missing treatment indicator (assumed to be equal to
 #' 0 if missing 1 if observed).
-#' @param a_0 A vector of fixed treatment values at which to return marginal 
+#' @param a_0 A vector of fixed treatment values at which to return marginal
 #' mean estimates.
-#' @param family A \code{family} object equal to either \code{binomial()} or 
-#' \code{gaussian()}, to be passed to the \code{SuperLearner} or \code{glm} 
+#' @param family A \code{family} object equal to either \code{binomial()} or
+#' \code{gaussian()}, to be passed to the \code{SuperLearner} or \code{glm}
 #' function.
-#' @param stratify A \code{boolean} indicating whether to estimate the outcome 
-#' regression separately for different values of \code{A} (if \code{TRUE}) or to 
+#' @param stratify A \code{boolean} indicating whether to estimate the outcome
+#' regression separately for different values of \code{A} (if \code{TRUE}) or to
 #' pool across \code{A} (if \code{FALSE}).
-#' @param SL_Q A vector of characters or a list describing the Super Learner 
-#' library to be used for the outcome regression. See 
+#' @param SL_Q A vector of characters or a list describing the Super Learner
+#' library to be used for the outcome regression. See
 #' \code{link{SuperLearner::SuperLearner}} for details.
-#' @param SL_g A vector of characters describing the super learner library to be 
-#' used for each of the propensity score regressions (\code{DeltaA}, \code{A}, 
-#' and \code{DeltaY}). To use the same library for each of the regressions (or 
-#' if there is no missing data in \code{A} nor \code{Y}), a single library may 
[…]

### 2017-09-27 R/adaptive_iptw.R
update vignette for futures

@@ -131,8 +131,8 @@ adaptive_iptw <- function(W, A, Y,
   if (!parallel) {
     future::plan(future::sequential)
   } else {
+    doFuture::registerDoFuture()
     if (all(c("sequential", "uniprocess") %in% class(future::plan()))) {
-      doFuture::registerDoFuture()
       future::plan(future::multiprocess)
     }
   }

### 2017-09-27 R/drtmle.R

@@ -183,8 +183,8 @@ drtmle <- function(Y, A, W,
   if (!parallel) {
     future::plan(future::sequential)
   } else {
+    doFuture::registerDoFuture()
     if (all(c("sequential", "uniprocess") %in% class(future::plan()))) {
-      doFuture::registerDoFuture()
       future::plan(future::multiprocess)
     }
   }

### 2017-09-27 R/drtmle.R
add option for user input Qn and gn

@@ -77,6 +77,20 @@ globalVariables(c("v", "%dopar%"))
 #' @param parallel A boolean indicating whether to use \code{foreach}
 #' to estimate nuisance parameters in parallel. Only useful if there is a 
 #' registered parallel backend and \code{cvFolds > 1}.
+#' @param Qn An optional list of outcome regression estimates. If specified, 
+#' the function will ignore the nuisance parameter estimation specified by 
+#' \code{SL_Q} and \code{glm_Q}. The entries in the list should correspond to 
+#' the outcome regression evaluated at \code{A} and the observed values of \code{W},
+#' with order determined by the input to \code{a_0} (e.g., if \code{a_0 = c(0,1)} then
+#' \code{Qn[[1]]} should be outcome regression at \code{A} = 0 and \code{Qn[[2]]} should
+#' be outcome regression at \code{A} = 1). 
+#' @param gn An optional list of propensity score estimates. If specified, 
+#' the function will ignore the nuisance parameter estimation specified by 
+#' \code{SL_g} and \code{glm_g}. The entries in the list should correspond to 
+#' the propensity for the observed values of \code{W},
+#' with order determined by the input to \code{a_0} (e.g., if \code{a_0 = c(0,1)} then
+#' \code{gn[[1]]} should be propensity of \code{A} = 0 and \code{Qn[[2]]} should
+#' be propensity of \code{A} = 1). 
 #' @param ... Other options (not currently used).
 #' 
 #' @return An object of class \code{"drtmle"}.
@@ -168,6 +182,8 @@ drtmle <- function(Y, A, W,
                    verbose = FALSE,
                    Qsteps = 2,
                    parallel = FALSE,
+                   Qn = NULL,
+                   gn = NULL,
                    ...){
   call <- match.call()
   # if cvFolds non-null split data into cvFolds pieces
@@ -180,62 +196,64 @@ drtmle <- function(Y, A, W,
   #-------------------------------
   # estimate propensity score
   #-------------------------------
-  if(!parallel){
-    gnOut <- lapply(X = validRows, FUN = estimateG,
-                    A = A, W = W, DeltaA = DeltaA, DeltaY = DeltaY,
[…]

### 2017-09-27 R/drtmle.R
allow cvFolds to enter as vector

@@ -73,7 +73,9 @@ globalVariables(c("v", "%dopar%"))
 #' more stable in simulations and is the default. 
 #' @param cvFolds A numeric equal to the number of folds to be used in 
 #' cross-validated fitting of nuisance parameters. If \code{cvFolds = 1}, no 
-#' cross-validation is used.
+#' cross-validation is used. Alternatively, \code{cvFolds} may be entered as a 
+#' vector of fold assignments for observations, in which case its length should be
+#' the same length as \code{Y}. 
 #' @param parallel A boolean indicating whether to use \code{foreach}
 #' to estimate nuisance parameters in parallel. Only useful if there is a 
 #' registered parallel backend and \code{cvFolds > 1}.
@@ -188,9 +190,18 @@ drtmle <- function(Y, A, W,
   call <- match.call()
   # if cvFolds non-null split data into cvFolds pieces
   n <- length(Y)
-  if(cvFolds!=1){
+  if(length(cvFolds) > 1){
+    stopifnot(length(cvFolds) == length(Y))
+    # comes in as vector of fold assignments
+    # split up into a list of id's
+    validRows <- sapply(sort(unique(validRows)), function(f){ 
+      which(validRows == f) 
+    })
+  }else if(cvFolds!=1){
+    # split data up
     validRows <- split(sample(1:n), rep(1:cvFolds, length = n))       
   }else{
+    # no cross-validation
     validRows <- list(1:n)
   }
   #-------------------------------

### 2017-09-29 README.Rmd

@@
 
-`drtmle` is an R package that computes marginal means of an outcome under fixed levels of a treatment.
-The package computes targeted minimum loss-based (TMLE) estimators that are doubly robust, not only with respect to consistency, but also with respect to asymptotic normality, as discussed in Benkeser, Carone, van der Laan \& Gilbert, 2017 (_accepted Biometrika_; [working paper](http://biostats.bepress.com/ucbbiostat/paper356/)).
-This property facilitates construction of doubly-robust confidence intervals and hypothesis tests.
+`drtmle` is an R package that computes marginal means of an outcome under fixed
+levels of a treatment. The package computes targeted minimum loss-based (TMLE)
+estimators that are doubly robust, not only with respect to consistency, but
+also with respect to asymptotic normality, as discussed in Benkeser, Carone, van
+der Laan \& Gilbert, 2017 (_accepted Biometrika_; [working
+paper](http://biostats.bepress.com/ucbbiostat/paper356/)). This property
+facilitates construction of doubly-robust confidence intervals and hypothesis
+tests.
 
-The package additionally includes methods for computing valid confidence intervals for an inverse probability of treatment weighted (IPTW) estimator of the average treatment effect when the propensity score is estimated via super learning, as discussed in [van der Laan, 2014](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
+The package additionally includes methods for computing valid confidence
+intervals for an inverse probability of treatment weighted (IPTW) estimator of
+the average treatment effect when the propensity score is estimated via super
+learning, as discussed in
+[van der Laan, 2014](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
 
@@
 
-## Use
+## Usage
 
-### Doubly-robust inference for average treatment effect
+### Doubly-robust inference for the average treatment effect
 
-Suppose the data consist of a vector of baseline covariates (`W`), a multi-level treatment assignment (`A`), and a continuous or binary-valued outcome (`Y`).
-The function `drtmle` may be used to estimate $E[E(Y | A=a_0, W)]$ for user-selected values of $a_0$ (via option `a_0`).
-The resulting targeted minimum loss-based estimates are doubly robust with respect to both consistency and asymptotic normality.
-The function computes doubly robust covariance estimates that can be used to construct doubly robust confidence intervals for marginal means and contrasts between means.
-A simple example on simulated data is shown below.
-We refer users to the vignette for more information and further examples.
+Suppose the data consist of a vector of baseline covariates (`W`), a multi-level
+treatment assignment (`A`), and a continuous or binary-valued outcome (`Y`). The
+function `drtmle` may be used to estimate $E[E(Y \mid A = a_0, W)]$ for
+user-selected values of $a_0$ (via option `a_0`). The resulting targeted minimum
+loss-based estimates are doubly robust with respect to both consistency and
+asymptotic normality. The function computes doubly robust covariance estimates
+that can be used to construct doubly robust confidence intervals for marginal
+means and contrasts between means. A simple example on simulated data is shown
+below. We refer users to the vignette for more information and further examples.
 
@@
 fit1 <- drtmle(W = W, A = A, Y = Y, # input data
-              a_0 = c(0,1), # return estimates for A = 0 and A = 1
-              SL_Q = "SL.npreg", # use kernel regression for E(Y | A=a, W)
+               a_0 = c(0, 1), # return estimates for A = 0 and A = 1
+               SL_Q = "SL.npreg", # use kernel regression for E(Y | A = a, W)
                glm_g = "W1 + W2", # use misspecified main terms glm for E(A | W)
@@
 # get confidence intervals for ate
-ci_ate1 <- ci(fit1,contrast = c(-1,1))
+ci_ate1 <- ci(fit1,contrast = c(-1, 1))
 # print the output
@@
 
-The package additionally includes a function for computing valid confidence intervals about an inverse probability of treatment weight (IPTW) estimator when super learning is used to estimate the propensity score.
+The package additionally includes a function for computing valid confidence
+intervals about an inverse probability of treatment weight (IPTW) estimator when
+super learning is used to estimate the propensity score.
 
@@
 # fit iptw
-fit2 <- adaptive_iptw(Y = Y, A = A, W = W, a_0 = c(0,1),
-               SL_g = c("SL.glm","SL.mean","SL.step.interaction"),
+fit2 <- adaptive_iptw(Y = Y, A = A, W = W, a_0 = c(0, 1),
+                      SL_g = c("SL.glm", "SL.mean", "SL.step.interaction"),
                       SL_Qr = "SL.npreg")
@@
 # compute a confidence interval for the ate
-ci_ate2 <- ci(fit2, contrast = c(-1,1))
+ci_ate2 <- ci(fit2, contrast = c(-1, 1))
 # print the output
@@
           year  = {2017},
-          author = {Benkeser, David C and Carone, Marco and van der Laan, Mark J and Gilbert, Peter B},
-          title = {Doubly-robust nonparametric inference on the average treatment effect},
+          author = {Benkeser, David C and Carone, Marco and van der Laan, Mark J
+            and Gilbert, Peter B},
+          title = {Doubly-robust nonparametric inference on the average
+            treatment effect},
           journal = {Biometrika}

### 2017-12-12 README.Rmd
update readme

@@
 estimators that are doubly robust, not only with respect to consistency, but
-also with respect to asymptotic normality, as discussed in Benkeser, Carone, van
-der Laan \& Gilbert, 2017 (_accepted Biometrika_; [working
-paper](http://biostats.bepress.com/ucbbiostat/paper356/)). This property
+also with respect to asymptotic normality, as discussed in [Benkeser, et al (2017)](https://doi.org/10.1093/biomet/asx053). This property
 facilitates construction of doubly-robust confidence intervals and hypothesis
@@
 means and contrasts between means. A simple example on simulated data is shown
-below. We refer users to the vignette for more information and further examples.
+below. We refer users to [the vignette](https://benkeser.github.io/drtmle/articles/using_drtmle.html) for more information and further examples.
 
@@
 # load packages
-library(drtmle, quietly = TRUE)
-library(SuperLearner, quietly = TRUE)
+library(drtmle)
+library(SuperLearner)
 
@@
 After using the `drtmle` R package, please cite the following:
-
+  @Manual{drtmlepackage,
+    title = {drtmle: Doubly-Robust Nonparametric Estimation and Inference},
+    author = {David Benkeser},
+    note = {R package version 1.0.0},
+    doi = {10.5281/zenodo.844836},
+  }
   @article{benkeser2017improved,
@@
       treatment effect},
-          journal = {Biometrika}
+    journal = {Biometrika},
+    volume = {104}, number = {4},
+    pages = {863–880},
+    doi = {10.1093/biomet/asx053}
   }

### 2017-12-12 README.Rmd
update readme

@@
 learning, as discussed in
-[van der Laan, 2014](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
+[van der Laan (2014)](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
 
@@
 After using the `drtmle` R package, please cite the following:
+
     @Manual{drtmlepackage,
@@
     }
+
     @article{benkeser2017improved,

### 2017-12-19 R/adaptive_iptw.R
remove lingering foreach dependencies. bump version. update news. update docs. fix vignette.

@@ -1,17 +1,15 @@
-globalVariables(c("v"))
-
 #' Compute asymptotically linear IPTW estimators with super learning
 #' for the propensity score
 #'
 #' @param W A \code{data.frame} of named covariates
-#' @param A A vector of binary treatment assignment (assumed to be equal to 0 or
+#' @param A A \code{numeric} vector of binary treatment assignment (assumed to be equal to 0 or
 #'  1)
-#' @param Y A numeric of continuous or binary outcomes.
-#' @param DeltaY Indicator of missing outcome (assumed to be equal to 0 if
+#' @param Y A \code{numeric} numeric of continuous or binary outcomes.
+#' @param DeltaY A \code{numeric} indicator of missing outcome (assumed to be equal to 0 if
 #' missing 1 if observed)
-#' @param DeltaA Indicator of missing treatment (assumed to be equal to 0 if
+#' @param DeltaA A \code{numeric} indicator of missing treatment (assumed to be equal to 0 if
 #' missing 1 if observed)
-#' @param a_0 A vector of fixed treatment values at which to return marginal
+#' @param a_0 A vector of \code{numeric} treatment values at which to return marginal
 #' mean estimates.
 #' @param stratify A \code{boolean} indicating whether to estimate the missing
 #' outcome regression separately
@@ -49,7 +47,7 @@ globalVariables(c("v"))
 #' cross-validated fitting of nuisance parameters. If \code{cvFolds = 1}, no
 #' cross-validation is used.
 #' @param parallel A boolean indicating whether to use parallelization based on
-#' \code{foreach}, \code{future}, and \code{doFuture} to estimate nuisance
+#' \code{future} to estimate nuisance
 #' parameters in parallel. Only useful if \code{cvFolds > 1}. By default, a
 #' \code{multiprocess} evaluation scheme is invoked, using forked R processes
 #' (if supported on the OS) and background R sessions otherwise. Users may also
@@ -83,7 +81,6 @@ globalVariables(c("v"))
 #' }
 #'
 #' @importFrom plyr llply laply
-#' @importFrom foreach foreach "%dopar%"
 #' @importFrom future plan future_lapply
 #' @importFrom doFuture registerDoFuture
 #' @importFrom stats cov
@@ -142,24 +139,13 @@ adaptive_iptw <- function(W, A, Y,
   #-------------------------------
   # estimate propensity score
   #-------------------------------
-  if (!parallel) {
[…]

### 2017-12-19 R/drtmle.R

@@ -1,8 +1,7 @@
 #' TMLE estimate of the average treatment effect with doubly-robust inference
 #'
 #' @param W A \code{data.frame} of named covariates.
-#' @param A A \code{numeric} vector of discrete-valued treatment assignment (assumed to be
-#' equal to 0 or 1).
+#' @param A A \code{numeric} vector of discrete-valued treatment assignment.
 #' @param Y A \code{numeric} continuous or binary outcomes.
 #' @param DeltaY A \code{numeric} vector of missing outcome indicator (assumed to be equal to 0
 #' if missing 1 if observed).

### 2017-12-19 R/method_fixes.R

@@ -58,6 +58,8 @@ method.CC_LS_mod <- function()
 #' Issue reported to \code{SuperLearner} maintainers. This function will
 #' be deprecated when a more robust fix in the \code{SuperLearner} package
 #' is implemented.
+#' @importFrom SuperLearner trimLogit
+#' @importFrom stats plogis
 #' @export
 method.CC_nloglik_mod <- function () 
 {
@@ -65,7 +67,7 @@ method.CC_nloglik_mod <- function ()
         if (sum(coef != 0) == 0) {
             stop("All metalearner coefficients are zero, cannot compute prediction.")
         }
-        plogis(trimLogit(predY[, coef != 0], trim = control$trimLogit) %*% 
+        stats::plogis(SuperLearner::trimLogit(predY[, coef != 0], trim = control$trimLogit) %*% 
             matrix(coef[coef != 0]))
     }
     computeCoef = function(Z, Y, libraryNames, obsWeights, control, 
@@ -76,10 +78,10 @@ method.CC_nloglik_mod <- function ()
             warning(paste0("Algorithm ", colDup, " is duplicated. Setting weight to 0."))
             modZ <- modZ[,-colDup]
         }
-        modlogitZ = trimLogit(modZ, control$trimLogit)
-        logitZ = trimLogit(Z, control$trimLogit)
+        modlogitZ = SuperLearner::trimLogit(modZ, control$trimLogit)
+        logitZ = SuperLearner::trimLogit(Z, control$trimLogit)
         cvRisk <- apply(logitZ, 2, function(x) -sum(2 * obsWeights * 
-            ifelse(Y, plogis(x, log.p = TRUE), plogis(x, log.p = TRUE, 
+            ifelse(Y, stats::plogis(x, log.p = TRUE), stats::plogis(x, log.p = TRUE, 
                 lower.tail = FALSE))))
         names(cvRisk) <- libraryNames
         obj_and_grad <- function(y, x, w = NULL) {
@@ -87,12 +89,12 @@ method.CC_nloglik_mod <- function ()
             x <- x
             function(beta) {
                 xB <- x %*% cbind(beta)
-                loglik <- y * plogis(xB, log.p = TRUE) + (1 - 
-                  y) * plogis(xB, log.p = TRUE, lower.tail = FALSE)
+                loglik <- y * stats::plogis(xB, log.p = TRUE) + (1 - 
+                  y) * stats::plogis(xB, log.p = TRUE, lower.tail = FALSE)
[…]

### 2017-12-22 README.Rmd
update biometrika link in readme

@@
 estimators that are doubly robust, not only with respect to consistency, but
-also with respect to asymptotic normality, as discussed in [Benkeser, et al (2017)](https://doi.org/10.1093/biomet/asx053). This property
+also with respect to asymptotic normality, as discussed in [Benkeser, et al (2017)](https://academic.oup.com/biomet/article-abstract/104/4/863/4554445). This property
 facilitates construction of doubly-robust confidence intervals and hypothesis

### 2018-02-05 README.Rmd
update readme example

@@
 # simulate simple data structure
-set.seed(1234)
+set.seed(12345)
 n <- 200
 W <- data.frame(W1 = runif(n,-2,2), W2 = rbinom(n,1,0.5))
-A <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W2))
-Y <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W2 + A))
+A <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W1*W$W2))
+Y <- rbinom(n, 1, plogis(-2 + W$W1 - 2*W$W1*W$W2 + A))
 
@@
 # get confidence intervals for marginal means
+# truth is E[Y(1)] = 0.29, E[Y(0)] = 0.15
 ci_fit1 <- ci(fit1)
@@
 # get confidence intervals for ate
-ci_ate1 <- ci(fit1,contrast = c(-1, 1))
+# truth is E[Y(1)] - E[Y(0)] = 0.14
+ci_ate1 <- ci(fit1, contrast = c(-1, 1))
 # print the output

### 2018-03-08 R/fluctuate.R
bump version. add earth instead of gam to vignette to avoid cran warnings. try new fixes in fluctuateG

@@ -72,7 +72,10 @@ fluctuateG <- function(Y, A, W, DeltaY, DeltaA, a_0, gn, Qrn, tolg, coefTol = 1e
         family = "binomial"
       )
     )
-    if (!fm$converged | abs(fm$coef) > coefTol) {
+    if(is.na(fm$coefficients)){
+      fm$coefficients <- Inf
+    }
+    if (!fm$converged | abs(fm$coefficients) > coefTol) {
       suppressWarnings(
         fm <- stats::glm(
           thisA ~ -1 + offset(off) + H1,
@@ -80,13 +83,16 @@ fluctuateG <- function(Y, A, W, DeltaY, DeltaA, a_0, gn, Qrn, tolg, coefTol = 1e
           family = "binomial"
         )
       )
-      if (!fm$converged | abs(fm$coef) > coefTol) {
-        # warning("No sane fluctuation found for G this iteration. Check mean of IC.")
+      if(is.na(fm$coefficients)){
+        fm$coefficients <- Inf
+      }
+      if (!fm$converged | abs(fm$coefficients) > coefTol) {
+        fm$coefficients <- 0
       }
     }
     pred <- stats::predict(fm, type = "response")
     pred[pred < tolg] <- tolg
-    list(est = pred, eps = fm$coef)
+    list(est = pred, eps = fm$coefficients)
   }, SIMPLIFY = FALSE)
   gnStar
 }
@@ -135,7 +141,7 @@ fluctuateQ2 <- function(Y, A, W, DeltaY, DeltaA,
         data = data.frame(Y = Y, off = off, H2 = H2), family = "binomial"
       )
     )
-    if (!fm$converged | abs(max(fm$coef)) > coefTol) {
+    if (!fm$converged | abs(max(fm$coefficients)) > coefTol) {
       # if it doesn't converge, try with no starting values
       suppressWarnings(
         fm <- stats::glm(
@@ -144,7 +150,7 @@ fluctuateQ2 <- function(Y, A, W, DeltaY, DeltaA,
           family = "binomial"
         )
       )
-      if (!fm$converged | abs(max(fm$coef)) > coefTol) {
+      if (!fm$converged | abs(max(fm$coefficients)) > coefTol) {
         # warning("No sane fluctuation found. Proceeding using current estimates.")
         if (reduction == "univariate") {
           return(list(est = Q, eps = rep(0, 2)))
[…]

### 2018-03-08 R/method_fixes.R

@@ -17,25 +17,36 @@ method.CC_LS_mod <- function() {
       d <- crossprod(wX, wY)
       A <- cbind(rep(1, ncol(wX)), diag(ncol(wX)))
       bvec <- c(1, rep(0, ncol(wX)))
-      fit <- quadprog::solve.QP(
+
+      fit <- tryCatch({quadprog::solve.QP(
         Dmat = D, dvec = d, Amat = A,
         bvec = bvec, meq = 1
-      )
+      )}, error = function(e){
+        "fit error"
+      })
       invisible(fit)
     }
-    colDup <- which(duplicated(round(Z, 8), MARGIN = 2))
+    colDup <- which(duplicated(round(Z, 5), MARGIN = 2))
     modZ <- Z
     if (length(colDup) > 0) {
       warning(paste0("Algorithm ", colDup, " is duplicated. Setting weight to 0."))
       modZ <- modZ[, -colDup]
     }
     fit <- compute(x = modZ, y = Y, wt = obsWeights)
-    coef <- fit$solution
-    if (length(colDup) > 0) {
-      ind <- c(seq_along(coef), colDup - 0.5)
-      coef <- c(coef, rep(0, length(colDup)))
-      coef <- coef[order(ind)]
+    if(class(fit) == "character"){
+      warning(paste0("Error in solve.QP; returning discrete SL weighting instead."))
+      coef <- rep(0, ncol(Z))
+      idx <- which.min(cvRisk)
+      coef[idx[1]] <- 1
+    }else{
+     coef <- fit$solution      
+      if (length(colDup) > 0) {
+        ind <- c(seq_along(coef), colDup - 0.5)
+        coef <- c(coef, rep(0, length(colDup)))
+        coef <- coef[order(ind)]
+      }
     }
+
     if (any(is.na(coef))) {
       warning("Some algorithms have weights of NA, setting to 0.")
       coef[is.na(coef)] <- 0
@@ -75,7 +86,7 @@ method.CC_nloglik_mod <- function() {
   }
   computeCoef <- function(Z, Y, libraryNames, obsWeights, control,
                           verbose, ...) {
-    colDup <- which(duplicated(round(Z, 8), MARGIN = 2))
+    colDup <- which(duplicated(round(Z, 5), MARGIN = 2))
     modZ <- Z
     if (length(colDup) > 0) {
       warning(paste0("Algorithm ", colDup, " is duplicated. Setting weight to 0."))

### 2018-05-07 R/drtmle.R
remove storing epsilon

@@ -392,7 +392,6 @@ drtmle <- function(Y, A, W,
   if ("Q" %in% guard) QrnStar <- Qrn
   gnStar <- gn
   PnDQnStar <- PnDgnStar <- PnDnoStar <- Inf
-  eps <- list(Inf)
   ct <- 0
 
   # fluctuate
@@ -411,12 +410,6 @@ drtmle <- function(Y, A, W,
       gnStar <- plyr::llply(gnStarOut, function(x) {
         unlist(x$est)
       })
-      epsg <- plyr::laply(gnStarOut, function(x) {
-        x$eps
-      })
-    } else {
-      epsg <- NA
-    }
 
     # fluctuate QnStar
     if ("g" %in% guard) {
@@ -451,9 +444,6 @@ drtmle <- function(Y, A, W,
         QnStar <- plyr::llply(QnStarOut, function(x) {
           unlist(x$est)
         })
-        epsQ <- plyr::laply(QnStarOut, function(x) {
-          x$eps
-        })
       } else if (Qsteps == 2) {
         # do the extra targeting
         QnStarOut2 <- fluctuateQ2(
@@ -475,11 +465,6 @@ drtmle <- function(Y, A, W,
         QnStar <- plyr::llply(QnStarOut1, function(x) {
           unlist(x[[1]])
         })
-
-        # for later retrieval of fluct coefficients
-        epsQ <- mapply(q1 = QnStarOut1, q2 = QnStarOut2, function(q1, q2) {
-          c(q1$eps, q2$eps)
-        })
       }
     } else {
       QnStarOut <- fluctuateQ1(
@@ -489,9 +474,6 @@ drtmle <- function(Y, A, W,
       )
       QnStar <- plyr::llply(QnStarOut, function(x) {
         unlist(x[[1]])
-      })
-      epsQ <- plyr::laply(QnStarOut, function(x) {
-        x$eps
       })
     }
 
@@ -517,9 +499,6 @@ drtmle <- function(Y, A, W,
       # obtain list of reduced dimension regression fits
       QrnMod <- QrnValid[seq(2, length(QrnValid), 2)]
     }
-
-    # get fluctuation parameters
-    eps <- c(epsQ, epsg)
 
     # tmle estimates
     psi_t <- lapply(QnStar, mean)

### 2018-05-07 R/drtmle.R
add back in }

@@ -412,6 +412,7 @@ drtmle <- function(Y, A, W,
       gnStar <- plyr::llply(gnStarOut, function(x) {
         unlist(x$est)
       })
+    }
 
     # fluctuate QnStar
     if ("g" %in% guard) {

### 2018-05-07 R/adaptive_iptw.R
remove eps from adapt iptw. remove from cat in drtmle

@@ -251,7 +251,6 @@ adaptive_iptw <- function(W, A, Y,
   gnStar <- gn
   QrnStar <- Qrn
   PnDgnStar <- Inf
-  eps <- list(Inf)
   ct <- 0
   # fluctuate
   while (max(abs(unlist(PnDgnStar))) > tolIC & ct < maxIter) {
@@ -265,9 +264,6 @@ adaptive_iptw <- function(W, A, Y,
     )
     gnStar <- plyr::llply(gnStarOut, function(x) {
       unlist(x$est)
-    })
-    eps <- plyr::laply(gnStarOut, function(x) {
-      x$eps
     })
     # re-estimate reduced dimension regression
     QrnStarOut <- future.apply::future_lapply(
@@ -297,7 +293,6 @@ adaptive_iptw <- function(W, A, Y,
     )
     PnDgnStar <- future.apply::future_lapply(DngoStar, mean)
     if (verbose) {
-      cat("TMLE Iteration", ct, "=", round(unlist(eps), 5), "\n")
       cat("Mean of IC       =", round(unlist(PnDgnStar), 10), "\n")
     }
   }

### 2018-05-07 R/drtmle.R

@@ -529,7 +529,6 @@ drtmle <- function(Y, A, W,
       PnDgnStar <- lapply(DngoStar, mean)
     }
     if (verbose) {
-      cat("TMLE Iteration", ct, "=", round(unlist(eps), 5), "\n")
       cat("Mean of IC       =", round(c(
         unlist(PnDnoStar),
         unlist(PnDQnStar),

### 2018-05-08 R/utils.R
update doc for plot

@@ -110,7 +110,8 @@ print.wald_test.adaptive_iptw <- function(x, digits = 3, ...) {
 #' @param nPoints Number of points to plot lines (increase for less bumpy plot,
 #'  decrease for faster evaluation).
 #' @param a_0 For what value of a_0 should the plot be made for?
-#' @param ... More arguments (not currently used) TO DO: Pass to plot? to lines?
+#' @param ask Boolean indicating whether R should ask to show each plot
+#' @param ... More arguments passed to \code{plot}
 #' @export
 #' @method plot drtmle
 #' @importFrom graphics axis lines par plot
@@ -135,9 +136,10 @@ print.wald_test.adaptive_iptw <- function(x, digits = 3, ...) {
 #' \dontrun{plot(fit1)}
 #
 plot.drtmle <- function(x, nPoints = 500,
+                        ask = TRUE,
                         a_0 = x$a_0[1], ...) {
   # ask to see next plot
-  par(ask = TRUE)
+  par(ask = ask)
   # check if returnModels is null
   if (is.null(x$QrnMod) & is.null(x$grnMod)) {
     stop("Plot function only works if returnModels = TRUE.")
@@ -185,7 +187,7 @@ plot.drtmle <- function(x, nPoints = 500,
     type = "n", xlim = xl, ylim = yl,
     xaxt = "n", yaxt = "n", bty = "n",
     xlab = expression(g[n](W)),
-    ylab = expression("E[Y-" * Q[n](W) * " | " * g[n](W) * "]")
+    ylab = expression("E[Y-" * Q[n](W) * " | " * g[n](W) * "]"), ...
   )
   # add axes
   axis(side = 1)
@@ -231,7 +233,7 @@ plot.drtmle <- function(x, nPoints = 500,
       xaxt = "n", yaxt = "n", bty = "n",
       xlab = expression(Q[n](W)),
       ylab = expression("E[{" * A - g[n](W) * "} / " * g[n](W) * "} | " *
-        Q[n](W) * "]")
+        Q[n](W) * "]", ...)
     )
     # add axes
     axis(side = 1)
@@ -265,7 +267,8 @@ plot.drtmle <- function(x, nPoints = 500,
       type = "n", xlim = xl, ylim = yl,
       xaxt = "n", yaxt = "n", bty = "n",
       xlab = expression(Q[n](W)),
-      ylab = expression("E[A | " * Q[n](W) * "]")
[…]

### 2018-05-08 README.Rmd
update readmes

@@
                                    # misspecification of outcome regression
-               SL_gr = "SL.npreg" # use kernel regression to guard against
+               SL_gr = "SL.npreg", # use kernel regression to guard against
                                   # misspecification of propensity score
+               returnModels = TRUE # for visualizing fits later
               )
@@
 ci_ate1
+```
+
+This method requires estimation of additional univariate regressions to ensure doubly robust confidence intervals and hypothesis tests.
+The method for estimation are input via `SL.Qr` and `SL.gr` or `glm.Qr` and `glm.gr` if parametric models are desired).
+These additional fits can be visualized by the `plot` method for `drtmle`.
+```r
+layout(t(1:3))
+plot(fit1, ask = FALSE)
 ```

### 2018-05-08 R/utils.R
fix plot label

@@ -232,7 +232,7 @@ plot.drtmle <- function(x, nPoints = 500,
       type = "n", xlim = xl, ylim = yl,
       xaxt = "n", yaxt = "n", bty = "n",
       xlab = expression(Q[n](W)),
-      ylab = expression("E[{" * A - g[n](W) * "} / " * g[n](W) * "} | " *
+      ylab = expression("E[{" * A - g[n](W) * "} / " * g[n](W) * " | " *
         Q[n](W) * "]", ...)
     )
     # add axes

### 2018-10-26 R/drtmle.R
add use_future option, resolve model returning issues

@@ -96,6 +96,8 @@
 #'  the input to \code{a_0} (e.g., if \code{a_0 = c(0,1)} then \code{gn[[1]]}
 #'  should be propensity of \code{A} = 0 and \code{gn[[2]]} should be propensity
 #'  of \code{A} = 1).
+#' @param use_future Boolean indicating whether to use \code{future_lapply} or
+#' instead to just use lapply. The latter can be easier to run down errors. 
 #' @param ... Other options (not currently used).
 #'
 #' @return An object of class \code{"drtmle"}.
@@ -200,6 +202,7 @@ drtmle <- function(Y, A, W,
                    future_hpc = NULL,
                    Qn = NULL,
                    gn = NULL,
+                   use_future = FALSE,
                    ...) {
   call <- match.call()
   # if cvFolds non-null split data into cvFolds pieces
@@ -246,14 +249,25 @@ drtmle <- function(Y, A, W,
   # estimate propensity score
   # -------------------------------
   if (is.null(gn)) {
-    gnOut <- future.apply::future_lapply(
-      X = validRows, FUN = estimateG, A = A,
-      W = W, DeltaA = DeltaA, DeltaY = DeltaY,
-      tolg = tolg, verbose = verbose,
-      stratify = stratify,
-      returnModels = returnModels, SL_g = SL_g,
-      glm_g = glm_g, a_0 = a_0
-    )
+    if(use_future){
+      gnOut <- future.apply::future_lapply(
+        X = validRows, FUN = estimateG, A = A,
+        W = W, DeltaA = DeltaA, DeltaY = DeltaY,
+        tolg = tolg, verbose = verbose,
+        stratify = stratify,
+        returnModels = returnModels, SL_g = SL_g,
+        glm_g = glm_g, a_0 = a_0
+      )
+    }else{
+      gnOut <- lapply(
+        X = validRows, FUN = estimateG, A = A,
+        W = W, DeltaA = DeltaA, DeltaY = DeltaY,
+        tolg = tolg, verbose = verbose,
+        stratify = stratify,
+        returnModels = returnModels, SL_g = SL_g,
+        glm_g = glm_g, a_0 = a_0
+      )
+    }
     # re-order predictions
[…]

### 2018-10-26 tests/testthat/test-drtmle.R

@@ -424,3 +424,86 @@ test_that("drtmle executes as expected with stratify = FALSE", {
   expect_true(is.numeric(fit9$aiptw_c$est))
   expect_true(is.numeric(fit9$aiptw_c$cov))
 })
+
+
+# --------------------------------------------------------------------
+
+test_that("drtmle executes when user inputs Qn and gn and returnModels = TRUE", {
+  set.seed(123456)
+  n <- 100
+  W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+  A <- rbinom(n, 1, plogis(W$W1 - W$W2))
+  Y <- rnorm(n, W$W1 * W$W2 * A, 2)
+  Qn <- list(rnorm(n), rnorm(n))
+  tmp <- runif(n)
+  gn <- list(tmp, 1 - tmp)
+
+    fit9 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = FALSE,
+    Qn = Qn, gn = gn,
+    SL_Qr = "SL.glm",
+    SL_gr = "SL.glm",
+    returnModels = TRUE,
+    guard = c("Q", "g"),
+    reduction = "univariate",
+    Qsteps = 2
+  )
+  expect_true(is.numeric(fit9$gcomp$est))
+  expect_true(is.numeric(fit9$tmle$est))
+  expect_true(is.numeric(fit9$tmle$est))
+  expect_true(is.numeric(fit9$tmle$cov))
+  expect_true(is.numeric(fit9$drtmle$est))
+  expect_true(is.numeric(fit9$drtmle$cov))
+  expect_true(is.numeric(fit9$aiptw$est))
+  expect_true(is.numeric(fit9$aiptw$cov))
+  expect_true(is.numeric(fit9$aiptw_c$est))
+  expect_true(is.numeric(fit9$aiptw_c$cov))
+}
+
+# --------------------------------------------------------------------
+
+test_that("GitHub error #16 resolves", {
+  set.seed(123456)
+  X <- runif(100, 0, 1)
+
+Q <- X
+
+g <- exp(X) / (1 + exp(X))
+
+A <- rbinom(100, 1, g)
+
+Y <- runif(Q, -0.1, 0.1)
+
+X <- as.data.frame(X)
+
+a <- drtmle(W = X, A = A, Y = Y, a_0 = 1, glm_Q = 'X', 
[…]

### 2018-10-26 tests/testthat/test-drtmle.R
add missing )

@@ -460,7 +460,7 @@ test_that("drtmle executes when user inputs Qn and gn and returnModels = TRUE", 
   expect_true(is.numeric(fit9$aiptw$cov))
   expect_true(is.numeric(fit9$aiptw_c$est))
   expect_true(is.numeric(fit9$aiptw_c$cov))
-}
+})
 
 # --------------------------------------------------------------------
 
@@ -506,4 +506,4 @@ b <- drtmle(W = X, A = A, Y = Y, a_0 = 1, glm_Q = 'X',
   expect_true(is.numeric(b$aiptw_c$est))
   expect_true(is.numeric(b$aiptw_c$cov))
 
-}
+})

### 2018-10-26 R/drtmle.R
handle naming of user input Qn and gn better

@@ -207,6 +207,11 @@ drtmle <- function(Y, A, W,
   call <- match.call()
   # if cvFolds non-null split data into cvFolds pieces
   n <- length(Y)
+  # save user input Qn and gn, because these values
+  # get overwritten later.
+  Qn_user <- Qn
+  gn_user <- gn
+
   if (length(cvFolds) > 1) {
     stopifnot(length(cvFolds) == length(Y))
     # comes in as vector of fold assignments
@@ -705,10 +710,10 @@ drtmle <- function(Y, A, W,
 
   # tack on models if requested
   if (returnModels) {
-    if(is.null(Qn)){
+    if(is.null(Qn_user)){
       out$QnMod <- QnMod
     }
-    if(is.null(gn)){
+    if(is.null(gn_user)){
       out$gnMod <- gnMod
     }
     if('Q' %in% guard){

### 2018-10-26 README.Rmd
remove errant ,

@@
       note = {R package version 1.0.0},
-      doi = {10.5281/zenodo.844836},
+      doi = {10.5281/zenodo.844836}
     }

### 2018-10-26 R/drtmle.R
change default to use future

@@ -202,7 +202,7 @@ drtmle <- function(Y, A, W,
                    future_hpc = NULL,
                    Qn = NULL,
                    gn = NULL,
-                   use_future = FALSE,
+                   use_future = TRUE,
                    ...) {
   call <- match.call()
   # if cvFolds non-null split data into cvFolds pieces

### 2018-10-26 tests/testthat/test-drtmle.R
add test for bypassing future_lapply

@@ -163,6 +163,28 @@ test_that("drtmle executes as expected with stratify = TRUE", {
   expect_true(is.numeric(fit5$aiptw$cov))
   expect_true(is.numeric(fit5$aiptw_c$est))
   expect_true(is.numeric(fit5$aiptw_c$cov))
+    fit5 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = TRUE,
+    SL_Q = "SL.glm",
+    SL_g = "SL.glm",
+    SL_Qr = "SL.glm",
+    SL_gr = "SL.glm",
+    guard = c("Q", "g"),
+    reduction = "bivariate",
+    use_future = FALSE,
+  )
+  expect_true(is.numeric(fit5$gcomp$est))
+  expect_true(is.numeric(fit5$tmle$est))
+  expect_true(is.numeric(fit5$tmle$est))
+  expect_true(is.numeric(fit5$tmle$cov))
+  expect_true(is.numeric(fit5$drtmle$est))
+  expect_true(is.numeric(fit5$drtmle$cov))
+  expect_true(is.numeric(fit5$aiptw$est))
+  expect_true(is.numeric(fit5$aiptw$cov))
+  expect_true(is.numeric(fit5$aiptw_c$est))
+  expect_true(is.numeric(fit5$aiptw_c$cov))
   # univariate reduction with
   # single SL + stratify
   fit6 <- drtmle(

### 2018-10-26 tests/testthat/test-drtmle.R
remove comma

@@ -173,7 +173,7 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     SL_gr = "SL.glm",
     guard = c("Q", "g"),
     reduction = "bivariate",
-    use_future = FALSE,
+    use_future = FALSE
   )
   expect_true(is.numeric(fit5$gcomp$est))
   expect_true(is.numeric(fit5$tmle$est))

### 2019-07-17 tests/testthat/test-drtmle-repeatsl.R
add to vignette. update site.

@@ -1,0 +1,460 @@
+# install.packages("~/Dropbox/R/drtmle",repos=NULL,type="source")
+library(drtmle)
+library(SuperLearner)
+context("Testing drtmle function with repeated super learners")
+test_that("drtmle executes as expected with parallel = TRUE", {
+  skip_on_os("windows") # Windows doesn't support multicore (esp. Appveyor CI)
+  set.seed(123456)
+  n <- 200
+  W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+  A <- rbinom(n, 1, plogis(W$W1 - W$W2))
+  Y <- rnorm(n, W$W1 * W$W2 * A, 2)
+
+  # univariate reduction with
+  # all GLMs + stratify
+  fit1 <- drtmle(
+    W = W, A = A, Y = Y,
+    parallel = TRUE,
+    family = gaussian(),
+    stratify = TRUE,
+    glm_Q = "W1 + W2",
+    glm_g = "W1 + W2",
+    glm_Qr = "gn",
+    glm_gr = "Qn",
+    guard = c("Q", "g"),
+    reduction = "univariate"
+  )
+
+  expect_true(is.numeric(fit1$gcomp$est))
+  expect_true(is.numeric(fit1$tmle$est))
+  expect_true(is.numeric(fit1$tmle$est))
+  expect_true(is.numeric(fit1$tmle$cov))
+  expect_true(is.numeric(fit1$drtmle$est))
+  expect_true(is.numeric(fit1$drtmle$cov))
+  expect_true(is.numeric(fit1$aiptw$est))
+  expect_true(is.numeric(fit1$aiptw$cov))
+  expect_true(is.numeric(fit1$aiptw_c$est))
+  expect_true(is.numeric(fit1$aiptw_c$cov))
+})
+test_that("drtmle executes as expected with stratify = TRUE", {
+  set.seed(123456)
+  n <- 200
+  W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+  A <- rbinom(n, 1, plogis(W$W1 - W$W2))
+  Y <- rnorm(n, W$W1 * W$W2 * A, 2)
+
+  # univariate reduction with
+  # all GLMs + stratify
+  fit1 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = TRUE,
+    glm_Q = "W1 + W2",
[…]

### 2019-07-22 R/estimate.R
update readme. multi treatment SL uses my version of SL methods

@@ -169,7 +169,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
             Y = as.numeric(trainA[include] == a),
             X = trainW[include, , drop = FALSE], newX = validW,
             family = stats::binomial(), SL.library = SL_g$A,
-            verbose = verbose
+            verbose = verbose, method = tmp_method.CC_nloglik()
           )
           # get predictions
           tmp_pred <- tmp_fm$SL.pred

### 2019-08-26 R/utils.R
update vignette. fix multiple SL + cvFolds > 1 code

@@ -355,8 +355,9 @@ extract_models <- function(a_list){
 #' Make list of rows in each validation fold. 
 #' @param cvFolds Numeric number of cv folds
 #' @param n Number of observations
+#' @param n_SL Number of super learners to fit
 #' @param ... Other arguments
-make_validRows <- function(cvFolds, n, ...){
+make_validRows <- function(cvFolds, n, n_SL, ...){
   if (length(cvFolds) > 1) {
     stopifnot(length(cvFolds) == n)
     # comes in as vector of fold assignments
@@ -371,7 +372,7 @@ make_validRows <- function(cvFolds, n, ...){
     # no cross-validation
     validRows <- list(seq_len(n))
   }
-  return(validRows)
+  return(rep(validRows, n_SL))
 }
 
 #' Temporary fix for convex combination method mean squared error

### 2019-08-26 R/utils.R
undo idiocy of last commit

@@ -372,7 +372,7 @@ make_validRows <- function(cvFolds, n, n_SL, ...){
     # no cross-validation
     validRows <- list(seq_len(n))
   }
-  return(rep(validRows, n_SL))
+  return(validRows)
 }
 
 #' Temporary fix for convex combination method mean squared error

### 2019-08-26 R/utils.R
fix docs

@@ -355,9 +355,8 @@ extract_models <- function(a_list){
 #' Make list of rows in each validation fold. 
 #' @param cvFolds Numeric number of cv folds
 #' @param n Number of observations
-#' @param n_SL Number of super learners to fit
 #' @param ... Other arguments
-make_validRows <- function(cvFolds, n, n_SL, ...){
+make_validRows <- function(cvFolds, n, ...){
   if (length(cvFolds) > 1) {
     stopifnot(length(cvFolds) == n)
     # comes in as vector of fold assignments

### 2020-01-09 R/estimate.R
space del

@@ -143,7 +143,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
           family = stats::binomial(), SL.library = SL_g$A,
           verbose = verbose, method = tmp_method.CC_nloglik()
         ))
-        gn_A <- vector(mode = "list", length = 2)
+        gn_A <- vector(mode = "list", length = 2)        
         gn_A[[1]] <- fm_A[[1]]$SL.predict
         gn_A[[2]] <- 1 - gn_A[[1]]
         # name for this model

### 2020-01-09 README.Rmd
update readme. multi treatment SL uses my version of SL methods

@@
 
-__Author:__ [David Benkeser](https://www.benkeserstatistics.com/)
+__Author:__ [David Benkeser](https://www.sph.emory.edu/faculty/profile/#!dbenkes)
 
@@
 estimators that are doubly robust, not only with respect to consistency, but
-also with respect to asymptotic normality, as discussed in [Benkeser, et al (2017)](https://academic.oup.com/biomet/article-abstract/104/4/863/4554445). This property
-facilitates construction of doubly-robust confidence intervals and hypothesis
-tests.
+also with respect to asymptotic normality, as discussed in [Benkeser, et al.
+(2017)](https://academic.oup.com/biomet/article-abstract/104/4/863/4554445).
+This property facilitates construction of doubly-robust confidence intervals and
+hypothesis tests.
 
@@
 learning, as discussed in
-[van der Laan (2014)](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
+[van der Laan
+(2014)](https://www.degruyter.com/downloadpdf/j/ijb.2014.10.issue-1/ijb-2012-0038/ijb-2012-0038.pdf).
 
@@
 means and contrasts between means. A simple example on simulated data is shown
-below. We refer users to [the vignette](https://benkeser.github.io/drtmle/articles/using_drtmle.html) for more information and further examples.
+below. We refer users to [the
+vignette](https://benkeser.github.io/drtmle/articles/using_drtmle.html) for more
+information and further examples.
 
@@
 
-This method requires estimation of additional univariate regressions to ensure doubly robust confidence intervals and hypothesis tests.
-The method for estimation are input via `SL.Qr` and `SL.gr` or `glm.Qr` and `glm.gr` if parametric models are desired).
-These additional fits can be visualized by the `plot` method for `drtmle`.
+This method requires estimation of additional univariate regressions to ensure
+doubly robust confidence intervals and hypothesis tests. The method for
+estimation are input via `SL.Qr` and `SL.gr` or `glm.Qr` and `glm.gr` if
+parametric models are desired). These additional fits can be visualized by the
+`plot` method for `drtmle`.
 ```r
@@
 
-&copy; 2016-2017 [David C. Benkeser](http://www.benkeserstatistics.com)
+&copy; 2016-2019 [David C. Benkeser](https://www.sph.emory.edu/faculty/profile/#!dbenkes)
 
@@
 
-Copyright (c) 2016-2017 David C. Benkeser
+Copyright (c) 2016-2019 David C. Benkeser
 

### 2020-01-09 R/adaptive_iptw.R
preparing for cran resubmission

@@ -106,11 +106,13 @@
 #' A <- rbinom(n, 1, plogis(W$W1 - W$W2))
 #' Y <- rbinom(n, 1, plogis(W$W1 * W$W2 * A))
 #' # fit iptw with maxIter = 1 to run fast
+#' \donttest{
 #' fit1 <- adaptive_iptw(
 #'   W = W, A = A, Y = Y, a_0 = c(1, 0),
 #'   SL_g = c("SL.glm", "SL.mean", "SL.step"),
 #'   SL_Qr = "SL.npreg", maxIter = 1
 #' )
+#' }
 adaptive_iptw <- function(W, A, Y,
                           DeltaY = as.numeric(!is.na(Y)),
                           DeltaA = as.numeric(!is.na(A)),

### 2020-01-09 R/confint.R

@@ -49,6 +49,7 @@ ci <- function(...) {
 #' W <- data.frame(W1 = runif(n), W2 = rnorm(n))
 #' A <- rbinom(n, 1, plogis(W$W1 - W$W2))
 #' Y <- rbinom(n, 1, plogis(W$W1 * W$W2 * A))
+#' \donttest{
 #' # fit drtmle with maxIter = 1 to run fast
 #' fit1 <- drtmle(
 #'   W = W, A = A, Y = Y, a_0 = c(1, 0),
@@ -83,6 +84,7 @@ ci <- function(...) {
 #'   }
 #' )
 #' ci_RR <- ci(fit1, contrast = myContrast)
+#' }
 ci.drtmle <- function(object, est = c("drtmle"), level = 0.95,
                       contrast = NULL, ...) {
   if (class(object) != "drtmle") {

### 2020-01-09 R/drtmle.R

@@ -36,8 +36,10 @@
 #' @param glm_g A list of characters describing the formulas to be used
 #'  for each of the propensity score regressions (\code{DeltaA}, \code{A}, and
 #'  \code{DeltaY}). To use the same formula for each of the regressions (or if
-#'  there is no missing data in \code{A} nor \code{Y}), a single character
-#'  formula may be input.
+#'  there are no missing data in \code{A} nor \code{Y}), a single character
+#'  formula may be input. In general the formulas can reference any variable in 
+#'  \code{colnames(W)}, unless \code{adapt_g = TRUE} in which case the formulas
+#'  should reference variables \code{QaW} where \code{a} takes values in \code{a_0}.
 #' @param glm_Qr A character describing a formula to be used in the call to
 #'  \code{glm} for reduced-dimension outcome regression. Ignored if
 #'  \code{SL_Qr!=NULL}. The formula should use the variable name \code{'gn'}.
@@ -45,6 +47,10 @@
 #'  \code{glm} for the reduced-dimension propensity score. Ignored if
 #'  \code{SL_gr!=NULL}. The formula should use the variable name \code{'Qn'} and
 #'  \code{'gn'} if \code{reduction='bivariate'} and \code{'Qn'} otherwise.
+#' @param adapt_g A boolean indicating whether the propensity score should be 
+#'  outcome adaptive. If \code{TRUE} then the propensity score is estimated as the
+#'  regression of \code{A} onto covariates \code{QaW} for \code{a} in each value
+#'  contained in \code{a_0}. See vignette for more details. 
 #' @param guard A character vector indicating what pattern of misspecifications
 #'  to guard against. If \code{guard} contains \code{"Q"}, then the TMLE guards
 #'  against misspecification of the outcome regression by estimating the
@@ -196,6 +202,7 @@ drtmle <- function(Y, A, W,
                    n_SL = 1,
                    glm_Q = NULL, glm_g = NULL,
                    glm_Qr = NULL, glm_gr = NULL,
+                   adapt_g = FALSE, 
                    guard = c("Q", "g"),
                    reduction = "univariate",
                    returnModels = FALSE,
@@ -220,7 +227,11 @@ drtmle <- function(Y, A, W,
 
   # check if any additional targeting is performed
   extra_targeting <- !is.null(guard)
-
[…]

### 2020-01-09 R/estimate.R

@@ -28,12 +28,16 @@
 #'  mean estimates.
 #' @param validRows A \code{list} of length \code{cvFolds} containing the row
 #'  indexes of observations to include in validation fold.
+#' @param Qn A \code{list} of estimates of the outcome regression for each value
+#'  in \code{a_0}. Only needed if \code{adapt_g = TRUE}. 
+#' @param adapt_g A boolean indicating whether propensity score is adaptive
+#'  to outcome regression. 
 #' @importFrom SuperLearner SuperLearner trimLogit All
 #' @importFrom stats predict glm as.formula
 #
 estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
                       stratify = FALSE, validRows = NULL, verbose = FALSE,
-                      returnModels = FALSE) {
+                      returnModels = FALSE, Qn = NULL, adapt_g = FALSE) {
   if (is.null(SL_g) & is.null(glm_g)) {
     stop("Specify Super Learner library or GLM formula for g")
   }
@@ -50,14 +54,28 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
     trainDeltaA <- DeltaA[-validRows]
     trainDeltaY <- DeltaY[-validRows]
     trainA <- A[-validRows]
-    trainW <- W[-validRows, , drop = FALSE]
-    validW <- W[validRows, , drop = FALSE]
+    if(!adapt_g){
+      trainW <- W[-validRows, , drop = FALSE]
+      validW <- W[validRows, , drop = FALSE]
+    }else{
+      allW <- data.frame(Reduce(cbind, Qn))
+      trainW <- allW[-validRows, , drop = FALSE]
+      validW <- allW[validRows, , drop = FALSE]
+      colnames(trainW) <- paste0("Q", a_0, "W")
+      colnames(validW) <- paste0("Q", a_0, "W")
+    }
     validA <- A[validRows]
     validDeltaA <- DeltaA[validRows]
     validDeltaY <- DeltaY[validRows]
   } else {
     trainA <- validA <- A
-    trainW <- validW <- W
+    if(!adapt_g){
+      trainW <- validW <- W
+    }else{
+      trainW <- validW <- data.frame(Reduce(cbind, Qn))
+      colnames(trainW) <- paste0("Q", a_0, "W")
+      colnames(validW) <- paste0("Q", a_0, "W")
+    }
[…]

### 2020-01-09 R/utils.R

@@ -127,6 +127,7 @@ print.wald_test.adaptive_iptw <- function(x, digits = 3, ...) {
 #' A <- rbinom(n, 1, plogis(W$W1 - W$W2))
 #' Y <- rbinom(n, 1, plogis(W$W1 * W$W2 * A))
 #' # fit drtmle with maxIter = 1 to run fast
+#' \donttest{
 #' fit1 <- drtmle(
 #'   W = W, A = A, Y = Y, a_0 = c(1, 0),
 #'   family = binomial(),
@@ -137,7 +138,7 @@ print.wald_test.adaptive_iptw <- function(x, digits = 3, ...) {
 #'   maxIter = 1, returnModels = TRUE
 #' )
 #' # plot the reduced-dimension regression fits (not run)
-#' \dontrun{
+#' 
 #' plot(fit1)
 #' }
 #' #

### 2020-01-09 tests/testthat/test-drtmle.R

@@ -532,3 +532,32 @@ test_that("GitHub error #16 resolves", {
   expect_true(is.numeric(b$aiptw_c$est))
   expect_true(is.numeric(b$aiptw_c$cov))
 })
+
+test_that("drtmle executes with adapt_g", {
+  set.seed(123456)
+  n <- 100
+  W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+  A <- rbinom(n, 1, plogis(W$W1 - W$W2))
+  Y <- rnorm(n, W$W1 * W$W2 * A, 2)
+  
+  fit10 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = FALSE,
+    adapt_g = TRUE, 
+    glm_Q = ".^2", glm_g = ".",
+    a_0 = c(0,1),
+    returnModels = TRUE,
+    guard = c("Q", "g"),
+    reduction = "univariate"
+  )
+
+  # fit externally as well
+  Qmod <- glm(Y ~ .^2 , data = data.frame(A = A, W))
+  Qnframe <- data.frame(Q0W = predict(Qmod, newdata = data.frame(A = 0, W)),
+                        Q1W = predict(Qmod, newdata = data.frame(A = 1, W)))
+  gmod <- glm(A ~ . , data = Qnframe, family = binomial())
+  gn <- list(1 - gmod$fitted.values,  gmod$fitted.values)
+  expect_true(all(fit10$nuisance_drtmle$gn[[1]] - gn[[1]] < 1e-4))
+  expect_true(all(fit10$nuisance_drtmle$gn[[2]] - gn[[2]] < 1e-4))
+})

### 2020-12-28 R/adaptive_iptw.R
prepping for v1.1.0 submission

@@ -45,15 +45,6 @@
 #' @param cvFolds A numeric equal to the number of folds to be used in
 #'  cross-validated fitting of nuisance parameters. If \code{cvFolds = 1}, no
 #'  cross-validation is used.
-#' @param parallel A logical indicating whether to use parallelization based on
-#'  \code{future} to estimate nuisance parameters in parallel. Only useful if
-#'  \code{cvFolds > 1}. By default, a \code{multiprocess} evaluation scheme is
-#'  invoked, using forked R processes (if supported on the OS) and background R
-#'  sessions otherwise. Users may also register their own backends using the
-#'  \code{future.batchtools} package.
-#' @param future_hpc A character string identifying a high-performance computing
-#'  backend to be used with parallelization. This should match exactly one of
-#'  the options available from the \code{future.batchtools} package.
 #' @param gn An optional list of propensity score estimates. If specified, the
 #'  function will ignore the nuisance parameter estimation specified by
 #'  \code{SL_g} and \code{glm_g}. The entries in the list should correspond to
@@ -89,9 +80,7 @@
 #'  \item{\code{call}}{The call to \code{adaptive_iptw}.}
 #' }
 #'
-#' @importFrom future plan
 #' @importFrom future.apply future_lapply
-#' @importFrom doFuture registerDoFuture
 #' @importFrom stats cov
 #'
 #' @export
@@ -131,8 +120,6 @@ adaptive_iptw <- function(W, A, Y,
                           tolIC = 1 / length(Y),
                           tolg = 1e-2,
                           cvFolds = 1,
-                          parallel = FALSE,
-                          future_hpc = NULL,
                           gn = NULL,
                           ...) {
   call <- match.call()
@@ -143,30 +130,7 @@ adaptive_iptw <- function(W, A, Y,
   } else {
     validRows <- list(seq_len(n))
   }
-  # use futures with foreach if parallel mode
-  if (!parallel) {
-    future::plan(future::transparent)
-  } else {
-    doFuture::registerDoFuture()
-    if (all(c("sequential", "uniprocess") %in% class(future::plan())) &
-      is.null(future_hpc)) {
-      future::plan(future::multiprocess)
-    } else if (!is.null(future_hpc)) {
-      if (future_hpc == "batchtools_torque") {
[…]

### 2020-12-28 R/drtmle.R

@@ -31,6 +31,33 @@
 #' @param n_SL Number of repeated Super Learners to run (default 1) for the
 #'  each nuisance parameter. Repeat Super Learners more times to obtain more stable
 #'  inference.
+#' @param avg_over If multiple Super Learners are run, on which scale should the
+#' results be aggregated. Options include: \code{"SL"} = 
+#' repeated nuisance parameter estimates are averaged before subsequently 
+#' generating a single vector of point estimates based on the averaged models;
+#' \code{"drtmle"} = repeated vectors of point estimates are generated and 
+#' averaged. Both can be specified, recognizing that this adds considerable
+#' computational expense. In this case, the final estimates are the average 
+#' of \code{n_SL} point estimates where each is built by averaging \code{n_SL} 
+#' fits. If \code{NULL}, no averaging is performed (in which case \code{n_SL} 
+#' should be set equal to 1). 
+#' @param se_cv Should cross-validated nuisance parameter estimates be used 
+#' for computing standard errors? 
+#' Options are \code{"none"} = no cross-validation is performed; \code{"partial"} = 
+#' only applicable if Super Learner is used for nuisance parameter estimates; 
+#' \code{"full"} = full cross-validation is performed. See vignette for further 
+#' details. Ignored if \code{cvFolds > 1}, since then
+#' cross-validated nuisance parameter estimates are used by default and it is 
+#' assumed that you want full cross-validated standard errors. 
+#' @param se_cvFolds If cross-validated nuisance parameter estimates are used
+#' to compute standard errors, how many folds should be used in this computation. 
+#' If \code{se_cv = "partial"}, then this option sets the number of folds used
+#' by the \code{SuperLearner} fitting procedure. 
+#' @param targeted_se A boolean indicating whether the targeted nuisance 
+#' parameters should be used in standard error computation or the initial 
+#' estimators. If \code{se_cv} is not set to \code{"none"}, this option is 
+#' ignored and standard errors are computed based on non-targeted, cross-validated 
+#' nuisance parameter fits. 
 #' @param glm_Q A character describing a formula to be used in the call to
[…]

### 2020-12-28 R/estimate.R

@@ -32,12 +32,25 @@
 #'  in \code{a_0}. Only needed if \code{adapt_g = TRUE}. 
 #' @param adapt_g A boolean indicating whether propensity score is adaptive
 #'  to outcome regression. 
+#' @param se_cv Should cross-validated nuisance parameter estimates be used 
+#' for computing standard errors? 
+#' Options are \code{"none"} = no cross-validation is performed; \code{"partial"} = 
+#' only applicable if Super Learner is used for nuisance parameter estimates; 
+#' \code{"full"} = full cross-validation is performed. See vignette for further 
+#' details. Ignored if \code{cvFolds > 1}, since then
+#' cross-validated nuisance parameter estimates are used by default and it is 
+#' assumed that you want full cross-validated standard errors. 
+#' @param se_cvFolds If cross-validated nuisance parameter estimates are used
+#' to compute standard errors, how many folds should be used in this computation. 
+#' If \code{se_cv = "partial"}, then this option sets the number of folds used
+#' by the \code{SuperLearner} fitting procedure. 
 #' @importFrom SuperLearner SuperLearner trimLogit All
 #' @importFrom stats predict glm as.formula
 #
 estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
                       stratify = FALSE, validRows = NULL, verbose = FALSE,
-                      returnModels = FALSE, Qn = NULL, adapt_g = FALSE) {
+                      returnModels = FALSE, Qn = NULL, adapt_g = FALSE,
+                      se_cv = "none", se_cvFolds = 10) {
   if (is.null(SL_g) & is.null(glm_g)) {
     stop("Specify Super Learner library or GLM formula for g")
   }
@@ -80,6 +93,8 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
     trainDeltaY <- validDeltaY <- DeltaY
   }
 
+  partial_cv <- se_cv == "partial"
+
   if (!is.null(SL_g)) {
     # check for names in SL_g
     namedSL_g <- c("DeltaA", "A", "DeltaY") %in% names(SL_g)
@@ -110,10 +125,17 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
           Y = trainDeltaA,
           X = trainW, newX = validW, family = stats::binomial(),
[…]

### 2020-12-28 R/utils.R

@@ -288,9 +288,6 @@ plot.drtmle <- function(x, nPoints = 500,
   }
 }
 
-# SHOULD MAKE THIS FUNCTION TAKE AS INPUT THE NUMBER OF SUPER LEARNERS AND
-# RETURN AS OUTPUT SOMETHING REORDERED/AVERAGED AS NECESSARY
-
 #' Helper function to reorder lists according to cvFolds
 #'
 #' @param a_list Structured list of nuisance parameters
@@ -301,14 +298,17 @@ plot.drtmle <- function(x, nPoints = 500,
 #' @param n_SL Number of super learners. If >1, then predictions
 #' are averaged
 #' @param n Sample size
+#' @param for_se_cv Is this being used to average over 
+#' cross-validated standard errors? Affects index of \code{a_list}. 
 reorder_list <- function(a_list,
                          a_0,
                          validRows,
                          n_SL = 1,
                          grn_ind = FALSE,
-                         n) {
+                         n,
+                         for_se_cv = FALSE) {
   n_cvFolds <- length(validRows) / n_SL
-
+  est_index <- ifelse(for_se_cv, 3, 1)
 
   reduced_outList <- vector(mode = "list", length = length(a_0))
 
@@ -326,7 +326,7 @@ reorder_list <- function(a_list,
       recursive = FALSE, use.names = FALSE
     )
     # this is in 0/1 format
-    outListUnOrd <- do.call(Map, c(c, outListValid[seq(1, length(outListValid), 2)]))
+    outListUnOrd <- do.call(Map, c(c, outListValid[seq(est_index, length(outListValid), length(a_list[[1]]))]))
     outList <- vector(mode = "list", length = length(a_0))
     if (!grn_ind) {
       for (i in seq_along(a_0)) {
@@ -359,7 +359,7 @@ reorder_list <- function(a_list,
 #' @param a_list Structured list of nuisance parameters
 extract_models <- function(a_list) {
   outListValid <- unlist(a_list, recursive = FALSE, use.names = FALSE)
-  outListValid[seq(2, length(outListValid), 2)]
+  outListValid[seq(2, length(outListValid), 3)]
 }
 
 #' Make list of rows in each validation fold.
@@ -431,12 +431,12 @@ tmp_method.CC_LS <- function() {
     tol <- 4
     dupCols <- which(duplicated(round(Z, tol), MARGIN = 2))
     anyDupCols <- length(dupCols) > 0
-    if (anyDupCols) {
-      warning(paste0(
[…]

### 2020-12-28 tests/testthat/test-drtmle-averaging.R

@@ -1,0 +1,235 @@
+# devtools::load_all("~/Dropbox/R/drtmle")
+library(drtmle)
+library(SuperLearner)
+context("Testing drtmle function with repeated super learners when averaging on drtmle scale")
+
+test_that("drtmle executes as expected with stratify = TRUE", {
+  set.seed(123456)
+  n <- 200
+  W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+  A <- rbinom(n, 1, plogis(W$W1 - W$W2))
+  Y <- rnorm(n, W$W1 * W$W2 * A, 2)
+
+  # univariate reduction with
+  # all GLMs + stratify
+  fit1 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = TRUE,
+    glm_Q = "W1 + W2",
+    glm_g = "W1 + W2",
+    glm_Qr = "gn",
+    glm_gr = "Qn",
+    guard = c("Q", "g"),
+    reduction = "univariate"
+  )
+
+  expect_true(is.numeric(fit1$gcomp$est))
+  expect_true(is.numeric(fit1$tmle$est))
+  expect_true(is.numeric(fit1$tmle$est))
+  expect_true(is.numeric(fit1$tmle$cov))
+  expect_true(is.numeric(fit1$drtmle$est))
+  expect_true(is.numeric(fit1$drtmle$cov))
+  expect_true(is.numeric(fit1$aiptw$est))
+  expect_true(is.numeric(fit1$aiptw$cov))
+  expect_true(is.numeric(fit1$aiptw_c$est))
+  expect_true(is.numeric(fit1$aiptw_c$cov))
+
+  # bivariate reduction with
+  # all GLMs + stratify
+  fit2 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = TRUE,
+    glm_Q = "W1 + W2",
+    glm_g = "W1 + W2",
+    glm_Qr = "gn",
+    glm_gr = "Qn + gn",
+    guard = c("Q", "g"),
+    reduction = "bivariate"
+  )
+
+  expect_true(is.numeric(fit2$gcomp$est))
+  expect_true(is.numeric(fit2$tmle$est))
+  expect_true(is.numeric(fit2$tmle$est))
+  expect_true(is.numeric(fit2$tmle$cov))
[…]

### 2020-12-28 tests/testthat/test-drtmle-missingAY.R

@@ -11,9 +11,9 @@ test_that("drtmle executes as expected with stratify = TRUE", {
   set.seed(123456)
   n <- 200
   W <- data.frame(W1 = runif(n), W2 = rnorm(n))
-  DeltaA <- 1 - rbinom(n, 1, plogis(-4 + W$W1 - W$W2))
+  DeltaA <- 1 - rbinom(n, 1, plogis(-2 + W$W1 - W$W2))
   A <- rbinom(n, 1, plogis(W$W1 - W$W2))
-  DeltaY <- 1 - rbinom(n, 1, plogis(-4 + W$W1 - A))
+  DeltaY <- 1 - rbinom(n, 1, plogis(-2 + W$W1 - A))
   Y <- rnorm(n, W$W1 * W$W2 * A, 2)
   Y[DeltaY == 0] <- NA
   A[DeltaA == 0] <- NA
@@ -85,7 +85,7 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
-    reduction = "univariate"
+    reduction = "univariate", use_future = FALSE
   )
   expect_true(is.numeric(fit3$gcomp$est))
   expect_true(is.numeric(fit3$tmle$est))
@@ -176,9 +176,9 @@ test_that("drtmle executes as expected with stratify = FALSE", {
   set.seed(123456)
   n <- 200
   W <- data.frame(W1 = runif(n), W2 = rnorm(n))
-  DeltaA <- 1 - rbinom(n, 1, plogis(-4 + W$W1 - W$W2))
+  DeltaA <- 1 - rbinom(n, 1, plogis(-2 + W$W1 - W$W2))
   A <- rbinom(n, 1, plogis(W$W1 - W$W2))
-  DeltaY <- 1 - rbinom(n, 1, plogis(-4 + W$W1 - A))
+  DeltaY <- 1 - rbinom(n, 1, plogis(-2 + W$W1 - A))
   Y <- rnorm(n, W$W1 * W$W2 * A, 2)
   Y[DeltaY == 0] <- NA
   A[DeltaA == 0] <- NA
@@ -383,9 +383,9 @@ test_that("drtmle executes as expected with stratify = FALSE", {
   set.seed(123456)
[…]

### 2020-12-28 tests/testthat/test-drtmle-repeatsl.R

@@ -46,6 +46,7 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     glm_Qr = "gn",
     glm_gr = "Qn + gn",
     guard = c("Q", "g"),
+    n_SL = 2, avg_over = "SL",
     reduction = "bivariate"
   )
 
@@ -71,8 +72,10 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
-    n_SL = 2,
-    reduction = "univariate"
+    n_SL = 2, avg_over = "SL",
+    reduction = "univariate",
+    returnModels = TRUE,
+    use_future = FALSE
   )
   expect_true(is.numeric(fit3$gcomp$est))
   expect_true(is.numeric(fit3$tmle$est))
@@ -96,7 +99,7 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
-    n_SL = 2,
+    n_SL = 2, avg_over = "SL",
     reduction = "bivariate"
   )
   expect_true(is.numeric(fit4$gcomp$est))
@@ -122,7 +125,7 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     SL_Qr = "SL.glm",
     SL_gr = "SL.glm",
     guard = c("Q", "g"),
-    n_SL = 2,
+    n_SL = 2, avg_over = "SL",
     reduction = "bivariate",
     returnModels = TRUE, use_future = FALSE
   )
@@ -151,7 +154,7 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     SL_g = "SL.glm",
     SL_Qr = "SL.glm",
     SL_gr = "SL.glm",
-    n_SL = 2,
+    n_SL = 2, avg_over = "SL",
     guard = c("Q", "g"),
     reduction = "univariate"
   )
@@ -242,7 +245,7 @@ test_that("drtmle executes as expected with stratify = FALSE", {
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
-    n_SL = 2,
+    n_SL = 2, avg_over = "SL",
     reduction = "bivariate"
   )
[…]

### 2020-12-28 tests/testthat/test-drtmle.R

@@ -99,8 +99,8 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     W = W, A = A, Y = Y,
     family = gaussian(),
     stratify = TRUE,
-    SL_Q = c("SL.glm", "SL.step"),
-    SL_g = c("SL.glm", "SL.step"),
+    SL_Q = c("SL.glm", "SL.mean"),
+    SL_g = c("SL.glm", "SL.mean"),
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
@@ -123,8 +123,8 @@ test_that("drtmle executes as expected with stratify = TRUE", {
     W = W, A = A, Y = Y,
     family = gaussian(),
     stratify = TRUE,
-    SL_Q = c("SL.glm", "SL.step"),
-    SL_g = c("SL.glm", "SL.step"),
+    SL_Q = c("SL.glm", "SL.mean"),
+    SL_g = c("SL.glm", "SL.mean"),
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
@@ -276,8 +276,8 @@ test_that("drtmle executes as expected with stratify = FALSE", {
     W = W, A = A, Y = Y,
     family = gaussian(),
     stratify = FALSE,
-    SL_Q = c("SL.glm", "SL.step"),
-    SL_g = c("SL.glm", "SL.step"),
+    SL_Q = c("SL.glm", "SL.mean"),
+    SL_g = c("SL.glm", "SL.mean"),
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
@@ -300,8 +300,8 @@ test_that("drtmle executes as expected with stratify = FALSE", {
     W = W, A = A, Y = Y,
     family = gaussian(),
     stratify = FALSE,
-    SL_Q = c("SL.glm", "SL.step"),
-    SL_g = c("SL.glm", "SL.step"),
+    SL_Q = c("SL.glm", "SL.mean"),
+    SL_g = c("SL.glm", "SL.mean"),
     SL_Qr = c("SL.glm", "SL.mean"),
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),

### 2020-12-28 tests/testthat/test-edge.R

@@ -13,7 +13,7 @@ test_that("drtmle executes as expected when only one value of gn", {
 
   # univariate reduction with
   # all GLMs + stratify
-  expect_warning(fit1 <- drtmle(
+  fit1 <- drtmle(
     W = W, A = A, Y = Y,
     cvFolds = 1, maxIter = 2,
     family = gaussian(),
@@ -24,7 +24,7 @@ test_that("drtmle executes as expected when only one value of gn", {
     glm_gr = "Qn",
     guard = c("Q", "g"),
     reduction = "univariate"
-  ))
+  )
   expect_true(is.numeric(fit1$gcomp$est))
   expect_true(is.numeric(fit1$tmle$est))
   expect_true(is.numeric(fit1$tmle$est))

### 2020-12-28 tests/testthat/test-fluctuate.R

@@ -40,20 +40,10 @@ test_that("Fail safe kicks in", {
       returnModels = returnModels, SL_g = SL_g,
       glm_g = glm_g, a_0 = a_0
     )
-  } else {
-    gnOut <- foreach::foreach(v = 1:cvFolds, .packages = "SuperLearner") %dopar% {
-      estimateG(
-        A = A, W = W, DeltaA = DeltaA, DeltaY = DeltaY,
-        tolg = tolg, verbose = verbose, stratify = stratify,
-        returnModels = returnModels, SL_g = SL_g,
-        glm_g = glm_g, a_0 = a_0, validRows = validRows[[v]]
-      )
-    }
-  }
-
+  } 
   # # re-order predictions
   gnValid <- unlist(gnOut, recursive = FALSE, use.names = FALSE)
-  gnUnOrd <- do.call(Map, c(c, gnValid[seq(1, length(gnValid), 2)]))
+  gnUnOrd <- do.call(Map, c(c, gnValid[seq(1, length(gnValid), 3)]))
   gn <- vector(mode = "list", length = length(a_0))
   for (i in 1:length(a_0)) {
     gn[[i]] <- rep(NA, n)
@@ -69,21 +59,14 @@ test_that("Fail safe kicks in", {
       Y = Y, A = A, W = W, DeltaA = DeltaA, DeltaY = DeltaY,
       verbose = verbose, returnModels = returnModels,
       SL_Q = SL_Q, a_0 = a_0, stratify = stratify,
-      glm_Q = glm_Q, family = family
+      glm_Q = glm_Q, family = family,
+      se_cv = "none", se_cvFolds = 10
     )
-  } else {
-    QnOut <- foreach::foreach(v = 1:cvFolds, .packages = "SuperLearner") %dopar% {
-      estimateQ(
-        Y = Y, A = A, W = W, DeltaA = DeltaA, DeltaY = DeltaY,
-        verbose = verbose, returnModels = returnModels,
-        SL_Q = SL_Q, a_0 = a_0, glm_Q = glm_Q, family = family,
-        stratify = stratify, validRows = validRows[[v]]
-      )
-    }
-  }
+  } 
+
   # re-order predictions
   QnValid <- unlist(QnOut, recursive = FALSE, use.names = FALSE)
-  QnUnOrd <- do.call(Map, c(c, QnValid[seq(1, length(QnValid), 2)]))
[…]

### 2020-12-28 tests/testthat/test-multileveltrt.R

@@ -53,7 +53,7 @@ test_that("drtmle executes as expected with multiple treatment levels", {
     family = gaussian(),
     stratify = TRUE,
     glm_Q = "W1 + W2",
-    SL_g = c("SL.step", "SL.glm"),
+    SL_g = c("SL.mean", "SL.glm"),
     glm_Qr = "gn",
     glm_gr = "Qn",
     guard = c("Q", "g"),
@@ -119,7 +119,7 @@ test_that("adaptive_iptw executes as expected with multiple treatment levels", {
   # all GLMs + stratify
   fit1 <- adaptive_iptw(
     W = W, A = A, Y = Y, a_0 = c(0, 1, 2),
-    SL_g = c("SL.step", "SL.step.interaction"),
+    SL_g = c("SL.mean", "SL.glm"),
     returnModels = TRUE,
     glm_Qr = "gn"
   )

### 2020-12-28 tests/testthat/test-partial-cv.R

@@ -1,0 +1,422 @@
+# devtools::load_all("~/Dropbox/R/drtmle")
+library(drtmle)
+library(SuperLearner)
+context("Testing drtmle function with partial CV")
+
+test_that("Same point estimates, different variance estimates, single treatment", {
+  set.seed(123456)
+  n <- 200
+  W <- data.frame(W1 = runif(n), W2 = rnorm(n))
+  A <- rbinom(n, 1, plogis(W$W1 - W$W2))
+  Y <- rnorm(n, W$W1 * W$W2 * A, 2)
+
+  # univariate reduction with
+  # all SL + stratify
+  set.seed(1234)
+  fit1 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = TRUE,
+    SL_Q = c("SL.glm", "SL.mean"),
+    SL_g = c("SL.glm", "SL.mean"),
+    SL_Qr = c("SL.glm", "SL.mean"),
+    SL_gr = c("SL.glm", "SL.mean"),
+    guard = c("Q", "g"),
+    reduction = "univariate",
+    returnModels = TRUE
+  )
+
+  set.seed(1234)
+  fit2 <- drtmle(
+    W = W, A = A, Y = Y,
+    family = gaussian(),
+    stratify = TRUE,
+    SL_Q = c("SL.glm", "SL.mean"),
+    SL_g = c("SL.glm", "SL.mean"),
+    SL_Qr = c("SL.glm", "SL.mean"),
+    SL_gr = c("SL.glm", "SL.mean"),
+    guard = c("Q", "g"),
+    se_cv = "partial", se_cvFolds = 10,
+    targeted_se = FALSE, 
+    reduction = "univariate",
+    returnModels = TRUE,
+    use_future = FALSE
+  )
+
+  expect_true(is.numeric(fit2$gcomp$est))
+  expect_true(is.numeric(fit2$tmle$est))
+  expect_true(is.numeric(fit2$tmle$est))
+  expect_true(is.numeric(fit2$tmle$cov))
+  expect_true(is.numeric(fit2$drtmle$est))
+  expect_true(is.numeric(fit2$drtmle$cov))
+  expect_true(is.numeric(fit2$aiptw$est))
+  expect_true(is.numeric(fit2$aiptw$cov))
[…]

### 2020-12-28 tests/testthat/test-partial-cv.R
try to fix test error for appveyor

@@ -323,7 +323,8 @@ test_that("Same point estimates, different variance estimates, missing data trea
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = TRUE,
+    use_future = FALSE
   )
 
   set.seed(1234)
@@ -378,7 +379,8 @@ test_that("Same point estimates, different variance estimates, missing data trea
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = FALSE,
+    use_future = FALSE
   )
 
   set.seed(1234)
@@ -394,7 +396,7 @@ test_that("Same point estimates, different variance estimates, missing data trea
     se_cv = "partial", se_cvFolds = 10,
     targeted_se = FALSE, 
     reduction = "univariate",
-    returnModels = TRUE,
+    returnModels = FALSE,
     use_future = FALSE
   )
 

### 2020-12-28 tests/testthat/test-partial-cv.R
remove future from all partial cv tests

@@ -78,7 +78,8 @@ test_that("Same point estimates, different variance estimates, single treatment"
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = TRUE,
+    use_future = FALSE
   )
 
   set.seed(1234)
@@ -142,7 +143,8 @@ test_that("Same point estimates, different variance estimates, multi treatment",
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = TRUE,
+    use_future = FALSE
   )
 
   set.seed(1234)
@@ -197,7 +199,8 @@ test_that("Same point estimates, different variance estimates, multi treatment",
     SL_gr = c("SL.glm", "SL.mean"),
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = TRUE,
+    use_future = FALSE
   )
 
   set.seed(1234)
@@ -260,7 +263,8 @@ test_that("drtmle executes as expected with multiple sl and multiple tx lvl", {
     glm_gr = "Qn",
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = TRUE,
+    use_future = FALSE
   )
 
   # univariate reduction with
@@ -276,7 +280,8 @@ test_that("drtmle executes as expected with multiple sl and multiple tx lvl", {
     n_SL = 2, avg_over = "drtmle", 
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = TRUE,
+    use_future = FALSE
   )
   expect_true(all(abs(fit2$drtmle$est - fit1$drtmle$est) < 1e-5))
   expect_true(all(abs(fit2$tmle$est - fit1$tmle$est) < 1e-5))
@@ -295,7 +300,8 @@ test_that("drtmle executes as expected with multiple sl and multiple tx lvl", {
     n_SL = 1, avg_over = "drtmle", 
     guard = c("Q", "g"),
     reduction = "univariate",
-    returnModels = TRUE
+    returnModels = TRUE,
+    use_future = FALSE
   )
   expect_true(all(abs(fit3$drtmle$est - fit1$drtmle$est) < 1e-5))
   expect_true(all(abs(fit3$tmle$est - fit1$tmle$est) < 1e-5))

### 2021-06-02 R/drtmle.R
turn off future by default

@@ -237,7 +237,7 @@ drtmle <- function(Y, A, W,
                    Qsteps = 2,
                    Qn = NULL,
                    gn = NULL,
-                   use_future = TRUE,
+                   use_future = FALSE,
                    ...) {
   call <- match.call()
   n <- length(Y)

### 2021-06-03 R/drtmle.R
update docs

@@ -695,7 +695,8 @@ drtmle <- function(Y, A, W,
               glm_Qr = glm_Qr,
               family = stats::gaussian(),
               SL_Qr = SL_Qr, a_0 = a_0,
-              returnModels = returnModels
+              returnModels = returnModels,
+              future.seed = TRUE
             )
           } else {
             QrnStarOut <- lapply(

### 2021-06-03 paper/using_drtmle_jss.Rmd
catch a typo; add orcids

---
documentclass: jss
author:
  - name: David Benkeser
    affiliation: Emory University
    address: |
      | 1518 Clifton Road, NE
      | Mailstop: 002-3AA
      | Atlanta, Georgia, 30322, U.S.A.
    email: \email{benkeser@emory.edu}
    orcdi: 0000-0002-1019-8343
  - name: Nima S. Hejazi
    affiliation: University of California, Berkeley
    email: \email{nhejazi@berkeley.edu}
    orcid: 0000-0002-7127-2789
title:
  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
  plain: "Doubly-Robust Inference in R using drtmle"
  short: "\\pkg{drtmle}: Doubly-robust Inference"
keywords:
  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
preamble: >
  \usepackage{amsmath}
abstract: >
  Inverse probability of treatment weighted estimators and doubly-robust
  estimators (including augmented inverse probability of treatment weight and
  targeted minimum loss-based estimators) are widely used in causal inference to
  estimate and draw inference about the average effect of a treatment. As an
  intermediate step, these estimators require estimation of key nuisance
  parameters, which are often regression functions. Typically, regressions are
  estimated using maximum likelihood and parametric models. Confidence intervals
  and p-values may be computed based on standard asymptotic results, such as the
  central limit theorem, the delta method, and the nonparametric bootstrap.
  However, in high-dimensional settings maximum likelihood estimation often
  breaks down and standard procedures no longer yield correct inference.
  Instead, we may rely on adaptive estimators of nuisance parameters to
  construct flexible regression estimators.  However, use of adaptive estimators
  poses a challenge for performing statistical inference about an estimated
  treatment effect. While doubly-robust estimators facilitate inference when
  *all* relevant regression functions are consistently estimated, the same
  cannot be said when at least one estimator is inconsistent. \pkg{drtmle}
  implements doubly-robust confidence intervals and hypothesis tests about
  targeted minimum loss-based estimates of the average treatment effect. The
  package additionally implements an inverse probability of treatment weighted
  estimator that yields valid inference even when the probability of treatment
  is estimated via adaptive methods.
output: rticles::jss_article
bibliography: refs
---

```{r, setup, include=FALSE}
options(prompt = 'R> ', continue = '+ ')
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

# Introduction

In many fields researchers are interested in assessing the population-level
effect of a particular treatment on an outcome of interest. The treatment might
correspond to a drug, a potentially harmful exposure, or a policy intervention.
Often, the treatment cannot be randomized due to ethical or logistical reasons.
This has sparked great interest in developing statistical methodology to
estimate population-level effects from observational data.  Estimation of these
effects often requires accounting for putative confounding factors, that is,
factors related to whether a participant receives treatment and to the
participant's outcome. In many settings, researchers measure many potential
confounders, for example using administrative databases or genetic sequencing
technology. Due to the curse-of-dimensionality, common statistical estimation
techniques are not feasible for such high-dimensional data without restrictive
modeling assumptions. When these assumptions are violated, estimates of the
population-level effect of a treatment may have large bias.

This has sparked an interest in using adaptive estimation techniques from the
machine learning literature to control for high-dimensional confounders when
estimating treatment effects. However, facilitating proper inference (e.g.,
confidence intervals and hypothesis tests) about estimates of effects based on
adaptive estimators is challenging. In particular, standard causal inference
techniques, including both G-computation (GCOMP) estimators
[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
estimators [@horvitzthompson:1952:jasa, @robins:2000:epi], fail to yield valid
inference for common estimators of effects. More complex proposals have been
developed that *are capable* of utilizing adaptive methods to control for
confounding while still yielding valid inference. These techniques include
augmented inverse probability of treatment estimation (AIPTW)
[@robins:1994:jasa] and targeted minimum loss-based estimation (TMLE)
[@vdlrubin:2006:ijb]. Under assumptions, these estimators have limiting normal
distributions with estimable variance. The asymptotic variance estimates then
serve as the basis for constructing confidence intervals and hypothesis tests.

As an intermediate step, the AIPTW and TMLE estimators require estimation of
two key nuisance parameters: the probability of treatment as a function of
confounders (the propensity score, hereafter referred to as the *PS*) and the
average outcome as a function of confounders and treatment (the outcome
regression, *OR*). In addition to their desirable inferential properties, AIPTW
and TMLE estimators also enjoy the property of double robustness. That is, the
estimators are consistent for the population-level effect of interest if either
of the PS or OR is consistently estimated. This property gives the AIPTW and
TMLE estimators a natural appeal: inconsistency in the estimation of one
nuisance parameter may be mitigated by the consistent estimation of the other.
As such, these estimators have increased in popularity in recent years.
However, recent work has shown that the double robustness property does not
necessarily extend to inferential properties about these estimators when
adaptive estimators of the PS and OR are used
[@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
consistent estimation of *both* the PS *and* the OR. \citet{vdl:2014:ijb}
proposed new TMLE estimators have an asymptotic normal sampling distribution
with estimable variance under consistent estimation of *either* the PS *or* the
OR, paving the way for inference that is doubly-robust. \citet{benkeser:2017:biometrika}
proposed modifications of these estimators with similar robustness properties
and illustrated their practical performance.

The two proposed procedures that yield doubly-robust inference are quite
involved, notably involving iterative estimation of additional, complex
nuisance parameters. The `drtmle` package was motivated by the need for a
user-friendly tool to implement these methods. The package implements the TMLE
estimators of population-level effects proposed in \citet{vdl:2014:ijb} and
\citet{benkeser:2017:biometrika}, as well as several extensions including cross-validated
targeted minimum loss-based estimators (CVTMLE). Additionally, `drtmle`
implements an IPTW estimator that overcomes shortcomings of existing IPTW
implementations with respect to statistical inference. In particular, the
estimator allows an adaptive estimator of the PS to be used in construction of
the IPTW estimator, while yielding valid statistical inference about
population-level effects. Both the TMLE and IPTW estimators implemented in the
`drtmle` package are capable of estimating population-level effects for
discrete-valued treatments and the package  provides convenient utilities for
constructing confidence intervals and hypothesis tests about contrasts of the
mean outcome under different levels of treatment. This document explains some
of the key concepts underlying doubly-robust inference and  illustrates usage
of the `drtmle` package.

# Definition and estimators of the average treatment effect

## Parameters of interest

Suppose the observed data consist of $n$ independent and identically
distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
a vector of baseline covariates, $A\in\{0,1\}$ is a binary (for the sake of
simplicity) treatment, $Y$ is an outcome, and $P_0$ is the true data-generating
distribution. The parameters of interest are \begin{equation}
\psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
\mbox{for} \ a = 0,1 \ ,
\label{parameterDef}
\end{equation}
where $\bar{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
function of the covariates. The value $\psi_0(a)$ represents the marginal
(i.e., population-level)treatment-specific average outcome, hereafter referred
to as the *marginal mean* under treatment $A=a$. The marginal effect of the
treatment on the outcome can be assessed by comparing marginal means under
different treatment levels. For example, the average treatment effect is often
defined as $\psi_0 := \psi_0(1) - \psi_0(0)$. Under additional assumptions,
these parameters have richer interpretations as mean counterfactual outcomes
under the different treatments and contrasts between these means define causal
effects.

## G-computation estimators

An estimator of $\psi_0(a)$ may be motivated directly by (1). Suppose we have
available an estimate of the OR, $\bar{Q}_n$, and an estimate of the
distribution function of baseline covariates, $Q_{W,n}$. An estimator of
$\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
\bar{Q}_0(a,w) dQ_{W,n}(w).$ Often the empirical distribution function of
covariates is used so that this estimator can be equivalently written as
\begin{equation*}
\psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
\end{equation*}
This estimator is commonly referred to as the G-computation (GCOMP) estimator.
Inference for GCOMP estimators based on parametric OR estimators may be
facilitated through the nonparametric bootstrap. However, if adaptive
approaches are used to estimate the OR, inference may not be available without
further modification to the estimator.

## Inverse probability of treatment estimators

An alternative estimator of the treatment-specific population-level mean may be
obtained by noting that (1) can be equivalently written as
\begin{align}
E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
          W)\biggr\} \notag  \\
          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
          \label{iptwID}
\end{align}
We use $g_0(a \mid W) := Pr_0(A = a \mid W)$ to denote the propensity score.
Suppose we had available $g_n$, an estimate of the PS. Equation (2) motivates
the estimator \begin{equation*}
\psi_{n,IPTW}(a) := \frac{1}{n}\sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid
W_i)} Y_i \ ,
\end{equation*}
which is referred to as the inverse probability of treatment (IPTW) estimator.
Inference for IPTW estimators based on parametric PS estimators may be
facilitated through the nonparametric bootstrap. If adaptive approaches are
used to estimate the PS, inference may not be available without further
modification of the propensity score estimator, as we discuss below.

## Doubly-robust estimators

The GCOMP and IPTW estimators suffer from two important shortcomings. The first
is a lack of robustness. That is, validity of GCOMP estimators relies on
consistent estimation of the OR, while validity of the IPTW relies on
consistent estimation of the PS. In practice, we may not know which of the OR
or PS is simpler to consistently estimate and thus may not know which of the
two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
estimators lack a basis for statistical inference.

This motivates a new class of estimators that combine both the GCOMP and IPTW
estimators. One example is the augmented IPTW (AIPTW) estimator
\begin{equation*}
\psi_{n,AIPTW}(a) = \psi_{n,IPTW}(a) + \frac{1}{n} \sum\limits_{i=1}^n \biggl\{
1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \  \bar{Q}_n(a, W_i) \ ,
\end{equation*}

which adds an augmentation term to the IPTW estimator. Equivalently, the
estimator can be viewed as adding an augmentation to the GCOMP estimator
\begin{equation} \label{oneStepEst}
\psi_{n,AIPTW}(a) = \psi_{n,GCOMP}(a) + \frac{1}{n} \sum\limits_{i=1}^n
\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
\end{equation}
The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
estimators. The estimator exhibits double-robustness in that it is consistent
for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
is consistent for the true PS. Furthermore, if both the OR and PS are
consistently estimated and regularity conditions hold, one can establish a
limiting normal distribution for the AIPTW estimator. Simple estimators of the
variance of this limiting distribution can be derived that may be used to
construct Wald-style confidence intervals and hypothesis tests.

A more recent proposal to improve robustness and inferential properties of the
GCOMP and IPTW estimators utilizes targeted minimum loss-based estimation
(TMLE). TMLE is a general methodology that may be used to modify an estimator
of a regression function in such a way so as to ensure that the modified
estimator satisfies user-selected equations. In the present problem, one can
show that if an estimator of the OR $\bar{Q}_n^*$ satisfies \begin{equation}
\label{meanEIFZero}
  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
  \bar{Q}_n^*(a, W_i)\} = 0 \ ,
\end{equation}
then the GCOMP estimator based on $\bar{Q}_n^*$ will have the same asymptotic
properties as the AIPTW estimator. That is, the estimator will be doubly-robust
and, provided both the OR and PS are consistently estimated and regularity
conditions are satisfied, will have a normal limiting distribution. TMLE
provides a means of mapping an initial estimator of the OR into an estimator
that satisfies (4). In addition to possessing the desirable asymptotic
properties, TMLE estimators have been shown to outperform AIPTW estimators in
finite samples [@porter:2011:ijb]. We refer readers to \citet{tlbook:2011}
for more about TMLE.

# Doubly-robust inference

\citet{vdl:2014:ijb} demonstrated that the double robustness property of AIPTW
and TMLE estimators does not extend to their normal limiting distribution if
adaptive estimators of the OR and PS are used to construct the estimators.
However, in settings with high-dimensional and continuous-valued covariates,
adaptive estimation is likely necessary in order to obtain a consistent
estimate of either the OR or PS, and thus may be necessary to obtain minimally
biased estimates of the marginal means of interest. \citet{vdl:2014:ijb} derived
modified TMLE estimators that are doubly-robust not only with respect to
consistency, but also with respect to asymptotic normality. Furthermore, he
provided closed-form, doubly-robust variance estimators that may be used to
construct doubly-robust confidence intervals and hypothesis tests. \citet{benkeser:2017:biometrika} proposed simplifications of the \citet{vdl:2014:ijb} estimators and
demonstrated the practical performance of estimators via simulation.

The key to the theory underlying these results is finding estimators of the OR
and PS that simultaneously satisfy equation (4) in addition to two additional
equations. These equations involve additional regression parameters that we
refer to as the reduced outcome regression (R-OR) and the reduced propensity
score (R-PS), defined respectively as \begin{align*}
  \bar{Q}_{r,0n}(a,w) &:= E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \\
  g_{r,0n}(a \mid w) &:= Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
  g_n(W)=g_n(w)\right\} \ .
\end{align*}
In words, the R-OR is the regression of the residual from the outcome
regression onto the estimated PS amongst observations with $A = a$, while the
R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
note that a key feature of these reduced-dimension regressions is that they are
low-dimensional regardless of the dimension of $W$ and they do not depend on
the true OR nor PS. Thus, these regressions may be consistently estimated at
reasonably fast rates irrespective of the dimension of $W$ and irrespective of
whether the OR or PS are consistently estimated. Based on these
reduced-dimension regressions, \citet{vdl:2014:ijb} showed that if estimates of
the OR $\bar{Q}_n^*$, PS $g_n^*$, R-OR $\bar{Q}_{r,n}^*$, and R-PS $g_{r,n}^*$
satisfied \begin{align}
& \frac{1}{n} \sum\limits_{i=1}^n  \frac{I(A_i = a)}{g_n^*(a \mid W_i)} \{Y_i -
\bar{Q}_n^*(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
\{I(A_i = a) - g_n^*(a \mid W_i)\} = 0 \ , \ \mbox{and} \label{tmleAsolve} \\
&\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n}^*(a \mid W_i)}
\biggl\{\frac{g_{r,n}^*(a \mid W_i) - g_n^*(a \mid W_i)}{g_n^*(a \mid W_i)}
\biggr\} \{Y_i - \bar{Q}_n^*(a, W_i) \} = 0 \ , \label{tmleQsolve}
\end{align}
then the GCOMP estimator based on $\bar{Q}_n^*$ would be doubly-robust not only
with respect to consistency, but also with respect to its limiting
distribution. The author additionally proposed a TMLE algorithm to map initial
estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy these key
equations.

\citet{benkeser:2017:biometrika} demonstrated alternative equations that can be satisfied
to achieve this form of double-robustness. In particular, they formulated an
alternative version of the R-PS, \begin{align*}
g_{r,0,1}(a \mid w) := Pr_0\{A = a \mid \bar{Q}_n(W) = \bar{Q}_n(w) \} \ ,
\end{align*}

which we refer to as R-PS1. They also introduced an additional regression of a
weighted residual from the PS on the OR, \begin{align*}
g_{r,0,2}(a \mid w) := E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}{g_n(a \mid
W)} \ \bigg\vert \ \bar{Q}_n(W) = \bar{Q}_n(w) \biggr\} \ ,
\end{align*}

which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
both univariate regressions, while R-PS is a bivariate regression. \citet{benkeser:2017:biometrika} suggested that this may make R-PS1 and R-PS2 easier to estimate
consistently than R-PS. Further, they showed that if estimators or the OR, PS,
and R-OR satisfy equations (5)-(6), and if additionally, estimators of the OR,
R-PS-1 $g_{r,n,1}^*$, and R-PS-2 $g_{r,n,2}^*$ satisfy $$
\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^*(a \mid W_i)}
g_{r,n,2}^*(a \mid W_i) \ \{Y_i - \bar{Q}_n^*(a,W_i)\} = 0 \ ,
$$
then the GCOMP estimator based on this $\bar{Q}_n^*$ also enjoys a
doubly-robust limiting distribution. \citet{benkeser:2017:biometrika} provided a TMLE
algorithm that produced such estimates and provided closed-form, doubly-robust
variance estimates.

\textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
methods for constructing doubly-robust (with respect to consistency) estimators
of our parameter of interest. Theoretically, these two frameworks are closely
related and practically the two estimators are generally seen as competitors
for estimating marginal means. In equation (3), we were able to generate a
doubly-robust estimator by adding a term to the GCOMP estimator. This term is
exactly the left-hand-side of the key equation (4), the key equation that is
solved by the standard TMLE. This is no coincidence; this term is crucial in
studying the asymptotic properties of the GCOMP, AIPTW, and TMLE estimators
(for discussion, see \citet{benkeser:2017:biometrika} Section 2.2). Recall that a TMLE
estimator achieves doubly-robust consistency by satisfying equation (4), while
the AIPTW achieves doubly-robust consistency by combining the GCOMP estimator
with the left-hand-side of this term. Now, we have argued that a TMLE estimator
may achieve doubly-robust limiting behavior if it additionally satisfies
equations (6) and (7). It is thus sensible to wonder whether an AIPTW-style
estimator could be constructed by combining the left-hand-side of these two
additional equations with the AIPTW estimator. \citet{benkeser:2017:biometrika} showed
that, surprisingly, this AIPTW-style estimator does not achieve the same
doubly-robust properties as the TMLE estimator. Thus, while AIPTW and TMLE are
generally competitors for producing estimators doubly-robust with respect to
consistency, for doubly-robust inference, TMLE is preferred.

# Implementation of doubly-robust inference

The main function of the package is the eponymous `drtmle` function. This
function estimates the treatment-specific marginal mean for user-specified
levels of a discrete-valued treatment and computes a doubly-robust covariance
matrix for these estimates. The `drtmle` function implements either the
estimators of \citet{vdl:2014:ijb} (if `reduction = "bivariate"`) or the
improved estimators of \citet{benkeser:2017:biometrika} (`reduction = "univariate"`). The
package includes utility functions for combining these estimates into estimates
of average treatment effects, as well as other contrasts, as discussed below.

The main data arguments for `drtmle` are `W, A, Y`, which, as above, represent
the covariates, a discrete-valued treatment, and an outcome, respectively.
Missing data are allowed in `A` and `Y`; this is discussed further in a section
below. Beyond the data arguments, it is necessary to specify how to estimate
each of the OR, PS, R-OR, and R-PS (or R-PS1, R-PS2). The package provides
flexibility in how these regression parameters may be estimated. Once the user
has specified the data arguments and how to estimate each regression parameter,
the function proceeds as follows:
  1. estimate the OR via the internal `estimateQ` function;
  2. estimate the PS via the internal `estimateG` function;
  3. estimate reduced-dimension regression via `estimateQrn` and
     `estimategrn` functions;
  4. iteratively modify initial estimates of the OR and PS via the `fluctuateQ`
      and `fluctuateG` functions until equations (5)-(7) are approximately
      solved;
  5. compute the TMLE estimates and covariance estimates based on the final OR.
We refer interested readers to \citet{benkeser:2017:biometrika} for a theoretical
derivation of how the iterative modification of initial OR and PS is performed.

In addition to returning doubly-robust parameter and covariance estimates, the
`drtmle` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
modified AIPTW estimator discussed in the remark above. While doubly-robust
inference based on these estimators is not available, the estimators are
provided for comparison. In the following subsections, we look at various
options for making calls to `drtmle`.

## Estimating regressions 
```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
```
While the estimators computed by `drtmle` are consistent and asymptotically
normal under inconsistent estimation of either the OR or PS, it is nevertheless
advisable to strive for consistent estimation of *both* the OR and PS. If both
of the regression parameters are consistently estimated then the estimators
computed by `drtmle` achieve the nonparametric efficiency bound, resulting in
narrower confidence intervals and more powerful hypothesis tests. In order that
the estimates of the OR and PS have the greatest chance of consistency, the
`drtmle` function facilitates estimation using adaptive estimation techniques.
In fact, there are three ways to estimate each of the regression parameters:
generalized linear models, super learning, and a user-specified estimation
technique. We demonstrate each of these approaches in turn on a simple
simulated data set. The true parameter values of interest are $\psi_0(0)=$ `r
round(EY0, 2)` and $\psi_0(1)=$ `r round(EY1, 2)`.

```{r}
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 1, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
```

### Using generalized linear models to estimate regressions

Generalized linear models are perhaps the most popular means of estimating
regression functions. Because of this legacy, these estimators have been
included as an option for estimating the regression parameters. The user
specifies the right-hand-side of a regression formula as a character in the
`glm_Q`, `glm_g`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS, R-OR,
and R-PS. These regressions are fit by calling the `glm` function in base R.
Thus, the model specification should be able to be parsed as a valid formula,
as required by `glm`. Recall that the R-OR parameter involves a regression onto
the estimated propensity. Thus, the formula that is input via `glm_Qr` should
assume data with a single column named `"gn"`. Similarly, the R-PS1 and R-PS2
that are computed for the estimators of \citet{benkeser:2017:biometrika} (the default
estimators, computed whenever `reduction = "univariate"`) involve regressions
onto the estimated outcome regression. Thus, the formula `glm_gr` should assume
data with a single column named `"Qn"` if `reduction = "univariate"`. If
instead, the estimators of \citet{vdl:2014:ijb} are preferred, we can set the
option `reduction = "bivariate"` and input a `glm_gr` formula that assumes data
with two columns named `"Qn"` and `"gn"`.

The code below illustrates how generalized linear models can be used to
estimate the regression parameters. In this call to `drtmle`, we specify
`family = binomial()` so that logistic regression is used to estimate the OR.
We specify `stratify = FALSE` to indicate that we wish to fit a single OR to
estimate $\bar{Q}_0(A,W)$, as opposed to fitting two separate regressions to
estimate $\bar{Q}_0(1,W)$ and $\bar{Q}_0(0,W)$. If the former fitting is
specified, then the OR regression is fit using a data frame that contains `W`
and `A`. Thus, both `colnames(W)` and `"A"` may appear in `glm_Q`. However, if
`stratify = TRUE` then the OR model is fit separately in observations with $A =
1$ and $A = 0$. In this case, the `glm_Q` option may not include `"A"`. We
illustrate non-stratified regression for both the univariate and bivariate
reductions.

```{r}
glm_fit_uni <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE)
glm_fit_uni

glm_fit_biv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE,
                      reduction = "bivariate")
glm_fit_biv
```

### Using super learner to estimate regressions

```{r, echo = FALSE, message = FALSE}
library(SuperLearner)
```
Super learning is a loss-based ensemble machine learning technique
[@vdl:2007:statapp] that is a generalization of stacking and stacked regression
[@wolpert:1992:nnet]breiman:1996:ml}. The user specifies many potential
candidate regression estimators, referred to as a *library* of candidate
estimators, and the super learner algorithm estimates the convex combination of
regression estimators that minimizes $V$-fold cross-validated risk based on a
user-selected loss function. Oracle inequalities show that the super learner
estimate performs essentially as well as the unknown best convex combination of
the candidate estimators [@vdvaart:2006:statdec]. Thus, the methodology may be
seen as an asymptotically optimal way of performing estimator selection.
Practically, the super learner provides the best chance for obtaining a
consistent estimator for both the OR and PS, and thereby an efficient estimator
of the marginal means of interest.

Super learning is implemented in the R package `SuperLearner`
[@SuperLearnerPackage] and `drtmle` provides the option to utilize super
learning for estimation of the OR, PS, and reduced-dimension regressions via
the `SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle` either
the generalized linear model options for regression modeling or the super
learner options should be invoked. If both are called, the function will
default to the super learner option. The inputs to the super learner options
are passed to the `SL.library` argument of the `SuperLearner` function of the
`SuperLearner` package. We refer readers to the `SuperLearner` help file for
information on how to properly specify a super learner library. In the code
below, we illustrate a call to `drtmle` with a relatively simple library. For
the OR and PS, the library includes a main-terms generalized linear model (via
the function `SL.glm` from `SuperLearner`) and an unadjusted
generalized linear model (`SL.mean`). For the reduced dimension regressions we
use a super learner library with a main terms generalized linear model
(`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the `earth`
package [@earthPackage]).

```{r, warning = FALSE, message = FALSE, cache = TRUE}
set.seed(1234)
sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                 SL_g = c("SL.glm","SL.mean"),
                 SL_Q = c("SL.glm","SL.mean"),
                 SL_gr = c("SL.glm", "SL.earth"),
                 SL_Qr = c("SL.glm", "SL.earth"),
                 stratify = FALSE)
sl_fit
```

The `drtmle` package also allows users to pass in their own functions for
estimating regression parameters via the `SL_` options. These functions must be
formatted properly for compatibility with `drtmle`. The necessary formatting is
identical the formatting required for writing a `SuperLearner` wrapper (see
`SuperLearner::write.SL.template()`). The `drtmle` package includes `SL.npreg`,
which is an example of how a user can specify a function for estimating
regression parameters. This functions fits a kernel regression estimator using
the `np` package [@npPackage]. Below we show the source code for this function.

```{r}
SL.npreg
```

In the above code, we see that a user-specified function should have options
`Y, X, newX, family, obsWeights, ...`, along with other options specific to the
function. The option `Y` corresponds to the outcome of the regression and `X`
to the variables onto which the outcome is regressed. The option `newX` should
be of the same class and dimension as `X` and is meant to contain new values of
variables at which to evaluate the regression fit. The `family` and
`obsWeights` options are not strictly necessary (indeed, they are not used by
`SL.npreg`), they are useful to include if one wishes to utilize the function
as part of a super learner library. The `SL.npreg` function proceeds by
checking whether there is sufficient variation in `Y` to justify fitting the
kernel regression (as specified by option `rangeThresh`), and if so, fits a
kernel regression with cross-validated bandwidth selection via the `npregbw`
function. The output is then formatted in a specific way so that `drtmle` is
able to retrieve the appropriate objects from the fitted regression. In
particular the output should be a list with named objects `pred` and `fit`; the
former including predictions made from the fitted regression on `newX`, the
latter including any information that may be needed to predict new values based
on the fitted object. A class is assigned to the `fit` object, so that an S3
`predict` method may later be used to predict new values based on the fitted
regression object. The user must also specify this `predict` method. Use
`drtmle:::predict.SL.npreg` to view the simple `predict` method for `SL.npreg`.

In the code below, we demonstrate a call to `drtmle` with `SL.npreg` used to
estimate each nuisance parameter. Here, we also illustrate the use of `stratify
= TRUE`, which fits a separate kernel regression of $Y$ onto $W$ in
observations with $A = 1$ and $A = 0$.

```{r, cache = TRUE, message = FALSE}
npreg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = "SL.npreg", SL_Q = "SL.npreg",
                    SL_gr = "SL.npreg", SL_Qr = "SL.npreg",
                    stratify = TRUE)
npreg_fit
```

### Reducing dependence on random seed

Some users may find it unsettling that when applying the super learner in real data 
analysis is that the results are dependent on the random seed that is set. To remove 
this dependence, one can instead average results over repeated super learner runs; more 
repeats will lead to less dependence on the random seed. There are generally two approaches
that can be used: (i) averaging on the scale of the nuisance parameters and (ii) averaging
on the scale of the point estimates.

For (i), suppose that we have $n_{SL}$ estimated super 
learners for the OR, $\bar{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS 
$g_{n,j}, j = 1, \dots, n_{SL}$. We can take the average over these fits, \[
\bar{Q}_n = \frac{1}{n_{SL}} \sum_{j = 1}^{n_{SL}} \bar{Q}_{n,j}
\] 
as our estimate of the OR (similarly for the PS). These averaged estimates are then used to 
compute the AIPTW or TMLE, as described above.

For(ii), suppose for example that we have $n_{SL}$ TMLE (similarly, AIPTW) estimates 
$\psi_{n,j}(1)$ of $\psi_0(1)$, each obtained based on one (or more) super learner of the OR and/or PS.
We can take the average over these point estimates, \[
\psi_n(1) = \frac{1}{n_{SL}} \sum_{j=1}^{n_{SL}} \psi_{n,j}(1)
\]
as our estimate of $\psi_{0}(1)$.

These feature is implemented via the `n_SL` and `avg_over` options, where the former specifies the
number of super learners to repeat and the latter specifies the scale(s) to average over. For example, if 
`avg_over = "drtmle"` and `n_SL = 2`, then two point estimates are computed, each based on a single 
super learner, and averaged to obtain a final estimate. If instead `avg_over = "SL"` and `n_SL = 2`, 
then two super learners are fit and averaged before a final point estimate is obtained. Finally, if
`avg_over = c("drtmle", "SL")` and `n_SL = 2` then averaging on *both* scales is performed. That is,
the final estimate is the average of two point estimates, each obtained using OR and PS estimates that
are averaged over two super learners.

Not surprisingly, averaging over multiple super learners can add considerable computational expense 
to calls to `drtmle`. The package is currently parallelized at the level of individual calls to 
`SuperLearner`. With parallelization over enough cores, repeated super 
learning should not add too much to the computational burden relative to a single 
call to `SuperLearner`.

### Outcome-adaptive propensity score

If the outcome-adaptive PS is desired, the user may specify this via 
the `adapt_g` option. In this case, the additional targeting steps needed for doubly
robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed above,
this estimator is *not* doubly robust. Note that when the `adapt-g = TRUE` the data
used for the PS estimation will be a `data.frame` with column names
`QaW` where `a` takes all the values specified in `a_0`. For example if `a_0 = c(0,1)`,
then the PS estimation data frame will have two columns called `Q0W`
and `Q1W`. This may be relevant for users writing their own `SuperLearner` wrappers
or those who make use of the `glm_g` setting, as illustrated below.

```{r, message = FALSE}
# outcome adaptive propensity score fit using glms
adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
                     adapt_g = TRUE)

# outcome adaptive propensity score fit using super learner
adaptg_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_Q = c("SL.glm", "SL.mean"), 
                     SL_g = c("SL.glm", "SL.mean"),
                     adapt_g = TRUE)
```

### Arbitrary user-specified regressions

The `drtmle` package also allows users to pass in their own estimated OR and PS
via the `Qn` and `gn` options, respectively. 
If `Qn` is specified by the user, `drtmle` will ignore the nuisance parameter 
estimation routines specified by `SL_Q` and `glm_Q` -- similarly for `gn` and
`SL_g`, `glm_g`. As noted in the package documentation, there is a specific format 
necessary for inputting `Qn` and `gn`. For `Qn`, the entries in the list should 
correspond to the OR evaluated at A and the observed values of `W`, 
with order determined by the input to `a_0`. For example, if `a_0 = c(0, 1)` then 
`Qn[[1]]` should be OR at $A = 0$ and `Qn[[2]]` should be outcome 
regression at $A = 1$. The example below illustrates this concept.

```{r, cache = TRUE, message = FALSE}
# fit a GLM for outcome regression outside of drtmle
out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())

# generate Qn list
Qn10 <- list(
  # first entry is predicted values setting A = 1
  predict(out_glm_fit, newdata = data.frame(A = 1, W), type = "response"),
  # second entry is predicted values setting A = 0
  predict(out_glm_fit, newdata = data.frame(A = 0, W), type = "response")
)

# pass this list to drtmle to avoid internal estimation of 
# outcome regression (note propensity score and reduced-dimension
# regressions are still estimated internally)
out_fit1 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn", Qn = Qn10,
                  # crucial to set a_0 to match Qn's construction!
                  a_0 = c(1,0))
out_fit1

# to be completely thorough let's re-order Qn10 and re-run
Qn01 <- list(Qn10[[2]], Qn10[[1]])

out_fit2 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn", 
                  # use re-ordered Qn list
                  Qn = Qn01,
                  # a_0 has to be reordered to match Qn01!
                  a_0 = c(0,1))
# should be the same as out_fit1, but re-ordered
out_fit2
```

Similarly, for `gn` we need to pass in properly ordered lists. For example, 
if `a_0 = c(0, 1)` then `gn[[1]]` should be propensity for receiving $A = 0$ 
and `gn[[2]]` should be the propensity for receiving $A = 1$. The example below 
illustrates this concept.

```{r, cache = TRUE, message = FALSE}
# fit a GLM for propensity score outside of drtmle
out_glm_fit <- glm(A ~ . , data = data.frame(W), family = binomial())

# get P(A = 1 | W) by calling predict
gn1W <- predict(out_glm_fit, newdata = data.frame(W), type = "response")

# generate gn list
gn10 <- list(
  # first entry is P(A = 1 | W)
  gn1W,
  # second entry is P(A = 0 | W) = 1 - P(A = 1 | W)
  1 - gn1W
)

# pass this list to drtmle to avoid internal estimation of 
# propensity score (note reduced-dimension regressions are 
# still estimated internally)
out_fit3 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                   glm_Qr = "gn", glm_gr = "Qn", 
                   Qn = Qn10, gn = gn10, 
                   # crucial to set a_0 to match Qn and gn's construction!
                   a_0 = c(1,0))
out_fit3

# to be completely thorough let's re-order gn10 and re-run
gn01 <- list(gn10[[2]], gn10[[1]])

out_fit4 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                   glm_Qr = "gn", glm_gr = "Qn", 
                   # use re-ordered Qn/gn list
                   Qn = Qn01, gn = gn01, 
                   # a_0 has to be reordered to match Qn01!
                   a_0 = c(0,1))
# should be the same as out_fit3, but re-ordered
out_fit4
```

This flexibility allows users to use any regression technique to externally
obtain initial estimates of nuisance parameters. The key step is making sure
that the ordering of `Qn` and `gn` is consistent with `a_0`.

## Standard error estimation

This section contains some more advanced mathematical material and may be omitted by
users interested only in implementation. For simplicity, we focus this discussion on the 
standard TMLE/AIPTW estimators. The ideas generalize immediately to the DRTMLE estimator, 
but the formulas are more complex.

Reported standard error estimates are based on an estimate the asymptotic variance of 
the estimators. This variance can be characterized by the estimators influence function.
For example, let $\psi_{n}(1)$ be an AIPTW estimate of $\psi_0(1)$. Under regularity 
conditions, \[
  \psi_{n}(1) - \psi_0(1) = \frac{1}{n} \sum_{i=1}^n D^*(\bar{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
\]
where \[
D^*(\bar{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}{g_0(1 \mid W_i)} \{ Y_i - \bar{Q}_0(1, W_i)\} + \bar{Q}_0(1, W_i) - \int \bar{Q}_0(1, w) dQ_{W,0}(w) \ .
\]
Thus, by the central limit theorem, $n^{1/2} \psi_{n}(1)$ converges in distribution to a Normal
distribution with mean $\psi_0(1)$ and variance \[
  \sigma^2_0 = E_0( [ D^*(\bar{Q}_0, g_0, Q_{W,0})(O) - E_0\{D^*(\bar{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
\]
An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical variance of $D_i = D^*(\bar{Q}_n, g_n, Q_{W,n})(O_i)$, \[
  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{ D_i - \frac{1}{n} \sum_{j=1}^n D_j \right\}^2 \ . 
\]
Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n / n^{1/2}$. This idea generalizes naturally to estimating the asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.

### Standard errors for TMLE

For TMLE-based estimators, it is standard practice to construct $D_i$ above using the *targeted* 
version of the OR (denoted by $\bar{Q}_n^*$ above). Recall that such estimates are obtained by modifying
an initial OR estimate in such a way as to ensure a particular (set of) equation(s) is (are) solved.
However, simulations have shown that better standard error estimates for TMLE may be achieved by instead
using the *initial* OR estimate in the computation of standard errors, as described in the previous
subsection. The `drtmle` function includes the `targeted_se` option to this end.

```{r, show_targeted_se, cache = TRUE}
glm_fit_uni_nontargeted_se <- 
  drtmle(W = W, A = A, Y = Y, family = binomial(),
         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
         glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE, 
         targeted_se = FALSE)
# compare to original call        
list(
  glm_fit_uni$drtmle$cov, # based on targeted OR
  glm_fit_uni_nontargeted_se$drtmle$cov # untargeted OR
)
```

### Cross-validated standard errors

In simulations, we often find that when employing machine learning to estimate the OR and/or PS, 
the estimated standard errors are *too small* in small to moderate-sized samples, leading to 
under-coverage of confidence intervals and possibly too-large type I errors of hypothesis tests that
are based on these standard error estimates. This should not be too surprising when studying the form
of the standard error estimates. In particular, they will include a term that looks like 
a (weighted) mean squared-error (MSE) of the OR. Because machine learning algorithms often set out 
to minimize empirical MSE as part of their training, we often find that the empirical MSE is biased 
too small relative to the true MSE. Thus, in order to more accurately estimate MSE of a machine learning
algorithm, we typically rely on cross-validation. The same trick can be used here to obtain a more
conservative standard error estimate.

Suppose we use $V$-fold cross-validation to obtained $V$ training-sample-specific estimates of the OR,
 $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v}, v = 1, \dots, V$. For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the validation fold. That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of $Y_i$. We define
 $g_{n,i}^{cv}$ similarly. Using these estimates, we can define \[
 D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} + \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1) 
 \]
and compute an estimate of $\sigma^2_0$ using \[
  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} - \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ . 
\]

Cross-validated standard errors can be obtained by setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1.

```{r, fit_w_full_cv, warning = FALSE, cache = TRUE}
glm_fit_uni_cv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         se_cv = "full", se_cvFolds = 2, stratify = FALSE,
                         targeted_se = FALSE) # needed for this to work
# compare variance to previously obtained
list(glm_fit_uni$drtmle$cov, glm_fit_uni_cv$drtmle$cov)
```

### Partially cross-validated standard errors

Obtaining fully cross-validated standard errors can be fairly time intensive when super learner is
used to estimate the OR and PS. In `drtmle` we give an option to compute standard error estimates that
are *partially cross-validated* and do not require any additional fitting. Here we describe briefly 
how these estimates are obtained and show calls to `drtmle` to obtain these estimates.

Suppose we have $K$ candidate regressions in our super learner library for both the OR and PS 
(in practice, the number could be different for the OR as for the PS). Recall that a super learner
estimate of for example $\bar{Q}_0(a, w)$ is obtained as a weighted combination of estimates 
from each candidate regression. Let $\bar{Q}_{n,k}$ denote the estimate of $\bar{Q}_{0}$ obtained 
by training algorithm $k$ using the full data set. The super learner estimate of $\bar{Q}_{0}(a,w)$ is \[
  \bar{Q}_{n,SL}(a, w) = \sum_{k = 1}^K \alpha_{n,k} \bar{Q}_{n,k}(a, w) \ , 
\]
where $\alpha_n = (\alpha_{n,1}, \dots, \alpha_{n,K})$ is a weight vector that is selected by minimizing
a cross-validated risk criteria. That is, each of the $K$ algorithms is trained according to a $V$-fold
cross-validation scheme, resulting in $V + 1$ fits of each algorithm (one in each of $V$ training folds
plus one obtained by fitting the algorithm using the full data). Let $\bar{Q}_{n,k,v}$ denote the $k$-th
algorithm trained using the $v$-th training fold. For $i = 1,\dots,n$, let $\bar{Q}_{n,k,i}^{cv}(1)$ 
denote the estimate of $\bar{Q}_0(1, W_i)$ implied by the fitted algorithm $\bar{Q}_{n,k,v}$, which 
was obtained when observation $i$ was in the validation sample. That is, for observations with
$A_i = 1$, $\bar{Q}_{n,k,i}^{cv}(1)$ is an out of sample prediction of $Y_i$ based on the
$k$-th algorithm. Now, we can construct a *partially cross-validated* super learner prediction as follows. For $i = 1, \dots, n$ let \[
  \bar{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k} \bar{Q}_{n,k,i}^{cv}(1)
\]
be the estimate obtained by ensembling the training-set-specific algorithm fits based on $\alpha_n$. 
We can use these estimates to produce our *partially cross-validated* standard error estimate as follows. We define \[
 D_i^{pcv} = \frac{I(A_i = 1)}{g_{n,SL,i}^{cv}} \{ Y_i - \bar{Q}_{n,SL,i}^{cv}(1)\} + \bar{Q}_{n,SL,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,SL,i}^{cv}(1) 
 \]
and compute an estimate of $\sigma^2_0$ using \[
  \sigma^2_{n,pcv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{pcv} - \frac{1}{n} \sum_{j=1}^n D_j^{pcv} \right\}^2 \ . 
\]

Note that is not a proper cross-validated estimate, since observation $i$ is used to compute the 
weight vector $\alpha_n$. However, such estimates can be obtained based on the output of a single 
round of super learning, thereby providing improved standard error estimates at minimal additional
computational cost.

Here, we demonstrate how to obtain such estimates using `drtmle`.

```{r, demo_pcv, cache = TRUE, warning = FALSE}
set.seed(1234)
sl_fit_pcv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                     SL_g = c("SL.glm","SL.mean"),
                     SL_Q = c("SL.glm","SL.mean"),
                     SL_gr = c("SL.glm", "SL.earth"),
                     SL_Qr = c("SL.glm", "SL.earth"),
                     stratify = FALSE,
                     se_cv = "partial")
# compare estimates
# pt estimates should be same
# variance should be larger for pcv
list(
  sl_fit, # no cv
  sl_fit_pcv # partial cv
)
```

## Confidence intervals

The `drtmle` package contains methods that compute doubly-robust confidence
intervals. The `ci` method computes confidence intervals about the estimated
marginal means in a fitted `drtmle` object, in addition to contrasts between
those means. By default the method gives confidence intervals about each mean.

```{r}
ci(npreg_fit)
```

Alternatively the `contrast` option allows users to specify a linear
combination of marginal means, which can be used to estimate the average
treatment effect as we show below.

```{r}
ci(npreg_fit, contrast = c(1,-1))
```

The `contrast` option can also be input as a list of functions in order to
compute a confidence interval for a user-specified contrast. For example, we
might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
constructing Wald style confidence intervals for ratios, it is common to
compute the interval on the log scale and back-transform. That is, the
confidence interval is of the form
$$
\mbox{exp}\biggl[\mbox{log}\biggl\{ \frac{\psi_0(1)}{\psi_0(0)} \biggr\} \pm
z_{1-\alpha/2} \sigma^{\text{log}}_n \biggr] \ ,
$$
where $\sigma^{\text{log}}_n$ is the estimated standard error on the log-scale.
By the delta method, an estimate of this standard error is given by $$
\sigma^{\text{log}}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
$$
where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
`drtmle` and $\nabla g(\psi)$ is the gradient of
$\mbox{log}\{\psi(1)/\psi(0)\}$. To obtain this confidence interval, we may use
the following code.

```{r}
riskRatio <- list(f = function(eff){ log(eff) },
                  f_inv = function(eff){ exp(eff) },
                  h = function(est){ est[1]/est[2] },
                  fh_grad =  function(est){ c(1/est[1],-1/est[2]) })
ci(npreg_fit, contrast = riskRatio)
```

More generally this method of inputting the `contrast` option to the `ci`
method can be used to compute confidence intervals of the form
$$
f^{-1}\bigl\{ f(h(\psi_n)) \pm z_{1-\alpha/2} \nabla f(h(\psi_n))^T \Sigma_n
\nabla f(h(\psi_n)) \bigr\} \ ,
$$
where $f$ (`contrast$f`) is the transformation of the confidence interval,
$f^{-1}$ (`contrast$f_inv`) is the back-transformation of the interval, $h$
(`contrast$h`) is the contrast of marginal means, and $\nabla f(h(\psi_n))$
(`contrast$fh_grad`) defines the gradient of the transformed contrast at the
estimated marginal means.

We note that this method of computing confidence intervals can also be used to
obtain a transformed confidence interval for a particular marginal mean. For
example, in our running example $Y$ is binary. Thus, we might wish to enforce
that our confidence interval be computed on a scale which ensures that the
confidence limits are bounded between zero and one. To that end, we can
consider an interval of the form construct an interval on the logit scale and
back-transform,
$$
\mbox{expit}\biggl[ \mbox{log}\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
\pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
$$
This confidence interval may be implemented as follows.

```{r}
logitMean <- list(f = function(eff){ qlogis(eff) },
                  f_inv = function(eff){ plogis(eff) },
                  h = function(est){ est[1] },
                  fh_grad = function(est){ c(1/(est[1] - est[1]^2), 0) })
ci(npreg_fit, contrast = logitMean)
```

## Hypothesis tests

Doubly-robust Wald-style hypothesis tests may be implemented using the
`wald_test` method. As with confidence intervals, there are three options for
testing. First, we may test the null hypothesis that the marginal means equal a
fixed value (or values). Below, we test the null hypothesis that $\psi_0(1) =
0.5$ and that $\psi_0(0) = 0.6$ by inputting a vector to the `null` option.

```{r}
wald_test(npreg_fit, null = c(0.5, 0.6))
```

As with the `ci` method, the `wald_test` method allows users to test linear
combinations of marginal means by inputting a vector of ones and negative ones
in the `contrast` option. We can use this to test the null hypothesis of no
average treatment effect or to test the null hypothesis that the average
treatment effect equals 0.05 (by specifying the `null` option)

```{r}
wald_test(npreg_fit, contrast = c(1, -1))
wald_test(npreg_fit, contrast = c(1, -1), null = 0.05)
```

The `wald_test` method also allows for user-specified contrasts to be tested
using syntax similar to the `ci` method. The only difference syntactically is
that it is not strictly necessary to specify `f_inv`, as the hypothesis test is
performed on the transformed scale. That is, we can generally test the null
hypothesis that $f(h(\psi_0))$ equals $f_0$ (the function $f$ applied to the
value passed to `null`) using the Wald statistic $$
Z_n := \frac{\{f(h(\psi_n)) - f_0\}}{\nabla f(h(\psi_n))^T \Sigma_n \nabla
f(h(\psi_n))} \ .
$$
We can use the `riskRatio` contrast defined previously to test the null
hypothesis that $\psi_0(1)/\psi_0(0) = 1$.

```{r}
wald_test(npreg_fit, contrast = riskRatio, null = 1)
```

## Missing data

The `drtmle` function supports missing data in `A` and `Y`. Such missing data
are common in practice and, if ignored, can bias estimates of the marginal
means of interest. The estimators previously discussed can be modified to allow
this missingness to depend on covariates. Let $\Delta_A$ and $\Delta_Y$ be
indicators that $A$ and $Y$ are observed, respectively. We can redefine the OR
to be $\bar{Q}_0(a,w) := E_0(\Delta_Y Y \mid \Delta_A A = a, \Delta_A = 1,
\Delta_Y = 1, W = w)$ and similarly redefine the PS to be \begin{align}
g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
\notag \\
&= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
\label{modPS} \\
&\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w) \ \notag.
\end{align}
We introduce missing values to `A` and `Y` in our running example.

```{r}
DeltaA <- rbinom(n, 1, plogis(2 + W$W1))
DeltaY <- rbinom(n, 1, plogis(2 + W$W2 + A))
A[DeltaA == 0] <- NA
Y[DeltaY == 0] <- NA
```

The only modification of the call to `drtmle` is in how the PS estimation is
performed. The options `glm_g` and `SL_g` are still used to fit generalized
linear models or a super learner, respectively. However, the code allows for
separate regression fits for each of the three components of the PS in (8). To
perform separate generalized linear model fits for each of the three
components, `glm_g` should be a named list with entries `"DeltaA"`, `"A"`, and
`"DeltaY"`. Respectively these should provide a regression formula for the
regression of $\Delta_A$ on $W$, $A$ on $W$ in observations with $\Delta_A =
1$, and a regression of $\Delta_Y$ on $W$ and $A$ in observations with
$\Delta_A = \Delta_Y = 1$ (if `stratify = FALSE`) or a regression of $\Delta_Y$
on $W$ to be fit in observations with $\Delta_A = \Delta_Y = 1$ and $A = a$ for
each $a$ specified in option `a_0` (if `stratify = TRUE`). If only a single
formula is passed to `glm_g`, it is used for each of the three regressions. We
illustrate a call to `drtmle` with missing data and generalized linear model
fits below.

```{r, cache = TRUE}
glm_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                       glm_g = list(DeltaA = "W1 + W2", A = "W1 + W2",
                                    DeltaY = "W1 + W2 + A"),
                       glm_Q = "W1 + W2*A", glm_Qr = "gn",
                       glm_gr = "Qn", family = binomial())
glm_fit_wNAs
```

It is possible to mix generalized linear models and super learning when fitting
each of the components of the PS. In the below example we use the wrapper
function `SL.glm`, which fits a main terms logistic regression, to fit the
regression of $\Delta_A$ on $W$, use `SL.npreg` to fit the regression of $A$ on
$W$, and use a super learner with library `SL.glm`, `SL.mean`, and `SL.earth` to
fit the regression of $\Delta_Y$ on $W$ and $A$.

```{r, cache = TRUE}
mixed_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         SL_g = list(DeltaA = "SL.glm", A = "SL.npreg",
                         DeltaY = c("SL.glm", "SL.mean", "SL.earth")),
                         glm_Q = "W1 + W2*A", glm_Qr = "gn",
                         glm_gr = "Qn", family = binomial())
mixed_fit_wNAs
```

We can also fit the PS regressions externally and pass in via `gn`, though this
process is slightly more complicated than when no missing data are present. 
The basic idea is to fit one regression for (i) the indicator of missing $A$, 
(ii) $A$ itself, and (iii) the indicator of missing the outcome. The three PS's 
are then multiplied together and an appropriately ordered list is constructed
as before. We illustrate with the following simple example.

```{r, cache = TRUE}
# first regress indicator of missing A on W
fit_DeltaA <- glm(DeltaA ~ . , data = W, family = binomial())

# get estimated propensity for observing A
ps_DeltaA <- predict(fit_DeltaA, type = "response")

# now regress A on W | DeltaA = 1
fit_A <- glm(A[DeltaA == 1] ~ . , data = W[DeltaA == 1, , drop = FALSE], 
             family = binomial())

# get estimated propensity for receiving A = 1
ps_A1 <- predict(fit_A, newdata = W, type = "response")
# propensity for receiving A = 0
ps_A0 <- 1 - ps_A1

# now regress DeltaY on A + W | DeltaA = 1. Here we are pooling over 
# values of A (i.e., this is what drtmle does if stratify = FALSE). 
# We could also fit two regressions, one of DeltaY ~ W | DeltaA = 1, A = 1
# and one of DeltaY ~ W | DeltaA = 1, A = 0 (this is what drtmle does if
# stratify = TRUE). 
fit_DeltaY <- glm(DeltaY[DeltaA == 1] ~ . , 
                  data = data.frame(A = A, W)[DeltaA == 1, ], 
                  family = binomial())

# get estimated propensity for observing outcome if A = 1
ps_DeltaY_A1 <- predict(fit_DeltaY, newdata = data.frame(A = 1, W), 
                        type = "response")

# get estimated propensity for observing outcome if A = 0
ps_DeltaY_A0 <- predict(fit_DeltaY, newdata = data.frame(A = 0, W), 
                        type = "response")

# now combine all results into a single propensity score 
gn <- list(
  # propensity for DeltaA = 1, A = 1, DeltaY = 1
  ps_DeltaA * ps_A1 * ps_DeltaY_A1,
  # propensity for DeltaA = 1, A = 0, DeltaY = 1
  ps_DeltaA * ps_A0 * ps_DeltaY_A0
)

# pass in this gn to drtmle
out_fit_ps <-  drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                      # make sure a_0/gn are ordered properly!
                      gn = gn, a_0 = c(1, 0),
                      glm_Q = "W1 + W2*A", glm_Qr = "gn",
                      glm_gr = "Qn", family = binomial())
out_fit_ps
```

## Multi-level treatments

```{r, echo = FALSE, message = FALSE}
library(drtmle)
set.seed(1234)
N <- 1e6
W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
Y2 <- rbinom(N, 1, plogis(W$W1 + 2*W$W2))
Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
Y0 <- rbinom(N, 1, plogis(W$W1))
EY1 <- mean(Y1)
EY0 <- mean(Y0)
EY2 <- mean(Y2)
```

So far in our examples, we have only considered binary treatments. However,
`drtmle` is equipped to handle treatments with an arbitrary number of discrete
values. Suppose that $A$ assumes values in a set $\mathcal{A}$, the PS
regression estimation is modified to ensure that
$$
  \sum\limits_{a \in \mathcal{A}} g_n(a \mid w) = 1 \ \mbox{for all $w$} \ .
$$
If this condition is true, we say that the estimates of the PS are compatible.
Many regression methodologies that have been developed work well with binary
outcomes, but have not been extended to multi-level outcomes. To ensure that
users of `drtmle` have access to the large number of binary regression
techniques when estimating the PS, we have taken an alternative approach to
estimation of the propensity for each level of the treatment. Specifically,
rather than fitting a single multinomial-type regression, we fit a series of
binomial regressions. As a concrete example, consider the case of no missing
data and where $\mathcal{A} = \{0,1,2\}$. We can ensure compatible estimates of
the PS by generating an estimate $g_n(0 \mid w)$ of $Pr_0(A = 0 \mid W = w)$
and $\tilde{g}_n(1 \mid w)$ of $Pr_0(A = 1 \mid A > 0, W = w)$. The latter
estimate may be generated by regression $I(A = 1)$ on $W$ in observations with
$A > 0$. Because the outcome is binary, this regression may be estimated using
any technique suitable for binary outcomes. By simple rules of conditional
probability, we know that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
Pr_0(A = 0 \mid W)\}$ and thus an estimate of $g_0(1 \mid w)$ can be computed
as $g_n(1 \mid w) = \tilde{g}_n(1 \mid w) \{1 - g_n(0 \mid w)\}$. Finally, we
let $g_n(2 \mid w) = 1 - g_n(0 \mid w) - g_n(1 \mid w)$, which ensures
compatibility of the estimates for each level of covariates. This approach
generalizes to an arbitrary number of treatment levels. Below, we provide an
illustration using simulated data with a treatment that has three levels. The
true parameter values in this simulation are $\psi_0(0)=$ `r round(EY0,2)`,
$\psi_0(1)=$ `r round(EY1, 2)`, and $\psi_0(2)=$ `r round(EY2,2)`.

```{r, cache = TRUE, message = FALSE}
set.seed(1234)
n <- 200
W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
A <- rbinom(n, 2, plogis(W$W1 + W$W2))
Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
```

The multi-level PS can still be estimated using any of the three techniques
discussed previously. Here, we illustrate with simple generalized linear
models.

```{r, cache = TRUE, message = FALSE}
glm_fit_multiA <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
                         glm_gr = "Qn", glm_Qr = "gn",
                         family = binomial(), a_0 = c(0,1,2))
glm_fit_multiA
```

Above, we used the option `a_0` to specify the levels of the treatment at which
we were interested in estimating marginal means. There is no need for `a_0` to
contain all unique values of `A`. For example, certain levels of the treatment
may not be of scientific interest, but nevertheless we would like to use these
data to help estimate the OR and PS (by setting `stratify = FALSE`).

The confidence interval and testing procedures extend to multi-level treatments
with minor modifications. We can still obtain a confidence interval and
hypothesis test for each of the marginal means by not specifying a `contrast`.

```{r, cache = TRUE}
ci(glm_fit_multiA)
wald_test(glm_fit_multiA, null = c(0.4, 0.5, 0.6))
```

Alternatively, we can specify a `contrast` to estimate confidence intervals
about the average treatment effect of $A=1$ vs. $A=0$. Calls to `wald_test` can
also be made.

```{r, cache = TRUE}
ci(glm_fit_multiA, contrast = c(-1, 1, 0))
```

Similarly, we can compute confidence intervals comparing the average treatment
effect of $A = 2$ vs. $A = 0$.

```{r, cache = TRUE}
ci(glm_fit_multiA, contrast = c(-1, 0, 1))
```

The user-specified contrasts are also available. For example, we can modify the
`riskRatio` object we used previously for a two-level treatment to compute the
risk ratio comparing $A=1$ to $A=0$ and comparing $A = 2$ to $A = 0$.

```{r, cache = TRUE}
riskRatio_1v0 <- list(f = function(eff){ log(eff) },
                      f_inv = function(eff){ exp(eff) },
                      h = function(est){ est[2]/est[1] },
                      fh_grad =  function(est){ c(1/est[2], -1/est[1], 0) })
ci(glm_fit_multiA, contrast = riskRatio_1v0)
```

Similarly, we can compute confidence intervals for the risk ratio comparing the
marginal mean for $A=2$ to $A=1$.

```{r, cache = TRUE}
riskRatio_2v0 <- list(f = function(eff){ log(eff) },
                      f_inv = function(eff){ exp(eff) },
                      h = function(est){ est[3]/est[1] },
                      fh_grad =  function(est){ c(0, -1/est[1], 1/est[3]) })
ci(glm_fit_multiA, contrast = riskRatio_2v0)
```

## Cross-validated regression estimates

When considering adaptive estimation techniques, one runs the risk of
overfitting the initial regressions. While the cross validation used by super
learning attempts to guard against overfitting, there are no guarantees in
finite-samples. To definitively avoid such overfitting, one can use
cross-validated estimates of the regression parameters. Theoretically, the use
of cross-validated regression weakens the assumptions necessary to achieve
asymptotic normality [@zheng:2010:working]. This is true of the regular TMLE and
AIPTW, as well as of the TMLE estimator with doubly-robust inference. The
`drtmle` function implements cross-validated estimation of the regression
parameters through the `cvFolds` option, as shown in the code below, where we
continue with the multi-level treatment example.

```{r, cache = TRUE}
cv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
                    SL_g = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_Q = c("SL.glm", "SL.glm.interaction", "SL.earth"),
                    SL_gr = c("SL.glm", "SL.mean"),
                    SL_Qr = c("SL.glm", "SL.mean"),
                    stratify = FALSE, cvFolds = 2, a_0 = c(0, 1, 2))
cv_sl_fit
```

Cross-validation may significantly increase the computation time necessary for
running `drtmle`. Parallelized regression fitting is available by setting the
option `use_future = TRUE`. Since, internally, all computations are carried out
using futures (with the API provided by the "future" R package), setting this
option amounts only to parallelizing with futures [@futurePackage] via
`future.apply::lapply`. Selecting this option results in parallelization of the
estimation of the OR, PS, and reduced-dimension regressions.

If no adjustments are made prior to invoking the `drtmle` function, futures are
computed sequentially. It is up to the user to specify appropriate system-specific 
future back-ends that may better suit their problem (e.g., using the `future.batchtools` 
package [@futurebatchtoolsPackage]). Below we
demonstrate registration using an ad-hoc cluster on the local system. The choice
of back-end is flexible, and submission of the job to a wide variety
of schedulers (e.g., SLURM, TORQUE, etc.) is also permitted. Interested readers
are invited to consult the documentation of the `future.batchtools` package for
further information.

```{r, cache = TRUE, message = FALSE, eval = FALSE}
## Commented out to avoid build errors
# library(future.batchtools)
# library(parallel)
# cl <- makeCluster(2, type = "SOCK")
# plan(cluster, workers = cl)
# clusterEvalQ(cl, library("SuperLearner"))
# pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
#                      SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
#                      SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
#                      SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
#                      SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
#                      stratify = FALSE, a_0 = c(0,1,2),
#                      cvFolds = 2, use_future = TRUE)
```

# Additional methods included in drtmle

## Inference for IPTW using adaptive PS estimators

Doubly-robust estimators of marginal means are appealing in their robustness and
efficiency properties. However, researchers may prefer the IPTW estimator for
its intuitive derivation and implementation. Heretofore, inference for IPTW
estimators precluded the use of adaptive estimators of the PS. However, \citet{vdl:2014:ijb} proposed modified IPTW estimators that allow for adaptive PS
estimation. Similarly as above, this estimator is based on an additional
regression parameter, $\tilde{Q}_{r,0n}(a,w) := E_0\{Y \mid A = a, g_n(W) =
g_n(w)\}$, that is the regression of the outcome $Y$ on the estimated propensity
amongst observations with $A=a$. The author showed that if $g_n^*(a \mid w)$, an
estimator of the propensity for treatment $A=a$, satisfied that
$$ \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,$$
then the IPTW estimator based on $g_n^*$ has an asymptotic normal distribution
with an estimable variance. In fact, \citet{vdl:2014:ijb} proposed a TMLE
algorithm that mapped an initial PS estimate into an estimate that satisfies
this key equation and variance estimators that can be used to generate
Wald-style confidence intervals and hypothesis tests.

An estimator with equivalent asymptotic properties based on a PS estimate,
$g_n(a \mid w)$, can be computed as
$$
  \psi_{n,IPTW}(a) = \psi_{n,IPTW}(a) - \frac{1}{n} \sum\limits_{i=1}^n
  \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)} \ \{ I(A_i = a) - g_n(a
  \mid W = W_i)\} \ .
$$
Both this estimator and the TMLE-based IPTW estimator above are implemented in
`drtmle`.

### Implementation of inference for adaptive IPTW

The IPTW estimators based on adaptive estimation of the PS are implemented in
the `adaptive_iptw` function. The `adaptive_iptw` function shares many options
with `drtmle`. However, as the IPTW estimator does not rely on an estimate of
the OR, these options are omitted. We continue with the multilevel treatment
example.

```{r, cache = TRUE}
npreg_iptw_multiA <- adaptive_iptw(W = W, A = A, Y = Y, stratify = FALSE,
                                   SL_g = "SL.npreg", SL_Qr = "SL.npreg",
                                   family = binomial(), a_0 = c(0, 1, 2))
npreg_iptw_multiA
```

The `"adaptive_iptw"` class contains identical methods for computing confidence
intervals and Wald tests as the `"drtmle"` class and we refer readers back to
confidence intervals and hypothesis tests section for reminders on the various
options for these tests. By default, these methods are implemented for the
`iptw_tmle` estimator of \citet{vdl:2014:ijb}. We expect that this estimator
will have improved finite-sample behavior relative to `iptw_os` (the
alternative estimator described in the previous subsection), though further
study is needed.

As with `drtmle`, the `adaptive_iptw` function allows missing data in `A` and
`Y`. Similarly, `adaptive_iptw` also allows for parallelized, cross-validated
estimation of the regression parameters via `cvFolds` and `parallel` options.
As with the estimators implemented in `drtmle`, cross-validation allows for
valid inference under weaker assumptions.

## Outcome-adaptive propensity score estimation

We have recently proposed 
estimators of the average treatment effect to be used in settings where this 
effect is weakly identifiable, as evidenced by small estimated propensity scores. 
In these settings doubly robust estimators may exhibit non-robust behavior
in small samples. The proposed solution was to target an alternative quantity 
in the propensity estimation step. If we are interested in estimating $\psi_0(a)$
for $a = 0, 1$, we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$,
which describes the conditional probability of receiving treatment $A = a$ given
the value of the two outcome regressions. For this reason, we call this quantity
an *outcome-adaptive propensity score*. We have shown that all of the usual 
asymptotics for TMLE carry through when an estimator of this quantity is 
substituted for the true propensity score. The resultant TMLE is *super-efficient*
meaning that it will generally have greater asymptotic efficiency than a standard
TMLE. Simulations indicate that this approach also delivers more stable behavior
in settings where estimated propensities become very small. However, the estimator 
is not doubly robust -- the asymptotics rely on the outcome regression being
consistently estimated at $n^{-1/4}$ rate. In this sense, this estimator trades off
robustness in the name of achieving greater stability and efficiency. See our
paper for further discussion.

# Discussion

The `drtmle` package was designed to provide a user-friendly implementation of
the TMLE algorithm that facilitates doubly-robust inference about marginal
means and average treatment effects. While the theory underlying doubly-robust
inference is complex, the implementation of the methods only requires users to
provide the data and specify how to estimate regressions. While estimation of
the OR and PS is familiar to those familiar with causal inference-related
estimation, the reduced-dimension regressions may appear strange. Indeed, it is
difficult to provide intuition for these parameters and more difficult to
determine what scientific background knowledge might guide users in how to
estimate these regression parameters. Therefore, we recommended estimating
these quantities adaptively, e.g., with the super learner. The super learner
library should contain several relatively smooth estimators such as `SL.mean`,
`SL.glm`, and `SL.gam`. In particular, including `SL.mean` (an unadjusted
generalized linear model) may be important due to the fact that, when the OR
and PS are consistently estimated, the reduced-dimension regressions are
identically equal to zero. Thus, this unadjusted regression estimator may do a
good job estimating the reduced-dimension regressions in situations where the
OR and PS are estimated well. In addition to relatively smooth estimators, we
recommend additionally including several adaptive estimators such as `SL.npreg`
or `SL.earth`.

Planned extensions of the `drtmle` package include extensions of longitudinal
settings involving treatments at multiple time points and inclusion of the
collaborative targeted maximum likelihood estimator (CTMLE) of \citet{vdl:2014:ijb} 
that also facilitates collaborative doubly-robust inference.

### 2021-06-03 JSS/using_drtmle_jss.Rmd
add jss files

--- 2021-06-03 paper/using_drtmle_jss.Rmd
@@
     email: \email{benkeser@emory.edu}
-    orcdi: 0000-0002-1019-8343
-  - name: Nima S. Hejazi
+  - name: Nima Hejazi
     affiliation: University of California, Berkeley
     email: \email{nhejazi@berkeley.edu}
-    orcid: 0000-0002-7127-2789
 title:
@@
   central limit theorem, the delta method, and the nonparametric bootstrap. 
-  However, in high-dimensional settings maximum likelihood estimation often
-  breaks down and standard procedures no longer yield correct inference.
-  Instead, we may rely on adaptive estimators of nuisance parameters to
-  construct flexible regression estimators.  However, use of adaptive estimators
-  poses a challenge for performing statistical inference about an estimated
-  treatment effect. While doubly-robust estimators facilitate inference when
-  *all* relevant regression functions are consistently estimated, the same
-  cannot be said when at least one estimator is inconsistent. \pkg{drtmle}
-  implements doubly-robust confidence intervals and hypothesis tests about
-  targeted minimum loss-based estimates of the average treatment effect. The
-  package additionally implements an inverse probability of treatment weighted
-  estimator that yields valid inference even when the probability of treatment
-  is estimated via adaptive methods.
+  However, in high-dimensional settings maximum likelihood estimation often breaks 
+  down and standard procedures no longer yield correct inference. Instead, we may 
+  rely on adaptive estimators of nuisance parameters to construct flexible regression estimators.
+  However, use of adaptive estimators poses a challenge for performing
+  statistical inference about an estimated treatment effect. While doubly-robust
+  estimators facilitate inference when *all* relevant regression functions are
+  consistently estimated, the same cannot be said when at least one estimator is
+  inconsistent. \pkg{drtmle} implements
+  doubly-robust confidence intervals and hypothesis tests about targeted minimum
+  loss-based estimates of the average treatment effect. The package additionally
+  implements an inverse probability of treatment weighted estimator that yields
+  valid inference even when the probability of treatment is estimated via
+  adaptive methods.
 output: rticles::jss_article
@@
 effect of a particular treatment on an outcome of interest. The treatment might
-correspond to a drug, a potentially harmful exposure, or a policy intervention.
-Often, the treatment cannot be randomized due to ethical or logistical reasons.
-This has sparked great interest in developing statistical methodology to
-estimate population-level effects from observational data.  Estimation of these
-effects often requires accounting for putative confounding factors, that is,
-factors related to whether a participant receives treatment and to the
-participant's outcome. In many settings, researchers measure many potential
-confounders, for example using administrative databases or genetic sequencing
-technology. Due to the curse-of-dimensionality, common statistical estimation
-techniques are not feasible for such high-dimensional data without restrictive
-modeling assumptions. When these assumptions are violated, estimates of the
-population-level effect of a treatment may have large bias.
+correspond to a drug, a potentially harmful exposure exposure, or a policy
+intervention. Often, the treatment may not be randomized due to ethical or
+logistical reasons. This has sparked great interest in developing statistical
+methodology to estimate population-level effects from observational data.
+Estimation of these effects often requires accounting for putative confounding
+factors, that is, factors related to whether a participant receives treatment
+and to the participant's outcome. In many settings, researchers measure many
+potential confounders, for example using administrative databases or genetic
+sequencing technology. Due to the curse-of-dimensionality, common statistical
+estimation techniques are not feasible for such high-dimensional data without
+restrictive modeling assumptions. When these assumptions are violated,
+estimates of the population-level effect of a treatment may have large bias.
 

### 2021-06-03 JSS/using_drtmle_jss.Rmd
update manuscript

@@
 effect of a particular treatment on an outcome of interest. The treatment might
-correspond to a drug, a potentially harmful exposure exposure, or a policy
-intervention. Often, the treatment may not be randomized due to ethical or
+correspond to a drug, a potentially harmful exposure, or a policy
+intervention. Often, the treatment cannot be randomized due to ethical or
 logistical reasons. This has sparked great interest in developing statistical
@@
 [@vdlrubin:2006:ijb]. Under assumptions, these estimators have limiting normal
-distributions with estimable variance. The asymptotic variance estimates then
+distributions with estimable variance. Estimates of the asymptotic variance then
 serve as the basis for constructing confidence intervals and hypothesis tests.
@@
 defined as $\psi_0 := \psi_0(1) - \psi_0(0)$. Under additional assumptions,
-these parameters have richer interpretations as mean counterfactual outcomes
+these parameters have interpretations as mean counterfactual outcomes
 under the different treatments and contrasts between these means define causal
@@
 $\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
-\bar{Q}_0(a,w) dQ_{W,n}(w).$ Often the empirical distribution function of
+\bar{Q}_0(a,w) dQ_{W,n}(w).$ Typically, the empirical distribution function of
 covariates is used so that this estimator can be equivalently written as
@@
 levels of a discrete-valued treatment and computes a doubly-robust covariance
-matrix for these estimates. The `drtmle` function implements either the
+matrix for these estimates. The function implements either the
 estimators of \citet{vdl:2014:ijb} (if `reduction = "bivariate"`) or the
@@
 the covariates, a discrete-valued treatment, and an outcome, respectively.
-Missing data are allowed in `A` and `Y`; this is discussed further in a section
-below. Beyond the data arguments, it is necessary to specify how to estimate
+Missing data are allowed in `A` and `Y`, as discussed below.
+Beyond the data arguments, it is necessary to specify how to estimate
 each of the OR, PS, R-OR, and R-PS (or R-PS1, R-PS2). The package provides
@@
 specifies the right-hand-side of a regression formula as a character in the
-`glm_Q`, `glm_g`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS, R-OR,
+`glm_g`, `glm_Q`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS, R-OR,
 and R-PS. These regressions are fit by calling the `glm` function in base R.
 Thus, the model specification should be able to be parsed as a valid formula,
-as required by `glm`. Recall that the R-OR parameter involves a regression onto
-the estimated propensity. Thus, the formula that is input via `glm_Qr` should
-assume data with a single column named `"gn"`. Similarly, the R-PS1 and R-PS2
+as required by `glm`. For the PS formula specified by `glm_g`, `colnames(W)` are 
+available to use in the formula. For the OR (`glm_Q`) available names
+of variables to be used in the formula depends on the `stratify` option.
+If `stratify = FALSE`, a single OR is fit to
+estimate $\bar{Q}_0(A,W)$.
+In this case, the OR regression is fit passing a `data.frame` into the `data` argument of `glm` that contains `W` and `A`.
+Thus, both `colnames(W)` and `"A"` may appear in the formula `glm_Q`.
+If `stratify = TRUE`, regressions are fit stratified by `A` to separately
+estimate $\bar{Q}_0(1,W)$ and $\bar{Q}_0(0,W)$. In this case, the `glm_Q` option 
+may not include `"A"`.
+
+To understand the variable names available for R-OR, recall that the parameter 
+involves a regression onto the estimated propensity. Thus, the formula that is 
+input via `glm_Qr` should
+assume a data with a single column, `"gn"`. 
+Similarly, the R-PS1 and R-PS2
 that are computed for the estimators of \citet{benkeser:2017:biometrika} (the default
@@
 data with a single column named `"Qn"` if `reduction = "univariate"`. If
-instead, the estimators of \citet{vdl:2014:ijb} are preferred, we can set the
-option `reduction = "bivariate"` and input a `glm_gr` formula that assumes data
-with two columns named `"Qn"` and `"gn"`.
+instead, the estimators of \citet{vdl:2014:ijb} are preferred (`reduction = "bivariate"`) the 
+`glm_gr` formula assumes data with two columns, `"Qn"` and `"gn"`.
 
@@
 `family = binomial()` so that logistic regression is used to estimate the OR.
-We specify `stratify = FALSE` to indicate that we wish to fit a single OR to
-estimate $\bar{Q}_0(A,W)$, as opposed to fitting two separate regressions to
-estimate $\bar{Q}_0(1,W)$ and $\bar{Q}_0(0,W)$. If the former fitting is
-specified, then the OR regression is fit using a data frame that contains `W`
-and `A`. Thus, both `colnames(W)` and `"A"` may appear in `glm_Q`. However, if
-`stratify = TRUE` then the OR model is fit separately in observations with $A =
-1$ and $A = 0$. In this case, the `glm_Q` option may not include `"A"`. We
-illustrate non-stratified regression for both the univariate and bivariate
+We illustrate non-stratified regression for both the univariate (default) and bivariate
 reductions.
@@
                       glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE,
+                      glm_gr = "Qn", glm_Qr = "gn + Qn", stratify = FALSE,
                       reduction = "bivariate")
@@
 Super learning is a loss-based ensemble machine learning technique
-[@vdl:2007:statapp] that is a generalization of stacking and stacked regression
-[@wolpert:1992:nnet]breiman:1996:ml}. The user specifies many potential
+[@vdl:2007:statapp] that generalizes stacking and stacked regression
+\citep{wolpert:1992:nnet,breiman:1996:ml}. The user specifies many potential
 candidate regression estimators, referred to as a *library* of candidate
@@
 
-Super learning is implemented in the R package `SuperLearner`
+Super learning is implemented in the R package \pkg{SuperLearner}
 [@SuperLearnerPackage] and `drtmle` provides the option to utilize super
@@
 Some users may find it unsettling that when applying the super learner in real data 
-analysis is that the results are dependent on the random seed that is set. To remove 
-this dependence, one can instead average results over repeated super learner runs; more 
+analysis the results are dependent on the random seed that is set. To remove 
+this dependence, one can average results over repeated super learner runs; more 
 repeats will lead to less dependence on the random seed. There are generally two approaches
@@
 
-For(ii), suppose for example that we have $n_{SL}$ TMLE (similarly, AIPTW) estimates 
+For (ii), suppose for example that we have $n_{SL}$ TMLE (similarly, AIPTW) estimates 
 $\psi_{n,j}(1)$ of $\psi_0(1)$, each obtained based on one (or more) super learner of the OR and/or PS.
@@
 call to `SuperLearner`.
-
-### Outcome-adaptive propensity score
-
-If the outcome-adaptive PS is desired, the user may specify this via 
-the `adapt_g` option. In this case, the additional targeting steps needed for doubly
-robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed above,
-this estimator is *not* doubly robust. Note that when the `adapt-g = TRUE` the data
-used for the PS estimation will be a `data.frame` with column names
-`QaW` where `a` takes all the values specified in `a_0`. For example if `a_0 = c(0,1)`,
-then the PS estimation data frame will have two columns called `Q0W`
-and `Q1W`. This may be relevant for users writing their own `SuperLearner` wrappers
-or those who make use of the `glm_g` setting, as illustrated below.
-
-```{r, message = FALSE}
-# outcome adaptive propensity score fit using glms
-adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
-                     adapt_g = TRUE)
-
-# outcome adaptive propensity score fit using super learner
-adaptg_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                     SL_Q = c("SL.glm", "SL.mean"), 
-                     SL_g = c("SL.glm", "SL.mean"),
-                     adapt_g = TRUE)
-```
 
@@
 used to estimate the OR and PS. In `drtmle` we give an option to compute standard error estimates that
-are *partially cross-validated* and do not require any additional fitting. Here we describe briefly 
-how these estimates are obtained and show calls to `drtmle` to obtain these estimates.
+are *partially cross-validated* and do not require any additional fitting beyond a standard super learner. 
+Here we describe briefly how these estimates are obtained and show calls to `drtmle` to obtain these estimates.
 
@@
 
+### Implementation of outcome-adaptive propensity score
+
+If the outcome-adaptive PS is desired, the user may specify this via 
+the `adapt_g` option. In this case, the additional targeting steps needed for doubly
+robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed above,
+this estimator is *not* doubly robust. Note that when the `adapt_g = TRUE` the data
+used for the PS estimation will be a `data.frame` with column names
+`QaW` where `a` takes all the values specified in `a_0`. For example if `a_0 = c(0,1)`,
+then the PS estimation data frame will have two columns called `Q0W`
+and `Q1W`. This may be relevant for users writing their own `SuperLearner` wrappers
+or those who make use of the `glm_g` setting.
+
+```{r, message = FALSE}
+# outcome adaptive propensity score fit using glms
+adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
+                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
+                     adapt_g = TRUE)
+
+# outcome adaptive propensity score fit using super learner
+adaptg_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
+                     SL_Q = c("SL.glm", "SL.mean"), 
+                     SL_g = c("SL.glm", "SL.mean"),
+                     adapt_g = TRUE)
+```
+
 # Discussion

### 2021-06-03 JSS/using_drtmle_jss.Rmd
update manuscript

@@
                       glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-                      glm_gr = "Qn", glm_Qr = "gn + Qn", stratify = FALSE,
+                      glm_gr = "Qn + gn", glm_Qr = "gn", stratify = FALSE,
                       reduction = "bivariate")
@@
 ```{r, message = FALSE}
+# use simplified data
+set.seed(1234)
+n <- 200
+W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
+A <- rbinom(n, 1, plogis(W$W1 + W$W2))
+Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
+
 # outcome adaptive propensity score fit using glms

### 2021-06-03 JSS/using_drtmle_jss.Rmd
porting over minor JSS changes from fork

@@
     email: \email{benkeser@emory.edu}
-  - name: Nima Hejazi
+    orcdi: 0000-0002-1019-8343
+  - name: Nima S. Hejazi
     affiliation: University of California, Berkeley
     email: \email{nhejazi@berkeley.edu}
+    orcdi: 0000-0002-7127-2789
 title:
@@
   central limit theorem, the delta method, and the nonparametric bootstrap.
-  However, in high-dimensional settings maximum likelihood estimation often breaks 
-  down and standard procedures no longer yield correct inference. Instead, we may 
-  rely on adaptive estimators of nuisance parameters to construct flexible regression estimators.
-  However, use of adaptive estimators poses a challenge for performing
-  statistical inference about an estimated treatment effect. While doubly-robust
-  estimators facilitate inference when *all* relevant regression functions are
-  consistently estimated, the same cannot be said when at least one estimator is
-  inconsistent. \pkg{drtmle} implements
-  doubly-robust confidence intervals and hypothesis tests about targeted minimum
-  loss-based estimates of the average treatment effect. The package additionally
-  implements an inverse probability of treatment weighted estimator that yields
-  valid inference even when the probability of treatment is estimated via
-  adaptive methods.
+  However, in high-dimensional settings maximum likelihood estimation often
+  breaks down and standard procedures no longer yield correct inference.
+  Instead, we may rely on adaptive estimators of nuisance parameters to
+  construct flexible regression estimators.  However, use of adaptive estimators
+  poses a challenge for performing statistical inference about an estimated
+  treatment effect. While doubly-robust estimators facilitate inference when
+  *all* relevant regression functions are consistently estimated, the same
+  cannot be said when at least one estimator is inconsistent. \pkg{drtmle}
+  implements doubly-robust confidence intervals and hypothesis tests about
+  targeted minimum loss-based estimates of the average treatment effect. The
+  package additionally implements an inverse probability of treatment weighted
+  estimator that yields valid inference even when the probability of treatment
+  is estimated via adaptive methods.
 output: rticles::jss_article
@@
 effect of a particular treatment on an outcome of interest. The treatment might
-correspond to a drug, a potentially harmful exposure, or a policy
-intervention. Often, the treatment cannot be randomized due to ethical or
-logistical reasons. This has sparked great interest in developing statistical
-methodology to estimate population-level effects from observational data.
-Estimation of these effects often requires accounting for putative confounding
-factors, that is, factors related to whether a participant receives treatment
-and to the participant's outcome. In many settings, researchers measure many
-potential confounders, for example using administrative databases or genetic
-sequencing technology. Due to the curse-of-dimensionality, common statistical
-estimation techniques are not feasible for such high-dimensional data without
-restrictive modeling assumptions. When these assumptions are violated,
-estimates of the population-level effect of a treatment may have large bias.
+correspond to a drug, a potentially harmful exposure, or a policy intervention.
+Often, the treatment cannot be randomized due to ethical or logistical reasons.
+This has sparked great interest in developing statistical methodology to
+estimate population-level effects from observational data.  Estimation of these
+effects often requires accounting for putative confounding factors, that is,
+factors related to whether a participant receives treatment and to the
+participant's outcome. In many settings, researchers measure many potential
+confounders, for example using administrative databases or genetic sequencing
+technology. Due to the curse-of-dimensionality, common statistical estimation
+techniques are not feasible for such high-dimensional data without restrictive
+modeling assumptions. When these assumptions are violated, estimates of the
+population-level effect of a treatment may have large bias.
 

### 2021-06-14 JSS/using_drtmle_jss.Rmd
edits up to std err

@@
   *all* relevant regression functions are consistently estimated, the same
-  cannot be said when at least one estimator is inconsistent. \pkg{drtmle}
-  implements doubly-robust confidence intervals and hypothesis tests about
-  targeted minimum loss-based estimates of the average treatment effect. The
-  package additionally implements an inverse probability of treatment weighted
-  estimator that yields valid inference even when the probability of treatment
-  is estimated via adaptive methods.
+  cannot be said when at least one nuisance estimator is inconsistent.
+  \pkg{drtmle} implements doubly-robust confidence intervals and hypothesis
+  tests for targeted minimum loss-based estimates of the average treatment
+  effect. The package additionally implements a targeted inverse probability of
+  treatment weighted estimator that yields valid inference even when the
+  probability of treatment is estimated via adaptive methods.
+header-includes:
+  - \usepackage{mathtools}
 output: rticles::jss_article
@@
 
-In many fields researchers are interested in assessing the population-level
+In many fields, researchers are interested in assessing the population-level
 effect of a particular treatment on an outcome of interest. The treatment might
@@
 confounders, for example using administrative databases or genetic sequencing
-technology. Due to the curse-of-dimensionality, common statistical estimation
+technology. Due to the curse of dimensionality, common statistical estimation
 techniques are not feasible for such high-dimensional data without restrictive
@@
 machine learning literature to control for high-dimensional confounders when
-estimating treatment effects. However, facilitating proper inference (e.g.,
+estimating treatment effects. However, facilitating proper inference (i.e.,
 confidence intervals and hypothesis tests) about estimates of effects based on
@@
 estimators [@horvitzthompson:1952:jasa, @robins:2000:epi], fail to yield valid
-inference for common estimators of effects. More complex proposals have been
-developed that *are capable* of utilizing adaptive methods to control for
-confounding while still yielding valid inference. These techniques include
-augmented inverse probability of treatment estimation (AIPTW)
-[@robins:1994:jasa] and targeted minimum loss-based estimation (TMLE)
-[@vdlrubin:2006:ijb]. Under assumptions, these estimators have limiting normal
-distributions with estimable variance. Estimates of the asymptotic variance then
-serve as the basis for constructing confidence intervals and hypothesis tests.
-
-As an intermediate step, the AIPTW and TMLE estimators require estimation of
+inference under adaptive estimates of nuisance parameters. More complex
+proposals have been developed that *are capable* of utilizing adaptive methods
+to control for confounding while still yielding valid inference. These
+techniques include augmented inverse probability of treatment weighted
+(AIPTW) estimation [@robins:1994:jasa] and targeted minimum loss-based (TML)
+estimation [@vdlrubin:2006:ijb]. Under standard assumptions, these estimators
+have limiting normal distributions with estimable variance. Estimates of the
+asymptotic variance serve as the basis for constructing confidence intervals and
+hypothesis tests.
+
+As an intermediate step, the AIPTW and TML estimators require estimation of
 two key nuisance parameters: the probability of treatment as a function of
@@
 regression, *OR*). In addition to their desirable inferential properties, AIPTW
-and TMLE estimators also enjoy the property of double robustness. That is, the
+and TML estimators also enjoy a *double robustness* property. That is, the
 estimators are consistent for the population-level effect of interest if either
 of the PS or OR is consistently estimated. This property gives the AIPTW and
-TMLE estimators a natural appeal: inconsistency in the estimation of one
+TML estimators a natural appeal: inconsistency in the estimation of one
 nuisance parameter may be mitigated by the consistent estimation of the other.
@@
 However, recent work has shown that the double robustness property does not
-necessarily extend to inferential properties about these estimators when
-adaptive estimators of the PS and OR are used
+necessarily extend to the inferential properties of these estimators when
+adaptive estimators of the PS and/or OR are used
 [@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
 consistent estimation of *both* the PS *and* the OR. \citet{vdl:2014:ijb}
-proposed new TMLE estimators have an asymptotic normal sampling distribution
+proposed new TML estimators that have an asymptotic normal sampling distribution
 with estimable variance under consistent estimation of *either* the PS *or* the
-OR, paving the way for inference that is doubly-robust. \citet{benkeser:2017:biometrika}
-proposed modifications of these estimators with similar robustness properties
-and illustrated their practical performance.
+OR, paving the way for inference that is doubly-robust.
+\citet{benkeser:2017:biometrika} proposed modifications of these estimators with
+similar robustness properties and illustrated their practical performance.
 
 The two proposed procedures that yield doubly-robust inference are quite
-involved, notably involving iterative estimation of additional, complex
+involved, notably requiring iterative estimation of additional, complex
 nuisance parameters. The `drtmle` package was motivated by the need for a
-user-friendly tool to implement these methods. The package implements the TMLE
+user-friendly tool to implement these methods. The package implements the TML
 estimators of population-level effects proposed in \citet{vdl:2014:ijb} and
-\citet{benkeser:2017:biometrika}, as well as several extensions including cross-validated
-targeted minimum loss-based estimators (CVTMLE). Additionally, `drtmle`
-implements an IPTW estimator that overcomes shortcomings of existing IPTW
-implementations with respect to statistical inference. In particular, the
+\citet{benkeser:2017:biometrika}, as well as several extensions including
+cross-validated targeted minimum loss-based estimation (CV--TMLE). Additionally,
+`drtmle` implements an IPTW estimator that overcomes shortcomings of existing
+IPTW implementations with respect to statistical inference. In particular, the
 estimator allows an adaptive estimator of the PS to be used in construction of
 the IPTW estimator, while yielding valid statistical inference about
-population-level effects. Both the TMLE and IPTW estimators implemented in the
+population-level effects. Both the TML and IPTW estimators implemented in the
 `drtmle` package are capable of estimating population-level effects for
-discrete-valued treatments and the package  provides convenient utilities for
-constructing confidence intervals and hypothesis tests about contrasts of the
-mean outcome under different levels of treatment. This document explains some
-of the key concepts underlying doubly-robust inference and  illustrates usage
-of the `drtmle` package.
+discrete-valued treatments; moreover, the package provides convenient utilities
+for constructing confidence intervals and hypothesis tests about contrasts of
+the mean outcome under different levels of treatment. This document explains
+some of the key concepts underlying doubly-robust inference and illustrates
+usage of the `drtmle` package.
 
@@
 distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
-a vector of baseline covariates, $A\in\{0,1\}$ is a binary (for the sake of
+a vector of baseline covariates, $A \in \{0,1\}$ is a binary (for the sake of
 simplicity) treatment, $Y$ is an outcome, and $P_0$ is the true data-generating
-distribution. The parameters of interest are \begin{equation}
+distribution. The parameters of interest are
+\begin{equation}
   \psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
@@
 function of the covariates. The value $\psi_0(a)$ represents the marginal
-(i.e., population-level)treatment-specific average outcome, hereafter referred
+(i.e., population-level) treatment-specific average outcome, hereafter referred
 to as the *marginal mean* under treatment $A=a$. The marginal effect of the
@@
 different treatment levels. For example, the average treatment effect is often
-defined as $\psi_0 := \psi_0(1) - \psi_0(0)$. Under additional assumptions,
-these parameters have interpretations as mean counterfactual outcomes
-under the different treatments and contrasts between these means define causal
-effects.
+defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
+assumptions, these parameters have interpretations as mean counterfactual
+outcomes under the different treatments, with contrasts between these means
+defining causal effects.
 
@@
 $\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
-\bar{Q}_0(a,w) dQ_{W,n}(w).$ Typically, the empirical distribution function of
-covariates is used so that this estimator can be equivalently written as
+\bar{Q}_0(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
+covariates is used so that this estimator can be equivalently expressed
 \begin{equation*}
@@
 \end{align}
-We use $g_0(a \mid W) := Pr_0(A = a \mid W)$ to denote the propensity score.
-Suppose we had available $g_n$, an estimate of the PS. Equation (2) motivates
-the estimator \begin{equation*}
-\psi_{n,IPTW}(a) := \frac{1}{n}\sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid
-W_i)} Y_i \ ,
+We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
+score. Suppose we had available $g_n$, an estimate of the PS. Equation (2)
+motivates the estimator
+\begin{equation*}
+  \psi_{n,IPTW}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
+    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
 \end{equation*}
-which is referred to as the inverse probability of treatment (IPTW) estimator.
-Inference for IPTW estimators based on parametric PS estimators may be
-facilitated through the nonparametric bootstrap. If adaptive approaches are
-used to estimate the PS, inference may not be available without further
-modification of the propensity score estimator, as we discuss below.
+which is referred to as the inverse probability of treatment weighted (IPTW)
+estimator. Like GCOMP estimators, inference for IPTW estimators based on
+parametric PS estimators may be facilitated through the nonparametric bootstrap.
+If adaptive approaches are used to estimate the PS, inference may not be
+available without further modification of the propensity score estimator, as we
+discuss below.
 
@@
 \end{equation*}
-
 which adds an augmentation term to the IPTW estimator. Equivalently, the
@@
 (TMLE). TMLE is a general methodology that may be used to modify an estimator
-of a regression function in such a way so as to ensure that the modified
-estimator satisfies user-selected equations. In the present problem, one can
-show that if an estimator of the OR $\bar{Q}_n^*$ satisfies \begin{equation}
-\label{meanEIFZero}
+of a regression function in a manner that ensures that the modified estimator
+satisfies user-selected equations. In the present problem, one can show that if
+an estimator of the OR $\bar{Q}_n^{\star}$ satisfies
+\begin{equation}\label{meanEIFZero}
   \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
-  \bar{Q}_n^*(a, W_i)\} = 0 \ ,
+  \bar{Q}_n^{\star}(a, W_i)\} = 0 \ ,
 \end{equation}
-then the GCOMP estimator based on $\bar{Q}_n^*$ will have the same asymptotic
-properties as the AIPTW estimator. That is, the estimator will be doubly-robust
-and, provided both the OR and PS are consistently estimated and regularity
-conditions are satisfied, will have a normal limiting distribution. TMLE
-provides a means of mapping an initial estimator of the OR into an estimator
-that satisfies (4). In addition to possessing the desirable asymptotic
-properties, TMLE estimators have been shown to outperform AIPTW estimators in
-finite samples [@porter:2011:ijb]. We refer readers to \citet{tlbook:2011}
-for more about TMLE.
+then the GCOMP estimator based on $\bar{Q}_n^{\star}$ will have the same
+asymptotic properties as the AIPTW estimator. That is, the estimator will be
+doubly-robust and, provided both the OR and PS are consistently estimated and
+regularity conditions are satisfied, will have a normal limiting distribution.
+TMLE provides a means of mapping an initial estimator of the OR into an
+estimator that satisfies (4). In addition to possessing the desirable asymptotic
+properties, TML estimators have been shown to outperform AIPTW estimators in
+finite samples [@porter:2011:ijb]. We refer readers to \citet{tlbook:2011} for
+more about TMLE.
 
@@
 \citet{vdl:2014:ijb} demonstrated that the double robustness property of AIPTW
-and TMLE estimators does not extend to their normal limiting distribution if
+and TML estimators does not extend to their normal limiting distribution if
 adaptive estimators of the OR and PS are used to construct the estimators.
@@
 biased estimates of the marginal means of interest. \citet{vdl:2014:ijb} derived
-modified TMLE estimators that are doubly-robust not only with respect to
-consistency, but also with respect to asymptotic normality. Furthermore, he
-provided closed-form, doubly-robust variance estimators that may be used to
-construct doubly-robust confidence intervals and hypothesis tests. \citet{benkeser:2017:biometrika} proposed simplifications of the \citet{vdl:2014:ijb} estimators and
-demonstrated the practical performance of estimators via simulation.
+modified TML estimators that are doubly-robust not only with respect to
+consistency, but also with respect to asymptotic normality. Furthermore, this
+work provided closed-form, doubly-robust variance estimators that may be used to
+construct doubly-robust confidence intervals and hypothesis tests.
+\citet{benkeser:2017:biometrika} proposed simplifications of the
+\citet{vdl:2014:ijb} estimators and demonstrated the practical performance of
+these estimators via simulation experiments.
 
@@
 refer to as the reduced outcome regression (R-OR) and the reduced propensity
-score (R-PS), defined respectively as \begin{align*}
-  \bar{Q}_{r,0n}(a,w) &:= E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
+score (R-PS), defined respectively as
+\begin{align*}
+  \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
   g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \\
-  g_{r,0n}(a \mid w) &:= Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
+  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
   g_n(W)=g_n(w)\right\} \ .
@@
 note that a key feature of these reduced-dimension regressions is that they are
-low-dimensional regardless of the dimension of $W$ and they do not depend on
-the true OR nor PS. Thus, these regressions may be consistently estimated at
-reasonably fast rates irrespective of the dimension of $W$ and irrespective of
-whether the OR or PS are consistently estimated. Based on these
+low-dimensional regardless of the dimension of $W$ and depend on neither the
+true OR nor the true PS. Thus, these regressions may be consistently estimated
+at reasonably fast rates irrespective of the dimension of $W$ and irrespective
+of whether the OR or PS are consistently estimated. Based on these
 reduced-dimension regressions, \citet{vdl:2014:ijb} showed that if estimates of
-the OR $\bar{Q}_n^*$, PS $g_n^*$, R-OR $\bar{Q}_{r,n}^*$, and R-PS $g_{r,n}^*$
-satisfied \begin{align}
-& \frac{1}{n} \sum\limits_{i=1}^n  \frac{I(A_i = a)}{g_n^*(a \mid W_i)} \{Y_i -
-\bar{Q}_n^*(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
+the OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
+R-PS $g_{r,n}^{\star}$ satisfied
+\begin{align}
+& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
+\{Y_i - \bar{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
 &\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
-\{I(A_i = a) - g_n^*(a \mid W_i)\} = 0 \ , \ \mbox{and} \label{tmleAsolve} \\
-&\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n}^*(a \mid W_i)}
-\biggl\{\frac{g_{r,n}^*(a \mid W_i) - g_n^*(a \mid W_i)}{g_n^*(a \mid W_i)}
-\biggr\} \{Y_i - \bar{Q}_n^*(a, W_i) \} = 0 \ , \label{tmleQsolve}
+\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
+\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
+\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
+\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
+{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \bar{Q}_n^{\star}(a, W_i) \} = 0 \ ,
+\label{tmleQsolve}
 \end{align}
-then the GCOMP estimator based on $\bar{Q}_n^*$ would be doubly-robust not only
-with respect to consistency, but also with respect to its limiting
+then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly-robust not
+only with respect to consistency, but also with respect to its limiting
 distribution. The author additionally proposed a TMLE algorithm to map initial
@@
 
-\citet{benkeser:2017:biometrika} demonstrated alternative equations that can be satisfied
-to achieve this form of double-robustness. In particular, they formulated an
-alternative version of the R-PS, \begin{align*}
-g_{r,0,1}(a \mid w) := Pr_0\{A = a \mid \bar{Q}_n(W) = \bar{Q}_n(w) \} \ ,
+\citet{benkeser:2017:biometrika} demonstrated alternative equations that can be
+satisfied to achieve this form of double-robustness. In particular, these
+authors formulated an alternative version of the R-PS,
+\begin{align*}
+  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \bar{Q}_n(W) =
+  \bar{Q}_n(w) \} \ ,
 \end{align*}
-
 which we refer to as R-PS1. They also introduced an additional regression of a
-weighted residual from the PS on the OR, \begin{align*}
-g_{r,0,2}(a \mid w) := E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}{g_n(a \mid
-W)} \ \bigg\vert \ \bar{Q}_n(W) = \bar{Q}_n(w) \biggr\} \ ,
+weighted residual from the PS on the OR,
+\begin{align*}
+  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
+  {g_n(a \mid W)} \ \bigg\vert \ \bar{Q}_n(W) = \bar{Q}_n(w) \biggr\} \ ,
 \end{align*}
-
 which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
-both univariate regressions, while R-PS is a bivariate regression. \citet{benkeser:2017:biometrika} suggested that this may make R-PS1 and R-PS2 easier to estimate
-consistently than R-PS. Further, they showed that if estimators or the OR, PS,
-and R-OR satisfy equations (5)-(6), and if additionally, estimators of the OR,
-R-PS-1 $g_{r,n,1}^*$, and R-PS-2 $g_{r,n,2}^*$ satisfy $$
-\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^*(a \mid W_i)}
-g_{r,n,2}^*(a \mid W_i) \ \{Y_i - \bar{Q}_n^*(a,W_i)\} = 0 \ ,
-$$
-then the GCOMP estimator based on this $\bar{Q}_n^*$ also enjoys a
-doubly-robust limiting distribution. \citet{benkeser:2017:biometrika} provided a TMLE
-algorithm that produced such estimates and provided closed-form, doubly-robust
-variance estimates.
+both univariate regressions, while R-PS is a bivariate regression.
+\citet{benkeser:2017:biometrika} suggested that this may make R-PS1 and R-PS2
+easier to estimate consistently than R-PS. Further, they showed that if
+estimators of the OR, PS, and R-OR satisfy equations (5)-(6), and if,
+additionally, estimators of the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2
+$g_{r,n,2}^{\star}$ satisfy
+\begin{equation*}
+\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
+(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
+\bar{Q}_n^{\star}(a,W_i)\} = 0 \ ,
+\end{equation*}
+then the GCOMP estimator based on this $\bar{Q}_n^{\star}$ also enjoys a
+doubly-robust limiting distribution. \citet{benkeser:2017:biometrika} provided a
+TMLE algorithm that produced such estimates and provided closed-form,
+doubly-robust variance estimates.
 
@@
 of our parameter of interest. Theoretically, these two frameworks are closely
-related and practically the two estimators are generally seen as competitors
+related and, practically, the two estimators are generally seen as competitors
 for estimating marginal means. In equation (3), we were able to generate a
@@
 solved by the standard TMLE. This is no coincidence; this term is crucial in
-studying the asymptotic properties of the GCOMP, AIPTW, and TMLE estimators
-(for discussion, see \citet{benkeser:2017:biometrika} Section 2.2). Recall that a TMLE
-estimator achieves doubly-robust consistency by satisfying equation (4), while
-the AIPTW achieves doubly-robust consistency by combining the GCOMP estimator
-with the left-hand-side of this term. Now, we have argued that a TMLE estimator
-may achieve doubly-robust limiting behavior if it additionally satisfies
-equations (6) and (7). It is thus sensible to wonder whether an AIPTW-style
-estimator could be constructed by combining the left-hand-side of these two
-additional equations with the AIPTW estimator. \citet{benkeser:2017:biometrika} showed
-that, surprisingly, this AIPTW-style estimator does not achieve the same
-doubly-robust properties as the TMLE estimator. Thus, while AIPTW and TMLE are
-generally competitors for producing estimators doubly-robust with respect to
-consistency, for doubly-robust inference, TMLE is preferred.
+studying the asymptotic properties of the GCOMP, AIPTW, and TML estimators
+(for discussion, see \citet{benkeser:2017:biometrika} Section 2.2). Recall that
+a TML estimator achieves doubly-robust consistency by satisfying equation (4),
+while the AIPTW achieves doubly-robust consistency by combining the GCOMP
+estimator with the left-hand-side of this term. Now, we have argued that a TML
+estimator may achieve doubly-robust limiting behavior if it additionally
+satisfies equations (6) and (7). It is thus sensible to wonder whether an
+AIPTW-style estimator could be constructed by combining the left-hand-side of
+these two additional equations with the AIPTW estimator.
+\citet{benkeser:2017:biometrika} showed that, surprisingly, this AIPTW-style
+estimator does not achieve the same doubly-robust properties as the TML
+estimator. Thus, while AIPTW and TMLE are generally competitors for producing
+estimators doubly-robust with respect to consistency, for doubly-robust
+inference, TMLE ought to be preferred.
 
@@
 estimators of \citet{vdl:2014:ijb} (if `reduction = "bivariate"`) or the
-improved estimators of \citet{benkeser:2017:biometrika} (`reduction = "univariate"`). The
-package includes utility functions for combining these estimates into estimates
-of average treatment effects, as well as other contrasts, as discussed below.
+improved estimators of \citet{benkeser:2017:biometrika}
+(`reduction = "univariate"`). The package includes utility functions for
+combining these estimates into estimates of average treatment effects, as well
+as other contrasts, as discussed below.
 
@@
 the covariates, a discrete-valued treatment, and an outcome, respectively.
-Missing data are allowed in `A` and `Y`, as discussed below.
-Beyond the data arguments, it is necessary to specify how to estimate
-each of the OR, PS, R-OR, and R-PS (or R-PS1, R-PS2). The package provides
-flexibility in how these regression parameters may be estimated. Once the user
-has specified the data arguments and how to estimate each regression parameter,
-the function proceeds as follows:
+Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
+arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
+and R-PS (or R-PS1, R-PS2). The package provides flexibility in how these
+regression parameters may be estimated. Once the user has specified the data
+arguments and how to estimate each regression parameter, the function proceeds
+as follows:
   1. estimate the OR via the internal `estimateQ` function;
   2. estimate the PS via the internal `estimateG` function;
-  3. estimate reduced-dimension regression via `estimateQrn` and
+  3. estimate reduced-dimension regression via the internal `estimateQrn` and
      `estimategrn` functions;
@@
       solved;
-  5. compute the TMLE estimates and covariance estimates based on the final OR.
-We refer interested readers to \citet{benkeser:2017:biometrika} for a theoretical
-derivation of how the iterative modification of initial OR and PS is performed.
+  5. compute the TML estimates and covariance estimates based on the final OR.
+We refer interested readers to \citet{benkeser:2017:biometrika} for a
+theoretical derivation of how the iterative modification of initial OR and PS is
+performed.
 
@@
 advisable to strive for consistent estimation of *both* the OR and PS. If both
-of the regression parameters are consistently estimated then the estimators
+of the regression parameters are consistently estimated, then the estimators
 computed by `drtmle` achieve the nonparametric efficiency bound, resulting in
-narrower confidence intervals and more powerful hypothesis tests. In order that
-the estimates of the OR and PS have the greatest chance of consistency, the
+narrower confidence intervals and more powerful hypothesis tests. In order for
+the estimates of the OR and PS to have the greatest chance of consistency, the
 `drtmle` function facilitates estimation using adaptive estimation techniques.
 In fact, there are three ways to estimate each of the regression parameters:
-generalized linear models, super learning, and a user-specified estimation
-technique. We demonstrate each of these approaches in turn on a simple
-simulated data set. The true parameter values of interest are $\psi_0(0)=$ `r
-round(EY0, 2)` and $\psi_0(1)=$ `r round(EY1, 2)`.
+generalized linear models, super learning
+[@vdl:2007:statapp; @SuperLearnerPackage], and a user-specified estimation
+technique. We demonstrate each of these approaches in turn on a simple simulated
+data set. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0,
+2)` and $\psi_0(1)=$ `r round(EY1, 2)`.
 
@@
 `glm_g`, `glm_Q`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS, R-OR,
-and R-PS. These regressions are fit by calling the `glm` function in base R.
+and R-PS. These regressions are fit by calling the `glm()` function in R [@R].
 Thus, the model specification should be able to be parsed as a valid formula,
-as required by `glm`. For the PS formula specified by `glm_g`, `colnames(W)` are 
-available to use in the formula. For the OR (`glm_Q`) available names
+as required by `glm()`. For the PS formula specified by `glm_g`, `colnames(W)`
+are available to use in the formula. For the OR (`glm_Q`) available names
 of variables to be used in the formula depends on the `stratify` option.
-If `stratify = FALSE`, a single OR is fit to
-estimate $\bar{Q}_0(A,W)$.
-In this case, the OR regression is fit passing a `data.frame` into the `data` argument of `glm` that contains `W` and `A`.
-Thus, both `colnames(W)` and `"A"` may appear in the formula `glm_Q`.
-If `stratify = TRUE`, regressions are fit stratified by `A` to separately
-estimate $\bar{Q}_0(1,W)$ and $\bar{Q}_0(0,W)$. In this case, the `glm_Q` option 
-may not include `"A"`.
+If `stratify = FALSE`, a single OR is fit to estimate $\bar{Q}_0(A,W)$. In this
+case, the OR regression is fit by passing a `data.frame` that contains `W` and
+`A` into the `data` argument of `glm()`. Thus, both `colnames(W)` and `"A"` may
+appear in the formula specified in `glm_Q`. If `stratify = TRUE`, regressions
+are fit stratifying on `A` to separately estimate $\bar{Q}_0(1,W)$ and
+$\bar{Q}_0(0,W)$. In this case, the `glm_Q` option may not include `"A"`.
 
 To understand the variable names available for R-OR, recall that the parameter
-involves a regression onto the estimated propensity. Thus, the formula that is 
-input via `glm_Qr` should
-assume a data with a single column, `"gn"`. 
-Similarly, the R-PS1 and R-PS2
-that are computed for the estimators of \citet{benkeser:2017:biometrika} (the default
-estimators, computed whenever `reduction = "univariate"`) involve regressions
-onto the estimated outcome regression. Thus, the formula `glm_gr` should assume
-data with a single column named `"Qn"` if `reduction = "univariate"`. If
-instead, the estimators of \citet{vdl:2014:ijb} are preferred (`reduction = "bivariate"`) the 
-`glm_gr` formula assumes data with two columns, `"Qn"` and `"gn"`.
+involves a regression onto the estimated propensity score. Thus, the formula
+that is input via `glm_Qr` should assume data with a single column, `"gn"`.
+Similarly, the R-PS1 and R-PS2 that are computed for the estimators of
+\citet{benkeser:2017:biometrika} (the default estimators, computed whenever
+`reduction = "univariate"`) involve regressions onto the estimated outcome
+regression. Thus, the formula `glm_gr` should assume data with a single column
+named `"Qn"` if `reduction = "univariate"`. Instead, if the estimators of
+\citet{vdl:2014:ijb} are preferred (`reduction = "bivariate"`) the `glm_gr`
+formula assumes data with two columns, `"Qn"` and `"gn"`.
 
@@
 `family = binomial()` so that logistic regression is used to estimate the OR.
-We illustrate non-stratified regression for both the univariate (default) and bivariate
-reductions.
+We illustrate non-stratified regression for both the univariate (default) and
+bivariate reductions below.
 
@@
 estimators, and the super learner algorithm estimates the convex combination of
-regression estimators that minimizes $V$-fold cross-validated risk based on a
-user-selected loss function. Oracle inequalities show that the super learner
-estimate performs essentially as well as the unknown best convex combination of
-the candidate estimators [@vdvaart:2006:statdec]. Thus, the methodology may be
-seen as an asymptotically optimal way of performing estimator selection.
-Practically, the super learner provides the best chance for obtaining a
-consistent estimator for both the OR and PS, and thereby an efficient estimator
-of the marginal means of interest.
+regression estimators that minimizes the $V$-fold cross-validated risk based on
+a user-selected loss function [@dudoit2005asymptotics].  Oracle inequalities
+show that the super learner estimate performs essentially as well as the unknown
+best convex combination of the candidate estimators [@vdvaart:2006:statdec].
+Thus, the methodology may be seen as an asymptotically optimal way of performing
+estimator selection.  Practically, the super learner provides the best chance
+for obtaining a consistent estimator for both the OR and PS, and thereby an
+efficient estimator of the marginal means of interest.
 
@@
 learner options should be invoked. If both are called, the function will
-default to the super learner option. The inputs to the super learner options
-are passed to the `SL.library` argument of the `SuperLearner` function of the
-`SuperLearner` package. We refer readers to the `SuperLearner` help file for
-information on how to properly specify a super learner library. In the code
-below, we illustrate a call to `drtmle` with a relatively simple library. For
-the OR and PS, the library includes a main-terms generalized linear model (via
-the function `SL.glm` from `SuperLearner`) and an unadjusted
-generalized linear model (`SL.mean`). For the reduced dimension regressions we
-use a super learner library with a main terms generalized linear model
-(`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the `earth`
-package [@earthPackage]).
+automatically opt to using the super learner option. The inputs to the super
+learner options are passed to the `SL.library` argument of the `SuperLearner`
+function of the `SuperLearner` package. We refer readers to the `SuperLearner`
+documentation for information on how to properly specify a super learner
+library. In the code below, we illustrate a call to `drtmle` with a relatively
+simple library. For the OR and PS, the library includes a main-terms generalized
+linear model (via the function `SL.glm` from `SuperLearner`) and an unadjusted
+generalized linear model (`SL.mean`). For the reduced-dimension regressions we
+use a super learner library with a main-terms generalized linear model
+(`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
+`earth` package [@earthPackage]).
 
@@
 formatted properly for compatibility with `drtmle`. The necessary formatting is
-identical the formatting required for writing a `SuperLearner` wrapper (see
+identical to the formatting required for writing a `SuperLearner` wrapper (see
 `SuperLearner::write.SL.template()`). The `drtmle` package includes `SL.npreg`,
@@
 `obsWeights` options are not strictly necessary (indeed, they are not used by
-`SL.npreg`), they are useful to include if one wishes to utilize the function
-as part of a super learner library. The `SL.npreg` function proceeds by
+`SL.npreg`), though these may be useful to include if one wishes to utilize the
+function as part of a super learner library. The `SL.npreg` function proceeds by
 checking whether there is sufficient variation in `Y` to justify fitting the
 kernel regression (as specified by option `rangeThresh`), and if so, fits a
-kernel regression with cross-validated bandwidth selection via the `npregbw`
+kernel regression with cross-validated bandwidth selection via the `npregbw()`
 function. The output is then formatted in a specific way so that `drtmle` is
@@
 
-Some users may find it unsettling that when applying the super learner in real data 
-analysis the results are dependent on the random seed that is set. To remove 
-this dependence, one can average results over repeated super learner runs; more 
-repeats will lead to less dependence on the random seed. There are generally two approaches
-that can be used: (i) averaging on the scale of the nuisance parameters and (ii) averaging
-on the scale of the point estimates.
-
-For (i), suppose that we have $n_{SL}$ estimated super 
-learners for the OR, $\bar{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS 
-$g_{n,j}, j = 1, \dots, n_{SL}$. We can take the average over these fits, \[
+Some users may find it unsettling that when applying the super learner in real
+data analysis the results are dependent on the random seed that is set. To
+remove this dependence, one can average results over repeated super learner
+runs; more repeats will lead to less dependence on the random seed. There are
+generally two approaches that can be used: (i) averaging on the scale of the
+nuisance parameters and (ii) averaging on the scale of the point estimates.
+
+For (i), suppose that we have $n_{SL}$ estimated super learners for the OR,
+$\bar{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS $g_{n,j}, j = 1, \dots,
+n_{SL}$. We can take the average over these fits,
+\begin{equation*}
   \bar{Q}_n = \frac{1}{n_{SL}} \sum_{j = 1}^{n_{SL}} \bar{Q}_{n,j}
-\] 
-as our estimate of the OR (similarly for the PS). These averaged estimates are then used to 
-compute the AIPTW or TMLE, as described above.
-
-For (ii), suppose for example that we have $n_{SL}$ TMLE (similarly, AIPTW) estimates 
-$\psi_{n,j}(1)$ of $\psi_0(1)$, each obtained based on one (or more) super learner of the OR and/or PS.
-We can take the average over these point estimates, \[
+\end{equation*}
+as our estimate of the OR (similarly for the PS). These averaged estimates are
+then used to compute the AIPTW or TMLE, as described above.
+
+For (ii), suppose for example that we have $n_{SL}$ TML (similarly, AIPTW)
+estimates $\psi_{n,j}(1)$ of $\psi_0(1)$, each obtained based on one (or more)
+super learners of the OR and/or PS. We can take the average over these point
+estimates,
+\begin{equation*}
   \psi_n(1) = \frac{1}{n_{SL}} \sum_{j=1}^{n_{SL}} \psi_{n,j}(1)
-\]
+\end{equation*}
 as our estimate of $\psi_{0}(1)$.
 
-These feature is implemented via the `n_SL` and `avg_over` options, where the former specifies the
-number of super learners to repeat and the latter specifies the scale(s) to average over. For example, if 
-`avg_over = "drtmle"` and `n_SL = 2`, then two point estimates are computed, each based on a single 
-super learner, and averaged to obtain a final estimate. If instead `avg_over = "SL"` and `n_SL = 2`, 
-then two super learners are fit and averaged before a final point estimate is obtained. Finally, if
-`avg_over = c("drtmle", "SL")` and `n_SL = 2` then averaging on *both* scales is performed. That is,
-the final estimate is the average of two point estimates, each obtained using OR and PS estimates that
-are averaged over two super learners.
-
-Not surprisingly, averaging over multiple super learners can add considerable computational expense 
-to calls to `drtmle`. The package is currently parallelized at the level of individual calls to 
-`SuperLearner`. With parallelization over enough cores, repeated super 
-learning should not add too much to the computational burden relative to a single 
-call to `SuperLearner`.
+This feature is implemented via the `n_SL` and `avg_over` options, where the
+former specifies the number of super learners to repeat and the latter specifies
+the scale(s) to average over. For example, if `avg_over = "drtmle"` and `n_SL =
+2`, then two point estimates are computed, each based on a single super learner,
+and averaged to obtain a final estimate. If instead `avg_over = "SL"` and `n_SL
+= 2`, then two super learners are fit and averaged before a final point estimate
+is obtained. Finally, if `avg_over = c("drtmle", "SL")` and `n_SL = 2` then
+averaging on *both* scales is performed. That is, the final estimate is the
+average of two point estimates, each obtained using OR and PS estimates that are
+averaged over two super learners.
+
+Not surprisingly, averaging over multiple super learners can add considerable
+computational expense to calls to `drtmle`. The package is currently
+parallelized at the level of individual calls to `SuperLearner`. With
+parallelization over enough cores, repeated super learning should not add too
+much to the computational burden relative to a single call to `SuperLearner`.
 
@@
 The `drtmle` package also allows users to pass in their own estimated OR and PS
-via the `Qn` and `gn` options, respectively. 
-If `Qn` is specified by the user, `drtmle` will ignore the nuisance parameter 
-estimation routines specified by `SL_Q` and `glm_Q` -- similarly for `gn` and
-`SL_g`, `glm_g`. As noted in the package documentation, there is a specific format 
-necessary for inputting `Qn` and `gn`. For `Qn`, the entries in the list should 
-correspond to the OR evaluated at A and the observed values of `W`, 
-with order determined by the input to `a_0`. For example, if `a_0 = c(0, 1)` then 
-`Qn[[1]]` should be OR at $A = 0$ and `Qn[[2]]` should be outcome 
-regression at $A = 1$. The example below illustrates this concept.
+via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the user,
+`drtmle` will ignore the nuisance parameter estimation routines specified by
+`SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As noted in the
+package documentation, there is a specific format necessary for inputting `Qn`
+and `gn`. For `Qn`, the entries in the list should correspond to the OR
+evaluated at `A` and the observed values of `W`, with order determined by the
+input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the OR
+at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
+example below illustrates this concept.
 
@@
 
-Similarly, for `gn` we need to pass in properly ordered lists. For example, 
-if `a_0 = c(0, 1)` then `gn[[1]]` should be propensity for receiving $A = 0$ 
-and `gn[[2]]` should be the propensity for receiving $A = 1$. The example below 
-illustrates this concept.
+Similarly, for `gn` we need to pass in properly ordered lists. For example, if
+`a_0 = c(0, 1)` then `gn[[1]]` should be the PS for receiving $A = 0$ and
+`gn[[2]]` should be the PS for receiving $A = 1$. The example below illustrates
+this concept.
 
@@
 
-This flexibility allows users to use any regression technique to externally
+This flexibility allows users to leverage any regression technique to externally
 obtain initial estimates of nuisance parameters. The key step is making sure
@@
 
-This section contains some more advanced mathematical material and may be omitted by
-users interested only in implementation. For simplicity, we focus this discussion on the 
-standard TMLE/AIPTW estimators. The ideas generalize immediately to the DRTMLE estimator, 
-but the formulas are more complex.
-
-Reported standard error estimates are based on an estimate the asymptotic variance of 
-the estimators. This variance can be characterized by the estimators influence function.
-For example, let $\psi_{n}(1)$ be an AIPTW estimate of $\psi_0(1)$. Under regularity 
-conditions, \[
-  \psi_{n}(1) - \psi_0(1) = \frac{1}{n} \sum_{i=1}^n D^*(\bar{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
-\]
-where \[
-D^*(\bar{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}{g_0(1 \mid W_i)} \{ Y_i - \bar{Q}_0(1, W_i)\} + \bar{Q}_0(1, W_i) - \int \bar{Q}_0(1, w) dQ_{W,0}(w) \ .
-\]
-Thus, by the central limit theorem, $n^{1/2} \psi_{n}(1)$ converges in distribution to a Normal
-distribution with mean $\psi_0(1)$ and variance \[
-  \sigma^2_0 = E_0( [ D^*(\bar{Q}_0, g_0, Q_{W,0})(O) - E_0\{D^*(\bar{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
-\]
-An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical variance of $D_i = D^*(\bar{Q}_n, g_n, Q_{W,n})(O_i)$, \[
-  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{ D_i - \frac{1}{n} \sum_{j=1}^n D_j \right\}^2 \ . 
-\]
-Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n / n^{1/2}$.
-This idea generalizes naturally to estimating the asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.
+This section contains some more advanced mathematical material and may be
+omitted by users interested only in implementation. For simplicity, we focus
+this discussion on the standard TMLE/AIPTW estimators. The ideas generalize
+immediately to the DRTMLE estimator, but the formulas are more complex.
+
+Reported standard error estimates are based on an estimate the asymptotic
+variance of the estimators. This variance can be characterized by the estimators
+influence function.  For example, let $\psi_{n}(1)$ be an AIPTW estimate of
+$\psi_0(1)$. Under regularity conditions,
+\begin{equation*}
+  \psi_{n}(1) - \psi_0(1) = \frac{1}{n} \sum_{i=1}^n
+    D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
+\end{equation*}
+where
+\begin{equation*}
+  D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}
+  {g_0(1 \mid W_i)} \{ Y_i - \bar{Q}_0(1, W_i)\} + \bar{Q}_0(1, W_i) -
+  \int \bar{Q}_0(1, w) dQ_{W,0}(w) \ .
+\end{equation*}
+Thus, by the central limit theorem, $n^{1/2} \psi_{n}(1)$ converges in
+distribution to a Normal distribution with mean $\psi_0(1)$ and variance
+\begin{equation*}
+  \sigma^2_0 = E_0( [ D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O) -
+  E_0\{D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
+\end{equation*}
+An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical
+variance of $D_i = D^{\star}(\bar{Q}_n, g_n, Q_{W,n})(O_i)$,
+\begin{equation*}
+  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{ D_i - \frac{1}{n} \sum_{j=1}^n
+  D_j \right\}^2 \ .
+\end{equation*}
+Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is
+$\sigma_n / n^{1/2}$. This idea generalizes naturally to estimating the
+asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.
 
@@
 settings involving treatments at multiple time points and inclusion of the
-collaborative targeted maximum likelihood estimator (CTMLE) of \citet{vdl:2014:ijb} 
-that also facilitates collaborative doubly-robust inference.
+collaborative targeted maximum likelihood estimator (C--TMLE) of
+\citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust
+inference.

### 2021-06-14 JSS/using_drtmle_jss.Rmd
touch-ups up through CIs

@@
 This section contains some more advanced mathematical material and may be
-omitted by users interested only in implementation. For simplicity, we focus
-this discussion on the standard TMLE/AIPTW estimators. The ideas generalize
-immediately to the DRTMLE estimator, but the formulas are more complex.
-
-Reported standard error estimates are based on an estimate the asymptotic
-variance of the estimators. This variance can be characterized by the estimators
-influence function.  For example, let $\psi_{n}(1)$ be an AIPTW estimate of
-$\psi_0(1)$. Under regularity conditions,
+skipped over by users interested only in implementation. For simplicity, we focus
+this discussion on the standard TML/AIPTW estimators. The ideas generalize
+immediately to the DR-TML estimator, but the formulas are more complex in this
+latter case.
+
+Reported standard error estimates are based on an estimate of the asymptotic
+variance of the estimators. This variance can be characterized by influence
+function of the estimators. For example, let $\psi_{n}(1)$ be an AIPTW estimate
+of $\psi_0(1)$. Under regularity conditions,
 \begin{equation*}
@@
 Thus, by the central limit theorem, $n^{1/2} \psi_{n}(1)$ converges in
-distribution to a Normal distribution with mean $\psi_0(1)$ and variance
+distribution to a normal distribution with mean $\psi_0(1)$ and variance
 \begin{equation*}
@@
 An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical
-variance of $D_i = D^{\star}(\bar{Q}_n, g_n, Q_{W,n})(O_i)$,
-\begin{equation*}
-  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{ D_i - \frac{1}{n} \sum_{j=1}^n
+variance of $D_i = D^{\star}(\bar{Q}_n, g_n, Q_{W,n})(O_i)$, that is,
+\begin{equation*}
+  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{D_i - \frac{1}{n} \sum_{j=1}^n
   D_j \right\}^2 \ .
 \end{equation*}
-Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is
-$\sigma_n / n^{1/2}$. This idea generalizes naturally to estimating the
+Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n /
+n^{1/2}$. Conveniently, this idea generalizes naturally to estimating the
 asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.
@@
 
-For TMLE-based estimators, it is standard practice to construct $D_i$ above using the *targeted* 
-version of the OR (denoted by $\bar{Q}_n^*$ above). Recall that such estimates are obtained by modifying
-an initial OR estimate in such a way as to ensure a particular (set of) equation(s) is (are) solved.
-However, simulations have shown that better standard error estimates for TMLE may be achieved by instead
-using the *initial* OR estimate in the computation of standard errors, as described in the previous
-subsection. The `drtmle` function includes the `targeted_se` option to this end.
+For TML estimators, it is standard practice to construct $D_i$ using the
+*targeted* version of the OR (denoted by $\bar{Q}_n^{\star}$ above). Recall that
+such estimates are obtained by modifying an initial OR estimate so as to ensure
+a particular (set of) equation(s) is (are) solved.  However, simulations have
+shown that better standard error estimates for TMLE may be achieved by instead
+using the *initial* OR estimate in the computation of standard errors, as
+described in the preceding subsection. To this end, The `drtmle` function
+includes the `targeted_se` option.
 
@@
 
-In simulations, we often find that when employing machine learning to estimate the OR and/or PS, 
-the estimated standard errors are *too small* in small to moderate-sized samples, leading to 
-under-coverage of confidence intervals and possibly too-large type I errors of hypothesis tests that
-are based on these standard error estimates. This should not be too surprising when studying the form
-of the standard error estimates. In particular, they will include a term that looks like 
-a (weighted) mean squared-error (MSE) of the OR. Because machine learning algorithms often set out 
-to minimize empirical MSE as part of their training, we often find that the empirical MSE is biased 
-too small relative to the true MSE. Thus, in order to more accurately estimate MSE of a machine learning
-algorithm, we typically rely on cross-validation. The same trick can be used here to obtain a more
-conservative standard error estimate.
-
-Suppose we use $V$-fold cross-validation to obtained $V$ training-sample-specific estimates of the OR,
- $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v}, v = 1, \dots, V$.
-For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the validation fold.
-That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of $Y_i$.
-We define
- $g_{n,i}^{cv}$ similarly. Using these estimates, we can define \[
- D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} + \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1) 
- \]
-and compute an estimate of $\sigma^2_0$ using \[
-  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} - \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ . 
-\]
-
-Cross-validated standard errors can be obtained by setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1.
+Simulation studies have revealed that, when employing machine learning to
+estimate the OR and/or PS, the estimated standard errors are often *too small*,
+in small to moderate-sized samples. Problematically, this leads directly to
+under-coverage of confidence intervals and possibly inflated Type-I error rates
+of hypothesis tests based on these standard error estimates. This should not be
+too surprising, as the phenomenon is related to the form of the standard error
+estimates. In particular, such estimates include a term that resembles a
+(weighted) mean squared error (MSE) of the OR. Since machine learning algorithms
+often set out to minimize empirical MSE as part of their training process, it
+follows that the empirical MSE often ends up biased towards zero relative to the
+true MSE. To more accurately estimate the MSE of a machine learning algorithm,
+cross-validation can be used. The same sample-splitting idea can be used to
+obtain more conservative standard error estimates.
+
+Suppose we use $V$-fold cross-validation to obtaine $V$ training-sample-specific
+estimates of the OR, $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v},
+v = 1, \dots, V$. For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the
+estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the
+validation fold. That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of
+$Y_i$. We define $g_{n,i}^{cv}$ similarly. Using these estimates, we can define
+\begin{equation*}
+ D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} +
+ \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1)
+\end{equation*}
+and compute an estimate of $\sigma^2_0$ using
+\begin{equation*}
+  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} -
+  \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ .
+\end{equation*}
+
+In `drtmle`, cross-validated standard error estimates can be obtained by setting
+`se_cv = "full"` and `se_cvFolds` to a number greater than 1. For example,
 
@@
 
-Obtaining fully cross-validated standard errors can be fairly time intensive when super learner is
-used to estimate the OR and PS. In `drtmle` we give an option to compute standard error estimates that
-are *partially cross-validated* and do not require any additional fitting beyond a standard super learner. 
-Here we describe briefly how these estimates are obtained and show calls to `drtmle` to obtain these estimates.
-
-Suppose we have $K$ candidate regressions in our super learner library for both the OR and PS 
-(in practice, the number could be different for the OR as for the PS). Recall that a super learner
-estimate of for example $\bar{Q}_0(a, w)$ is obtained as a weighted combination of estimates 
-from each candidate regression. Let $\bar{Q}_{n,k}$ denote the estimate of $\bar{Q}_{0}$ obtained 
-by training algorithm $k$ using the full data set. The super learner estimate of $\bar{Q}_{0}(a,w)$ is \[
+Obtaining fully cross-validated standard errors can be fairly time-intensive
+when super learner is used to estimate the OR and PS. In `drtmle`, we give an
+option to compute standard error estimates that are *partially cross-validated*
+and do not require any additional fitting beyond a standard super learner. Here,
+we describe briefly how these estimates are obtained and show calls to `drtmle`
+for obtaining such estimates.
+
+Suppose we have $K$ candidate regressions in our super learner library for both
+the OR and PS (in practice, the number could be different for the OR and for the
+PS). Recall that a super learner estimate of, for example, $\bar{Q}_0(a, w)$ is
+obtained as a weighted combination of estimates from each candidate regression.
+Let $\bar{Q}_{n,k}$ denote the estimate of $\bar{Q}_{0}$ obtained by training
+algorithm $k$ using the full data set. The super learner estimate of
+$\bar{Q}_{0}(a,w)$ is
+\begin{equation*}
   \bar{Q}_{n,SL}(a, w) = \sum_{k = 1}^K \alpha_{n,k} \bar{Q}_{n,k}(a, w) \ ,
-\]
-where $\alpha_n = (\alpha_{n,1}, \dots, \alpha_{n,K})$ is a weight vector that is selected by minimizing
-a cross-validated risk criteria. That is, each of the $K$ algorithms is trained according to a $V$-fold
-cross-validation scheme, resulting in $V + 1$ fits of each algorithm (one in each of $V$ training folds
-plus one obtained by fitting the algorithm using the full data). Let $\bar{Q}_{n,k,v}$ denote the $k$-th
-algorithm trained using the $v$-th training fold. For $i = 1,\dots,n$, let $\bar{Q}_{n,k,i}^{cv}(1)$ 
-denote the estimate of $\bar{Q}_0(1, W_i)$ implied by the fitted algorithm $\bar{Q}_{n,k,v}$, which 
-was obtained when observation $i$ was in the validation sample. That is, for observations with
-$A_i = 1$, $\bar{Q}_{n,k,i}^{cv}(1)$ is an out of sample prediction of $Y_i$ based on the
-$k$-th algorithm. Now, we can construct a *partially cross-validated* super learner prediction as follows. For $i = 1, \dots, n$ let \[
-  \bar{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k} \bar{Q}_{n,k,i}^{cv}(1)
-\]
-be the estimate obtained by ensembling the training-set-specific algorithm fits based on $\alpha_n$. 
-We can use these estimates to produce our *partially cross-validated* standard error estimate as follows. We define \[
- D_i^{pcv} = \frac{I(A_i = 1)}{g_{n,SL,i}^{cv}} \{ Y_i - \bar{Q}_{n,SL,i}^{cv}(1)\} + \bar{Q}_{n,SL,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,SL,i}^{cv}(1) 
- \]
-and compute an estimate of $\sigma^2_0$ using \[
-  \sigma^2_{n,pcv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{pcv} - \frac{1}{n} \sum_{j=1}^n D_j^{pcv} \right\}^2 \ . 
-\]
-
-Note that is not a proper cross-validated estimate, since observation $i$ is used to compute the 
-weight vector $\alpha_n$. However, such estimates can be obtained based on the output of a single 
-round of super learning, thereby providing improved standard error estimates at minimal additional
-computational cost.
-
-Here, we demonstrate how to obtain such estimates using `drtmle`.
+\end{equation*}
+where $\alpha_n = (\alpha_{n,1}, \dots, \alpha_{n,K})$ is a weight vector that
+is selected by minimizing a cross-validated risk criterion. That is, each of the
+$K$ algorithms is trained according to a $V$-fold cross-validation scheme,
+resulting in $V + 1$ fits of each algorithm (one in each of $V$ training folds,
+plus one obtained by fitting the algorithm using the full data). Let
+$\bar{Q}_{n,k,v}$ denote the $k$\textsuperscript{th} algorithm trained using the
+$v$\textsuperscript{th} training fold. For $i = 1, \dots, n$, let
+$\bar{Q}_{n,k,i}^{cv}(1)$ denote the estimate of $\bar{Q}_0(1, W_i)$ implied by
+the fitted algorithm $\bar{Q}_{n,k,v}$, which was obtained when observation $i$
+was in the validation sample. That is, for observations with $A_i = 1$,
+$\bar{Q}_{n,k,i}^{cv}(1)$ is an out-of-sample prediction of $Y_i$ based on the
+$k$\textsuperscript{th} algorithm. Now, we can construct a *partially
+cross-validated* super learner prediction as follows. For $i = 1, \dots, n$ let
+\begin{equation*}
+  \bar{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k}
+    \bar{Q}_{n,k,i}^{cv}(1)
+\end{equation*}
+be the estimate obtained by ensembling the training-set-specific algorithm fits
+based on $\alpha_n$. We can use these estimates to produce our *partially
+cross-validated* standard error estimate as follows. First, define
+\begin{equation*}
+  D_i^{pcv} = \frac{I(A_i = 1)}{g_{n,SL,i}^{cv}} \{ Y_i -
+    \bar{Q}_{n,SL,i}^{cv}(1)\} + \bar{Q}_{n,SL,i}^{cv}(1) -
+    \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,SL,i}^{cv}(1),
+\end{equation*}
+then, compute an estimate of $\sigma^2_0$ using
+\begin{equation*}
+  \sigma^2_{n,pcv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{pcv} -
+  \frac{1}{n} \sum_{j=1}^n D_j^{pcv} \right\}^2 \ .
+\end{equation*}
+
+Note that is not a properly cross-validated estimate, since observation $i$ is
+used to compute the weight vector $\alpha_n$. However, such estimates can be
+obtained based on the output of a single round of super learning, thereby
+providing improved standard error estimates at minimal additional computational
+cost.
+
+Below, we demonstrate how to obtain such estimates using `drtmle`.
 
@@
 settings involving treatments at multiple time points and inclusion of the
-collaborative targeted maximum likelihood estimator (C--TMLE) of
+collaborative targeted maximum likelihood estimator (C-TMLE) of
 \citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust

### 2021-06-15 JSS/using_drtmle_jss.Rmd
touch-ups to censoring

@@
 The `drtmle` package contains methods that compute doubly-robust confidence
-intervals. The `ci` method computes confidence intervals about the estimated
+intervals. The `ci()` method computes confidence intervals about the estimated
 marginal means in a fitted `drtmle` object, in addition to contrasts between
@@
 combination of marginal means, which can be used to estimate the average
-treatment effect as we show below.
+treatment effect, as we show below.
 
 ```{r}
-ci(npreg_fit, contrast = c(1,-1))
-```
-
-The `contrast` option can also be input as a list of functions in order to
+ci(npreg_fit, contrast = c(1, -1))
+```
+
+The `contrast` option also accepts as input a list of functions to be used to
 compute a confidence interval for a user-specified contrast. For example, we
 might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
-constructing Wald style confidence intervals for ratios, it is common to
-compute the interval on the log scale and back-transform. That is, the
-confidence interval is of the form
-$$
-\mbox{exp}\biggl[\mbox{log}\biggl\{ \frac{\psi_0(1)}{\psi_0(0)} \biggr\} \pm
-z_{1-\alpha/2} \sigma^{\text{log}}_n \biggr] \ ,
-$$
-where $\sigma^{\text{log}}_n$ is the estimated standard error on the log-scale.
-By the delta method, an estimate of this standard error is given by $$
-\sigma^{\text{log}}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
-$$
+constructing Wald-style confidence intervals for ratios, it is common to
+compute the interval on the logarithmic scale and subsequently back-transform.
+That is, the confidence interval takes the form
+\begin{equation*}
+  \exp \biggl[\log \biggl\{ \frac{\psi_0(1)}{\psi_0(0)} \biggr\} \pm
+  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
+\end{equation*}
+where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
+the delta method, an estimate of this standard error is given by
+\begin{equation*}
+  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
+\end{equation*}
 where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
-`drtmle` and $\nabla g(\psi)$ is the gradient of
-$\mbox{log}\{\psi(1)/\psi(0)\}$. To obtain this confidence interval, we may use
-the following code.
+`drtmle` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$. This
+confidence interval may be obtained as follows.
 
@@
                   f_inv = function(eff){ exp(eff) },
-                  h = function(est){ est[1]/est[2] },
-                  fh_grad =  function(est){ c(1/est[1],-1/est[2]) })
+                  h = function(est){ est[1] / est[2] },
+                  fh_grad =  function(est){ c(1 / est[1], -1 / est[2]) })
 ci(npreg_fit, contrast = riskRatio)
@@
 
-More generally this method of inputting the `contrast` option to the `ci`
+More generally, this method of inputting the `contrast` option to the `ci`
 method can be used to compute confidence intervals of the form
-$$
+\begin{equation*}
   f^{-1}\bigl\{ f(h(\psi_n)) \pm z_{1-\alpha/2} \nabla f(h(\psi_n))^T \Sigma_n
   \nabla f(h(\psi_n)) \bigr\} \ ,
-$$
+\end{equation*}
 where $f$ (`contrast$f`) is the transformation of the confidence interval,
 $f^{-1}$ (`contrast$f_inv`) is the back-transformation of the interval, $h$
-(`contrast$h`) is the contrast of marginal means, and $\nabla f(h(\psi_n))$
+(`contrast$h`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
 (`contrast$fh_grad`) defines the gradient of the transformed contrast at the
@@
 obtain a transformed confidence interval for a particular marginal mean. For
-example, in our running example $Y$ is binary. Thus, we might wish to enforce
-that our confidence interval be computed on a scale which ensures that the
-confidence limits are bounded between zero and one. To that end, we can
-consider an interval of the form construct an interval on the logit scale and
-back-transform,
-$$
-\mbox{expit}\biggl[ \mbox{log}\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
+example, recall that $Y$ is binary in our running example. Thus, we may wish to
+enforce that our confidence interval be computed on a scale ensuring that the
+estimated confidence limits fall within the unit interval $[0, 1]$. To that end,
+we can utilize an interval of the form:
+\begin{equation*}
+  \expit \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
   \pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
-$$
+\end{equation*}
 This confidence interval may be implemented as follows.
@@
                   h = function(est){ est[1] },
-                  fh_grad = function(est){ c(1/(est[1] - est[1]^2), 0) })
+                  fh_grad = function(est){ c(1 / (est[1] - est[1]^2), 0) })
 ci(npreg_fit, contrast = logitMean)
@@
 average treatment effect or to test the null hypothesis that the average
-treatment effect equals 0.05 (by specifying the `null` option)
+treatment effect equals a given reference value, e.g., $0.05$, by specifying
+the `null` option.
 
@@
 
-The `wald_test` method also allows for user-specified contrasts to be tested
-using syntax similar to the `ci` method. The only difference syntactically is
+The `wald_test()` method also allows for user-specified contrasts to be tested
+using syntax similar to the `ci()` method. The only difference syntactically is
 that it is not strictly necessary to specify `f_inv`, as the hypothesis test is
@@
 hypothesis that $f(h(\psi_0))$ equals $f_0$ (the function $f$ applied to the
-value passed to `null`) using the Wald statistic $$
-Z_n := \frac{\{f(h(\psi_n)) - f_0\}}{\nabla f(h(\psi_n))^T \Sigma_n \nabla
-f(h(\psi_n))} \ .
-$$
+value passed to `null`) using the Wald statistic
+\begin{equation*}
+  Z_n \coloneqq \frac{\{f(h(\psi_n)) - f_0\}}{\nabla f(h(\psi_n))^T \Sigma_n
+  \nabla f(h(\psi_n))} \ .
+\end{equation*}
 We can use the `riskRatio` contrast defined previously to test the null
-hypothesis that $\psi_0(1)/\psi_0(0) = 1$.
+hypothesis that $\psi_0(1) / \psi_0(0) = 1$.
 
@@
 
-The `drtmle` function supports missing data in `A` and `Y`. Such missing data
+The `drtmle()` function supports missing data in `A` and `Y`. Such missing data
 are common in practice and, if ignored, can bias estimates of the marginal
 means of interest. The estimators previously discussed can be modified to allow
-this missingness to depend on covariates. Let $\Delta_A$ and $\Delta_Y$ be
-indicators that $A$ and $Y$ are observed, respectively. We can redefine the OR
-to be $\bar{Q}_0(a,w) := E_0(\Delta_Y Y \mid \Delta_A A = a, \Delta_A = 1,
-\Delta_Y = 1, W = w)$ and similarly redefine the PS to be \begin{align}
+such missingness processes to depend on observed covariates. Let $\Delta_A$ and
+$\Delta_Y$ be indicators that $A$ and $Y$ are observed, respectively. We can
+redefine the OR to be $\bar{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
+A = a, \Delta_A = 1, \Delta_Y = 1, W = w)$ and similarly redefine the PS to be
+\begin{align}
   g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
@@
   \label{modPS} \\
-&\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w) \ \notag.
+  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w) \
+   \notag.
 \end{align}
-We introduce missing values to `A` and `Y` in our running example.
+
+To demonstrate, we artificially introduce missing values to `A` and `Y` in our
+running example.
 
@@
 
-The only modification of the call to `drtmle` is in how the PS estimation is
+The only modification of the call to `drtmle()` is in how the PS estimation is
 performed. The options `glm_g` and `SL_g` are still used to fit generalized
@@
 separate regression fits for each of the three components of the PS in (8). To
-perform separate generalized linear model fits for each of the three
-components, `glm_g` should be a named list with entries `"DeltaA"`, `"A"`, and
-`"DeltaY"`. Respectively these should provide a regression formula for the
-regression of $\Delta_A$ on $W$, $A$ on $W$ in observations with $\Delta_A =
-1$, and a regression of $\Delta_Y$ on $W$ and $A$ in observations with
-$\Delta_A = \Delta_Y = 1$ (if `stratify = FALSE`) or a regression of $\Delta_Y$
-on $W$ to be fit in observations with $\Delta_A = \Delta_Y = 1$ and $A = a$ for
-each $a$ specified in option `a_0` (if `stratify = TRUE`). If only a single
-formula is passed to `glm_g`, it is used for each of the three regressions. We
-illustrate a call to `drtmle` with missing data and generalized linear model
-fits below.
+perform separate generalized linear model fits for each of the three components,
+`glm_g` should be a named list with entries `"DeltaA"`, `"A"`, and `"DeltaY"`.
+Respectively, these should provide a formula expression for the regression of
+$\Delta_A$ on $W$, $A$ on $W$ in observations with $\Delta_A = 1$, and of
+$\Delta_Y$ on $W$ and $A$ in observations with $\Delta_A = \Delta_Y = 1$ (if
+`stratify = FALSE`) _or_ of $\Delta_Y$ on $W$ to be fit in observations with
+$\Delta_A = \Delta_Y = 1$ and $A = a$ for each $a$ specified in option `a_0` (if
+`stratify = TRUE`). If only a single formula is passed to `glm_g`, it is used
+for each of the three regressions. We illustrate a call to `drtmle()` with
+missing data and generalized linear model fits below.
 
@@
 It is possible to mix generalized linear models and super learning when fitting
-each of the components of the PS. In the below example we use the wrapper
+each of the components of the PS. In the example below, we use the wrapper
 function `SL.glm`, which fits a main terms logistic regression, to fit the
-regression of $\Delta_A$ on $W$, use `SL.npreg` to fit the regression of $A$ on
-$W$, and use a super learner with library `SL.glm`, `SL.mean`, and `SL.earth` to
+regression of $\Delta_A$ on $W$; use `SL.npreg` to fit the regression of $A$ on
+$W$; and use a super learner with library `SL.glm`, `SL.mean`, and `SL.earth` to
 fit the regression of $\Delta_Y$ on $W$ and $A$.
@@
 We can also fit the PS regressions externally and pass in via `gn`, though this
-process is slightly more complicated than when no missing data are present. 
-The basic idea is to fit one regression for (i) the indicator of missing $A$, 
-(ii) $A$ itself, and (iii) the indicator of missing the outcome. The three PS's 
-are then multiplied together and an appropriately ordered list is constructed
-as before. We illustrate with the following simple example.
+process is slightly more complicated than when no missing data are present. The
+basic idea is to fit one regression for (i) the indicator of missing $A$, (ii)
+$A$ itself, and (iii) the indicator of the outcome being missing. The three PS's
+are then multiplied together and an appropriately ordered list is constructed as
+before. We illustrate with the following simple example.
 
@@
 
-So far in our examples, we have only considered binary treatments. However,
+In our examples so far, we have only considered binary treatments. However,
 `drtmle` is equipped to handle treatments with an arbitrary number of discrete
@@
 regression estimation is modified to ensure that
-$$
+\begin{equation*}
   \sum\limits_{a \in \mathcal{A}} g_n(a \mid w) = 1 \ \mbox{for all $w$} \ .
-$$
+\end{equation*}
 If this condition is true, we say that the estimates of the PS are compatible.

### 2021-06-15 JSS/using_drtmle_jss.Rmd
finish first pass

@@
 \begin{equation*}
-  \expit \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
+  \text{expit} \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
   \pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
@@
 In our examples so far, we have only considered binary treatments. However,
-`drtmle` is equipped to handle treatments with an arbitrary number of discrete
-values. Suppose that $A$ assumes values in a set $\mathcal{A}$, the PS
-regression estimation is modified to ensure that
+\pkg{drtmle} is fully equipped to handle treatments with an arbitrary number of
+discrete values as well. Suppose that $A$ assumes values in a set $\mathcal{A}$
+of limited cardinality, the PS regression estimation is modified to ensure that
 \begin{equation*}
@@
 outcomes, but have not been extended to multi-level outcomes. To ensure that
-users of `drtmle` have access to the large number of binary regression
+users of \pkg{drtmle} have access to the large number of binary regression
 techniques when estimating the PS, we have taken an alternative approach to
@@
 and $\tilde{g}_n(1 \mid w)$ of $Pr_0(A = 1 \mid A > 0, W = w)$. The latter
-estimate may be generated by regression $I(A = 1)$ on $W$ in observations with
-$A > 0$. Because the outcome is binary, this regression may be estimated using
+estimate may be generated by regressing $I(A = 1)$ on $W$ in observations with
+$A > 0$. Since the outcome is binary, this regression may be estimated using
 any technique suitable for binary outcomes. By simple rules of conditional
-probability, we know that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
+probability, it holds that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
 Pr_0(A = 0 \mid W)\}$ and thus an estimate of $g_0(1 \mid w)$ can be computed
@@
 let $g_n(2 \mid w) = 1 - g_n(0 \mid w) - g_n(1 \mid w)$, which ensures
-compatibility of the estimates for each level of covariates. This approach
+compatibility of the estimates for each level of the covariates. This approach
 generalizes to an arbitrary number of treatment levels. Below, we provide an
@@
 Alternatively, we can specify a `contrast` to estimate confidence intervals
-about the average treatment effect of $A=1$ vs. $A=0$. Calls to `wald_test` can
-also be made.
+about the average treatment effect of $A=1$ versus $A=0$. Calls to `wald_test()`
+can also be made.
 
@@
 Similarly, we can compute confidence intervals comparing the average treatment
-effect of $A = 2$ vs. $A = 0$.
+effect of $A = 2$ versus $A = 0$.
 
@@
 
-Similarly, we can compute confidence intervals for the risk ratio comparing the
-marginal mean for $A=2$ to $A=1$.
+Just as straightforwardly, we can compute confidence intervals for the risk
+ratio comparing the marginal mean for $A=2$ to $A=1$.
 
@@
 When considering adaptive estimation techniques, one runs the risk of
-overfitting the initial regressions. While the cross validation used by super
-learning attempts to guard against overfitting, there are no guarantees in
-finite-samples. To definitively avoid such overfitting, one can use
-cross-validated estimates of the regression parameters. Theoretically, the use
-of cross-validated regression weakens the assumptions necessary to achieve
-asymptotic normality [@zheng:2010:working]. This is true of the regular TMLE and
-AIPTW, as well as of the TMLE estimator with doubly-robust inference. The
-`drtmle` function implements cross-validated estimation of the regression
-parameters through the `cvFolds` option, as shown in the code below, where we
-continue with the multi-level treatment example.
+overfitting the initial nuisance regressions. While the layer of
+cross-validation built into super learning attempts to guard against
+overfitting, there are no guarantees in finite samples. To definitively avoid
+such overfitting, one can use cross-validated estimates of the nuisance
+regressions. Theoretically, the use of cross-validated regression weakens the
+assumptions necessary to achieve asymptotic normality [@zheng:2010:working].
+This is true of the regular TML and AIPTW estimators, as well as of the TML
+estimator with doubly-robust inference. The `drtmle()` function implements
+cross-validated estimation of the regression parameters through the `cvFolds`
+option, as shown in the code below, where we continue on with the multi-level
+treatment example.
 
@@
 Cross-validation may significantly increase the computation time necessary for
-running `drtmle`. Parallelized regression fitting is available by setting the
-option `use_future = TRUE`. Since, internally, all computations are carried out
-using futures (with the API provided by the "future" R package), setting this
-option amounts only to parallelizing with futures [@futurePackage] via
-`future.apply::lapply`. Selecting this option results in parallelization of the
-estimation of the OR, PS, and reduced-dimension regressions.
-
-If no adjustments are made prior to invoking the `drtmle` function, futures are
-computed sequentially. It is up to the user to specify appropriate system-specific 
-future back-ends that may better suit their problem (e.g., using the `future.batchtools` 
-package [@futurebatchtoolsPackage]). Below we
-demonstrate registration using an ad-hoc cluster on the local system. The choice
-of back-end is flexible, and submission of the job to a wide variety
-of schedulers (e.g., SLURM, TORQUE, etc.) is also permitted. Interested readers
-are invited to consult the documentation of the `future.batchtools` package for
-further information.
-
-```{r, cache = TRUE, message = FALSE, eval = FALSE}
-## Commented out to avoid build errors
-# library(future.batchtools)
-# library(parallel)
-# cl <- makeCluster(2, type = "SOCK")
-# plan(cluster, workers = cl)
-# clusterEvalQ(cl, library("SuperLearner"))
-# pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-#                      SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
-#                      SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
-#                      SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
-#                      SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
-#                      stratify = FALSE, a_0 = c(0,1,2),
-#                      cvFolds = 2, use_future = TRUE)
+running `drtmle()`. Parallelized nuisance regression fitting is available by
+setting the option `use_future = TRUE`. Since, internally, all computations are
+carried out using futures (with the API provided by the \pkg{future} R package),
+setting this option amounts only to parallelizing with futures [@futurePackage]
+via `future.apply::lapply()`; when `use_future = FALSE`, base `lapply()` is
+simply used instead.  Together, setting these options results in parallelization
+of the estimation of the OR, PS, and reduced-dimension regressions.
+
+If no adjustments are made prior to invoking the `drtmle()` function, futures
+are computed sequentially, as per the defaults of the \pkg{future} package. It
+is left to the user to specify appropriate system-specific future backends that
+may better suit their problem (e.g., using the \pkg{future.batchtools} package
+[@futurebatchtoolsPackage] or even simply setting a parallelized
+`future::plan()`). Below, we demonstrate registration using an ad-hoc cluster on
+the local system. The choice of backend is flexible, and submission of the job
+to a wide variety of schedulers (e.g., SLURM, TORQUE) is also permitted.
+Interested readers are invited to consult the documentation of the
+\pkg{future.batchtools} package for further information.
+
+```{r, message = FALSE, eval = FALSE}
+library(future.batchtools)
+library(parallel)
+cl <- makeCluster(2, type = "SOCK")
+plan(cluster, workers = cl)
+clusterEvalQ(cl, library("SuperLearner"))
+pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
+                     SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
+                     SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
+                     SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
+                     SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
+                     stratify = FALSE, a_0 = c(0,1,2),
+                     cvFolds = 2, use_future = TRUE)
 ```
@@
 its intuitive derivation and implementation. Heretofore, inference for IPTW
-estimators precluded the use of adaptive estimators of the PS. However, \citet{vdl:2014:ijb} proposed modified IPTW estimators that allow for adaptive PS
-estimation. Similarly as above, this estimator is based on an additional
-regression parameter, $\tilde{Q}_{r,0n}(a,w) := E_0\{Y \mid A = a, g_n(W) =
-g_n(w)\}$, that is the regression of the outcome $Y$ on the estimated propensity
-amongst observations with $A=a$. The author showed that if $g_n^*(a \mid w)$, an
-estimator of the propensity for treatment $A=a$, satisfied that
-$$ \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
-W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,$$
-then the IPTW estimator based on $g_n^*$ has an asymptotic normal distribution
-with an estimable variance. In fact, \citet{vdl:2014:ijb} proposed a TMLE
-algorithm that mapped an initial PS estimate into an estimate that satisfies
-this key equation and variance estimators that can be used to generate
+estimators precluded the use of adaptive estimators of the PS. However,
+\citet{vdl:2014:ijb} proposed modified IPTW estimators that allow for adaptive
+PS estimation. Similarly as above, this estimator is based on an additional
+regression parameter, $\tilde{Q}_{r,0n}(a,w) \coloneqq E_0\{Y \mid A = a, g_n(W)
+= g_n(w)\}$, that is, the regression of the outcome $Y$ on the estimated
+propensity score amongst observations with $A=a$. The author showed that if
+$g_n^{\star}(a \mid w)$, an estimator of the propensity for treatment $A=a$,
+satisfied that
+\begin{equation*}
+  \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
+    W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,
+\end{equation*}
+then the IPTW estimator based on $g_n^{\star}$ would have an asymptotic normal
+distribution with estimable variance. In fact, \citet{vdl:2014:ijb} proposed a
+TMLE algorithm that mapped an initial PS estimate into an estimate that
+satisfies this key equation and variance estimators that can be used to generate
 Wald-style confidence intervals and hypothesis tests.
@@
 $g_n(a \mid w)$, can be computed as
-$$
+\begin{equation*}
   \psi_{n,IPTW}(a) = \psi_{n,IPTW}(a) - \frac{1}{n} \sum\limits_{i=1}^n
@@
   \mid W = W_i)\} \ .
-$$
+\end{equation*}
 Both this estimator and the TMLE-based IPTW estimator above are implemented in
-`drtmle`.
+\pkg{drtmle}.
 
@@
 The IPTW estimators based on adaptive estimation of the PS are implemented in
-the `adaptive_iptw` function. The `adaptive_iptw` function shares many options
-with `drtmle`. However, as the IPTW estimator does not rely on an estimate of
-the OR, these options are omitted. We continue with the multilevel treatment
-example.
+the `adaptive_iptw()` function. The `adaptive_iptw()` function shares many
+options with `drtmle()`; however, as the IPTW estimator does not rely on an
+estimate of the OR, options related to the OR are omitted. We demonstrate below,
+while continuing with the multi-level treatment example.
 
@@
 
-The `"adaptive_iptw"` class contains identical methods for computing confidence
-intervals and Wald tests as the `"drtmle"` class and we refer readers back to
-confidence intervals and hypothesis tests section for reminders on the various
-options for these tests. By default, these methods are implemented for the
-`iptw_tmle` estimator of \citet{vdl:2014:ijb}. We expect that this estimator
-will have improved finite-sample behavior relative to `iptw_os` (the
-alternative estimator described in the previous subsection), though further
-study is needed.
-
-As with `drtmle`, the `adaptive_iptw` function allows missing data in `A` and
-`Y`. Similarly, `adaptive_iptw` also allows for parallelized, cross-validated
-estimation of the regression parameters via `cvFolds` and `parallel` options.
-As with the estimators implemented in `drtmle`, cross-validation allows for
-valid inference under weaker assumptions.
+The `"adaptive_iptw"` S3 class contains identical methods for computing
+confidence intervals and Wald-style tests as the `"drtmle"` class, thus readers
+are referred back to the section on confidence intervals and hypothesis tests
+for reminders on the various options for these procedures. By default, these
+methods are implemented for the `iptw_tmle()` estimator of \citet{vdl:2014:ijb}.
+We conjecture that this estimator will have improved finite-sample behavior
+relative to `iptw_os()` (the alternative estimator described above), though
+further study in numerical experiments is warranted.
+
+As with `drtmle()`, the `adaptive_iptw()` function allows missing data in `A`
+and `Y`. Similarly, `adaptive_iptw()` also allows for parallelized,
+cross-validated estimation of the regression parameters via the `cvFolds` and
+`parallel` options.  As with the estimators implemented by `drtmle()`,
+cross-validation allows for valid inference under weaker assumptions.
 
@@
 
-We have recently proposed 
-estimators of the average treatment effect to be used in settings where this 
-effect is weakly identifiable, as evidenced by small estimated propensity scores. 
-In these settings doubly robust estimators may exhibit non-robust behavior
-in small samples. The proposed solution was to target an alternative quantity 
-in the propensity estimation step. If we are interested in estimating $\psi_0(a)$
-for $a = 0, 1$, we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$,
-which describes the conditional probability of receiving treatment $A = a$ given
-the value of the two outcome regressions. For this reason, we call this quantity
-an *outcome-adaptive propensity score*. We have shown that all of the usual 
+We have recently proposed estimators of the average treatment effect to be used
+in settings where this effect is weakly identifiable, as evidenced by estimated
+propensity scores of small magnitude.  In these settings, doubly robust
+estimators may exhibit non-robust behavior in small samples. The proposed
+solution was to target an alternative quantity in the propensity score
+estimation step. If we are interested in estimating $\psi_0(a)$ for $a = 0, 1$,
+we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$, which
+describes the conditional probability of receiving treatment $A = a$ given the
+value of the two outcome regressions. For this reason, we call this quantity an
+*outcome-adaptive propensity score*. We have shown that all of the usual
 asymptotics for TMLE carry through when an estimator of this quantity is
-substituted for the true propensity score. The resultant TMLE is *super-efficient*
-meaning that it will generally have greater asymptotic efficiency than a standard
-TMLE. Simulations indicate that this approach also delivers more stable behavior
-in settings where estimated propensities become very small. However, the estimator 
-is not doubly robust -- the asymptotics rely on the outcome regression being
-consistently estimated at $n^{-1/4}$ rate. In this sense, this estimator trades off
-robustness in the name of achieving greater stability and efficiency. See our
-paper for further discussion.
+substituted for the true propensity score. The resultant TML estimator is
+*super-efficient*, meaning that it will generally have greater asymptotic
+efficiency than a standard TML estimator. Simulation experiments indicate that
+this approach also delivers more stable behavior in settings where estimated
+propensity scores become very small. However, the estimator is not doubly robust
+-- the asymptotics rely on the outcome regression being consistently estimated
+at $n^{-1/4}$ rate. In this sense, this estimator trades off robustness against
+achieving greater stability and efficiency. See \citet{ju2020robust} for further
+discussion.
 
@@
 
-If the outcome-adaptive PS is desired, the user may specify this via
-the `adapt_g` option. In this case, the additional targeting steps needed for doubly
-robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed above,
-this estimator is *not* doubly robust. Note that when the `adapt_g = TRUE` the data
-used for the PS estimation will be a `data.frame` with column names
-`QaW` where `a` takes all the values specified in `a_0`. For example if `a_0 = c(0,1)`,
-then the PS estimation data frame will have two columns called `Q0W`
-and `Q1W`. This may be relevant for users writing their own `SuperLearner` wrappers
-or those who make use of the `glm_g` setting.
+If the outcome-adaptive PS is desired, the user may specify this via the
+`adapt_g` option. In this case, the additional targeting steps needed for doubly
+robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed
+above, this estimator is *not* doubly robust. Note that when `adapt_g = TRUE`,
+the data used for PS estimation will be a `data.frame` with column names `QaW`
+where `a` takes all the values specified in `a_0`. For example if `a_0 =
+c(0,1)`, then the PS estimation data frame will have two columns called `Q0W`
+and `Q1W`. This may be relevant for users writing their own \pkg{SuperLearner}
+wrapper functions or those who make use of the `glm_g` setting.
 
@@
 
-The `drtmle` package was designed to provide a user-friendly implementation of
-the TMLE algorithm that facilitates doubly-robust inference about marginal
+The \pkg{drtmle} package was designed to provide a user-friendly implementation
+of the TMLE algorithm that facilitates doubly-robust inference about marginal
 means and average treatment effects. While the theory underlying doubly-robust
 inference is complex, the implementation of the methods only requires users to
-provide the data and specify how to estimate regressions. While estimation of
-the OR and PS is familiar to those familiar with causal inference-related
-estimation, the reduced-dimension regressions may appear strange. Indeed, it is
-difficult to provide intuition for these parameters and more difficult to
-determine what scientific background knowledge might guide users in how to
-estimate these regression parameters. Therefore, we recommended estimating
-these quantities adaptively, e.g., with the super learner. The super learner
-library should contain several relatively smooth estimators such as `SL.mean`,
-`SL.glm`, and `SL.gam`. In particular, including `SL.mean` (an unadjusted
-generalized linear model) may be important due to the fact that, when the OR
-and PS are consistently estimated, the reduced-dimension regressions are
-identically equal to zero. Thus, this unadjusted regression estimator may do a
-good job estimating the reduced-dimension regressions in situations where the
-OR and PS are estimated well. In addition to relatively smooth estimators, we
-recommend additionally including several adaptive estimators such as `SL.npreg`
-or `SL.earth`.
-
-Planned extensions of the `drtmle` package include extensions of longitudinal
-settings involving treatments at multiple time points and inclusion of the
-collaborative targeted maximum likelihood estimator (C-TMLE) of
+provide the data and specify how to estimate nuisance regressions. Though
+estimation of the OR and PS is familiar to those with backgrounds in causal
+inference-related estimation, the reduced-dimension regressions may appear
+strange. Indeed, it is difficult to provide intuition for these parameters and
+more difficult still to determine what scientific background knowledge might
+guide users in how to estimate these regression parameters. Therefore, we
+recommended estimating these quantities adaptively, e.g., with the super
+learner. The super learner library should contain several relatively smooth
+estimators such as `SL.mean`, `SL.glm`, and `SL.gam`. In particular, including
+`SL.mean` (an unadjusted generalized linear model) may be important due to the
+fact that, when the OR and PS are consistently estimated, the reduced-dimension
+regressions are identically equal to zero. Thus, this unadjusted regression
+estimator may do a good job estimating the reduced-dimension regressions in
+situations where the OR and PS are estimated well. In addition to relatively
+smooth estimators, we recommend additionally including several adaptive
+estimators such as `SL.npreg` or `SL.earth`.
+
+Planned extensions of the \pkg{drtmle} package include extensions to
+longitudinal settings involving treatments at multiple time points and inclusion
+of the collaborative targeted maximum likelihood estimator (C-TMLE) of
 \citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust

### 2021-06-15 JSS/using_drtmle_jss.Rmd
final touch-ups

@@
     affiliation: University of California, Berkeley
+    address: |
+      | 2121 Berkeley Way,
+      | Berkeley, California, 94704, U.S.A.
     email: \email{nhejazi@berkeley.edu}
@@
 The two proposed procedures that yield doubly-robust inference are quite
-involved, notably requiring iterative estimation of additional, complex
-nuisance parameters. The `drtmle` package was motivated by the need for a
+involved, notably requiring iterative estimation of additional, complex nuisance
+parameters. The \pkg{drtmle} package was motivated by the need for a
 user-friendly tool to implement these methods. The package implements the TML
@@
 \citet{benkeser:2017:biometrika}, as well as several extensions including
-cross-validated targeted minimum loss-based estimation (CV--TMLE). Additionally,
-`drtmle` implements an IPTW estimator that overcomes shortcomings of existing
-IPTW implementations with respect to statistical inference. In particular, the
-estimator allows an adaptive estimator of the PS to be used in construction of
-the IPTW estimator, while yielding valid statistical inference about
-population-level effects. Both the TML and IPTW estimators implemented in the
-`drtmle` package are capable of estimating population-level effects for
+cross-validated targeted minimum loss-based estimation (CV-TMLE). Additionally,
+\pkg{drtmle} implements an IPTW estimator that overcomes shortcomings of
+existing IPTW implementations with respect to statistical inference. In
+particular, the estimator allows an adaptive estimator of the PS to be used in
+construction of the IPTW estimator, while yielding valid statistical inference
+about population-level effects. Both the TML and IPTW estimators implemented in
+the \pkg{drtmle} package are capable of estimating population-level effects for
 discrete-valued treatments; moreover, the package provides convenient utilities
@@
 some of the key concepts underlying doubly-robust inference and illustrates
-usage of the `drtmle` package.
+usage of the \pkg{drtmle} package.
 
@@
 
-The main function of the package is the eponymous `drtmle` function. This
+The main function of the package is the eponymous `drtmle()` function. This
 function estimates the treatment-specific marginal mean for user-specified
@@
 
-The main data arguments for `drtmle` are `W, A, Y`, which, as above, represent
+The main data arguments for `drtmle()` are `W, A, Y`, which, as above, represent
 the covariates, a discrete-valued treatment, and an outcome, respectively.
@@
 as follows:
-  1. estimate the OR via the internal `estimateQ` function;
-  2. estimate the PS via the internal `estimateG` function;
-  3. estimate reduced-dimension regression via the internal `estimateQrn` and
-     `estimategrn` functions;
-  4. iteratively modify initial estimates of the OR and PS via the `fluctuateQ`
-      and `fluctuateG` functions until equations (5)-(7) are approximately
-      solved;
+  1. estimate the OR via the internal `estimateQ()` function;
+  2. estimate the PS via the internal `estimateG()` function;
+  3. estimate reduced-dimension regression via the internal `estimateQrn()` and
+     `estimategrn()` functions;
+  4. iteratively modify initial estimates of the OR and PS via the
+     `fluctuateQ()` and `fluctuateG()` functions until equations (5)-(7) are
+     approximately solved;
   5. compute the TML estimates and covariance estimates based on the final OR.
@@
 In addition to returning doubly-robust parameter and covariance estimates, the
-`drtmle` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
+`drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
 modified AIPTW estimator discussed in the remark above. While doubly-robust
@@
 provided for comparison. In the following subsections, we look at various
-options for making calls to `drtmle`.
+options for making calls to `drtmle()`.
 
@@
 ```
-While the estimators computed by `drtmle` are consistent and asymptotically
+While the estimators computed by `drtmle()` are consistent and asymptotically
 normal under inconsistent estimation of either the OR or PS, it is nevertheless
@@
 of the regression parameters are consistently estimated, then the estimators
-computed by `drtmle` achieve the nonparametric efficiency bound, resulting in
+computed by `drtmle()` achieve the nonparametric efficiency bound, resulting in
 narrower confidence intervals and more powerful hypothesis tests. In order for
 the estimates of the OR and PS to have the greatest chance of consistency, the
-`drtmle` function facilitates estimation using adaptive estimation techniques.
+`drtmle()` function facilitates estimation using adaptive estimation techniques.
 In fact, there are three ways to estimate each of the regression parameters:
@@
 The code below illustrates how generalized linear models can be used to
-estimate the regression parameters. In this call to `drtmle`, we specify
+estimate the regression parameters. In this call to `drtmle()`, we specify
 `family = binomial()` so that logistic regression is used to estimate the OR.
@@
 Super learning is implemented in the R package \pkg{SuperLearner}
-[@SuperLearnerPackage] and `drtmle` provides the option to utilize super
-learning for estimation of the OR, PS, and reduced-dimension regressions via
-the `SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle` either
-the generalized linear model options for regression modeling or the super
-learner options should be invoked. If both are called, the function will
-automatically opt to using the super learner option. The inputs to the super
-learner options are passed to the `SL.library` argument of the `SuperLearner`
-function of the `SuperLearner` package. We refer readers to the `SuperLearner`
-documentation for information on how to properly specify a super learner
-library. In the code below, we illustrate a call to `drtmle` with a relatively
-simple library. For the OR and PS, the library includes a main-terms generalized
-linear model (via the function `SL.glm` from `SuperLearner`) and an unadjusted
-generalized linear model (`SL.mean`). For the reduced-dimension regressions we
-use a super learner library with a main-terms generalized linear model
-(`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
-`earth` package [@earthPackage]).
+[@SuperLearnerPackage] and `drtmle()` provides the option to utilize super
+learning for estimation of the OR, PS, and reduced-dimension regressions via the
+`SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle()` either the
+generalized linear model options for regression modeling or the super learner
+options should be invoked. If both are called, the function will automatically
+opt to using the super learner option. The inputs to the super learner options
+are passed to the `SL.library` argument of the `SuperLearner()` function of the
+\pkg{SuperLearner} package. We refer readers to the documentation of the
+`SuperLearner()` function for information on how to properly specify a super
+learner library. In the code below, we illustrate a call to `drtmle()` with a
+relatively simple library. For the OR and PS, the library includes a main-terms
+generalized linear model (via the function `SL.glm` from \pkg{SuperLearner}) and
+an unadjusted generalized linear model (`SL.mean`). For the reduced-dimension
+regressions we use a super learner library with a main-terms generalized linear
+model (`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
+\pkg{earth} package [@earthPackage]).
 
@@
 
-The `drtmle` package also allows users to pass in their own functions for
+The \pkg{drtmle} package also allows users to pass in their own functions for
 estimating regression parameters via the `SL_` options. These functions must be
-formatted properly for compatibility with `drtmle`. The necessary formatting is
-identical to the formatting required for writing a `SuperLearner` wrapper (see
-`SuperLearner::write.SL.template()`). The `drtmle` package includes `SL.npreg`,
-which is an example of how a user can specify a function for estimating
-regression parameters. This functions fits a kernel regression estimator using
-the `np` package [@npPackage]. Below we show the source code for this function.
+formatted properly for compatibility with `drtmle()`. The necessary formatting
+is identical to the formatting required for writing a `SuperLearner` wrapper
+(see `SuperLearner::write.SL.template()`). The \pkg{drtmle} package includes
+`SL.npreg`, which is an example of how a user can specify a function for
+estimating regression parameters. This functions fits a kernel regression
+estimator using the \pkg{np} package [@npPackage]. Below we show the source code
+for this function.
 
@@
 kernel regression with cross-validated bandwidth selection via the `npregbw()`
-function. The output is then formatted in a specific way so that `drtmle` is
+function. The output is then formatted in a specific way so that `drtmle()` is
 able to retrieve the appropriate objects from the fitted regression. In
@@
 on the fitted object. A class is assigned to the `fit` object, so that an S3
-`predict` method may later be used to predict new values based on the fitted
-regression object. The user must also specify this `predict` method. Use
-`drtmle:::predict.SL.npreg` to view the simple `predict` method for `SL.npreg`.
-
-In the code below, we demonstrate a call to `drtmle` with `SL.npreg` used to
+`predict()` method may later be used to predict new values based on the fitted
+regression object. The user must also specify this `predict()` method. Use
+`drtmle:::predict.SL.npreg()` to view the simple `predict()` method for
+`SL.npreg`.
+
+In the code below, we demonstrate a call to `drtmle()` with `SL.npreg` used to
 estimate each nuisance parameter. Here, we also illustrate the use of `stratify
@@
 Not surprisingly, averaging over multiple super learners can add considerable
-computational expense to calls to `drtmle`. The package is currently
-parallelized at the level of individual calls to `SuperLearner`. With
+computational expense to calls to `drtmle()`. The package is currently
+parallelized at the level of individual calls to `SuperLearner()`. With
 parallelization over enough cores, repeated super learning should not add too
-much to the computational burden relative to a single call to `SuperLearner`.
+much to the computational burden relative to a single call to `SuperLearner()`.
 
@@
 
-The `drtmle` package also allows users to pass in their own estimated OR and PS
-via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the user,
-`drtmle` will ignore the nuisance parameter estimation routines specified by
-`SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As noted in the
-package documentation, there is a specific format necessary for inputting `Qn`
-and `gn`. For `Qn`, the entries in the list should correspond to the OR
-evaluated at `A` and the observed values of `W`, with order determined by the
-input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the OR
-at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
+The \pkg{drtmle} package also allows users to pass in their own estimated OR and
+PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
+user, `drtmle()` will ignore the nuisance parameter estimation routines
+specified by `SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As
+noted in the package documentation, there is a specific format necessary for
+inputting `Qn` and `gn`. For `Qn`, the entries in the list should correspond to
+the OR evaluated at `A` and the observed values of `W`, with order determined by
+the input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the
+OR at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
 example below illustrates this concept.
@@
 This section contains some more advanced mathematical material and may be
-skipped over by users interested only in implementation. For simplicity, we focus
-this discussion on the standard TML/AIPTW estimators. The ideas generalize
-immediately to the DR-TML estimator, but the formulas are more complex in this
-latter case.
+skipped over by users interested only in implementation. For simplicity, we
+focus this discussion on the standard TML and AIPTW estimators. The ideas
+generalize immediately to the DR-TML estimator, but the formulas are more
+complex in this latter case.
 
@@
 using the *initial* OR estimate in the computation of standard errors, as
-described in the preceding subsection. To this end, The `drtmle` function
+described in the preceding subsection. To this end, The `drtmle()` function
 includes the `targeted_se` option.
@@
 
-Suppose we use $V$-fold cross-validation to obtaine $V$ training-sample-specific
+Suppose we use $V$-fold cross-validation to obtain $V$ training-sample-specific
 estimates of the OR, $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v},
@@
 
-In `drtmle`, cross-validated standard error estimates can be obtained by setting
-`se_cv = "full"` and `se_cvFolds` to a number greater than 1. For example,
+In \pkg{drtmle}, cross-validated standard error estimates can be obtained by
+setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1. For
+example,
 
@@
 Obtaining fully cross-validated standard errors can be fairly time-intensive
-when super learner is used to estimate the OR and PS. In `drtmle`, we give an
-option to compute standard error estimates that are *partially cross-validated*
-and do not require any additional fitting beyond a standard super learner. Here,
-we describe briefly how these estimates are obtained and show calls to `drtmle`
-for obtaining such estimates.
+when super learner is used to estimate the OR and PS. In \pkg{drtmle}, we give
+an option to compute standard error estimates that are *partially
+cross-validated* and do not require any additional fitting beyond a standard
+super learner. Here, we describe briefly how these estimates are obtained and
+show calls to `drtmle()` for obtaining such estimates.
 
@@
 
-Below, we demonstrate how to obtain such estimates using `drtmle`.
+Below, we demonstrate how to obtain such estimates using `drtmle()`.
 
@@
 
-The `drtmle` package contains methods that compute doubly-robust confidence
+The \pkg{drtmle} package contains methods that compute doubly-robust confidence
 intervals. The `ci()` method computes confidence intervals about the estimated
-marginal means in a fitted `drtmle` object, in addition to contrasts between
-those means. By default the method gives confidence intervals about each mean.
+marginal means in a fitted object of S3 class `"drtmle"`, in addition to
+contrasts between those means. By default, the method gives confidence intervals
+about each mean.
 
@@
 where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
-`drtmle` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$. This
-confidence interval may be obtained as follows.
+`drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
+This confidence interval may be obtained as follows.
 
@@
 
-More generally, this method of inputting the `contrast` option to the `ci`
+More generally, this manner of inputting the `contrast` option to the `ci()`
 method can be used to compute confidence intervals of the form
@@
 Doubly-robust Wald-style hypothesis tests may be implemented using the
-`wald_test` method. As with confidence intervals, there are three options for
+`wald_test()` method. As with confidence intervals, there are three options for
 testing. First, we may test the null hypothesis that the marginal means equal a
@@
 
-As with the `ci` method, the `wald_test` method allows users to test linear
+As with the `ci()` method, the `wald_test()` method allows users to test linear
 combinations of marginal means by inputting a vector of ones and negative ones

### 2021-06-16 JSS/using_drtmle_jss.Rmd
add renv; cosmetic tweaks

@@
 knitr::opts_chunk$set(warning=FALSE, message=FALSE)
+renv::activate(here::here())
+library(nloptr)
 ```
@@
 $\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
-\bar{Q}_0(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
+\bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
 covariates is used so that this estimator can be equivalently expressed
@@
 score (R-PS), defined respectively as
-\begin{align*}
+\begin{align}
   \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
-  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \\
+  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qbarr} \\
   g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
-  g_n(W)=g_n(w)\right\} \ .
-\end{align*}
+  g_n(W)=g_n(w)\right\} \ . \label{eqn:gr_bivariate}
+\end{align}
 In words, the R-OR is the regression of the residual from the outcome
@@
 [@vdl:2007:statapp; @SuperLearnerPackage], and a user-specified estimation
-technique. We demonstrate each of these approaches in turn on a simple simulated
+technique. We demonstrate each of these approaches in turn on a  simulated
 data set. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0,
@@
 are fit stratifying on `A` to separately estimate $\bar{Q}_0(1,W)$ and
-$\bar{Q}_0(0,W)$. In this case, the `glm_Q` option may not include `"A"`.
-
-To understand the variable names available for R-OR, recall that the parameter
+$\bar{Q}_0(0,W)$. In this case, the `glm_Q` formula may not include `"A"`.
+
+To understand the variable names available for R-OR, recall from (\ref{eqn:qbarr}) that the parameter
 involves a regression onto the estimated propensity score. Thus, the formula
@@
 \citet{vdl:2014:ijb} are preferred (`reduction = "bivariate"`) the `glm_gr` 
-formula assumes data with two columns, `"Qn"` and `"gn"`.
+formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes data with two columns, `"Qn"` and `"gn"`.
 
@@
 regression object. The user must also specify this `predict()` method. Use
-`drtmle:::predict.SL.npreg()` to view the simple `predict()` method for
+`drtmle:::predict.SL.npreg()` to view the `predict()` method for
 `SL.npreg`.
@@
 \end{equation*}
-where $f$ (`contrast$f`) is the transformation of the confidence interval,
-$f^{-1}$ (`contrast$f_inv`) is the back-transformation of the interval, $h$
-(`contrast$h`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
-(`contrast$fh_grad`) defines the gradient of the transformed contrast at the
+where $f$ (`contrast$f()`) is the transformation of the confidence interval,
+$f^{-1}$ (`contrast$f_inv()`) is the back-transformation of the interval, $h$
+(`contrast$h()`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
+(`contrast$fh_grad()`) defines the gradient of the transformed contrast at the
 estimated marginal means.
@@
 using syntax similar to the `ci()` method. The only difference syntactically is
-that it is not strictly necessary to specify `f_inv`, as the hypothesis test is
+that it is not strictly necessary to specify `f_inv()`, as the hypothesis test is
 performed on the transformed scale. That is, we can generally test the null
@@
 are then multiplied together and an appropriately ordered list is constructed as
-before. We illustrate with the following simple example.
+before. We illustrate with the following example.
 
@@
 $A > 0$. Since the outcome is binary, this regression may be estimated using
-any technique suitable for binary outcomes. By simple rules of conditional
+any technique suitable for binary outcomes. By rules of conditional
 probability, it holds that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
@@
 The multi-level PS can still be estimated using any of the three techniques
-discussed previously. Here, we illustrate with simple generalized linear
+discussed previously. Here, we illustrate with generalized linear
 models.
@@
 
-Just as straightforwardly, we can compute confidence intervals for the risk
+Similarly, we can compute confidence intervals for the risk
 ratio comparing the marginal mean for $A=2$ to $A=1$.

### 2021-06-25 JSS/using_drtmle_jss.Rmd
minor fix
updates to manuscript. move some sections to appendix

@@
   tests for targeted minimum loss-based estimates of the average treatment
-  effect. The package additionally implements a targeted inverse probability of
-  treatment weighted estimator that yields valid inference even when the
-  probability of treatment is estimated via adaptive methods.
+  effect, in addition to several other recently-proposed estimators of the 
+  average treatment effect.
 header-includes:
@@
 
-This has sparked an interest in using adaptive estimation techniques from the
+This has led to increased interest in adopting adaptive estimation techniques from the
 machine learning literature to control for high-dimensional confounders when
@@
 confidence intervals and hypothesis tests) about estimates of effects based on
-adaptive estimators is challenging. In particular, standard causal inference
+adaptive estimators is challenging. In particular, the use of standard causal inference
 techniques, including both G-computation (GCOMP) estimators
 [@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
-estimators [@horvitzthompson:1952:jasa, @robins:2000:epi], fail to yield valid
-inference under adaptive estimates of nuisance parameters. More complex
-proposals have been developed that *are capable* of utilizing adaptive methods
-to control for confounding while still yielding valid inference. These
-techniques include augmented inverse probability of treatment weighted
-(AIPTW) estimation [@robins:1994:jasa] and targeted minimum loss-based (TML)
-estimation [@vdlrubin:2006:ijb]. Under standard assumptions, these estimators
-have limiting normal distributions with estimable variance. Estimates of the
-asymptotic variance serve as the basis for constructing confidence intervals and
-hypothesis tests.
+estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when utilizing
+adaptive estimators. In particular, with rare exception, if cross-validation is used
+to select tuning parameters for the nuisance parameters, estimates of the population-level
+effect will generally be too biased to achieve standard $n^{1/2}$-rate asymptotics. Instead,
+*undersmoothing* must be used to select tuning parameters to control the bias 
+[@undersmoothed_hal]. However, this can prove quite challenging in practice and 
+few guidelines are available.
+
+A more natural framework for integrating adaptive methods into nuisance regression 
+estimation are doubly robust approaches, including augmented inverse probability of 
+treatment weighted (AIPTW) estimation [@robins:1994:jasa] and targeted minimum 
+loss-based (TML) estimation [@vdlrubin:2006:ijb]. The asymptotics of these estimators
+are characterized in part by the convergence rates of nuisance parameter estimates with
+respect to a proper norm. This fact makes the use of cross-validation more appealing for
+selection of tuning parameters for estimators of nuisance parameters, thereby providing 
+a more straightforward means of obtaining an estimate of the effect of interest that
+enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield closed form
+standard error estimates, which can serve as the basis for constructing confidence 
+intervals and hypothesis tests.
 
@@
 nuisance parameter may be mitigated by the consistent estimation of the other.
-As such, these estimators have increased in popularity in recent years.
 However, recent work has shown that the double robustness property does not
@@
 distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
-a vector of baseline covariates, $A \in \{0,1\}$ is a binary (for the sake of
-simplicity) treatment, $Y$ is an outcome, and $P_0$ is the true data-generating
+a vector of baseline covariates, $A \in \{0,1\}$ is a binary treatment, 
+$Y$ is an outcome, and $P_0$ is the true data-generating
 distribution. The parameters of interest are
@@
 approaches are used to estimate the OR, inference may not be available without
-further modification to the estimator.
+further modification to the estimator (e.g., undersmoothing as discussed above).
 
@@
 machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
-estimators lack a basis for statistical inference.
+estimators generally lack a basis for statistical inference.
 
@@
 \citet{vdl:2014:ijb} estimators and demonstrated the practical performance of
-these estimators via simulation experiments.
+these estimators via simulation.
 
@@
                     stratify = TRUE)
-npreg_fit
-```
+```
+
+In the Appendix, we describe an additional means of using `drtmle()` with 
+OR and PS estimates that have been obtained externally.
 
@@
 much to the computational burden relative to a single call to `SuperLearner()`.
-
-### Arbitrary user-specified regressions
-
-The \pkg{drtmle} package also allows users to pass in their own estimated OR and
-PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
-user, `drtmle()` will ignore the nuisance parameter estimation routines
-specified by `SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As
-noted in the package documentation, there is a specific format necessary for
-inputting `Qn` and `gn`. For `Qn`, the entries in the list should correspond to
-the OR evaluated at `A` and the observed values of `W`, with order determined by
-the input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the
-OR at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
-example below illustrates this concept.
-
-```{r, cache = TRUE, message = FALSE}
-# fit a GLM for outcome regression outside of drtmle
-out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())
-
-# generate Qn list
-Qn10 <- list(
-  # first entry is predicted values setting A = 1
-  predict(out_glm_fit, newdata = data.frame(A = 1, W), type = "response"),
-  # second entry is predicted values setting A = 0
-  predict(out_glm_fit, newdata = data.frame(A = 0, W), type = "response")
-)
-
-# pass this list to drtmle to avoid internal estimation of
-# outcome regression (note propensity score and reduced-dimension
-# regressions are still estimated internally)
-out_fit1 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn", Qn = Qn10,
-                  # crucial to set a_0 to match Qn's construction!
-                  a_0 = c(1,0))
-out_fit1
-
-# to be completely thorough let's re-order Qn10 and re-run
-Qn01 <- list(Qn10[[2]], Qn10[[1]])
-
-out_fit2 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn",
-                  # use re-ordered Qn list
-                  Qn = Qn01,
-                  # a_0 has to be reordered to match Qn01!
-                  a_0 = c(0,1))
-# should be the same as out_fit1, but re-ordered
-out_fit2
-```
-
-Similarly, for `gn` we need to pass in properly ordered lists. For example, if
-`a_0 = c(0, 1)` then `gn[[1]]` should be the PS for receiving $A = 0$ and
-`gn[[2]]` should be the PS for receiving $A = 1$. The example below illustrates
-this concept.
-
-```{r, cache = TRUE, message = FALSE}
-# fit a GLM for propensity score outside of drtmle
-out_glm_fit <- glm(A ~ . , data = data.frame(W), family = binomial())
-
-# get P(A = 1 | W) by calling predict
-gn1W <- predict(out_glm_fit, newdata = data.frame(W), type = "response")
-
-# generate gn list
-gn10 <- list(
-  # first entry is P(A = 1 | W)
-  gn1W,
-  # second entry is P(A = 0 | W) = 1 - P(A = 1 | W)
-  1 - gn1W
-)
-
-# pass this list to drtmle to avoid internal estimation of
-# propensity score (note reduced-dimension regressions are
-# still estimated internally)
-out_fit3 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                   glm_Qr = "gn", glm_gr = "Qn",
-                   Qn = Qn10, gn = gn10,
-                   # crucial to set a_0 to match Qn and gn's construction!
-                   a_0 = c(1,0))
-out_fit3
-
-# to be completely thorough let's re-order gn10 and re-run
-gn01 <- list(gn10[[2]], gn10[[1]])
-
-out_fit4 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                   glm_Qr = "gn", glm_gr = "Qn",
-                   # use re-ordered Qn/gn list
-                   Qn = Qn01, gn = gn01,
-                   # a_0 has to be reordered to match Qn01!
-                   a_0 = c(0,1))
-# should be the same as out_fit3, but re-ordered
-out_fit4
-```
-
-This flexibility allows users to leverage any regression technique to externally
-obtain initial estimates of nuisance parameters. The key step is making sure
-that the ordering of `Qn` and `gn` is consistent with `a_0`.
 
@@
 # compare variance to previously obtained
-list(glm_fit_uni$drtmle$cov, glm_fit_uni_cv$drtmle$cov)
+list(
+  `No CV` = glm_fit_uni$drtmle$cov, 
+  `CV` = glm_fit_uni_cv$drtmle$cov
+)
 ```
@@
 list(
-  sl_fit, # no cv
-  sl_fit_pcv # partial cv
+  `No-CV` = sl_fit$cov, # no cv
+  `Partial CV` = sl_fit_pcv$cov # partial cv
 )
@@
 \begin{equation*}
-  \exp \biggl[\log \biggl\{ \frac{\psi_0(1)}{\psi_0(0)} \biggr\} \pm
+  \exp \biggl[\log \biggl\{ \frac{\psi_n(1)}{\psi_n(0)} \biggr\} \pm
   z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
@@
                          glm_gr = "Qn", family = binomial())
-mixed_fit_wNAs
 ```
@@
 
-The user-specified contrasts are also available. For example, we can modify the
+User-specified contrasts are also available. For example, we can modify the
 `riskRatio` object we used previously for a two-level treatment to compute the
@@
 
-When considering adaptive estimation techniques, one runs the risk of
-overfitting the initial nuisance regressions. While the layer of
-cross-validation built into super learning attempts to guard against
-overfitting, there are no guarantees in finite samples. To definitively avoid
-such overfitting, one can use cross-validated estimates of the nuisance
+To avoid overfitting initial regressions, one can use cross-validated 
+estimates of the nuisance
 regressions. Theoretically, the use of cross-validated regression weakens the
@@
 cross-validated estimation of the regression parameters through the `cvFolds`
-option, as shown in the code below, where we continue on with the multi-level
-treatment example.
+option, as shown in the code below.
 
@@
                     stratify = FALSE, cvFolds = 2, a_0 = c(0, 1, 2))
-cv_sl_fit
 ```
@@
 
-# Additional methods included in drtmle
-
-## Inference for IPTW using adaptive PS estimators
+# Discussion
+
+The \pkg{drtmle} package was designed to provide a user-friendly implementation
+of the TMLE algorithm that facilitates doubly-robust inference about marginal
+means and average treatment effects. While the theory underlying doubly-robust
+inference is complex, the implementation of the methods only requires users to
+provide the data and specify how to estimate nuisance regressions. Though
+estimation of the OR and PS is familiar to those with backgrounds in causal
+inference-related estimation, the reduced-dimension regressions may appear
+strange. Indeed, it is difficult to provide intuition for these parameters and
+more difficult still to determine what scientific background knowledge might
+guide users in how to estimate these regression parameters. Therefore, we
+recommended estimating these quantities adaptively, e.g., with the super
+learner. The super learner library should contain several relatively smooth
+estimators such as `SL.mean`, `SL.glm`, and `SL.gam`. In particular, including
+`SL.mean` (an unadjusted generalized linear model) may be important due to the
+fact that, when the OR and PS are consistently estimated, the reduced-dimension
+regressions are identically equal to zero. Thus, this unadjusted regression
+estimator may do a good job estimating the reduced-dimension regressions in
+situations where the OR and PS are estimated well. In addition to relatively
+smooth estimators, we recommend additionally including several adaptive
+estimators such as `SL.npreg` or `SL.earth`.
+
+Planned extensions of the \pkg{drtmle} package include extensions to
+longitudinal settings involving treatments at multiple time points and inclusion
+of the collaborative targeted maximum likelihood estimator (C-TMLE) of
+\citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust
+inference.
+
+# Appendix
+
+## Arbitrary user-specified regressions
+
+The \pkg{drtmle} package also allows users to pass in their own estimated OR and
+PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
+user, `drtmle()` will ignore the nuisance parameter estimation routines
+specified by `SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As
+noted in the package documentation, there is a specific format necessary for
+inputting `Qn` and `gn`. For `Qn`, the entries in the list should correspond to
+the OR evaluated at `A` and the observed values of `W`, with order determined by
+the input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the
+OR at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
+example below illustrates this concept.
+
+```{r, cache = TRUE, message = FALSE}
+# use simplified data
+set.seed(1234)
+n <- 200
+W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
+A <- rbinom(n, 1, plogis(W$W1 + W$W2))
+Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
+
+# fit a GLM for outcome regression outside of drtmle
+out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())
+
+# generate Qn list
+Qn10 <- list(
+  # first entry is predicted values setting A = 1
+  predict(out_glm_fit, newdata = data.frame(A = 1, W), type = "response"),
+  # second entry is predicted values setting A = 0
+  predict(out_glm_fit, newdata = data.frame(A = 0, W), type = "response")
+)
+
+# pass this list to drtmle to avoid internal estimation of
+# outcome regression (note propensity score and reduced-dimension
+# regressions are still estimated internally)
+out_fit1 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
+                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn", Qn = Qn10,
+                  # crucial to set a_0 to match Qn's construction!
+                  a_0 = c(1,0))
+out_fit1
+
+# to be completely thorough let's re-order Qn10 and re-run
+Qn01 <- list(Qn10[[2]], Qn10[[1]])
+
+out_fit2 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
+                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn",
+                  # use re-ordered Qn list
+                  Qn = Qn01,
+                  # a_0 has to be reordered to match Qn01!
+                  a_0 = c(0,1))
+# should be the same as out_fit1, but re-ordered
+out_fit2
+```
+
+Similarly, for `gn` we need to pass in properly ordered lists. For example, if
+`a_0 = c(0, 1)` then `gn[[1]]` should be the PS for receiving $A = 0$ and
+`gn[[2]]` should be the PS for receiving $A = 1$. The example below illustrates
+this concept.
+
+```{r, cache = TRUE, message = FALSE}
+# fit a GLM for propensity score outside of drtmle
+out_glm_fit <- glm(A ~ . , data = data.frame(W), family = binomial())
+
+# get P(A = 1 | W) by calling predict
+gn1W <- predict(out_glm_fit, newdata = data.frame(W), type = "response")
+
+# generate gn list
+gn10 <- list(
+  # first entry is P(A = 1 | W)
+  gn1W,
+  # second entry is P(A = 0 | W) = 1 - P(A = 1 | W)
+  1 - gn1W
+)
+
+# pass this list to drtmle to avoid internal estimation of
+# propensity score (note reduced-dimension regressions are
+# still estimated internally)
+out_fit3 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
+                   glm_Qr = "gn", glm_gr = "Qn",
+                   Qn = Qn10, gn = gn10,
+                   # crucial to set a_0 to match Qn and gn's construction!
+                   a_0 = c(1,0))
+out_fit3
+
+# to be completely thorough let's re-order gn10 and re-run
+gn01 <- list(gn10[[2]], gn10[[1]])
+
+out_fit4 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
+                   glm_Qr = "gn", glm_gr = "Qn",
+                   # use re-ordered Qn/gn list
+                   Qn = Qn01, gn = gn01,
+                   # a_0 has to be reordered to match Qn01!
+                   a_0 = c(0,1))
+# should be the same as out_fit3, but re-ordered
+out_fit4
+```
+
+This flexibility allows users to leverage any regression technique to externally
+obtain initial estimates of nuisance parameters. The key step is making sure
+that the ordering of `Qn` and `gn` is consistent with `a_0`.
+
+## Additional methods included in drtmle
+
+### Inference for IPTW using adaptive PS estimators
 
@@
 
-### Implementation of inference for adaptive IPTW
+#### Implementation of inference for adaptive IPTW
 
@@
 options with `drtmle()`; however, as the IPTW estimator does not rely on an
-estimate of the OR, options related to the OR are omitted. We demonstrate below,
-while continuing with the multi-level treatment example.
+estimate of the OR, options related to the OR are omitted.
 
@@
                                    SL_g = "SL.npreg", SL_Qr = "SL.npreg",
-                                   family = binomial(), a_0 = c(0, 1, 2))
+                                   family = binomial(), a_0 = c(0, 1))
 npreg_iptw_multiA
@@
 
-## Outcome-adaptive propensity score estimation
+### Outcome-adaptive propensity score estimation
 
@@
 
-### Implementation of outcome-adaptive propensity score
+#### Implementation of outcome-adaptive propensity score
 
@@
 ```{r, message = FALSE}
-# use simplified data
-set.seed(1234)
-n <- 200
-W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
-A <- rbinom(n, 1, plogis(W$W1 + W$W2))
-Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
-
 # outcome adaptive propensity score fit using glms
@@
 ```
-
-# Discussion
-
-The \pkg{drtmle} package was designed to provide a user-friendly implementation
-of the TMLE algorithm that facilitates doubly-robust inference about marginal
-means and average treatment effects. While the theory underlying doubly-robust
-inference is complex, the implementation of the methods only requires users to
-provide the data and specify how to estimate nuisance regressions. Though
-estimation of the OR and PS is familiar to those with backgrounds in causal
-inference-related estimation, the reduced-dimension regressions may appear
-strange. Indeed, it is difficult to provide intuition for these parameters and
-more difficult still to determine what scientific background knowledge might
-guide users in how to estimate these regression parameters. Therefore, we
-recommended estimating these quantities adaptively, e.g., with the super
-learner. The super learner library should contain several relatively smooth
-estimators such as `SL.mean`, `SL.glm`, and `SL.gam`. In particular, including
-`SL.mean` (an unadjusted generalized linear model) may be important due to the
-fact that, when the OR and PS are consistently estimated, the reduced-dimension
-regressions are identically equal to zero. Thus, this unadjusted regression
-estimator may do a good job estimating the reduced-dimension regressions in
-situations where the OR and PS are estimated well. In addition to relatively
-smooth estimators, we recommend additionally including several adaptive
-estimators such as `SL.npreg` or `SL.earth`.
-
-Planned extensions of the \pkg{drtmle} package include extensions to
-longitudinal settings involving treatments at multiple time points and inclusion
-of the collaborative targeted maximum likelihood estimator (C-TMLE) of
-\citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust
-inference.

### 2021-06-25 JSS/using_drtmle_jss.Rmd

@@
 renv::activate(here::here())
-library(nloptr)
+library(nloptr); library(earth)
 ```

### 2022-04-20 R/estimate.R
fix input to family in partial_cv_preds in estimateG

@@ -134,7 +134,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
         # get partially cross-validated estimates
         if(partial_cv){
           gn_DeltaA_se <- partial_cv_preds(a_0 = NULL, fit_sl = fm_DeltaA,
-                                           family = family,
+                                           family = stats::binomial(),
                                            easy = TRUE)
         }
       } else if (!is.list(SL_g$DeltaA) & length(SL_g$DeltaA) == 1) {
@@ -192,7 +192,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
         if(partial_cv){
           gn_A_se <- vector(mode = "list", length = 2)        
           gn_A_se[[1]] <- partial_cv_preds(fit_sl = fm_A[[1]], a_0 = NULL,
-                                           family = family,
+                                           family = stats::binomial(),
                                            W = validW, include = trainDeltaA == 1,
                                            easy = all(trainDeltaA == 1))
           gn_A_se[[2]] <- 1 - gn_A_se[[1]]
@@ -229,7 +229,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
           tmp_pred <- tmp_fm$SL.pred
           if(partial_cv){
             tmp_pred_se <- partial_cv_preds(fit_sl = tmp_fm, a_0 = NULL,
-                                            family = family,
+                                            family = stats::binomial(),
                                             W = validW, include = include,
                                             easy = all(include))
           }
@@ -418,7 +418,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
             gn_DeltaY[[a_ct]] <- fm_DeltaY[[a_ct]]$SL.predict
             if(partial_cv){
               gn_DeltaY_se[[a_ct]] <- partial_cv_preds(fit_sl = fm_DeltaY[[a_ct]],
-                                                       family = family,
+                                                       family = stats::binomial(),
                                                        a_0 = NULL, include = include & include2, 
                                                        W = validW, easy = all(include & include2))
             }
@@ -450,7 +450,7 @@ estimateG <- function(A, W, DeltaY, DeltaA, SL_g, glm_g, a_0, tolg,
             )$pred
             if(partial_cv){
[…]

### 2022-06-23 paper/using_drtmle_obstudies.Rmd
attempt paper conversion

--- 2021-06-25 JSS/using_drtmle_jss.Rmd
@@
 ---
-documentclass: jss
-author:
-  - name: David Benkeser
-    affiliation: Emory University
-    address: |
-      | 1518 Clifton Road, NE
-      | Mailstop: 002-3AA
-      | Atlanta, Georgia, 30322, U.S.A.
-    email: \email{benkeser@emory.edu}
-    orcdi: 0000-0002-1019-8343
-  - name: Nima S. Hejazi
-    affiliation: University of California, Berkeley
-    address: |
-      | 2121 Berkeley Way,
-      | Berkeley, California, 94704, U.S.A.
-    email: \email{nhejazi@berkeley.edu}
-    orcdi: 0000-0002-7127-2789
-title:
-  formatted: "Doubly-Robust Inference in R using \\pkg{drtmle}"
-  plain: "Doubly-Robust Inference in R using drtmle"
-  short: "\\pkg{drtmle}: Doubly-robust Inference"
-keywords:
-  formatted: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
-  plain: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
-preamble: >
-  \usepackage{amsmath}
+documentclass: article
+classoption: [twoside,11pt]
+author: David Benkeser and Nima S. Hejazi
+title: "Doubly-Robust Inference in R using \\texttt{drtmle}"
+keywords: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
+output:
+  latex_document:
+    extra_dependencies:
+      obs_study_style: null
+      amsmath: null
+      mathtools: null
+      fancyvrb: null
+      inputenc: ["utf8"]
+bibliography: refs.bib
+header-includes: >
+  \providecommand{\tightlist}{%
+    \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+  \newcommand{\dataset}{{\cal D}}
+  \newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
+  \DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
+  \DefineVerbatimEnvironment{Soutput}{Verbatim}{}
+  \DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
+  \newenvironment{Schunk}{}{}
+  \DefineVerbatimEnvironment{Code}{Verbatim}{}
+  \DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
+  \DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
+  \newenvironment{CodeChunk}{}{}
+  \setkeys{Gin}{width=0.8\textwidth}
 abstract: >
@@
   cannot be said when at least one nuisance estimator is inconsistent.
-  \pkg{drtmle} implements doubly-robust confidence intervals and hypothesis
+  \texttt{drtmle} implements doubly-robust confidence intervals and hypothesis
   tests for targeted minimum loss-based estimates of the average treatment
@@
   average treatment effect.
-header-includes:
-  - \usepackage{mathtools}
-output: rticles::jss_article
-bibliography: refs
 ---
@@
 [@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
-consistent estimation of *both* the PS *and* the OR. \citet{vdl:2014:ijb}
+consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb
 proposed new TML estimators that have an asymptotic normal sampling distribution
@@
 OR, paving the way for inference that is doubly-robust.
-\citet{benkeser:2017:biometrika} proposed modifications of these estimators with
+@benkeser:2017:biometrika proposed modifications of these estimators with
 similar robustness properties and illustrated their practical performance.
@@
 involved, notably requiring iterative estimation of additional, complex nuisance
-parameters. The \pkg{drtmle} package was motivated by the need for a
+parameters. The \texttt{drtmle} package was motivated by the need for a
 user-friendly tool to implement these methods. The package implements the TML
-estimators of population-level effects proposed in \citet{vdl:2014:ijb} and
-\citet{benkeser:2017:biometrika}, as well as several extensions including
+estimators of population-level effects proposed in @vdl:2014:ijb and
+@benkeser:2017:biometrika, as well as several extensions including
 cross-validated targeted minimum loss-based estimation (CV-TMLE). Additionally,
-\pkg{drtmle} implements an IPTW estimator that overcomes shortcomings of
+\texttt{drtmle} implements an IPTW estimator that overcomes shortcomings of
 existing IPTW implementations with respect to statistical inference. In
@@
 about population-level effects. Both the TML and IPTW estimators implemented in
-the \pkg{drtmle} package are capable of estimating population-level effects for
+the \texttt{drtmle} package are capable of estimating population-level effects for
 discrete-valued treatments; moreover, the package provides convenient utilities
@@
 some of the key concepts underlying doubly-robust inference and illustrates
-usage of the \pkg{drtmle} package.
+usage of the \texttt{drtmle} package.
 
@@
 properties, TML estimators have been shown to outperform AIPTW estimators in
-finite samples [@porter:2011:ijb]. We refer readers to \citet{tlbook:2011} for
+finite samples [@porter:2011:ijb]. We refer readers to @tlbook:2011 for
 more about TMLE.
@@
 
-\citet{vdl:2014:ijb} demonstrated that the double robustness property of AIPTW
+@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
 and TML estimators does not extend to their normal limiting distribution if
@@
 estimate of either the OR or PS, and thus may be necessary to obtain minimally
-biased estimates of the marginal means of interest. \citet{vdl:2014:ijb} derived
+biased estimates of the marginal means of interest. @vdl:2014:ijb derived
 modified TML estimators that are doubly-robust not only with respect to
@@
 construct doubly-robust confidence intervals and hypothesis tests.
-\citet{benkeser:2017:biometrika} proposed simplifications of the
-\citet{vdl:2014:ijb} estimators and demonstrated the practical performance of
-these estimators via simulation.
+@benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
+estimators and demonstrated the practical performance of these estimators via
+simulation.
 
@@
 of whether the OR or PS are consistently estimated. Based on these
-reduced-dimension regressions, \citet{vdl:2014:ijb} showed that if estimates of
-the OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
+reduced-dimension regressions, @{vdl:2014:ijb showed that if estimates of the
+OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
 R-PS $g_{r,n}^{\star}$ satisfied
@@
 
-\citet{benkeser:2017:biometrika} demonstrated alternative equations that can be
+@benkeser:2017:biometrika demonstrated alternative equations that can be
 satisfied to achieve this form of double-robustness. In particular, these
@@
 both univariate regressions, while R-PS is a bivariate regression.
-\citet{benkeser:2017:biometrika} suggested that this may make R-PS1 and R-PS2
+@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
 easier to estimate consistently than R-PS. Further, they showed that if
@@
 then the GCOMP estimator based on this $\bar{Q}_n^{\star}$ also enjoys a
-doubly-robust limiting distribution. \citet{benkeser:2017:biometrika} provided a
+doubly-robust limiting distribution. @benkeser:2017:biometrika provided a
 TMLE algorithm that produced such estimates and provided closed-form,
@@
 studying the asymptotic properties of the GCOMP, AIPTW, and TML estimators
-(for discussion, see \citet{benkeser:2017:biometrika} Section 2.2). Recall that
+(for discussion, see @benkeser:2017:biometrika Section 2.2). Recall that
 a TML estimator achieves doubly-robust consistency by satisfying equation (4),
@@
 these two additional equations with the AIPTW estimator.
-\citet{benkeser:2017:biometrika} showed that, surprisingly, this AIPTW-style
-estimator does not achieve the same doubly-robust properties as the TML
-estimator. Thus, while AIPTW and TMLE are generally competitors for producing
-estimators doubly-robust with respect to consistency, for doubly-robust
-inference, TMLE ought to be preferred.
+@benkeser:2017:biometrika showed that, surprisingly, this AIPTW-style estimator
+does not achieve the same doubly-robust properties as the TML estimator. Thus,
+while AIPTW and TMLE are generally competitors for producing estimators
+doubly-robust with respect to consistency, for doubly-robust inference, TMLE
+ought to be preferred.
 
@@
 matrix for these estimates. The function implements either the
-estimators of \citet{vdl:2014:ijb} (if `reduction = "bivariate"`) or the
-improved estimators of \citet{benkeser:2017:biometrika}
+estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the
+improved estimators of @benkeser:2017:biometrika
 (`reduction = "univariate"`). The package includes utility functions for
@@
   5. compute the TML estimates and covariance estimates based on the final OR.
-We refer interested readers to \citet{benkeser:2017:biometrika} for a
+We refer interested readers to @benkeser:2017:biometrika for a
 theoretical derivation of how the iterative modification of initial OR and PS is
@@
 
-To understand the variable names available for R-OR, recall from (\ref{eqn:qbarr}) that the parameter
-involves a regression onto the estimated propensity score. Thus, the formula
-that is input via `glm_Qr` should assume data with a single column, `"gn"`.
-Similarly, the R-PS1 and R-PS2 that are computed for the estimators of
-\citet{benkeser:2017:biometrika} (the default estimators, computed whenever
-`reduction = "univariate"`) involve regressions onto the estimated outcome
-regression. Thus, the formula `glm_gr` should assume data with a single column
-named `"Qn"` if `reduction = "univariate"`. Instead, if the estimators of
-\citet{vdl:2014:ijb} are preferred (`reduction = "bivariate"`) the `glm_gr` 
-formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes data with two columns, `"Qn"` and `"gn"`.
+To understand the variable names available for R-OR, recall from
+(\ref{eqn:qbarr}) that the parameter involves a regression onto the estimated
+propensity score. Thus, the formula that is input via `glm_Qr` should assume
+data with a single column, `"gn"`.  Similarly, the R-PS1 and R-PS2 that are
+computed for the estimators of @benkeser:2017:biometrika (the default
+estimators, computed whenever `reduction = "univariate"`) involve regressions
+onto the estimated outcome regression. Thus, the formula `glm_gr` should assume
+data with a single column named `"Qn"` if `reduction = "univariate"`. Instead,
+if the estimators of @vdl:2014:ijb are preferred (`reduction = "bivariate"`) the
+`glm_gr` formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes data
+with two columns, `"Qn"` and `"gn"`.
 
@@
 
-Super learning is implemented in the R package \pkg{SuperLearner}
+Super learning is implemented in the R package \texttt{SuperLearner}
 [@SuperLearnerPackage] and `drtmle()` provides the option to utilize super
@@
 are passed to the `SL.library` argument of the `SuperLearner()` function of the
-\pkg{SuperLearner} package. We refer readers to the documentation of the
+\texttt{SuperLearner} package. We refer readers to the documentation of the
 `SuperLearner()` function for information on how to properly specify a super
@@
 relatively simple library. For the OR and PS, the library includes a main-terms
-generalized linear model (via the function `SL.glm` from \pkg{SuperLearner}) and
+generalized linear model (via the function `SL.glm` from \texttt{SuperLearner}) and
 an unadjusted generalized linear model (`SL.mean`). For the reduced-dimension
@@
 model (`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
-\pkg{earth} package [@earthPackage]).
+\texttt{earth} package [@earthPackage]).
 
@@
 
-The \pkg{drtmle} package also allows users to pass in their own functions for
+The \texttt{drtmle} package also allows users to pass in their own functions for
 estimating regression parameters via the `SL_` options. These functions must be
@@
 is identical to the formatting required for writing a `SuperLearner` wrapper
-(see `SuperLearner::write.SL.template()`). The \pkg{drtmle} package includes
+(see `SuperLearner::write.SL.template()`). The \texttt{drtmle} package includes
 `SL.npreg`, which is an example of how a user can specify a function for
 estimating regression parameters. This functions fits a kernel regression
-estimator using the \pkg{np} package [@npPackage]. Below we show the source code
+estimator using the \texttt{np} package [@npPackage]. Below we show the source code
 for this function.
@@
 
-In \pkg{drtmle}, cross-validated standard error estimates can be obtained by
+In \texttt{drtmle}, cross-validated standard error estimates can be obtained by
 setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1. For
@@
 Obtaining fully cross-validated standard errors can be fairly time-intensive
-when super learner is used to estimate the OR and PS. In \pkg{drtmle}, we give
+when super learner is used to estimate the OR and PS. In \texttt{drtmle}, we give
 an option to compute standard error estimates that are *partially
@@
 
-The \pkg{drtmle} package contains methods that compute doubly-robust confidence
+The \texttt{drtmle} package contains methods that compute doubly-robust confidence
 intervals. The `ci()` method computes confidence intervals about the estimated
@@
 In our examples so far, we have only considered binary treatments. However,
-\pkg{drtmle} is fully equipped to handle treatments with an arbitrary number of
+\texttt{drtmle} is fully equipped to handle treatments with an arbitrary number of
 discrete values as well. Suppose that $A$ assumes values in a set $\mathcal{A}$
@@
 outcomes, but have not been extended to multi-level outcomes. To ensure that
-users of \pkg{drtmle} have access to the large number of binary regression
+users of \texttt{drtmle} have access to the large number of binary regression
 techniques when estimating the PS, we have taken an alternative approach to
@@
 setting the option `use_future = TRUE`. Since, internally, all computations are
-carried out using futures (with the API provided by the \pkg{future} R package),
+carried out using futures (with the API provided by the \texttt{future} R package),
 setting this option amounts only to parallelizing with futures [@futurePackage]
@@
 If no adjustments are made prior to invoking the `drtmle()` function, futures
-are computed sequentially, as per the defaults of the \pkg{future} package. It
+are computed sequentially, as per the defaults of the \texttt{future} package. It
 is left to the user to specify appropriate system-specific future backends that
-may better suit their problem (e.g., using the \pkg{future.batchtools} package
+may better suit their problem (e.g., using the \texttt{future.batchtools} package
 [@futurebatchtoolsPackage] or even simply setting a parallelized
@@
 Interested readers are invited to consult the documentation of the
-\pkg{future.batchtools} package for further information.
+\texttt{future.batchtools} package for further information.
 
@@
 
-The \pkg{drtmle} package was designed to provide a user-friendly implementation
+The \texttt{drtmle} package was designed to provide a user-friendly implementation
 of the TMLE algorithm that facilitates doubly-robust inference about marginal
@@
 
-Planned extensions of the \pkg{drtmle} package include extensions to
+Planned extensions of the \texttt{drtmle} package include extensions to
 longitudinal settings involving treatments at multiple time points and inclusion
 of the collaborative targeted maximum likelihood estimator (C-TMLE) of
-\citet{vdl:2014:ijb} that also facilitates collaborative doubly-robust
-inference.
+@vdl:2014:ijb that also facilitates collaborative doubly-robust inference.
 
@@
 
-The \pkg{drtmle} package also allows users to pass in their own estimated OR and
+The \texttt{drtmle} package also allows users to pass in their own estimated OR and
 PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
@@
 estimators precluded the use of adaptive estimators of the PS. However,
-\citet{vdl:2014:ijb} proposed modified IPTW estimators that allow for adaptive
+@vdl:2014:ijb proposed modified IPTW estimators that allow for adaptive
 PS estimation. Similarly as above, this estimator is based on an additional
@@
 then the IPTW estimator based on $g_n^{\star}$ would have an asymptotic normal
-distribution with estimable variance. In fact, \citet{vdl:2014:ijb} proposed a
+distribution with estimable variance. In fact, @vdl:2014:ijb proposed a
 TMLE algorithm that mapped an initial PS estimate into an estimate that
@@
 Both this estimator and the TMLE-based IPTW estimator above are implemented in
-\pkg{drtmle}.
+\texttt{drtmle}.
 
@@
 for reminders on the various options for these procedures. By default, these
-methods are implemented for the `iptw_tmle()` estimator of \citet{vdl:2014:ijb}.
-We conjecture that this estimator will have improved finite-sample behavior
+methods are implemented for the `iptw_tmle()` estimator of @vdl:2014:ijb.  We
+conjecture that this estimator will have improved finite-sample behavior
 relative to `iptw_os()` (the alternative estimator described above), though
@@
 at $n^{-1/4}$ rate. In this sense, this estimator trades off robustness against
-achieving greater stability and efficiency. See \citet{ju2020robust} for further
+achieving greater stability and efficiency. See @ju2020robust for further
 discussion.
@@
 c(0,1)`, then the PS estimation data frame will have two columns called `Q0W`
-and `Q1W`. This may be relevant for users writing their own \pkg{SuperLearner}
+and `Q1W`. This may be relevant for users writing their own \texttt{SuperLearner}
 wrapper functions or those who make use of the `glm_g` setting.

### 2022-06-24 paper/preamble.tex
work on header

--- 2022-06-23 paper/using_drtmle_obstudies.Rmd
@@
----
-documentclass: article
-classoption: [twoside,11pt]
-author: David Benkeser and Nima S. Hejazi
-title: "Doubly-Robust Inference in R using \\texttt{drtmle}"
-keywords: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
-output:
-  latex_document:
-    extra_dependencies:
-      obs_study_style: null
-      amsmath: null
-      mathtools: null
-      fancyvrb: null
-      inputenc: ["utf8"]
-bibliography: refs.bib
-header-includes: >
+\documentclass[twoside,11pt]{article}
+
+% Any additional packages needed should be included after obs_study_style.
+% Note that obs_study_style.sty includes epsfig, amssymb, natbib and graphicx,
+% and defines many common macros, such as 'proof' and 'example'.
+%
+% It also sets the bibliographystyle to plainnat; for more information on
+% natbib citation styles, see the natbib documentation, a copy of which
+
+\usepackage[utf8]{inputenc}
+\usepackage{mathtools}
+\usepackage{amsmath}
+\usepackage{fancyvrb}
+\usepackage{obs_study_style}
+
+% Definitions of handy macros can go here
 \providecommand{\tightlist}{%
@@
 \setkeys{Gin}{width=0.8\textwidth}
-abstract: >
+
+% For papers submitted for review, just fill in author names
+% For accepted papers, heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}
+\heading{}{}{}{}{}{David Benkeser and Nima S.~Hejazi}
+
+% Short headings should be running head and authors last names
+
+\ShortHeadings{\texttt{drtmle}: Doubly-robust Inference}{Benkeser and Hejazi}
+\firstpageno{1}
+
+\begin{document}
+
+\title{Doubly-Robust Inference in R using \texttt{drtmle}}
+
+\author{
+  \name David Benkeser \email benkeser@emory.edu \\
+  \addr Department of Biostatistics \& Bioinformatics\\
+  Rollins School of Public Health\\
+  Emory University\\
+  1518 Clifton Road, N.E., Atlanta, GA 30322
+  \AND
+  \name Nima S.~Hejazi \email nhejazi@hsph.harvard.edu \\
+  \addr Department of Biostatistics\\
+  T.H.~Chan School of Public Health\\
+  Harvard University \\
+  677 Huntington Avenue, Boston, MA 02115
+}
+
+\maketitle
+
+\begin{abstract}
   Inverse probability of treatment weighted estimators and doubly-robust
@@
   average treatment effect.
----
+\end{abstract}
 
-```{r, setup, include=FALSE}
-options(prompt = 'R> ', continue = '+ ')
-knitr::opts_chunk$set(warning=FALSE, message=FALSE)
-renv::activate(here::here())
-library(nloptr); library(earth)
-```
-
-# Introduction
-
-In many fields, researchers are interested in assessing the population-level
-effect of a particular treatment on an outcome of interest. The treatment might
-correspond to a drug, a potentially harmful exposure, or a policy intervention.
-Often, the treatment cannot be randomized due to ethical or logistical reasons.
-This has sparked great interest in developing statistical methodology to
-estimate population-level effects from observational data. Estimation of these
-effects often requires accounting for putative confounding factors, that is,
-factors related to whether a participant receives treatment and to the
-participant's outcome. In many settings, researchers measure many potential
-confounders, for example using administrative databases or genetic sequencing
-technology. Due to the curse of dimensionality, common statistical estimation
-techniques are not feasible for such high-dimensional data without restrictive
-modeling assumptions. When these assumptions are violated, estimates of the
-population-level effect of a treatment may have large bias.
-
-This has led to increased interest in adopting adaptive estimation techniques from the
-machine learning literature to control for high-dimensional confounders when
-estimating treatment effects. However, facilitating proper inference (i.e.,
-confidence intervals and hypothesis tests) about estimates of effects based on
-adaptive estimators is challenging. In particular, the use of standard causal inference
-techniques, including both G-computation (GCOMP) estimators
-[@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
-estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when utilizing
-adaptive estimators. In particular, with rare exception, if cross-validation is used
-to select tuning parameters for the nuisance parameters, estimates of the population-level
-effect will generally be too biased to achieve standard $n^{1/2}$-rate asymptotics. Instead,
-*undersmoothing* must be used to select tuning parameters to control the bias 
-[@undersmoothed_hal]. However, this can prove quite challenging in practice and 
-few guidelines are available.
-
-A more natural framework for integrating adaptive methods into nuisance regression 
-estimation are doubly robust approaches, including augmented inverse probability of 
-treatment weighted (AIPTW) estimation [@robins:1994:jasa] and targeted minimum 
-loss-based (TML) estimation [@vdlrubin:2006:ijb]. The asymptotics of these estimators
-are characterized in part by the convergence rates of nuisance parameter estimates with
-respect to a proper norm. This fact makes the use of cross-validation more appealing for
-selection of tuning parameters for estimators of nuisance parameters, thereby providing 
-a more straightforward means of obtaining an estimate of the effect of interest that
-enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield closed form
-standard error estimates, which can serve as the basis for constructing confidence 
-intervals and hypothesis tests.
-
-As an intermediate step, the AIPTW and TML estimators require estimation of
-two key nuisance parameters: the probability of treatment as a function of
-confounders (the propensity score, hereafter referred to as the *PS*) and the
-average outcome as a function of confounders and treatment (the outcome
-regression, *OR*). In addition to their desirable inferential properties, AIPTW
-and TML estimators also enjoy a *double robustness* property. That is, the
-estimators are consistent for the population-level effect of interest if either
-of the PS or OR is consistently estimated. This property gives the AIPTW and
-TML estimators a natural appeal: inconsistency in the estimation of one
-nuisance parameter may be mitigated by the consistent estimation of the other.
-However, recent work has shown that the double robustness property does not
-necessarily extend to the inferential properties of these estimators when
-adaptive estimators of the PS and/or OR are used
-[@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
-consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb
-proposed new TML estimators that have an asymptotic normal sampling distribution
-with estimable variance under consistent estimation of *either* the PS *or* the
-OR, paving the way for inference that is doubly-robust.
-@benkeser:2017:biometrika proposed modifications of these estimators with
-similar robustness properties and illustrated their practical performance.
-
-The two proposed procedures that yield doubly-robust inference are quite
-involved, notably requiring iterative estimation of additional, complex nuisance
-parameters. The \texttt{drtmle} package was motivated by the need for a
-user-friendly tool to implement these methods. The package implements the TML
-estimators of population-level effects proposed in @vdl:2014:ijb and
-@benkeser:2017:biometrika, as well as several extensions including
-cross-validated targeted minimum loss-based estimation (CV-TMLE). Additionally,
-\texttt{drtmle} implements an IPTW estimator that overcomes shortcomings of
-existing IPTW implementations with respect to statistical inference. In
-particular, the estimator allows an adaptive estimator of the PS to be used in
-construction of the IPTW estimator, while yielding valid statistical inference
-about population-level effects. Both the TML and IPTW estimators implemented in
-the \texttt{drtmle} package are capable of estimating population-level effects for
-discrete-valued treatments; moreover, the package provides convenient utilities
-for constructing confidence intervals and hypothesis tests about contrasts of
-the mean outcome under different levels of treatment. This document explains
-some of the key concepts underlying doubly-robust inference and illustrates
-usage of the \texttt{drtmle} package.
-
-# Definition and estimators of the average treatment effect
-
-## Parameters of interest
-
-Suppose the observed data consist of $n$ independent and identically
-distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
-a vector of baseline covariates, $A \in \{0,1\}$ is a binary treatment, 
-$Y$ is an outcome, and $P_0$ is the true data-generating
-distribution. The parameters of interest are
-\begin{equation}
-  \psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
-  \mbox{for} \ a = 0,1 \ ,
-  \label{parameterDef}
-\end{equation}
-where $\bar{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
-outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
-function of the covariates. The value $\psi_0(a)$ represents the marginal
-(i.e., population-level) treatment-specific average outcome, hereafter referred
-to as the *marginal mean* under treatment $A=a$. The marginal effect of the
-treatment on the outcome can be assessed by comparing marginal means under
-different treatment levels. For example, the average treatment effect is often
-defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
-assumptions, these parameters have interpretations as mean counterfactual
-outcomes under the different treatments, with contrasts between these means
-defining causal effects.
-
-## G-computation estimators
-
-An estimator of $\psi_0(a)$ may be motivated directly by (1). Suppose we have
-available an estimate of the OR, $\bar{Q}_n$, and an estimate of the
-distribution function of baseline covariates, $Q_{W,n}$. An estimator of
-$\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
-\bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
-covariates is used so that this estimator can be equivalently expressed
-\begin{equation*}
-  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
-\end{equation*}
-This estimator is commonly referred to as the G-computation (GCOMP) estimator.
-Inference for GCOMP estimators based on parametric OR estimators may be
-facilitated through the nonparametric bootstrap. However, if adaptive
-approaches are used to estimate the OR, inference may not be available without
-further modification to the estimator (e.g., undersmoothing as discussed above).
-
-## Inverse probability of treatment estimators
-
-An alternative estimator of the treatment-specific population-level mean may be
-obtained by noting that (1) can be equivalently written as
-\begin{align}
-  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
-          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
-          W)\biggr\} \notag  \\
-          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A,
-          W)\biggr\} \notag  \\
-          &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
-          \label{iptwID}
-\end{align}
-We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
-score. Suppose we had available $g_n$, an estimate of the PS. Equation (2)
-motivates the estimator
-\begin{equation*}
-  \psi_{n,IPTW}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
-    \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
-\end{equation*}
-which is referred to as the inverse probability of treatment weighted (IPTW)
-estimator. Like GCOMP estimators, inference for IPTW estimators based on
-parametric PS estimators may be facilitated through the nonparametric bootstrap.
-If adaptive approaches are used to estimate the PS, inference may not be
-available without further modification of the propensity score estimator, as we
-discuss below.
-
-## Doubly-robust estimators
-
-The GCOMP and IPTW estimators suffer from two important shortcomings. The first
-is a lack of robustness. That is, validity of GCOMP estimators relies on
-consistent estimation of the OR, while validity of the IPTW relies on
-consistent estimation of the PS. In practice, we may not know which of the OR
-or PS is simpler to consistently estimate and thus may not know which of the
-two estimators to prefer. Furthermore, if adaptive estimators (e.g., based on
-machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
-estimators generally lack a basis for statistical inference.
-
-This motivates a new class of estimators that combine both the GCOMP and IPTW
-estimators. One example is the augmented IPTW (AIPTW) estimator
-\begin{equation*}
-\psi_{n,AIPTW}(a) = \psi_{n,IPTW}(a) + \frac{1}{n} \sum\limits_{i=1}^n \biggl\{
-1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \  \bar{Q}_n(a, W_i) \ ,
-\end{equation*}
-which adds an augmentation term to the IPTW estimator. Equivalently, the
-estimator can be viewed as adding an augmentation to the GCOMP estimator
-\begin{equation} \label{oneStepEst}
-\psi_{n,AIPTW}(a) = \psi_{n,GCOMP}(a) + \frac{1}{n} \sum\limits_{i=1}^n
-\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
-\end{equation}
-The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
-estimators. The estimator exhibits double-robustness in that it is consistent
-for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
-is consistent for the true PS. Furthermore, if both the OR and PS are
-consistently estimated and regularity conditions hold, one can establish a
-limiting normal distribution for the AIPTW estimator. Simple estimators of the
-variance of this limiting distribution can be derived that may be used to
-construct Wald-style confidence intervals and hypothesis tests.
-
-A more recent proposal to improve robustness and inferential properties of the
-GCOMP and IPTW estimators utilizes targeted minimum loss-based estimation
-(TMLE). TMLE is a general methodology that may be used to modify an estimator
-of a regression function in a manner that ensures that the modified estimator
-satisfies user-selected equations. In the present problem, one can show that if
-an estimator of the OR $\bar{Q}_n^{\star}$ satisfies
-\begin{equation}\label{meanEIFZero}
-  \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
-  \bar{Q}_n^{\star}(a, W_i)\} = 0 \ ,
-\end{equation}
-then the GCOMP estimator based on $\bar{Q}_n^{\star}$ will have the same
-asymptotic properties as the AIPTW estimator. That is, the estimator will be
-doubly-robust and, provided both the OR and PS are consistently estimated and
-regularity conditions are satisfied, will have a normal limiting distribution.
-TMLE provides a means of mapping an initial estimator of the OR into an
-estimator that satisfies (4). In addition to possessing the desirable asymptotic
-properties, TML estimators have been shown to outperform AIPTW estimators in
-finite samples [@porter:2011:ijb]. We refer readers to @tlbook:2011 for
-more about TMLE.
-
-# Doubly-robust inference
-
-@vdl:2014:ijb demonstrated that the double robustness property of AIPTW
-and TML estimators does not extend to their normal limiting distribution if
-adaptive estimators of the OR and PS are used to construct the estimators.
-However, in settings with high-dimensional and continuous-valued covariates,
-adaptive estimation is likely necessary in order to obtain a consistent
-estimate of either the OR or PS, and thus may be necessary to obtain minimally
-biased estimates of the marginal means of interest. @vdl:2014:ijb derived
-modified TML estimators that are doubly-robust not only with respect to
-consistency, but also with respect to asymptotic normality. Furthermore, this
-work provided closed-form, doubly-robust variance estimators that may be used to
-construct doubly-robust confidence intervals and hypothesis tests.
-@benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
-estimators and demonstrated the practical performance of these estimators via
-simulation.
-
-The key to the theory underlying these results is finding estimators of the OR
-and PS that simultaneously satisfy equation (4) in addition to two additional
-equations. These equations involve additional regression parameters that we
-refer to as the reduced outcome regression (R-OR) and the reduced propensity
-score (R-PS), defined respectively as
-\begin{align}
-  \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
-  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qbarr} \\
-  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
-  g_n(W)=g_n(w)\right\} \ . \label{eqn:gr_bivariate}
-\end{align}
-In words, the R-OR is the regression of the residual from the outcome
-regression onto the estimated PS amongst observations with $A = a$, while the
-R-PS is the propensity for treatment $A = a$ given the estimated OR and PS. We
-note that a key feature of these reduced-dimension regressions is that they are
-low-dimensional regardless of the dimension of $W$ and depend on neither the
-true OR nor the true PS. Thus, these regressions may be consistently estimated
-at reasonably fast rates irrespective of the dimension of $W$ and irrespective
-of whether the OR or PS are consistently estimated. Based on these
-reduced-dimension regressions, @{vdl:2014:ijb showed that if estimates of the
-OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
-R-PS $g_{r,n}^{\star}$ satisfied
-\begin{align}
-& \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
-\{Y_i - \bar{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
-&\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
-\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
-\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
-\frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
-\biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
-{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \bar{Q}_n^{\star}(a, W_i) \} = 0 \ ,
-\label{tmleQsolve}
-\end{align}
-then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly-robust not
-only with respect to consistency, but also with respect to its limiting
-distribution. The author additionally proposed a TMLE algorithm to map initial
-estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy these key
-equations.
-
-@benkeser:2017:biometrika demonstrated alternative equations that can be
-satisfied to achieve this form of double-robustness. In particular, these
-authors formulated an alternative version of the R-PS,
-\begin{align*}
-  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \bar{Q}_n(W) =
-  \bar{Q}_n(w) \} \ ,
-\end{align*}
-which we refer to as R-PS1. They also introduced an additional regression of a
-weighted residual from the PS on the OR,
-\begin{align*}
-  g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
-  {g_n(a \mid W)} \ \bigg\vert \ \bar{Q}_n(W) = \bar{Q}_n(w) \biggr\} \ ,
-\end{align*}
-which we refer to as R-PS2. The key feature of R-PS1 and R-PS2 is that they are
-both univariate regressions, while R-PS is a bivariate regression.
-@benkeser:2017:biometrika suggested that this may make R-PS1 and R-PS2
-easier to estimate consistently than R-PS. Further, they showed that if
-estimators of the OR, PS, and R-OR satisfy equations (5)-(6), and if,
-additionally, estimators of the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2
-$g_{r,n,2}^{\star}$ satisfy
-\begin{equation*}
-\frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_{r,n,1}^{\star}
-(a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
-\bar{Q}_n^{\star}(a,W_i)\} = 0 \ ,
-\end{equation*}
-then the GCOMP estimator based on this $\bar{Q}_n^{\star}$ also enjoys a
-doubly-robust limiting distribution. @benkeser:2017:biometrika provided a
-TMLE algorithm that produced such estimates and provided closed-form,
-doubly-robust variance estimates.
-
-\textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
-methods for constructing doubly-robust (with respect to consistency) estimators
-of our parameter of interest. Theoretically, these two frameworks are closely
-related and, practically, the two estimators are generally seen as competitors
-for estimating marginal means. In equation (3), we were able to generate a
-doubly-robust estimator by adding a term to the GCOMP estimator. This term is
-exactly the left-hand-side of the key equation (4), the key equation that is
-solved by the standard TMLE. This is no coincidence; this term is crucial in
-studying the asymptotic properties of the GCOMP, AIPTW, and TML estimators
-(for discussion, see @benkeser:2017:biometrika Section 2.2). Recall that
-a TML estimator achieves doubly-robust consistency by satisfying equation (4),
-while the AIPTW achieves doubly-robust consistency by combining the GCOMP
-estimator with the left-hand-side of this term. Now, we have argued that a TML
-estimator may achieve doubly-robust limiting behavior if it additionally
-satisfies equations (6) and (7). It is thus sensible to wonder whether an
-AIPTW-style estimator could be constructed by combining the left-hand-side of
-these two additional equations with the AIPTW estimator.
-@benkeser:2017:biometrika showed that, surprisingly, this AIPTW-style estimator
-does not achieve the same doubly-robust properties as the TML estimator. Thus,
-while AIPTW and TMLE are generally competitors for producing estimators
-doubly-robust with respect to consistency, for doubly-robust inference, TMLE
-ought to be preferred.
-
-# Implementation of doubly-robust inference
-
-The main function of the package is the eponymous `drtmle()` function. This
-function estimates the treatment-specific marginal mean for user-specified
-levels of a discrete-valued treatment and computes a doubly-robust covariance
-matrix for these estimates. The function implements either the
-estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the
-improved estimators of @benkeser:2017:biometrika
-(`reduction = "univariate"`). The package includes utility functions for
-combining these estimates into estimates of average treatment effects, as well
-as other contrasts, as discussed below.
-
-The main data arguments for `drtmle()` are `W, A, Y`, which, as above, represent
-the covariates, a discrete-valued treatment, and an outcome, respectively.
-Missing data are allowed in `A` and `Y`, as discussed below. Beyond the data
-arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
-and R-PS (or R-PS1, R-PS2). The package provides flexibility in how these
-regression parameters may be estimated. Once the user has specified the data
-arguments and how to estimate each regression parameter, the function proceeds
-as follows:
-  1. estimate the OR via the internal `estimateQ()` function;
-  2. estimate the PS via the internal `estimateG()` function;
-  3. estimate reduced-dimension regression via the internal `estimateQrn()` and
-     `estimategrn()` functions;
-  4. iteratively modify initial estimates of the OR and PS via the
-     `fluctuateQ()` and `fluctuateG()` functions until equations (5)-(7) are
-     approximately solved;
-  5. compute the TML estimates and covariance estimates based on the final OR.
-We refer interested readers to @benkeser:2017:biometrika for a
-theoretical derivation of how the iterative modification of initial OR and PS is
-performed.
-
-In addition to returning doubly-robust parameter and covariance estimates, the
-`drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
-modified AIPTW estimator discussed in the remark above. While doubly-robust
-inference based on these estimators is not available, the estimators are
-provided for comparison. In the following subsections, we look at various
-options for making calls to `drtmle()`.
-
-## Estimating regressions
-```{r, echo = FALSE, message = FALSE}
-library(drtmle)
-set.seed(1234)
-N <- 1e6
-W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
-Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
-Y0 <- rbinom(N, 1, plogis(W$W1))
-EY1 <- mean(Y1)
-EY0 <- mean(Y0)
-```
-While the estimators computed by `drtmle()` are consistent and asymptotically
-normal under inconsistent estimation of either the OR or PS, it is nevertheless
-advisable to strive for consistent estimation of *both* the OR and PS. If both
-of the regression parameters are consistently estimated, then the estimators
-computed by `drtmle()` achieve the nonparametric efficiency bound, resulting in
-narrower confidence intervals and more powerful hypothesis tests. In order for
-the estimates of the OR and PS to have the greatest chance of consistency, the
-`drtmle()` function facilitates estimation using adaptive estimation techniques.
-In fact, there are three ways to estimate each of the regression parameters:
-generalized linear models, super learning
-[@vdl:2007:statapp; @SuperLearnerPackage], and a user-specified estimation
-technique. We demonstrate each of these approaches in turn on a  simulated
-data set. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0,
-2)` and $\psi_0(1)=$ `r round(EY1, 2)`.
-
-```{r}
-set.seed(1234)
-n <- 200
-W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
-A <- rbinom(n, 1, plogis(W$W1 + W$W2))
-Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
-```
-
-### Using generalized linear models to estimate regressions
-
-Generalized linear models are perhaps the most popular means of estimating
-regression functions. Because of this legacy, these estimators have been
-included as an option for estimating the regression parameters. The user
-specifies the right-hand-side of a regression formula as a character in the
-`glm_g`, `glm_Q`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS, R-OR,
-and R-PS. These regressions are fit by calling the `glm()` function in R [@R].
-Thus, the model specification should be able to be parsed as a valid formula,
-as required by `glm()`. For the PS formula specified by `glm_g`, `colnames(W)`
-are available to use in the formula. For the OR (`glm_Q`) available names
-of variables to be used in the formula depends on the `stratify` option.
-If `stratify = FALSE`, a single OR is fit to estimate $\bar{Q}_0(A,W)$. In this
-case, the OR regression is fit by passing a `data.frame` that contains `W` and
-`A` into the `data` argument of `glm()`. Thus, both `colnames(W)` and `"A"` may
-appear in the formula specified in `glm_Q`. If `stratify = TRUE`, regressions
-are fit stratifying on `A` to separately estimate $\bar{Q}_0(1,W)$ and
-$\bar{Q}_0(0,W)$. In this case, the `glm_Q` formula may not include `"A"`.
-
-To understand the variable names available for R-OR, recall from
-(\ref{eqn:qbarr}) that the parameter involves a regression onto the estimated
-propensity score. Thus, the formula that is input via `glm_Qr` should assume
-data with a single column, `"gn"`.  Similarly, the R-PS1 and R-PS2 that are
-computed for the estimators of @benkeser:2017:biometrika (the default
-estimators, computed whenever `reduction = "univariate"`) involve regressions
-onto the estimated outcome regression. Thus, the formula `glm_gr` should assume
-data with a single column named `"Qn"` if `reduction = "univariate"`. Instead,
-if the estimators of @vdl:2014:ijb are preferred (`reduction = "bivariate"`) the
-`glm_gr` formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes data
-with two columns, `"Qn"` and `"gn"`.
-
-The code below illustrates how generalized linear models can be used to
-estimate the regression parameters. In this call to `drtmle()`, we specify
-`family = binomial()` so that logistic regression is used to estimate the OR.
-We illustrate non-stratified regression for both the univariate (default) and
-bivariate reductions below.
-
-```{r}
-glm_fit_uni <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE)
-glm_fit_uni
-
-glm_fit_biv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                      glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-                      glm_gr = "Qn + gn", glm_Qr = "gn", stratify = FALSE,
-                      reduction = "bivariate")
-glm_fit_biv
-```
-
-### Using super learner to estimate regressions
-
-```{r, echo = FALSE, message = FALSE}
-library(SuperLearner)
-```
-Super learning is a loss-based ensemble machine learning technique
-[@vdl:2007:statapp] that generalizes stacking and stacked regression
-\citep{wolpert:1992:nnet,breiman:1996:ml}. The user specifies many potential
-candidate regression estimators, referred to as a *library* of candidate
-estimators, and the super learner algorithm estimates the convex combination of
-regression estimators that minimizes the $V$-fold cross-validated risk based on
-a user-selected loss function [@dudoit2005asymptotics].  Oracle inequalities
-show that the super learner estimate performs essentially as well as the unknown
-best convex combination of the candidate estimators [@vdvaart:2006:statdec].
-Thus, the methodology may be seen as an asymptotically optimal way of performing
-estimator selection. Practically, the super learner provides the best chance
-for obtaining a consistent estimator for both the OR and PS, and thereby an
-efficient estimator of the marginal means of interest.
-
-Super learning is implemented in the R package \texttt{SuperLearner}
-[@SuperLearnerPackage] and `drtmle()` provides the option to utilize super
-learning for estimation of the OR, PS, and reduced-dimension regressions via the
-`SL_Q`, `SL_g`, `SL_Qr`, and `SL_gr` options. When calling `drtmle()` either the
-generalized linear model options for regression modeling or the super learner
-options should be invoked. If both are called, the function will automatically
-opt to using the super learner option. The inputs to the super learner options
-are passed to the `SL.library` argument of the `SuperLearner()` function of the
-\texttt{SuperLearner} package. We refer readers to the documentation of the
-`SuperLearner()` function for information on how to properly specify a super
-learner library. In the code below, we illustrate a call to `drtmle()` with a
-relatively simple library. For the OR and PS, the library includes a main-terms
-generalized linear model (via the function `SL.glm` from \texttt{SuperLearner}) and
-an unadjusted generalized linear model (`SL.mean`). For the reduced-dimension
-regressions we use a super learner library with a main-terms generalized linear
-model (`SL.glm`) and adaptive regression splines (`SL.earth`, which utilizes the
-\texttt{earth} package [@earthPackage]).
-
-```{r, warning = FALSE, message = FALSE, cache = TRUE}
-set.seed(1234)
-sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                 SL_g = c("SL.glm","SL.mean"),
-                 SL_Q = c("SL.glm","SL.mean"),
-                 SL_gr = c("SL.glm", "SL.earth"),
-                 SL_Qr = c("SL.glm", "SL.earth"),
-                 stratify = FALSE)
-sl_fit
-```
-
-The \texttt{drtmle} package also allows users to pass in their own functions for
-estimating regression parameters via the `SL_` options. These functions must be
-formatted properly for compatibility with `drtmle()`. The necessary formatting
-is identical to the formatting required for writing a `SuperLearner` wrapper
-(see `SuperLearner::write.SL.template()`). The \texttt{drtmle} package includes
-`SL.npreg`, which is an example of how a user can specify a function for
-estimating regression parameters. This functions fits a kernel regression
-estimator using the \texttt{np} package [@npPackage]. Below we show the source code
-for this function.
-
-```{r}
-SL.npreg
-```
-
-In the above code, we see that a user-specified function should have options
-`Y, X, newX, family, obsWeights, ...`, along with other options specific to the
-function. The option `Y` corresponds to the outcome of the regression and `X`
-to the variables onto which the outcome is regressed. The option `newX` should
-be of the same class and dimension as `X` and is meant to contain new values of
-variables at which to evaluate the regression fit. The `family` and
-`obsWeights` options are not strictly necessary (indeed, they are not used by
-`SL.npreg`), though these may be useful to include if one wishes to utilize the
-function as part of a super learner library. The `SL.npreg` function proceeds by
-checking whether there is sufficient variation in `Y` to justify fitting the
-kernel regression (as specified by option `rangeThresh`), and if so, fits a
-kernel regression with cross-validated bandwidth selection via the `npregbw()`
-function. The output is then formatted in a specific way so that `drtmle()` is
-able to retrieve the appropriate objects from the fitted regression. In
-particular the output should be a list with named objects `pred` and `fit`; the
-former including predictions made from the fitted regression on `newX`, the
-latter including any information that may be needed to predict new values based
-on the fitted object. A class is assigned to the `fit` object, so that an S3
-`predict()` method may later be used to predict new values based on the fitted
-regression object. The user must also specify this `predict()` method. Use
-`drtmle:::predict.SL.npreg()` to view the `predict()` method for
-`SL.npreg`.
-
-In the code below, we demonstrate a call to `drtmle()` with `SL.npreg` used to
-estimate each nuisance parameter. Here, we also illustrate the use of `stratify
-= TRUE`, which fits a separate kernel regression of $Y$ onto $W$ in
-observations with $A = 1$ and $A = 0$.
-
-```{r, cache = TRUE, message = FALSE}
-npreg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                    SL_g = "SL.npreg", SL_Q = "SL.npreg",
-                    SL_gr = "SL.npreg", SL_Qr = "SL.npreg",
-                    stratify = TRUE)
-```
-
-In the Appendix, we describe an additional means of using `drtmle()` with 
-OR and PS estimates that have been obtained externally.
-
-### Reducing dependence on random seed
-
-Some users may find it unsettling that when applying the super learner in real
-data analysis the results are dependent on the random seed that is set. To
-remove this dependence, one can average results over repeated super learner
-runs; more repeats will lead to less dependence on the random seed. There are
-generally two approaches that can be used: (i) averaging on the scale of the
-nuisance parameters and (ii) averaging on the scale of the point estimates.
-
-For (i), suppose that we have $n_{SL}$ estimated super learners for the OR,
-$\bar{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS $g_{n,j}, j = 1, \dots,
-n_{SL}$. We can take the average over these fits,
-\begin{equation*}
-  \bar{Q}_n = \frac{1}{n_{SL}} \sum_{j = 1}^{n_{SL}} \bar{Q}_{n,j}
-\end{equation*}
-as our estimate of the OR (similarly for the PS). These averaged estimates are
-then used to compute the AIPTW or TMLE, as described above.
-
-For (ii), suppose for example that we have $n_{SL}$ TML (similarly, AIPTW)
-estimates $\psi_{n,j}(1)$ of $\psi_0(1)$, each obtained based on one (or more)
-super learners of the OR and/or PS. We can take the average over these point
-estimates,
-\begin{equation*}
-  \psi_n(1) = \frac{1}{n_{SL}} \sum_{j=1}^{n_{SL}} \psi_{n,j}(1)
-\end{equation*}
-as our estimate of $\psi_{0}(1)$.
-
-This feature is implemented via the `n_SL` and `avg_over` options, where the
-former specifies the number of super learners to repeat and the latter specifies
-the scale(s) to average over. For example, if `avg_over = "drtmle"` and `n_SL =
-2`, then two point estimates are computed, each based on a single super learner,
-and averaged to obtain a final estimate. If instead `avg_over = "SL"` and `n_SL
-= 2`, then two super learners are fit and averaged before a final point estimate
-is obtained. Finally, if `avg_over = c("drtmle", "SL")` and `n_SL = 2` then
-averaging on *both* scales is performed. That is, the final estimate is the
-average of two point estimates, each obtained using OR and PS estimates that are
-averaged over two super learners.
-
-Not surprisingly, averaging over multiple super learners can add considerable
-computational expense to calls to `drtmle()`. The package is currently
-parallelized at the level of individual calls to `SuperLearner()`. With
-parallelization over enough cores, repeated super learning should not add too
-much to the computational burden relative to a single call to `SuperLearner()`.
-
-## Standard error estimation
-
-This section contains some more advanced mathematical material and may be
-skipped over by users interested only in implementation. For simplicity, we
-focus this discussion on the standard TML and AIPTW estimators. The ideas
-generalize immediately to the DR-TML estimator, but the formulas are more
-complex in this latter case.
-
-Reported standard error estimates are based on an estimate of the asymptotic
-variance of the estimators. This variance can be characterized by influence
-function of the estimators. For example, let $\psi_{n}(1)$ be an AIPTW estimate
-of $\psi_0(1)$. Under regularity conditions,
-\begin{equation*}
-  \psi_{n}(1) - \psi_0(1) = \frac{1}{n} \sum_{i=1}^n
-    D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
-\end{equation*}
-where
-\begin{equation*}
-  D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}
-  {g_0(1 \mid W_i)} \{ Y_i - \bar{Q}_0(1, W_i)\} + \bar{Q}_0(1, W_i) -
-  \int \bar{Q}_0(1, w) dQ_{W,0}(w) \ .
-\end{equation*}
-Thus, by the central limit theorem, $n^{1/2} \psi_{n}(1)$ converges in
-distribution to a normal distribution with mean $\psi_0(1)$ and variance
-\begin{equation*}
-  \sigma^2_0 = E_0( [ D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O) -
-  E_0\{D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
-\end{equation*}
-An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical
-variance of $D_i = D^{\star}(\bar{Q}_n, g_n, Q_{W,n})(O_i)$, that is,
-\begin{equation*}
-  \sigma^2_n = \frac{1}{n} \sum_{i=1}^n \left\{D_i - \frac{1}{n} \sum_{j=1}^n
-  D_j \right\}^2 \ .
-\end{equation*}
-Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n /
-n^{1/2}$. Conveniently, this idea generalizes naturally to estimating the
-asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.
-
-### Standard errors for TMLE
-
-For TML estimators, it is standard practice to construct $D_i$ using the
-*targeted* version of the OR (denoted by $\bar{Q}_n^{\star}$ above). Recall that
-such estimates are obtained by modifying an initial OR estimate so as to ensure
-a particular (set of) equation(s) is (are) solved.  However, simulations have
-shown that better standard error estimates for TMLE may be achieved by instead
-using the *initial* OR estimate in the computation of standard errors, as
-described in the preceding subsection. To this end, The `drtmle()` function
-includes the `targeted_se` option.
-
-```{r, show_targeted_se, cache = TRUE}
-glm_fit_uni_nontargeted_se <-
-  drtmle(W = W, A = A, Y = Y, family = binomial(),
-         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-         glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE,
-         targeted_se = FALSE)
-# compare to original call
-list(
-  glm_fit_uni$drtmle$cov, # based on targeted OR
-  glm_fit_uni_nontargeted_se$drtmle$cov # untargeted OR
-)
-```
-
-### Cross-validated standard errors
-
-Simulation studies have revealed that, when employing machine learning to
-estimate the OR and/or PS, the estimated standard errors are often *too small*,
-in small to moderate-sized samples. Problematically, this leads directly to
-under-coverage of confidence intervals and possibly inflated Type-I error rates
-of hypothesis tests based on these standard error estimates. This should not be
-too surprising, as the phenomenon is related to the form of the standard error
-estimates. In particular, such estimates include a term that resembles a
-(weighted) mean squared error (MSE) of the OR. Since machine learning algorithms
-often set out to minimize empirical MSE as part of their training process, it
-follows that the empirical MSE often ends up biased towards zero relative to the
-true MSE. To more accurately estimate the MSE of a machine learning algorithm,
-cross-validation can be used. The same sample-splitting idea can be used to
-obtain more conservative standard error estimates.
-
-Suppose we use $V$-fold cross-validation to obtain $V$ training-sample-specific
-estimates of the OR, $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v},
-v = 1, \dots, V$. For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the
-estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the
-validation fold. That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of
-$Y_i$. We define $g_{n,i}^{cv}$ similarly. Using these estimates, we can define
-\begin{equation*}
- D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} +
- \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1)
-\end{equation*}
-and compute an estimate of $\sigma^2_0$ using
-\begin{equation*}
-  \sigma^2_{n,cv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{cv} -
-  \frac{1}{n} \sum_{j=1}^n D_j^{cv} \right\}^2 \ .
-\end{equation*}
-
-In \texttt{drtmle}, cross-validated standard error estimates can be obtained by
-setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1. For
-example,
-
-```{r, fit_w_full_cv, warning = FALSE, cache = TRUE}
-glm_fit_uni_cv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-                         glm_gr = "Qn", glm_Qr = "gn",
-                         se_cv = "full", se_cvFolds = 2, stratify = FALSE,
-                         targeted_se = FALSE) # needed for this to work
-# compare variance to previously obtained
-list(
-  `No CV` = glm_fit_uni$drtmle$cov, 
-  `CV` = glm_fit_uni_cv$drtmle$cov
-)
-```
-
-### Partially cross-validated standard errors
-
-Obtaining fully cross-validated standard errors can be fairly time-intensive
-when super learner is used to estimate the OR and PS. In \texttt{drtmle}, we give
-an option to compute standard error estimates that are *partially
-cross-validated* and do not require any additional fitting beyond a standard
-super learner. Here, we describe briefly how these estimates are obtained and
-show calls to `drtmle()` for obtaining such estimates.
-
-Suppose we have $K$ candidate regressions in our super learner library for both
-the OR and PS (in practice, the number could be different for the OR and for the
-PS). Recall that a super learner estimate of, for example, $\bar{Q}_0(a, w)$ is
-obtained as a weighted combination of estimates from each candidate regression.
-Let $\bar{Q}_{n,k}$ denote the estimate of $\bar{Q}_{0}$ obtained by training
-algorithm $k$ using the full data set. The super learner estimate of
-$\bar{Q}_{0}(a,w)$ is
-\begin{equation*}
-  \bar{Q}_{n,SL}(a, w) = \sum_{k = 1}^K \alpha_{n,k} \bar{Q}_{n,k}(a, w) \ ,
-\end{equation*}
-where $\alpha_n = (\alpha_{n,1}, \dots, \alpha_{n,K})$ is a weight vector that
-is selected by minimizing a cross-validated risk criterion. That is, each of the
-$K$ algorithms is trained according to a $V$-fold cross-validation scheme,
-resulting in $V + 1$ fits of each algorithm (one in each of $V$ training folds,
-plus one obtained by fitting the algorithm using the full data). Let
-$\bar{Q}_{n,k,v}$ denote the $k$\textsuperscript{th} algorithm trained using the
-$v$\textsuperscript{th} training fold. For $i = 1, \dots, n$, let
-$\bar{Q}_{n,k,i}^{cv}(1)$ denote the estimate of $\bar{Q}_0(1, W_i)$ implied by
-the fitted algorithm $\bar{Q}_{n,k,v}$, which was obtained when observation $i$
-was in the validation sample. That is, for observations with $A_i = 1$,
-$\bar{Q}_{n,k,i}^{cv}(1)$ is an out-of-sample prediction of $Y_i$ based on the
-$k$\textsuperscript{th} algorithm. Now, we can construct a *partially
-cross-validated* super learner prediction as follows. For $i = 1, \dots, n$ let
-\begin{equation*}
-  \bar{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k}
-    \bar{Q}_{n,k,i}^{cv}(1)
-\end{equation*}
-be the estimate obtained by ensembling the training-set-specific algorithm fits
-based on $\alpha_n$. We can use these estimates to produce our *partially
-cross-validated* standard error estimate as follows. First, define
-\begin{equation*}
-  D_i^{pcv} = \frac{I(A_i = 1)}{g_{n,SL,i}^{cv}} \{ Y_i -
-    \bar{Q}_{n,SL,i}^{cv}(1)\} + \bar{Q}_{n,SL,i}^{cv}(1) -
-    \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,SL,i}^{cv}(1),
-\end{equation*}
-then, compute an estimate of $\sigma^2_0$ using
-\begin{equation*}
-  \sigma^2_{n,pcv} = \frac{1}{n} \sum_{i=1}^n \left\{ D_i^{pcv} -
-  \frac{1}{n} \sum_{j=1}^n D_j^{pcv} \right\}^2 \ .
-\end{equation*}
-
-Note that is not a properly cross-validated estimate, since observation $i$ is
-used to compute the weight vector $\alpha_n$. However, such estimates can be
-obtained based on the output of a single round of super learning, thereby
-providing improved standard error estimates at minimal additional computational
-cost.
-
-Below, we demonstrate how to obtain such estimates using `drtmle()`.
-
-```{r, demo_pcv, cache = TRUE, warning = FALSE}
-set.seed(1234)
-sl_fit_pcv <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                     SL_g = c("SL.glm","SL.mean"),
-                     SL_Q = c("SL.glm","SL.mean"),
-                     SL_gr = c("SL.glm", "SL.earth"),
-                     SL_Qr = c("SL.glm", "SL.earth"),
-                     stratify = FALSE,
-                     se_cv = "partial")
-# compare estimates
-# pt estimates should be same
-# variance should be larger for pcv
-list(
-  `No-CV` = sl_fit$cov, # no cv
-  `Partial CV` = sl_fit_pcv$cov # partial cv
-)
-```
-
-## Confidence intervals
-
-The \texttt{drtmle} package contains methods that compute doubly-robust confidence
-intervals. The `ci()` method computes confidence intervals about the estimated
-marginal means in a fitted object of S3 class `"drtmle"`, in addition to
-contrasts between those means. By default, the method gives confidence intervals
-about each mean.
-
-```{r}
-ci(npreg_fit)
-```
-
-Alternatively the `contrast` option allows users to specify a linear
-combination of marginal means, which can be used to estimate the average
-treatment effect, as we show below.
-
-```{r}
-ci(npreg_fit, contrast = c(1, -1))
-```
-
-The `contrast` option also accepts as input a list of functions to be used to
-compute a confidence interval for a user-specified contrast. For example, we
-might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
-constructing Wald-style confidence intervals for ratios, it is common to
-compute the interval on the logarithmic scale and subsequently back-transform.
-That is, the confidence interval takes the form
-\begin{equation*}
-  \exp \biggl[\log \biggl\{ \frac{\psi_n(1)}{\psi_n(0)} \biggr\} \pm
-  z_{1-\alpha/2} \sigma^{\log}_n \biggr] \ ,
-\end{equation*}
-where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
-the delta method, an estimate of this standard error is given by
-\begin{equation*}
-  \sigma^{\log}_n = \nabla g(\psi_n)^T \Sigma_n \nabla g(\psi_n),
-\end{equation*}
-where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
-`drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
-This confidence interval may be obtained as follows.
-
-```{r}
-riskRatio <- list(f = function(eff){ log(eff) },
-                  f_inv = function(eff){ exp(eff) },
-                  h = function(est){ est[1] / est[2] },
-                  fh_grad =  function(est){ c(1 / est[1], -1 / est[2]) })
-ci(npreg_fit, contrast = riskRatio)
-```
-
-More generally, this manner of inputting the `contrast` option to the `ci()`
-method can be used to compute confidence intervals of the form
-\begin{equation*}
-  f^{-1}\bigl\{ f(h(\psi_n)) \pm z_{1-\alpha/2} \nabla f(h(\psi_n))^T \Sigma_n
-  \nabla f(h(\psi_n)) \bigr\} \ ,
-\end{equation*}
-where $f$ (`contrast$f()`) is the transformation of the confidence interval,
-$f^{-1}$ (`contrast$f_inv()`) is the back-transformation of the interval, $h$
-(`contrast$h()`) is the contrast of the marginal means, and $\nabla f(h(\psi_n))$
-(`contrast$fh_grad()`) defines the gradient of the transformed contrast at the
-estimated marginal means.
-
-We note that this method of computing confidence intervals can also be used to
-obtain a transformed confidence interval for a particular marginal mean. For
-example, recall that $Y$ is binary in our running example. Thus, we may wish to
-enforce that our confidence interval be computed on a scale ensuring that the
-estimated confidence limits fall within the unit interval $[0, 1]$. To that end,
-we can utilize an interval of the form:
-\begin{equation*}
-  \text{expit} \biggl[ \log\biggl\{ \frac{\psi_n(1)}{1 - \psi_n(1)} \biggr\}
-  \pm z_{1-\alpha/2} \sigma^{\text{logit}}_n \biggr] \ .
-\end{equation*}
-This confidence interval may be implemented as follows.
-
-```{r}
-logitMean <- list(f = function(eff){ qlogis(eff) },
-                  f_inv = function(eff){ plogis(eff) },
-                  h = function(est){ est[1] },
-                  fh_grad = function(est){ c(1 / (est[1] - est[1]^2), 0) })
-ci(npreg_fit, contrast = logitMean)
-```
-
-## Hypothesis tests
-
-Doubly-robust Wald-style hypothesis tests may be implemented using the
-`wald_test()` method. As with confidence intervals, there are three options for
-testing. First, we may test the null hypothesis that the marginal means equal a
-fixed value (or values). Below, we test the null hypothesis that $\psi_0(1) =
-0.5$ and that $\psi_0(0) = 0.6$ by inputting a vector to the `null` option.
-
-```{r}
-wald_test(npreg_fit, null = c(0.5, 0.6))
-```
-
-As with the `ci()` method, the `wald_test()` method allows users to test linear
-combinations of marginal means by inputting a vector of ones and negative ones
-in the `contrast` option. We can use this to test the null hypothesis of no
-average treatment effect or to test the null hypothesis that the average
-treatment effect equals a given reference value, e.g., $0.05$, by specifying
-the `null` option.
-
-```{r}
-wald_test(npreg_fit, contrast = c(1, -1))
-wald_test(npreg_fit, contrast = c(1, -1), null = 0.05)
-```
-
-The `wald_test()` method also allows for user-specified contrasts to be tested
-using syntax similar to the `ci()` method. The only difference syntactically is
-that it is not strictly necessary to specify `f_inv()`, as the hypothesis test is
-performed on the transformed scale. That is, we can generally test the null
-hypothesis that $f(h(\psi_0))$ equals $f_0$ (the function $f$ applied to the
-value passed to `null`) using the Wald statistic
-\begin{equation*}
-  Z_n \coloneqq \frac{\{f(h(\psi_n)) - f_0\}}{\nabla f(h(\psi_n))^T \Sigma_n
-  \nabla f(h(\psi_n))} \ .
-\end{equation*}
-We can use the `riskRatio` contrast defined previously to test the null
-hypothesis that $\psi_0(1) / \psi_0(0) = 1$.
-
-```{r}
-wald_test(npreg_fit, contrast = riskRatio, null = 1)
-```
-
-## Missing data
-
-The `drtmle()` function supports missing data in `A` and `Y`. Such missing data
-are common in practice and, if ignored, can bias estimates of the marginal
-means of interest. The estimators previously discussed can be modified to allow
-such missingness processes to depend on observed covariates. Let $\Delta_A$ and
-$\Delta_Y$ be indicators that $A$ and $Y$ are observed, respectively. We can
-redefine the OR to be $\bar{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
-A = a, \Delta_A = 1, \Delta_Y = 1, W = w)$ and similarly redefine the PS to be
-\begin{align}
-  g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
-  \notag \\
-  &= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
-  \label{modPS} \\
-  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w) \
-   \notag.
-\end{align}
-
-To demonstrate, we artificially introduce missing values to `A` and `Y` in our
-running example.
-
-```{r}
-DeltaA <- rbinom(n, 1, plogis(2 + W$W1))
-DeltaY <- rbinom(n, 1, plogis(2 + W$W2 + A))
-A[DeltaA == 0] <- NA
-Y[DeltaY == 0] <- NA
-```
-
-The only modification of the call to `drtmle()` is in how the PS estimation is
-performed. The options `glm_g` and `SL_g` are still used to fit generalized
-linear models or a super learner, respectively. However, the code allows for
-separate regression fits for each of the three components of the PS in (8). To
-perform separate generalized linear model fits for each of the three components,
-`glm_g` should be a named list with entries `"DeltaA"`, `"A"`, and `"DeltaY"`.
-Respectively, these should provide a formula expression for the regression of
-$\Delta_A$ on $W$, $A$ on $W$ in observations with $\Delta_A = 1$, and of
-$\Delta_Y$ on $W$ and $A$ in observations with $\Delta_A = \Delta_Y = 1$ (if
-`stratify = FALSE`) _or_ of $\Delta_Y$ on $W$ to be fit in observations with
-$\Delta_A = \Delta_Y = 1$ and $A = a$ for each $a$ specified in option `a_0` (if
-`stratify = TRUE`). If only a single formula is passed to `glm_g`, it is used
-for each of the three regressions. We illustrate a call to `drtmle()` with
-missing data and generalized linear model fits below.
-
-```{r, cache = TRUE}
-glm_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
-                       glm_g = list(DeltaA = "W1 + W2", A = "W1 + W2",
-                                    DeltaY = "W1 + W2 + A"),
-                       glm_Q = "W1 + W2*A", glm_Qr = "gn",
-                       glm_gr = "Qn", family = binomial())
-glm_fit_wNAs
-```
-
-It is possible to mix generalized linear models and super learning when fitting
-each of the components of the PS. In the example below, we use the wrapper
-function `SL.glm`, which fits a main terms logistic regression, to fit the
-regression of $\Delta_A$ on $W$; use `SL.npreg` to fit the regression of $A$ on
-$W$; and use a super learner with library `SL.glm`, `SL.mean`, and `SL.earth` to
-fit the regression of $\Delta_Y$ on $W$ and $A$.
-
-```{r, cache = TRUE}
-mixed_fit_wNAs <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
-                         SL_g = list(DeltaA = "SL.glm", A = "SL.npreg",
-                         DeltaY = c("SL.glm", "SL.mean", "SL.earth")),
-                         glm_Q = "W1 + W2*A", glm_Qr = "gn",
-                         glm_gr = "Qn", family = binomial())
-```
-
-We can also fit the PS regressions externally and pass in via `gn`, though this
-process is slightly more complicated than when no missing data are present. The
-basic idea is to fit one regression for (i) the indicator of missing $A$, (ii)
-$A$ itself, and (iii) the indicator of the outcome being missing. The three PS's
-are then multiplied together and an appropriately ordered list is constructed as
-before. We illustrate with the following example.
-
-```{r, cache = TRUE}
-# first regress indicator of missing A on W
-fit_DeltaA <- glm(DeltaA ~ . , data = W, family = binomial())
-
-# get estimated propensity for observing A
-ps_DeltaA <- predict(fit_DeltaA, type = "response")
-
-# now regress A on W | DeltaA = 1
-fit_A <- glm(A[DeltaA == 1] ~ . , data = W[DeltaA == 1, , drop = FALSE],
-             family = binomial())
-
-# get estimated propensity for receiving A = 1
-ps_A1 <- predict(fit_A, newdata = W, type = "response")
-# propensity for receiving A = 0
-ps_A0 <- 1 - ps_A1
-
-# now regress DeltaY on A + W | DeltaA = 1. Here we are pooling over
-# values of A (i.e., this is what drtmle does if stratify = FALSE).
-# We could also fit two regressions, one of DeltaY ~ W | DeltaA = 1, A = 1
-# and one of DeltaY ~ W | DeltaA = 1, A = 0 (this is what drtmle does if
-# stratify = TRUE).
-fit_DeltaY <- glm(DeltaY[DeltaA == 1] ~ . ,
-                  data = data.frame(A = A, W)[DeltaA == 1, ],
-                  family = binomial())
-
-# get estimated propensity for observing outcome if A = 1
-ps_DeltaY_A1 <- predict(fit_DeltaY, newdata = data.frame(A = 1, W),
-                        type = "response")
-
-# get estimated propensity for observing outcome if A = 0
-ps_DeltaY_A0 <- predict(fit_DeltaY, newdata = data.frame(A = 0, W),
-                        type = "response")
-
-# now combine all results into a single propensity score
-gn <- list(
-  # propensity for DeltaA = 1, A = 1, DeltaY = 1
-  ps_DeltaA * ps_A1 * ps_DeltaY_A1,
-  # propensity for DeltaA = 1, A = 0, DeltaY = 1
-  ps_DeltaA * ps_A0 * ps_DeltaY_A0
-)
-
-# pass in this gn to drtmle
-out_fit_ps <-  drtmle(W = W, A = A, Y = Y, stratify = FALSE,
-                      # make sure a_0/gn are ordered properly!
-                      gn = gn, a_0 = c(1, 0),
-                      glm_Q = "W1 + W2*A", glm_Qr = "gn",
-                      glm_gr = "Qn", family = binomial())
-out_fit_ps
-```
-
-## Multi-level treatments
-
-```{r, echo = FALSE, message = FALSE}
-library(drtmle)
-set.seed(1234)
-N <- 1e6
-W <- data.frame(W1 = runif(N), W2 = rbinom(N, 1, 0.5))
-Y2 <- rbinom(N, 1, plogis(W$W1 + 2*W$W2))
-Y1 <- rbinom(N, 1, plogis(W$W1 + W$W2))
-Y0 <- rbinom(N, 1, plogis(W$W1))
-EY1 <- mean(Y1)
-EY0 <- mean(Y0)
-EY2 <- mean(Y2)
-```
-
-In our examples so far, we have only considered binary treatments. However,
-\texttt{drtmle} is fully equipped to handle treatments with an arbitrary number of
-discrete values as well. Suppose that $A$ assumes values in a set $\mathcal{A}$
-of limited cardinality, the PS regression estimation is modified to ensure that
-\begin{equation*}
-  \sum\limits_{a \in \mathcal{A}} g_n(a \mid w) = 1 \ \mbox{for all $w$} \ .
-\end{equation*}
-If this condition is true, we say that the estimates of the PS are compatible.
-Many regression methodologies that have been developed work well with binary
-outcomes, but have not been extended to multi-level outcomes. To ensure that
-users of \texttt{drtmle} have access to the large number of binary regression
-techniques when estimating the PS, we have taken an alternative approach to
-estimation of the propensity for each level of the treatment. Specifically,
-rather than fitting a single multinomial-type regression, we fit a series of
-binomial regressions. As a concrete example, consider the case of no missing
-data and where $\mathcal{A} = \{0,1,2\}$. We can ensure compatible estimates of
-the PS by generating an estimate $g_n(0 \mid w)$ of $Pr_0(A = 0 \mid W = w)$
-and $\tilde{g}_n(1 \mid w)$ of $Pr_0(A = 1 \mid A > 0, W = w)$. The latter
-estimate may be generated by regressing $I(A = 1)$ on $W$ in observations with
-$A > 0$. Since the outcome is binary, this regression may be estimated using
-any technique suitable for binary outcomes. By rules of conditional
-probability, it holds that $g_0(1 \mid w) = Pr_0(A = 1 \mid A > 0, W = w)\{1 -
-Pr_0(A = 0 \mid W)\}$ and thus an estimate of $g_0(1 \mid w)$ can be computed
-as $g_n(1 \mid w) = \tilde{g}_n(1 \mid w) \{1 - g_n(0 \mid w)\}$. Finally, we
-let $g_n(2 \mid w) = 1 - g_n(0 \mid w) - g_n(1 \mid w)$, which ensures
-compatibility of the estimates for each level of the covariates. This approach
-generalizes to an arbitrary number of treatment levels. Below, we provide an
-illustration using simulated data with a treatment that has three levels. The
-true parameter values in this simulation are $\psi_0(0)=$ `r round(EY0,2)`,
-$\psi_0(1)=$ `r round(EY1, 2)`, and $\psi_0(2)=$ `r round(EY2,2)`.
-
-```{r, cache = TRUE, message = FALSE}
-set.seed(1234)
-n <- 200
-W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
-A <- rbinom(n, 2, plogis(W$W1 + W$W2))
-Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
-```
-
-The multi-level PS can still be estimated using any of the three techniques
-discussed previously. Here, we illustrate with generalized linear
-models.
-
-```{r, cache = TRUE, message = FALSE}
-glm_fit_multiA <- drtmle(W = W, A = A, Y = Y, stratify = FALSE,
-                         glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-                         glm_gr = "Qn", glm_Qr = "gn",
-                         family = binomial(), a_0 = c(0,1,2))
-glm_fit_multiA
-```
-
-Above, we used the option `a_0` to specify the levels of the treatment at which
-we were interested in estimating marginal means. There is no need for `a_0` to
-contain all unique values of `A`. For example, certain levels of the treatment
-may not be of scientific interest, but nevertheless we would like to use these
-data to help estimate the OR and PS (by setting `stratify = FALSE`).
-
-The confidence interval and testing procedures extend to multi-level treatments
-with minor modifications. We can still obtain a confidence interval and
-hypothesis test for each of the marginal means by not specifying a `contrast`.
-
-```{r, cache = TRUE}
-ci(glm_fit_multiA)
-wald_test(glm_fit_multiA, null = c(0.4, 0.5, 0.6))
-```
-
-Alternatively, we can specify a `contrast` to estimate confidence intervals
-about the average treatment effect of $A=1$ versus $A=0$. Calls to `wald_test()`
-can also be made.
-
-```{r, cache = TRUE}
-ci(glm_fit_multiA, contrast = c(-1, 1, 0))
-```
-
-Similarly, we can compute confidence intervals comparing the average treatment
-effect of $A = 2$ versus $A = 0$.
-
-```{r, cache = TRUE}
-ci(glm_fit_multiA, contrast = c(-1, 0, 1))
-```
-
-User-specified contrasts are also available. For example, we can modify the
-`riskRatio` object we used previously for a two-level treatment to compute the
-risk ratio comparing $A=1$ to $A=0$ and comparing $A = 2$ to $A = 0$.
-
-```{r, cache = TRUE}
-riskRatio_1v0 <- list(f = function(eff){ log(eff) },
-                      f_inv = function(eff){ exp(eff) },
-                      h = function(est){ est[2]/est[1] },
-                      fh_grad =  function(est){ c(1/est[2], -1/est[1], 0) })
-ci(glm_fit_multiA, contrast = riskRatio_1v0)
-```
-
-Similarly, we can compute confidence intervals for the risk
-ratio comparing the marginal mean for $A=2$ to $A=1$.
-
-```{r, cache = TRUE}
-riskRatio_2v0 <- list(f = function(eff){ log(eff) },
-                      f_inv = function(eff){ exp(eff) },
-                      h = function(est){ est[3]/est[1] },
-                      fh_grad =  function(est){ c(0, -1/est[1], 1/est[3]) })
-ci(glm_fit_multiA, contrast = riskRatio_2v0)
-```
-
-## Cross-validated regression estimates
-
-To avoid overfitting initial regressions, one can use cross-validated 
-estimates of the nuisance
-regressions. Theoretically, the use of cross-validated regression weakens the
-assumptions necessary to achieve asymptotic normality [@zheng:2010:working].
-This is true of the regular TML and AIPTW estimators, as well as of the TML
-estimator with doubly-robust inference. The `drtmle()` function implements
-cross-validated estimation of the regression parameters through the `cvFolds`
-option, as shown in the code below.
-
-```{r, cache = TRUE}
-cv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                    SL_g = c("SL.glm", "SL.glm.interaction", "SL.earth"),
-                    SL_Q = c("SL.glm", "SL.glm.interaction", "SL.earth"),
-                    SL_gr = c("SL.glm", "SL.mean"),
-                    SL_Qr = c("SL.glm", "SL.mean"),
-                    stratify = FALSE, cvFolds = 2, a_0 = c(0, 1, 2))
-```
-
-Cross-validation may significantly increase the computation time necessary for
-running `drtmle()`. Parallelized nuisance regression fitting is available by
-setting the option `use_future = TRUE`. Since, internally, all computations are
-carried out using futures (with the API provided by the \texttt{future} R package),
-setting this option amounts only to parallelizing with futures [@futurePackage]
-via `future.apply::lapply()`; when `use_future = FALSE`, base `lapply()` is
-simply used instead.  Together, setting these options results in parallelization
-of the estimation of the OR, PS, and reduced-dimension regressions.
-
-If no adjustments are made prior to invoking the `drtmle()` function, futures
-are computed sequentially, as per the defaults of the \texttt{future} package. It
-is left to the user to specify appropriate system-specific future backends that
-may better suit their problem (e.g., using the \texttt{future.batchtools} package
-[@futurebatchtoolsPackage] or even simply setting a parallelized
-`future::plan()`). Below, we demonstrate registration using an ad-hoc cluster on
-the local system. The choice of backend is flexible, and submission of the job
-to a wide variety of schedulers (e.g., SLURM, TORQUE) is also permitted.
-Interested readers are invited to consult the documentation of the
-\texttt{future.batchtools} package for further information.
-
-```{r, message = FALSE, eval = FALSE}
-library(future.batchtools)
-library(parallel)
-cl <- makeCluster(2, type = "SOCK")
-plan(cluster, workers = cl)
-clusterEvalQ(cl, library("SuperLearner"))
-pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                     SL_g = c("SL.glm", "SL.glm.interaction","SL.earth"),
-                     SL_Q = c("SL.glm", "SL.glm.interaction","SL.earth"),
-                     SL_gr = c("SL.glm", "SL.earth", "SL.mean"),
-                     SL_Qr = c("SL.glm", "SL.earth", "SL.mean"),
-                     stratify = FALSE, a_0 = c(0,1,2),
-                     cvFolds = 2, use_future = TRUE)
-```
-
-# Discussion
-
-The \texttt{drtmle} package was designed to provide a user-friendly implementation
-of the TMLE algorithm that facilitates doubly-robust inference about marginal
-means and average treatment effects. While the theory underlying doubly-robust
-inference is complex, the implementation of the methods only requires users to
-provide the data and specify how to estimate nuisance regressions. Though
-estimation of the OR and PS is familiar to those with backgrounds in causal
-inference-related estimation, the reduced-dimension regressions may appear
-strange. Indeed, it is difficult to provide intuition for these parameters and
-more difficult still to determine what scientific background knowledge might
-guide users in how to estimate these regression parameters. Therefore, we
-recommended estimating these quantities adaptively, e.g., with the super
-learner. The super learner library should contain several relatively smooth
-estimators such as `SL.mean`, `SL.glm`, and `SL.gam`. In particular, including
-`SL.mean` (an unadjusted generalized linear model) may be important due to the
-fact that, when the OR and PS are consistently estimated, the reduced-dimension
-regressions are identically equal to zero. Thus, this unadjusted regression
-estimator may do a good job estimating the reduced-dimension regressions in
-situations where the OR and PS are estimated well. In addition to relatively
-smooth estimators, we recommend additionally including several adaptive
-estimators such as `SL.npreg` or `SL.earth`.
-
-Planned extensions of the \texttt{drtmle} package include extensions to
-longitudinal settings involving treatments at multiple time points and inclusion
-of the collaborative targeted maximum likelihood estimator (C-TMLE) of
-@vdl:2014:ijb that also facilitates collaborative doubly-robust inference.
-
-# Appendix
-
-## Arbitrary user-specified regressions
-
-The \texttt{drtmle} package also allows users to pass in their own estimated OR and
-PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
-user, `drtmle()` will ignore the nuisance parameter estimation routines
-specified by `SL_Q` and `glm_Q` -- similarly for `gn` and `SL_g`, `glm_g`. As
-noted in the package documentation, there is a specific format necessary for
-inputting `Qn` and `gn`. For `Qn`, the entries in the list should correspond to
-the OR evaluated at `A` and the observed values of `W`, with order determined by
-the input to `a_0`. For example, if `a_0 = c(0, 1)` then `Qn[[1]]` should be the
-OR at $A = 0$ and `Qn[[2]]` should be the outcome regression at $A = 1$. The
-example below illustrates this concept.
-
-```{r, cache = TRUE, message = FALSE}
-# use simplified data
-set.seed(1234)
-n <- 200
-W <- data.frame(W1 = runif(n), W2 = rbinom(n, 1, 0.5))
-A <- rbinom(n, 1, plogis(W$W1 + W$W2))
-Y <- rbinom(n, 1, plogis(W$W1 + W$W2*A))
-
-# fit a GLM for outcome regression outside of drtmle
-out_glm_fit <- glm(Y ~ . , data = data.frame(A = A, W), family = binomial())
-
-# generate Qn list
-Qn10 <- list(
-  # first entry is predicted values setting A = 1
-  predict(out_glm_fit, newdata = data.frame(A = 1, W), type = "response"),
-  # second entry is predicted values setting A = 0
-  predict(out_glm_fit, newdata = data.frame(A = 0, W), type = "response")
-)
-
-# pass this list to drtmle to avoid internal estimation of
-# outcome regression (note propensity score and reduced-dimension
-# regressions are still estimated internally)
-out_fit1 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn", Qn = Qn10,
-                  # crucial to set a_0 to match Qn's construction!
-                  a_0 = c(1,0))
-out_fit1
-
-# to be completely thorough let's re-order Qn10 and re-run
-Qn01 <- list(Qn10[[2]], Qn10[[1]])
-
-out_fit2 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                  glm_g = ".", glm_Qr = "gn", glm_gr = "Qn",
-                  # use re-ordered Qn list
-                  Qn = Qn01,
-                  # a_0 has to be reordered to match Qn01!
-                  a_0 = c(0,1))
-# should be the same as out_fit1, but re-ordered
-out_fit2
-```
-
-Similarly, for `gn` we need to pass in properly ordered lists. For example, if
-`a_0 = c(0, 1)` then `gn[[1]]` should be the PS for receiving $A = 0$ and
-`gn[[2]]` should be the PS for receiving $A = 1$. The example below illustrates
-this concept.
-
-```{r, cache = TRUE, message = FALSE}
-# fit a GLM for propensity score outside of drtmle
-out_glm_fit <- glm(A ~ . , data = data.frame(W), family = binomial())
-
-# get P(A = 1 | W) by calling predict
-gn1W <- predict(out_glm_fit, newdata = data.frame(W), type = "response")
-
-# generate gn list
-gn10 <- list(
-  # first entry is P(A = 1 | W)
-  gn1W,
-  # second entry is P(A = 0 | W) = 1 - P(A = 1 | W)
-  1 - gn1W
-)
-
-# pass this list to drtmle to avoid internal estimation of
-# propensity score (note reduced-dimension regressions are
-# still estimated internally)
-out_fit3 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                   glm_Qr = "gn", glm_gr = "Qn",
-                   Qn = Qn10, gn = gn10,
-                   # crucial to set a_0 to match Qn and gn's construction!
-                   a_0 = c(1,0))
-out_fit3
-
-# to be completely thorough let's re-order gn10 and re-run
-gn01 <- list(gn10[[2]], gn10[[1]])
-
-out_fit4 <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                   glm_Qr = "gn", glm_gr = "Qn",
-                   # use re-ordered Qn/gn list
-                   Qn = Qn01, gn = gn01,
-                   # a_0 has to be reordered to match Qn01!
-                   a_0 = c(0,1))
-# should be the same as out_fit3, but re-ordered
-out_fit4
-```
-
-This flexibility allows users to leverage any regression technique to externally
-obtain initial estimates of nuisance parameters. The key step is making sure
-that the ordering of `Qn` and `gn` is consistent with `a_0`.
-
-## Additional methods included in drtmle
-
-### Inference for IPTW using adaptive PS estimators
-
-Doubly-robust estimators of marginal means are appealing in their robustness and
-efficiency properties. However, researchers may prefer the IPTW estimator for
-its intuitive derivation and implementation. Heretofore, inference for IPTW
-estimators precluded the use of adaptive estimators of the PS. However,
-@vdl:2014:ijb proposed modified IPTW estimators that allow for adaptive
-PS estimation. Similarly as above, this estimator is based on an additional
-regression parameter, $\tilde{Q}_{r,0n}(a,w) \coloneqq E_0\{Y \mid A = a, g_n(W)
-= g_n(w)\}$, that is, the regression of the outcome $Y$ on the estimated
-propensity score amongst observations with $A=a$. The author showed that if
-$g_n^{\star}(a \mid w)$, an estimator of the propensity for treatment $A=a$,
-satisfied that
-\begin{equation*}
-  \frac{1}{n} \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W =
-    W_i)} \ \{ I(A_i = a) - g_n(a \mid W = W_i)\} = 0 \ ,
-\end{equation*}
-then the IPTW estimator based on $g_n^{\star}$ would have an asymptotic normal
-distribution with estimable variance. In fact, @vdl:2014:ijb proposed a
-TMLE algorithm that mapped an initial PS estimate into an estimate that
-satisfies this key equation and variance estimators that can be used to generate
-Wald-style confidence intervals and hypothesis tests.
-
-An estimator with equivalent asymptotic properties based on a PS estimate,
-$g_n(a \mid w)$, can be computed as
-\begin{equation*}
-  \psi_{n,IPTW}(a) = \psi_{n,IPTW}(a) - \frac{1}{n} \sum\limits_{i=1}^n
-  \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)} \ \{ I(A_i = a) - g_n(a
-  \mid W = W_i)\} \ .
-\end{equation*}
-Both this estimator and the TMLE-based IPTW estimator above are implemented in
-\texttt{drtmle}.
-
-#### Implementation of inference for adaptive IPTW
-
-The IPTW estimators based on adaptive estimation of the PS are implemented in
-the `adaptive_iptw()` function. The `adaptive_iptw()` function shares many
-options with `drtmle()`; however, as the IPTW estimator does not rely on an
-estimate of the OR, options related to the OR are omitted.
-
-```{r, cache = TRUE}
-npreg_iptw_multiA <- adaptive_iptw(W = W, A = A, Y = Y, stratify = FALSE,
-                                   SL_g = "SL.npreg", SL_Qr = "SL.npreg",
-                                   family = binomial(), a_0 = c(0, 1))
-npreg_iptw_multiA
-```
-
-The `"adaptive_iptw"` S3 class contains identical methods for computing
-confidence intervals and Wald-style tests as the `"drtmle"` class, thus readers
-are referred back to the section on confidence intervals and hypothesis tests
-for reminders on the various options for these procedures. By default, these
-methods are implemented for the `iptw_tmle()` estimator of @vdl:2014:ijb.  We
-conjecture that this estimator will have improved finite-sample behavior
-relative to `iptw_os()` (the alternative estimator described above), though
-further study in numerical experiments is warranted.
-
-As with `drtmle()`, the `adaptive_iptw()` function allows missing data in `A`
-and `Y`. Similarly, `adaptive_iptw()` also allows for parallelized,
-cross-validated estimation of the regression parameters via the `cvFolds` and
-`parallel` options.  As with the estimators implemented by `drtmle()`,
-cross-validation allows for valid inference under weaker assumptions.
-
-### Outcome-adaptive propensity score estimation
-
-We have recently proposed estimators of the average treatment effect to be used
-in settings where this effect is weakly identifiable, as evidenced by estimated
-propensity scores of small magnitude.  In these settings, doubly robust
-estimators may exhibit non-robust behavior in small samples. The proposed
-solution was to target an alternative quantity in the propensity score
-estimation step. If we are interested in estimating $\psi_0(a)$ for $a = 0, 1$,
-we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$, which
-describes the conditional probability of receiving treatment $A = a$ given the
-value of the two outcome regressions. For this reason, we call this quantity an
-*outcome-adaptive propensity score*. We have shown that all of the usual
-asymptotics for TMLE carry through when an estimator of this quantity is
-substituted for the true propensity score. The resultant TML estimator is
-*super-efficient*, meaning that it will generally have greater asymptotic
-efficiency than a standard TML estimator. Simulation experiments indicate that
-this approach also delivers more stable behavior in settings where estimated
-propensity scores become very small. However, the estimator is not doubly robust
--- the asymptotics rely on the outcome regression being consistently estimated
-at $n^{-1/4}$ rate. In this sense, this estimator trades off robustness against
-achieving greater stability and efficiency. See @ju2020robust for further
-discussion.
-
-#### Implementation of outcome-adaptive propensity score
-
-If the outcome-adaptive PS is desired, the user may specify this via the
-`adapt_g` option. In this case, the additional targeting steps needed for doubly
-robust inference are "turned off" (i.e., `guard = NULL`), since, as discussed
-above, this estimator is *not* doubly robust. Note that when `adapt_g = TRUE`,
-the data used for PS estimation will be a `data.frame` with column names `QaW`
-where `a` takes all the values specified in `a_0`. For example if `a_0 =
-c(0,1)`, then the PS estimation data frame will have two columns called `Q0W`
-and `Q1W`. This may be relevant for users writing their own \texttt{SuperLearner}
-wrapper functions or those who make use of the `glm_g` setting.
-
-```{r, message = FALSE}
-# outcome adaptive propensity score fit using glms
-adaptg_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                     glm_Q = ".^2", glm_g = "Q1W + Q0W",
-                     adapt_g = TRUE)
-
-# outcome adaptive propensity score fit using super learner
-adaptg_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
-                     SL_Q = c("SL.glm", "SL.mean"),
-                     SL_g = c("SL.glm", "SL.mean"),
-                     adapt_g = TRUE)
-```
+\begin{keywords}
+  causal inference, machine learning, treatment effects, targeted minimum
+  loss estimation
+\end{keywords}

### 2022-06-24 paper/using_drtmle_obstudies.Rmd
work on header

@@
 ---
-documentclass: article
-classoption: [twoside,11pt]
-author: David Benkeser and Nima S. Hejazi
-title: "Doubly-Robust Inference in R using \\texttt{drtmle}"
-keywords: [causal inference, machine learning, treatment effects, targeted minimum loss estimation]
 output:
-  latex_document:
-    extra_dependencies:
-      obs_study_style: null
-      amsmath: null
-      mathtools: null
-      fancyvrb: null
-      inputenc: ["utf8"]
+  bookdown::pdf_document2:
+    includes:
+      in_header: preamble.tex
 bibliography: refs.bib
-header-includes: >
-  \providecommand{\tightlist}{%
-    \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-  \newcommand{\dataset}{{\cal D}}
-  \newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
-  \DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
-  \DefineVerbatimEnvironment{Soutput}{Verbatim}{}
-  \DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
-  \newenvironment{Schunk}{}{}
-  \DefineVerbatimEnvironment{Code}{Verbatim}{}
-  \DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
-  \DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
-  \newenvironment{CodeChunk}{}{}
-  \setkeys{Gin}{width=0.8\textwidth}
-abstract: >
-  Inverse probability of treatment weighted estimators and doubly-robust
-  estimators (including augmented inverse probability of treatment weight and
-  targeted minimum loss-based estimators) are widely used in causal inference to
-  estimate and draw inference about the average effect of a treatment. As an
-  intermediate step, these estimators require estimation of key nuisance
-  parameters, which are often regression functions. Typically, regressions are
-  estimated using maximum likelihood and parametric models. Confidence intervals
-  and p-values may be computed based on standard asymptotic results, such as the
-  central limit theorem, the delta method, and the nonparametric bootstrap.
-  However, in high-dimensional settings maximum likelihood estimation often
-  breaks down and standard procedures no longer yield correct inference.
-  Instead, we may rely on adaptive estimators of nuisance parameters to
-  construct flexible regression estimators. However, use of adaptive estimators
-  poses a challenge for performing statistical inference about an estimated
-  treatment effect. While doubly-robust estimators facilitate inference when
-  *all* relevant regression functions are consistently estimated, the same
-  cannot be said when at least one nuisance estimator is inconsistent.
-  \texttt{drtmle} implements doubly-robust confidence intervals and hypothesis
-  tests for targeted minimum loss-based estimates of the average treatment
-  effect, in addition to several other recently-proposed estimators of the
-  average treatment effect.
 ---
@@
 ```
+
+## References

### 2022-06-24 paper/preamble.tex
more edits to build

@@
 % natbib citation styles, see the natbib documentation, a copy of which
-
+\usepackage{obs_study_style}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{fancyvrb}
-\usepackage{obs_study_style}
+\usepackage[pdfstartview=FitH]{hyperref}
+\usepackage{bookmark}
 
@@
 \providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}
+}
 \newcommand{\dataset}{{\cal D}}
@@
 \author{
-  \name David Benkeser \email benkeser@emory.edu \\
+  \name David Benkeser \email benkeser@emory.edu\\
   \addr Department of Biostatistics \& Bioinformatics\\
@@
   Emory University\\
-  1518 Clifton Road, N.E., Atlanta, GA 30322
+  1518 Clifton Road, N.E.,\\
+  Atlanta, GA 30322
   \AND
-  \name Nima S.~Hejazi \email nhejazi@hsph.harvard.edu \\
+  \name Nima S.~Hejazi \email nhejazi@hsph.harvard.edu\\
   \addr Department of Biostatistics\\
   T.H.~Chan School of Public Health\\
-  Harvard University \\
-  677 Huntington Avenue, Boston, MA 02115
+  Harvard University\\
+  677 Huntington Avenue,\\
+  Boston, MA 02115
 }

### 2022-06-24 paper/using_drtmle_obstudies.Rmd
more edits to build

@@
 output:
-  bookdown::pdf_document2:
+  latex_fragment:
     includes:
@@
 
-This has led to increased interest in adopting adaptive estimation techniques from the
-machine learning literature to control for high-dimensional confounders when
-estimating treatment effects. However, facilitating proper inference (i.e.,
+This has led to increased interest in adopting adaptive estimation techniques
+from the machine learning literature to control for high-dimensional confounders
+when estimating treatment effects. However, facilitating proper inference (i.e.,
 confidence intervals and hypothesis tests) about estimates of effects based on
-adaptive estimators is challenging. In particular, the use of standard causal inference
-techniques, including both G-computation (GCOMP) estimators
+adaptive estimators is challenging. In particular, the use of standard causal
+inference techniques, including both G-computation (GCOMP) estimators
 [@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
-estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when utilizing
-adaptive estimators. In particular, with rare exception, if cross-validation is used
-to select tuning parameters for the nuisance parameters, estimates of the population-level
-effect will generally be too biased to achieve standard $n^{1/2}$-rate asymptotics. Instead,
-*undersmoothing* must be used to select tuning parameters to control the bias 
+estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when
+utilizing adaptive estimators. In particular, with rare exception, if
+cross-validation is used to select tuning parameters for the nuisance
+parameters, estimates of the population-level effect will generally be too
+biased to achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing*
+must be used to select tuning parameters to control the bias
 [@undersmoothed_hal]. However, this can prove quite challenging in practice and
@@
 
-A more natural framework for integrating adaptive methods into nuisance regression 
-estimation are doubly robust approaches, including augmented inverse probability of 
-treatment weighted (AIPTW) estimation [@robins:1994:jasa] and targeted minimum 
-loss-based (TML) estimation [@vdlrubin:2006:ijb]. The asymptotics of these estimators
-are characterized in part by the convergence rates of nuisance parameter estimates with
-respect to a proper norm. This fact makes the use of cross-validation more appealing for
-selection of tuning parameters for estimators of nuisance parameters, thereby providing 
-a more straightforward means of obtaining an estimate of the effect of interest that
-enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield closed form
-standard error estimates, which can serve as the basis for constructing confidence 
-intervals and hypothesis tests.
-
-As an intermediate step, the AIPTW and TML estimators require estimation of
-two key nuisance parameters: the probability of treatment as a function of
+A more natural framework for integrating adaptive methods into nuisance
+regression estimation are doubly robust approaches, including augmented inverse
+probability of treatment weighted (AIPTW) estimation [@robins:1994:jasa] and
+targeted minimum loss-based (TML) estimation [@vdlrubin:2006:ijb]. The
+asymptotics of these estimators are characterized in part by the convergence
+rates of nuisance parameter estimates with respect to a proper norm. This fact
+makes the use of cross-validation more appealing for selection of tuning
+parameters for estimators of nuisance parameters, thereby providing a more
+straightforward means of obtaining an estimate of the effect of interest that
+enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield
+closed form standard error estimates, which can serve as the basis for
+constructing confidence intervals and hypothesis tests.
+
+As an intermediate step, the AIPTW and TML estimators require estimation of two
+key nuisance parameters: the probability of treatment as a function of
 confounders (the propensity score, hereafter referred to as the *PS*) and the
@@
 estimators are consistent for the population-level effect of interest if either
-of the PS or OR is consistently estimated. This property gives the AIPTW and
-TML estimators a natural appeal: inconsistency in the estimation of one
-nuisance parameter may be mitigated by the consistent estimation of the other.
-However, recent work has shown that the double robustness property does not
-necessarily extend to the inferential properties of these estimators when
-adaptive estimators of the PS and/or OR are used
-[@vdl:2014:ijb,@benkeser:2017:biometrika]. Instead, valid inference requires
-consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb
-proposed new TML estimators that have an asymptotic normal sampling distribution
-with estimable variance under consistent estimation of *either* the PS *or* the
-OR, paving the way for inference that is doubly-robust.
-@benkeser:2017:biometrika proposed modifications of these estimators with
-similar robustness properties and illustrated their practical performance.
+of the PS or OR is consistently estimated. This property gives the AIPTW and TML
+estimators a natural appeal: inconsistency in the estimation of one nuisance
+parameter may be mitigated by the consistent estimation of the other. However,
+recent work has shown that the double robustness property does not necessarily
+extend to the inferential properties of these estimators when adaptive
+estimators of the PS and/or OR are used
+[@vdl:2014:ijb, @benkeser:2017:biometrika]. Instead, valid inference requires
+consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb proposed new
+TML estimators that have an asymptotic normal sampling distribution with
+estimable variance under consistent estimation of *either* the PS *or* the OR,
+paving the way for inference that is doubly-robust. @benkeser:2017:biometrika
+proposed modifications of these estimators with similar robustness properties
+and illustrated their practical performance.
 
@@
 
-Suppose the observed data consist of $n$ independent and identically
-distributed copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is
-a vector of baseline covariates, $A \in \{0,1\}$ is a binary treatment, 
-$Y$ is an outcome, and $P_0$ is the true data-generating
-distribution. The parameters of interest are
+Suppose the observed data consist of $n$ independent and identically distributed
+copies of the random variable $O = (W,A,Y) \sim P_0$. Here, $W$ is a vector of
+baseline covariates, $A \in \{0,1\}$ is a binary treatment, $Y$ is an outcome,
+and $P_0$ is the true data-generating distribution. The parameters of interest
+are
 \begin{equation}

### 2022-06-24 paper/using_drtmle_obstudies.Rmd
fix tex template

@@
 output:
-  latex_fragment:
-    includes:
-      in_header: preamble.tex
+  bookdown::pdf_document2:
+    template: obs_study_skeleton.tex
 bibliography: refs.bib
+abs_text: |
+  Inverse probability of treatment weighted estimators and doubly-robust
+  estimators (including augmented inverse probability of treatment weight and
+  targeted minimum loss-based estimators) are widely used in causal inference
+  to estimate and draw inference about the average effect of a treatment. As an
+  intermediate step, these estimators require estimation of key nuisance
+  parameters, which are often regression functions. Typically, regressions are
+  estimated using maximum likelihood and parametric models. Confidence
+  intervals and p-values may be computed based on standard asymptotic results,
+  such as the central limit theorem, the delta method, and the nonparametric
+  bootstrap. However, in high-dimensional settings maximum likelihood estimation
+  often breaks down and standard procedures no longer yield correct inference.
+  Instead, we may rely on adaptive estimators of nuisance parameters to
+  construct flexible regression estimators. However, use of adaptive estimators
+  poses a challenge for performing statistical inference about an estimated
+  treatment effect. While doubly-robust estimators facilitate inference when
+  *all* relevant regression functions are consistently estimated, the same
+  cannot be said when at least one nuisance estimator is inconsistent.
+  \texttt{drtmle} implements doubly-robust confidence intervals and hypothesis
+  tests for targeted minimum loss-based estimates of the average treatment
+  effect, in addition to several other recently-proposed estimators of the
+  average treatment effect.
+abs_keywords: |
+  causal inference, machine learning, semiparametric estimation, treatment
+  effects, multiple robustness, targeted minimum loss estimation
 ---
@@
 [@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
-estimators [@horvitzthompson:1952:jasa, @robins:2000:epi] is challenging when
+estimators [@horvitzthompson:1952:jasa; @robins:2000:epi] is challenging when
 utilizing adaptive estimators. In particular, with rare exception, if
@@
 estimators of the PS and/or OR are used
-[@vdl:2014:ijb, @benkeser:2017:biometrika]. Instead, valid inference requires
+[@vdl:2014:ijb; @benkeser:2017:biometrika]. Instead, valid inference requires
 consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb proposed new
@@
 [@vdl:2007:statapp] that generalizes stacking and stacked regression
-\citep{wolpert:1992:nnet,breiman:1996:ml}. The user specifies many potential
+[@wolpert:1992:nnet; @breiman:1996:ml]. The user specifies many potential
 candidate regression estimators, referred to as a *library* of candidate
@@
 In our examples so far, we have only considered binary treatments. However,
-\texttt{drtmle} is fully equipped to handle treatments with an arbitrary number of
-discrete values as well. Suppose that $A$ assumes values in a set $\mathcal{A}$
-of limited cardinality, the PS regression estimation is modified to ensure that
+\texttt{drtmle} is fully equipped to handle treatments with an arbitrary number
+of discrete values as well. Suppose that $A$ assumes values in a set
+$\mathcal{A}$ of limited cardinality, the PS regression estimation is modified
+to ensure that
 \begin{equation*}
@@
 
-## References
+# References

### 2022-06-24 paper/using_drtmle_obstudies.Rmd
add related packages; minor touch-ups

@@
     template: obs_study_skeleton.tex
+    latex_engine: xelatex
 bibliography: refs.bib
 abs_text: |
-  Inverse probability of treatment weighted estimators and doubly-robust
+  Inverse probability of treatment weighted estimators and doubly robust
   estimators (including augmented inverse probability of treatment weight and
-  targeted minimum loss-based estimators) are widely used in causal inference
-  to estimate and draw inference about the average effect of a treatment. As an
+  targeted minimum loss estimators) are widely used in causal inference to
+  estimate and draw inference about the average effect of a treatment. As an
   intermediate step, these estimators require estimation of key nuisance
   parameters, which are often regression functions. Typically, regressions are
-  estimated using maximum likelihood and parametric models. Confidence
-  intervals and p-values may be computed based on standard asymptotic results,
-  such as the central limit theorem, the delta method, and the nonparametric
-  bootstrap. However, in high-dimensional settings maximum likelihood estimation
-  often breaks down and standard procedures no longer yield correct inference.
+  estimated using maximum likelihood and parametric models. Confidence intervals
+  and p-values may be computed based on standard asymptotic results, such as the
+  central limit theorem, the delta method, and the nonparametric bootstrap.
+  However, in high-dimensional settings, maximum likelihood estimation often
+  breaks down and standard procedures no longer yield correct inference.
   Instead, we may rely on adaptive estimators of nuisance parameters to
@@
   poses a challenge for performing statistical inference about an estimated
-  treatment effect. While doubly-robust estimators facilitate inference when
+  treatment effect. While doubly robust estimators facilitate inference when
   *all* relevant regression functions are consistently estimated, the same
   cannot be said when at least one nuisance estimator is inconsistent.
-  \texttt{drtmle} implements doubly-robust confidence intervals and hypothesis
-  tests for targeted minimum loss-based estimates of the average treatment
-  effect, in addition to several other recently-proposed estimators of the
-  average treatment effect.
+  \texttt{drtmle} implements doubly robust confidence intervals and hypothesis
+  tests for targeted minimum loss estimates of the average treatment effect, in
+  addition to several other recently proposed estimators of the average
+  treatment effect.
 abs_keywords: |
@@
 estimate population-level effects from observational data. Estimation of these
-effects often requires accounting for putative confounding factors, that is,
+effects often requires accounting for putative confounding factors -- that is,
 factors related to whether a participant receives treatment and to the
@@
 modeling assumptions. When these assumptions are violated, estimates of the
-population-level effect of a treatment may have large bias.
+population-level effect of a treatment can suffer from a large degree of bias.
 
@@
 from the machine learning literature to control for high-dimensional confounders
-when estimating treatment effects. However, facilitating proper inference (i.e.,
+when estimating treatment effects. However, facilitating proper statistical
+inference (i.e.,
 confidence intervals and hypothesis tests) about estimates of effects based on
 adaptive estimators is challenging. In particular, the use of standard causal
-inference techniques, including both G-computation (GCOMP) estimators
+inference techniques, including both G-computation estimators
 [@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
 estimators [@horvitzthompson:1952:jasa; @robins:2000:epi] is challenging when
-utilizing adaptive estimators. In particular, with rare exception, if
+utilizing adaptive estimators. In particular, with rare exceptions, if
 cross-validation is used to select tuning parameters for the nuisance
-parameters, estimates of the population-level effect will generally be too
-biased to achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing*
-must be used to select tuning parameters to control the bias
-[@undersmoothed_hal]. However, this can prove quite challenging in practice and
-few guidelines are available.
+parameters, estimates of the population-level effect will be too biased to
+achieve standard $n^{1/2}$-rate asymptotics. Instead, *undersmoothing* must be
+used to select tuning parameters to control the bias [@undersmoothed_hal;
+@ertefaie2022nonparametric; @hejazi2022efficient]. However, this can prove quite
+challenging in practice, and few guidelines are available.
 
@@
 probability of treatment weighted (AIPTW) estimation [@robins:1994:jasa] and
-targeted minimum loss-based (TML) estimation [@vdlrubin:2006:ijb]. The
+targeted minimum loss (TML) estimation [@vdlrubin:2006:ijb]. The
 asymptotics of these estimators are characterized in part by the convergence
@@
 straightforward means of obtaining an estimate of the effect of interest that
-enjoys standard $n^{1/2}$ asymptotics. These asymptotic results also yield
-closed form standard error estimates, which can serve as the basis for
+enjoys standard $n^{1/2}$ asymptotics. Based on the efficient influence
+function, a key quantity in semiparametric theory, these asymptotic results also
+yield closed-form standard error estimates, which can serve as the basis for
 constructing confidence intervals and hypothesis tests.
@@
 of the PS or OR is consistently estimated. This property gives the AIPTW and TML
-estimators a natural appeal: inconsistency in the estimation of one nuisance
+estimators a natural appeal: Inconsistency in the estimation of one nuisance
 parameter may be mitigated by the consistent estimation of the other. However,
-recent work has shown that the double robustness property does not necessarily
-extend to the inferential properties of these estimators when adaptive
-estimators of the PS and/or OR are used
-[@vdl:2014:ijb; @benkeser:2017:biometrika]. Instead, valid inference requires
-consistent estimation of *both* the PS *and* the OR. @vdl:2014:ijb proposed new
-TML estimators that have an asymptotic normal sampling distribution with
-estimable variance under consistent estimation of *either* the PS *or* the OR,
-paving the way for inference that is doubly-robust. @benkeser:2017:biometrika
-proposed modifications of these estimators with similar robustness properties
-and illustrated their practical performance.
+recent work has shown that the double robustness property does not extend to the
+inferential properties of these estimators when adaptive estimators of the PS
+and/or OR are used [@vdl:2014:ijb; @benkeser:2017:biometrika]. Instead, valid
+inference requires consistent estimation of *both* the PS *and* the OR.
+@vdl:2014:ijb proposed new TML estimators that asymptotically attain a normal
+sampling distribution with estimable variance under consistent estimation of
+*either* the PS *or* the OR, paving the way for inference that is doubly-robust.
+@benkeser:2017:biometrika proposed modifications of these estimators with
+similar robustness properties and illustrated their practical performance.
 
@@
 involved, notably requiring iterative estimation of additional, complex nuisance
-parameters. The \texttt{drtmle} package was motivated by the need for a
-user-friendly tool to implement these methods. The package implements the TML
+parameters. The \texttt{drtmle} package was motivated by the need for a open
+source tool to implement these methods. The package implements the TML
 estimators of population-level effects proposed in @vdl:2014:ijb and
 @benkeser:2017:biometrika, as well as several extensions including
-cross-validated targeted minimum loss-based estimation (CV-TMLE). Additionally,
+cross-validated targeted minimum loss estimation (CV-TMLE). Additionally,
 \texttt{drtmle} implements an IPTW estimator that overcomes shortcomings of
@@
 about population-level effects. Both the TML and IPTW estimators implemented in
-the \texttt{drtmle} package are capable of estimating population-level effects for
-discrete-valued treatments; moreover, the package provides convenient utilities
-for constructing confidence intervals and hypothesis tests about contrasts of
-the mean outcome under different levels of treatment. This document explains
-some of the key concepts underlying doubly-robust inference and illustrates
-usage of the \texttt{drtmle} package.
+the \texttt{drtmle} package are capable of estimating population-level effects
+for discrete-valued treatments; moreover, the package provides convenient
+utilities for constructing confidence intervals and hypothesis tests about
+contrasts of the mean outcome under different levels of treatment. This
+manuscript explains some of the key concepts underlying doubly robust inference
+and illustrates usage of the \texttt{drtmle} package for the \texttt{R}
+language and environment for statistical computing [@R].
+
+In tandem with growing interest in and usage of doubly robust estimation
+approaches, open source software tools implementing these methods have
+proliferated. Notably extant software packages include the \texttt{tmle}
+\texttt{R} package [@gruber2012tmle], which implements TML estimators of the
+effects of static single time point interventions; the \texttt{ltmle} \texttt{R}
+package [@lendle2017ltmle], which implements TML estimators of the effects of
+multiple time point interventions, subject to time-dependent confounding and
+censoring, as well as estimators of the parameters of a longitudinal working
+marginal structural model; the \texttt{tmle3} \texttt{R} package
+[@coyle2021tmle3-rpkg], which serves as a generalizable engine for implementing
+TML estimators within the \texttt{tlverse} ecosystem [@vdl2022targeted],
+designed to provide a common and extensible framework for TML estimation of the
+effects of static, dynamic, and stochastic single time point interventions; and
+the \texttt{doubleml} \texttt{R} and \texttt{Python} packages
+[@bach2021doublemlR; @bach2022doublemlPy], which implement estimators of causal
+effects within the double machine learning framework [@chernozhukov2017double;
+@chernozhukov2018double], reyling upon Neymon orthogonality and sample-splitting
+for robustness. From among this non-exhaustive list, \texttt{drtmle} stands out
+as, to our knowledge, the only implementation of statistical procedures designed
+to attain doubly robust inference [@benkeser:2017:biometrika], uniquely
+positioning \texttt{drtmle} as a robust tool for the estimation of treatment
+effects in observational data settings.
 
@@
 
-An estimator of $\psi_0(a)$ may be motivated directly by (1). Suppose we have
-available an estimate of the OR, $\bar{Q}_n$, and an estimate of the
-distribution function of baseline covariates, $Q_{W,n}$. An estimator of
-$\psi_0(a)$ may be obtained by plugging these into (1), $\psi_n(a) = \int
-\bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical distribution function of
-covariates is used so that this estimator can be equivalently expressed
-\begin{equation*}
+An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
+Suppose we have available an estimate of the OR, $\bar{Q}_n$, and an estimate of
+the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
+$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
+$\psi_n(a) = \int \bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
+distribution function of covariates is used so that this estimator can be
+equivalently expressed
+\begin{equation}
   \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
-\end{equation*}
+  \label{eqn:gcomp}
+\end{equation}
 This estimator is commonly referred to as the G-computation (GCOMP) estimator.
@@
 An alternative estimator of the treatment-specific population-level mean may be
-obtained by noting that (1) can be equivalently written as
+obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
 \begin{align}
@@
 We use $g_0(a \mid W) \coloneqq Pr_0(A = a \mid W)$ to denote the propensity
-score. Suppose we had available $g_n$, an estimate of the PS. Equation (2)
-motivates the estimator
+score. Suppose we had available $g_n$, an estimate of the PS. Equation
+(\ref{eqn:iptwID}) motivates the estimator
 \begin{equation*}
@@
 estimator. Like GCOMP estimators, inference for IPTW estimators based on
-parametric PS estimators may be facilitated through the nonparametric bootstrap.
-If adaptive approaches are used to estimate the PS, inference may not be
-available without further modification of the propensity score estimator, as we
-discuss below.
-
-## Doubly-robust estimators
+parametric PS estimators may be facilitated through the nonparametric bootstrap;
+however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
+are constructed as solutions to an IPTW estimating equation, the empirical
+variance of which may be used for standard error estimation. If adaptive
+approaches are used to estimate the PS, inference may not be available without
+further modification of the propensity score estimator, whether via targeting,
+as we discuss below, or undersmoothing, as explored by
+@ertefaie2022nonparametric and @hejazi2022efficient.
+
+## Doubly robust estimators
 
@@
 machine learning) are used to estimate the OR or PS, then the GCOMP and IPTW
-estimators generally lack a basis for statistical inference.
+estimators generally lack a valid basis for asymptotic statistical inference.
 
@@
 The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
-estimators. The estimator exhibits double-robustness in that it is consistent
+estimators. The estimator exhibits _double robustness_ in that it is consistent
 for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
@@
 consistently estimated and regularity conditions hold, one can establish a
-limiting normal distribution for the AIPTW estimator. Simple estimators of the
+normal limiting distribution for the AIPTW estimator. Simple estimators of the
 variance of this limiting distribution can be derived that may be used to
@@
 A more recent proposal to improve robustness and inferential properties of the
-GCOMP and IPTW estimators utilizes targeted minimum loss-based estimation
-(TMLE). TMLE is a general methodology that may be used to modify an estimator
-of a regression function in a manner that ensures that the modified estimator
+GCOMP and IPTW estimators utilizes targeted minimum loss estimation (or TMLE).
+TMLE is a general methodology that may be used to modify an estimator of a
+regression function in a manner that ensures that the modified estimator
 satisfies user-selected equations. In the present problem, one can show that if
@@
 asymptotic properties as the AIPTW estimator. That is, the estimator will be
-doubly-robust and, provided both the OR and PS are consistently estimated and
+doubly robust and, provided both the OR and PS are consistently estimated and
 regularity conditions are satisfied, will have a normal limiting distribution.
 TMLE provides a means of mapping an initial estimator of the OR into an
-estimator that satisfies (4). In addition to possessing the desirable asymptotic
-properties, TML estimators have been shown to outperform AIPTW estimators in
-finite samples [@porter:2011:ijb]. We refer readers to @tlbook:2011 for
-more about TMLE.
-
-# Doubly-robust inference
+estimator that satisfies (\ref{oneStepEst}). In addition to possessing the
+desirable asymptotic properties, TML estimators have been shown to outperform
+AIPTW estimators in finite samples [@porter:2011:ijb]. We refer readers to
+@tlbook:2011 and @vdl2022targeted for further details about on the TMLE
+methodology.
+
+# Doubly robust inference
 
@@
 biased estimates of the marginal means of interest. @vdl:2014:ijb derived
-modified TML estimators that are doubly-robust not only with respect to
+modified TML estimators that are doubly robust not only with respect to
 consistency, but also with respect to asymptotic normality. Furthermore, this
-work provided closed-form, doubly-robust variance estimators that may be used to
-construct doubly-robust confidence intervals and hypothesis tests.
+work provided closed-form, doubly robust variance estimators that may be used to
+construct doubly robust confidence intervals and hypothesis tests.
 @benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
 estimators and demonstrated the practical performance of these estimators via
-simulation.
+simulation experiments.
 
 The key to the theory underlying these results is finding estimators of the OR
-and PS that simultaneously satisfy equation (4) in addition to two additional
-equations. These equations involve additional regression parameters that we
-refer to as the reduced outcome regression (R-OR) and the reduced propensity
-score (R-PS), defined respectively as
+and PS that simultaneously satisfy equation (\ref{oneStepEst}) in addition to
+two additional equations. These equations involve additional regression
+parameters that we refer to as the reduced outcome regression (R-OR) and the
+reduced propensity score (R-PS), defined respectively as
 \begin{align}
@@
 \end{align}
-then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly-robust not
-only with respect to consistency, but also with respect to its limiting
+then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly robust
+not only with respect to consistency, but also with respect to its limiting
 distribution. The author additionally proposed a TMLE algorithm to map initial
@@
 In fact, there are three ways to estimate each of the regression parameters:
-generalized linear models, super learning
-[@vdl:2007:statapp; @SuperLearnerPackage], and a user-specified estimation
-technique. We demonstrate each of these approaches in turn on a  simulated
-data set. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0,
-2)` and $\psi_0(1)=$ `r round(EY1, 2)`.
+generalized linear models, super learning [@vdl:2007:statapp;
+@SuperLearnerPackage], and a user-specified estimation technique. We demonstrate
+each of these approaches in turn on a  simulated data set. The true parameter
+values of interest are $\psi_0(0)=$ `r round(EY0, 2)` and $\psi_0(1)=$ `r
+round(EY1, 2)`.
 
@@
 
-To avoid overfitting initial regressions, one can use cross-validated 
-estimates of the nuisance
-regressions. Theoretically, the use of cross-validated regression weakens the
-assumptions necessary to achieve asymptotic normality [@zheng:2010:working].
-This is true of the regular TML and AIPTW estimators, as well as of the TML
-estimator with doubly-robust inference. The `drtmle()` function implements
-cross-validated estimation of the regression parameters through the `cvFolds`
-option, as shown in the code below.
+To avoid overfitting initial regressions, one can use cross-validated estimates
+of the nuisance regressions. Theoretically, the use of cross-validated
+regression weakens the assumptions necessary to achieve asymptotic normality
+[@zheng:2010:working]. This is true of the regular TML and AIPTW estimators, as
+well as of the TML estimator with doubly-robust inference. The `drtmle()`
+function implements cross-validated estimation of the regression parameters
+through the `cvFolds` option, as shown in the code below.
 
@@
 
+# Acknowledgements
+
+NSH is grateful for the support of the National Science Foundation (award no.
+DMS 2102840).
+
 # References

### 2022-06-25 paper/using_drtmle_obstudies.Rmd
update renv

@@
 \begin{equation}
-  \psi_{n,GCOMP}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i) \ .
-  \label{eqn:gcomp}
+  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i)
+  \ . \label{eqn:gcomp}
 \end{equation}
@@
 score. Suppose we had available $g_n$, an estimate of the PS. Equation
-(\ref{eqn:iptwID}) motivates the estimator
-\begin{equation*}
-  \psi_{n,IPTW}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
+(\ref{iptwID}) motivates the estimator
+\begin{equation*}
+  \psi_{n,\text{IPTW}}(a) \coloneqq \frac{1}{n}\sum\limits_{i=1}^n
     \frac{I(A_i = a)}{g_n(a \mid W_i)} Y_i \ ,
@@
 are constructed as solutions to an IPTW estimating equation, the empirical
-variance of which may be used for standard error estimation. If adaptive
+variance of which may also be used for standard error estimation. If adaptive
 approaches are used to estimate the PS, inference may not be available without
@@
 \begin{equation*}
-\psi_{n,AIPTW}(a) = \psi_{n,IPTW}(a) + \frac{1}{n} \sum\limits_{i=1}^n \biggl\{
-1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \  \bar{Q}_n(a, W_i) \ ,
+\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n} \sum_{i=1}^n
+  \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \ \bar{Q}_n(a, W_i)
+  \ ,
 \end{equation*}
@@
 \begin{equation} \label{oneStepEst}
-\psi_{n,AIPTW}(a) = \psi_{n,GCOMP}(a) + \frac{1}{n} \sum\limits_{i=1}^n
+\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n} \sum_{i=1}^n
 \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .

### 2022-06-25 paper/using_drtmle_obstudies.Rmd
complete

@@
 when estimating treatment effects. However, facilitating proper statistical
-inference (i.e.,
-confidence intervals and hypothesis tests) about estimates of effects based on
-adaptive estimators is challenging. In particular, the use of standard causal
-inference techniques, including both G-computation estimators
+inference (i.e., confidence intervals and hypothesis tests) about estimates of
+effects based on adaptive estimators is challenging. In particular, the use of
+standard causal inference techniques, including both G-computation estimators
 [@robins:1986:mathmod] and inverse probability of treatment weighted (IPTW)
@@
 \begin{equation}
-  \psi_0(a) = E_0\{\bar{Q}(a,W)\} = \int \bar{Q}_0(a,w) dQ_{W,0}(w) \ , \
-  \mbox{for} \ a = 0,1 \ ,
+  \psi_0(a) = E_0\{\overline{Q}(a,W)\} = \int \overline{Q}_0(a,w) dQ_{W,0}(w)
+  \ , \ \mbox{for} \ a = 0,1 \ ,
   \label{parameterDef}
 \end{equation}
-where $\bar{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
+where $\overline{Q}_0(a,w)=E_0 \left(Y \mid A=a, W=w\right)$ is the so-called
 outcome regression and $Q_{W,0}(w)=Pr_0\left(W\leq w\right)$ is the distribution
@@
 An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
-Suppose we have available an estimate of the OR, $\bar{Q}_n$, and an estimate of
+Suppose we have available an estimate of the OR, $\overline{Q}_n$, and an estimate of
 the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
 $\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
-$\psi_n(a) = \int \bar{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
+$\psi_n(a) = \int \overline{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
 distribution function of covariates is used so that this estimator can be
@@
 \begin{equation}
-  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n \bar{Q}_n(a, W_i)
-  \ . \label{eqn:gcomp}
+  \psi_{n,\text{GCOMP}}(a) = \frac{1}{n}\sum\limits_{i=1}^n
+  \overline{Q}_n(a, W_i) \ .
+  \label{eqn:gcomp}
 \end{equation}
@@
 \begin{align}
-  E_0\{\bar{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
+  E_0\{\overline{Q}(a,W)\} &= E_0\{E_0(Y \mid A = a, W)\} \notag \\
           &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} E_0(Y \mid A = a,
@@
 The GCOMP and IPTW estimators suffer from two important shortcomings. The first
-is a lack of robustness. That is, validity of GCOMP estimators relies on
+is a lack of robustness -- that is, validity of GCOMP estimators relies on
 consistent estimation of the OR, while validity of the IPTW relies on
@@
 This motivates a new class of estimators that combine both the GCOMP and IPTW
-estimators. One example is the augmented IPTW (AIPTW) estimator
-\begin{equation*}
-\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n} \sum_{i=1}^n
-  \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\} \ \bar{Q}_n(a, W_i)
-  \ ,
+estimators. One prominent example is the augmented IPTW (AIPTW) estimator
+\begin{equation*}
+  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{IPTW}}(a) + \frac{1}{n}
+  \sum\limits_{i=1}^n \biggl\{1 - \frac{I(A_i = a)}{g_n(a \mid W_i)} \biggr\}
+  \overline{Q}_n(a, W_i) \ ,
 \end{equation*}
@@
 \begin{equation} \label{oneStepEst}
-\psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n} \sum_{i=1}^n
-\frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i - \bar{Q}_n(a,W_i)\} \ .
+  \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n}
+  \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
+  \overline{Q}_n(a,W_i)\} \ .
 \end{equation}
 The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
-estimators. The estimator exhibits _double robustness_ in that it is consistent
-for $\psi_0(a)$ if either $\bar{Q}_n$ is consistent for the true OR *or* $g_n$
-is consistent for the true PS. Furthermore, if both the OR and PS are
-consistently estimated and regularity conditions hold, one can establish a
-normal limiting distribution for the AIPTW estimator. Simple estimators of the
-variance of this limiting distribution can be derived that may be used to
-construct Wald-style confidence intervals and hypothesis tests.
+estimators: It exhibits _double robustness_ in that it is consistent for
+$\psi_0(a)$ if either $\overline{Q}_n$ is consistent for the true OR *or* $g_n$ is
+consistent for the true PS. Furthermore, if both the OR and PS are consistently
+estimated and regularity conditions hold, one can establish a normal limiting
+distribution for the AIPTW estimator. Closed-form estimators of the variance of
+this limiting distribution can be derived that may be used to construct
+Wald-style confidence intervals and hypothesis tests.
 
@@
 satisfies user-selected equations. In the present problem, one can show that if
-an estimator of the OR $\bar{Q}_n^{\star}$ satisfies
+an estimator of the OR $\overline{Q}_n^{\star}$ satisfies
 \begin{equation}\label{meanEIFZero}
   \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
-  \bar{Q}_n^{\star}(a, W_i)\} = 0 \ ,
+  \overline{Q}_n^{\star}(a, W_i)\} = 0 \ ,
 \end{equation}
-then the GCOMP estimator based on $\bar{Q}_n^{\star}$ will have the same
+then the GCOMP estimator based on $\overline{Q}_n^{\star}$ will have the same
 asymptotic properties as the AIPTW estimator. That is, the estimator will be
@@
 construct doubly robust confidence intervals and hypothesis tests.
-@benkeser:2017:biometrika proposed simplifications of the @vdl:2014:ijb
+@benkeser:2017:biometrika proposed simplifications of @vdl:2014:ijb's
 estimators and demonstrated the practical performance of these estimators via
@@
 \begin{align}
-  \bar{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \bar{Q}_n(W) \mid A = a,
-  g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qbarr} \\
-  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \bar{Q}_n(W)=\bar{Q}_n(w),
-  g_n(W)=g_n(w)\right\} \ . \label{eqn:gr_bivariate}
+  \overline{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \overline{Q}_n(W) \mid
+  A = a, g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qoverliner} \\
+  g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \overline{Q}_n(W) =
+  \overline{Q}_n(w), g_n(W)=g_n(w)\right\} \ .
+  \label{eqn:gr_bivariate}
 \end{align}
@@
 of whether the OR or PS are consistently estimated. Based on these
-reduced-dimension regressions, @{vdl:2014:ijb showed that if estimates of the
-OR $\bar{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR $\bar{Q}_{r,n}^{\star}$, and
-R-PS $g_{r,n}^{\star}$ satisfied
+reduced-dimension regressions, @vdl:2014:ijb showed that if estimates of the
+OR $\overline{Q}_n^{\star}$, PS $g_n^{\star}$, R-OR
+$\overline{Q}_{r,n}^{\star}$, and R-PS $g_{r,n}^{\star}$ satisfied
 \begin{align}
 & \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
-\{Y_i - \bar{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
-&\frac{1}{n} \sum\limits_{i=1}^n \frac{\bar{Q}_{r,n}(a, W_i)}{g_n(a \mid W_i)}
-\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
+\{Y_i - \overline{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
+&\frac{1}{n} \sum\limits_{i=1}^n \frac{\overline{Q}_{r,n}(a, W_i)}
+{g_n(a \mid W_i)}\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
 \label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
@@
 \biggl\{\frac{g_{r,n}^{\star}(a \mid W_i) - g_n^{\star}(a \mid W_i)}
-{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \bar{Q}_n^{\star}(a, W_i) \} = 0 \ ,
+{g_n^{\star}(a \mid W_i)} \biggr\} \{Y_i - \overline{Q}_n^{\star}(a, W_i) \}
+= 0 \ ,
 \label{tmleQsolve}
 \end{align}
-then the GCOMP estimator based on $\bar{Q}_n^{\star}$ would be doubly robust
-not only with respect to consistency, but also with respect to its limiting
-distribution. The author additionally proposed a TMLE algorithm to map initial
-estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy these key
-equations.
+then the GCOMP estimator based on $\overline{Q}_n^{\star}$ would be doubly
+robust not only with respect to consistency, but also with respect to its
+limiting distribution. The author additionally proposed a TMLE algorithm to map
+initial estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy
+these key equations.
 
@@
 \begin{align*}
-  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \bar{Q}_n(W) =
-  \bar{Q}_n(w) \} \ ,
+  g_{r,0,1}(a \mid w) \coloneqq Pr_0\{A = a \mid \overline{Q}_n(W) =
+  \overline{Q}_n(w) \} \ ,
 \end{align*}
@@
   g_{r,0,2}(a \mid w) \coloneqq E_0\biggl\{\frac{I(A = a) - g_n(a \mid W)}
-  {g_n(a \mid W)} \ \bigg\vert \ \bar{Q}_n(W) = \bar{Q}_n(w) \biggr\} \ ,
+  {g_n(a \mid W)} \ \bigg\vert \ \overline{Q}_n(W) = \overline{Q}_n(w) \biggr\} \ ,
 \end{align*}
@@
 easier to estimate consistently than R-PS. Further, they showed that if
-estimators of the OR, PS, and R-OR satisfy equations (5)-(6), and if,
-additionally, estimators of the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2
-$g_{r,n,2}^{\star}$ satisfy
+estimators of the OR, PS, and R-OR satisfy equations
+(\ref{meanEIFZero})--(\ref{eqn:qoverliner}), and if, additionally, estimators of
+the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2 $g_{r,n,2}^{\star}$ satisfy
 \begin{equation*}
@@
 (a \mid W_i)} g_{r,n,2}^{\star}(a \mid W_i) \ \{Y_i -
-\bar{Q}_n^{\star}(a,W_i)\} = 0 \ ,
-\end{equation*}
-then the GCOMP estimator based on this $\bar{Q}_n^{\star}$ also enjoys a
-doubly-robust limiting distribution. @benkeser:2017:biometrika provided a
+\overline{Q}_n^{\star}(a,W_i)\} = 0 \ ,
+\end{equation*}
+then the GCOMP estimator based on this $\overline{Q}_n^{\star}$ also enjoys a
+doubly robust limiting distribution. @benkeser:2017:biometrika provided a
 TMLE algorithm that produced such estimates and provided closed-form,
-doubly-robust variance estimates.
+doubly robust variance estimates.
 
 \textbf{Remark: } In the previous section, we introduced AIPTW and TMLE as two
-methods for constructing doubly-robust (with respect to consistency) estimators
+methods for constructing doubly robust (with respect to consistency) estimators
 of our parameter of interest. Theoretically, these two frameworks are closely
 related and, practically, the two estimators are generally seen as competitors
-for estimating marginal means. In equation (3), we were able to generate a
-doubly-robust estimator by adding a term to the GCOMP estimator. This term is
-exactly the left-hand-side of the key equation (4), the key equation that is
-solved by the standard TMLE. This is no coincidence; this term is crucial in
-studying the asymptotic properties of the GCOMP, AIPTW, and TML estimators
-(for discussion, see @benkeser:2017:biometrika Section 2.2). Recall that
-a TML estimator achieves doubly-robust consistency by satisfying equation (4),
-while the AIPTW achieves doubly-robust consistency by combining the GCOMP
-estimator with the left-hand-side of this term. Now, we have argued that a TML
-estimator may achieve doubly-robust limiting behavior if it additionally
-satisfies equations (6) and (7). It is thus sensible to wonder whether an
-AIPTW-style estimator could be constructed by combining the left-hand-side of
-these two additional equations with the AIPTW estimator.
-@benkeser:2017:biometrika showed that, surprisingly, this AIPTW-style estimator
-does not achieve the same doubly-robust properties as the TML estimator. Thus,
-while AIPTW and TMLE are generally competitors for producing estimators
-doubly-robust with respect to consistency, for doubly-robust inference, TMLE
-ought to be preferred.
-
-# Implementation of doubly-robust inference
-
-The main function of the package is the eponymous `drtmle()` function. This
-function estimates the treatment-specific marginal mean for user-specified
-levels of a discrete-valued treatment and computes a doubly-robust covariance
-matrix for these estimates. The function implements either the
-estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the
-improved estimators of @benkeser:2017:biometrika
-(`reduction = "univariate"`). The package includes utility functions for
-combining these estimates into estimates of average treatment effects, as well
-as other contrasts, as discussed below.
+for estimating marginal means. In equation (\ref{oneStepEst}), we were able to
+generate a doubly robust estimator by adding a term to the GCOMP estimator. This
+term is exactly the left-hand side of the key equation (\ref{meanEIFZero}), the
+key equation
+that is solved by the standard TMLE. This is no coincidence -- this term is
+crucial in studying the asymptotic properties of the GCOMP, AIPTW, and TML
+estimators (for discussion, see @benkeser:2017:biometrika, Section 2.2). Recall
+that a TML estimator achieves doubly-robust consistency by satisfying equation
+(\ref{meanEIFZero}), while the AIPTW achieves doubly robust consistency by
+combining the GCOMP estimator with the left-hand side of this term. Now, we have
+argued that a TML estimator may achieve doubly robust limiting behavior if it
+additionally satisfies equations (\ref{eqn:qoverliner}) and
+(\ref{eqn:gr_bivariate}). It is thus sensible to wonder whether an AIPTW-style
+estimator could be constructed by combining the left-hand side of these two
+additional equations with the AIPTW estimator. @benkeser:2017:biometrika showed
+that, surprisingly, this AIPTW-style estimator does not achieve the same
+doubly robust properties as the TML estimator. Thus, while AIPTW and TMLE are
+generally competitors for producing estimators doubly robust with respect to
+consistency, for doubly robust inference, TML estimators are preferable.
+
+# Implementation of doubly robust inference
+
+The main workhorse of the \texttt{drtmle} package is the eponymous `drtmle()`
+function. This function estimates the treatment-specific marginal mean for
+user-specified levels of a discrete-valued treatment and computes a doubly
+robust covariance matrix for these estimates. The function implements either the
+estimators of @vdl:2014:ijb (if `reduction = "bivariate"`) or the improved
+estimators of @benkeser:2017:biometrika (if `reduction = "univariate"`). The
+package includes utility functions for combining these estimates into estimates
+of average treatment effects, as well as other contrasts, as discussed below.
 
@@
 arguments, it is necessary to specify how to estimate each of the OR, PS, R-OR,
-and R-PS (or R-PS1, R-PS2). The package provides flexibility in how these
+and R-PS (or R-PS1 and R-PS2). The package provides flexibility in how these
 regression parameters may be estimated. Once the user has specified the data
@@
   4. iteratively modify initial estimates of the OR and PS via the
-     `fluctuateQ()` and `fluctuateG()` functions until equations (5)-(7) are
-     approximately solved;
+     `fluctuateQ()` and `fluctuateG()` functions until equations
+     (\ref{meanEIFZero})--(\ref{eqn:gr_bivariate}) are approximately solved;
   5. compute the TML estimates and covariance estimates based on the final OR.
-We refer interested readers to @benkeser:2017:biometrika for a
-theoretical derivation of how the iterative modification of initial OR and PS is
-performed.
-
-In addition to returning doubly-robust parameter and covariance estimates, the
+We refer interested readers to @benkeser:2017:biometrika for a theoretical
+derivation of how the iterative modification of initial OR and PS is performed.
+
+In addition to returning doubly robust parameter and covariance estimates, the
 `drtmle()` function returns estimates of the GCOMP, AIPTW, standard TMLE, and the
-modified AIPTW estimator discussed in the remark above. While doubly-robust
+modified AIPTW estimator discussed in the remark above. While doubly robust
 inference based on these estimators is not available, the estimators are
-provided for comparison. In the following subsections, we look at various
-options for making calls to `drtmle()`.
+provided for comparison. In the following subsections, we demonstrate the use
+of these methods by examining various options for making calls to `drtmle()`.
 
@@
 of the regression parameters are consistently estimated, then the estimators
-computed by `drtmle()` achieve the nonparametric efficiency bound, resulting in
+computed by `drtmle()` achieve the nonparametric efficiency bound (that is,
+the asymptotic variance of the efficient influence function), resulting in
 narrower confidence intervals and more powerful hypothesis tests. In order for
@@
 @SuperLearnerPackage], and a user-specified estimation technique. We demonstrate
-each of these approaches in turn on a  simulated data set. The true parameter
-values of interest are $\psi_0(0)=$ `r round(EY0, 2)` and $\psi_0(1)=$ `r
-round(EY1, 2)`.
+each of these approaches in turn on a simulated data set in which the treatment
+is not randomized. The true parameter values of interest are $\psi_0(0)=$ `r
+round(EY0, 2)` and $\psi_0(1)=$ `r round(EY1, 2)`.
 
@@
 Generalized linear models are perhaps the most popular means of estimating
-regression functions. Because of this legacy, these estimators have been
+regression functions. On account of this legacy, these estimators have been
 included as an option for estimating the regression parameters. The user
-specifies the right-hand-side of a regression formula as a character in the
-`glm_g`, `glm_Q`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS, R-OR,
-and R-PS. These regressions are fit by calling the `glm()` function in R [@R].
-Thus, the model specification should be able to be parsed as a valid formula,
-as required by `glm()`. For the PS formula specified by `glm_g`, `colnames(W)`
-are available to use in the formula. For the OR (`glm_Q`) available names
-of variables to be used in the formula depends on the `stratify` option.
-If `stratify = FALSE`, a single OR is fit to estimate $\bar{Q}_0(A,W)$. In this
-case, the OR regression is fit by passing a `data.frame` that contains `W` and
-`A` into the `data` argument of `glm()`. Thus, both `colnames(W)` and `"A"` may
-appear in the formula specified in `glm_Q`. If `stratify = TRUE`, regressions
-are fit stratifying on `A` to separately estimate $\bar{Q}_0(1,W)$ and
-$\bar{Q}_0(0,W)$. In this case, the `glm_Q` formula may not include `"A"`.
-
-To understand the variable names available for R-OR, recall from
-(\ref{eqn:qbarr}) that the parameter involves a regression onto the estimated
-propensity score. Thus, the formula that is input via `glm_Qr` should assume
-data with a single column, `"gn"`.  Similarly, the R-PS1 and R-PS2 that are
-computed for the estimators of @benkeser:2017:biometrika (the default
+specifies the right-hand side of a regression formula as a character string
+in the `glm_g`, `glm_Q`, `glm_Qr`, and `glm_gr` options to estimate the OR, PS,
+R-OR, and R-PS. These regressions are fit by calling the `glm()` function from
+\texttt{R}'s base \texttt{stats} package [@R]. Thus, the model specification
+should be able to be parsed as a valid formula object, as required by `glm()`.
+For the PS formula specified by `glm_g`, `colnames(W)` are available to use in the
+formula. For the OR (`glm_Q`) available names of variables to be used in the
+formula depends on the `stratify` option. If `stratify = FALSE`, a single OR is
+fit to estimate $\overline{Q}_0(A,W)$. In this case, the OR regression is fit by
+passing a `data.frame` that contains `W` and `A` into the `data` argument of
+`glm()`. Thus, both `colnames(W)` and `"A"` may appear in the formula specified
+in `glm_Q`. If `stratify = TRUE`, regressions are fit stratifying on `A` to
+separately estimate $\overline{Q}_0(1,W)$ and $\overline{Q}_0(0,W)$. In this
+case, the `glm_Q` formula may not include `"A"`.
+
+To understand the variable names available for R-OR, recall from equation
+(\ref{eqn:qoverliner}) that the nuisance parameter involves a regression onto
+the estimated propensity score. Thus, the formula that is input via `glm_Qr`
+should assume data with a single column, `"gn"`.  Similarly, the R-PS1 and R-PS2
+that are computed for the estimators of @benkeser:2017:biometrika (the default
 estimators, computed whenever `reduction = "univariate"`) involve regressions
@@
 data with a single column named `"Qn"` if `reduction = "univariate"`. Instead,
-if the estimators of @vdl:2014:ijb are preferred (`reduction = "bivariate"`) the
-`glm_gr` formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes data
-with two columns, `"Qn"` and `"gn"`.
+if the estimators of @vdl:2014:ijb are preferred (`reduction = "bivariate"`),
+the `glm_gr` formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes
+data with two columns, `"Qn"` and `"gn"`.
 
@@
 [@wolpert:1992:nnet; @breiman:1996:ml]. The user specifies many potential
-candidate regression estimators, referred to as a *library* of candidate
+candidate regression algorithms, referred to as a *library* of candidate
 estimators, and the super learner algorithm estimates the convex combination of
@@
 estimating regression parameters. This functions fits a kernel regression
-estimator using the \texttt{np} package [@npPackage]. Below we show the source code
-for this function.
+estimator using the \texttt{np} package [@npPackage]. Below we show the source
+code for this function.
 
@@
 Some users may find it unsettling that when applying the super learner in real
-data analysis the results are dependent on the random seed that is set. To
-remove this dependence, one can average results over repeated super learner
-runs; more repeats will lead to less dependence on the random seed. There are
-generally two approaches that can be used: (i) averaging on the scale of the
-nuisance parameters and (ii) averaging on the scale of the point estimates.
+data analysis the results are dependent on the random seed that is set. While
+this is a natural artifact of cross-validation (sample-splitting) and adaptive
+estimation, one can average results over repeated super learner runs to reduce
+or remove this dependence -- more repeats will lead to less dependence on the
+random seed. There are generally two approaches that can be used: (i) averaging
+on the scale of the nuisance parameters and (ii) averaging on the scale of the
+point estimates.
 
 For (i), suppose that we have $n_{SL}$ estimated super learners for the OR,
-$\bar{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS $g_{n,j}, j = 1, \dots,
+$\overline{Q}_{n,j}, j = 1, \dots, n_{SL}$, and the PS $g_{n,j}, j = 1, \dots,
 n_{SL}$. We can take the average over these fits,
 \begin{equation*}
-  \bar{Q}_n = \frac{1}{n_{SL}} \sum_{j = 1}^{n_{SL}} \bar{Q}_{n,j}
+  \overline{Q}_n = \frac{1}{n_{SL}} \sum_{j = 1}^{n_{SL}} \overline{Q}_{n,j}
 \end{equation*}
@@
 This feature is implemented via the `n_SL` and `avg_over` options, where the
-former specifies the number of super learners to repeat and the latter specifies
-the scale(s) to average over. For example, if `avg_over = "drtmle"` and `n_SL =
-2`, then two point estimates are computed, each based on a single super learner,
-and averaged to obtain a final estimate. If instead `avg_over = "SL"` and `n_SL
-= 2`, then two super learners are fit and averaged before a final point estimate
-is obtained. Finally, if `avg_over = c("drtmle", "SL")` and `n_SL = 2` then
-averaging on *both* scales is performed. That is, the final estimate is the
-average of two point estimates, each obtained using OR and PS estimates that are
-averaged over two super learners.
-
-Not surprisingly, averaging over multiple super learners can add considerable
+former specifies the number of super learners to repeat and the latter
+specifies the scale(s) over which to average. For example, if `avg_over =
+"drtmle"` and `n_SL = 2`, then two point estimates are computed, each based on a
+single super learner, and averaged to obtain a final estimate. If instead
+`avg_over = "SL"` and `n_SL = 2`, then two super learners are fit and averaged
+before a final point estimate is obtained. Finally, if `avg_over = c("drtmle",
+"SL")` and `n_SL = 2` then averaging on *both* scales is performed. That is, the
+final estimate is the average of two point estimates, each obtained using OR and
+PS estimates that are averaged over two super learners.
+
+Unsurprisingly, averaging over multiple super learners can add considerable
 computational expense to calls to `drtmle()`. The package is currently
@@
 Reported standard error estimates are based on an estimate of the asymptotic
-variance of the estimators. This variance can be characterized by influence
-function of the estimators. For example, let $\psi_{n}(1)$ be an AIPTW estimate
-of $\psi_0(1)$. Under regularity conditions,
+variance of the estimators. This variance can be characterized by the
+(efficient) influence function of the estimators. For example, let
+$\psi_{n}(1)$ be an AIPTW estimate of $\psi_0(1)$. Under regularity conditions,
 \begin{equation*}
   \psi_{n}(1) - \psi_0(1) = \frac{1}{n} \sum_{i=1}^n
-    D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
+    D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O_i) + o_{p}(n^{-1/2}) \ ,
 \end{equation*}
@@
 \begin{equation*}
-  D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}
-  {g_0(1 \mid W_i)} \{ Y_i - \bar{Q}_0(1, W_i)\} + \bar{Q}_0(1, W_i) -
-  \int \bar{Q}_0(1, w) dQ_{W,0}(w) \ .
+  D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O_i) = \frac{I(A_i = 1)}
+  {g_0(1 \mid W_i)} \{ Y_i - \overline{Q}_0(1, W_i)\} + \overline{Q}_0(1, W_i)
+  - \int \overline{Q}_0(1, w) dQ_{W,0}(w) \ .
 \end{equation*}
@@
 \begin{equation*}
-  \sigma^2_0 = E_0( [ D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O) -
-  E_0\{D^{\star}(\bar{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
+  \sigma^2_0 = E_0( [ D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O) -
+  E_0\{D^{\star}(\overline{Q}_0, g_0, Q_{W,0})(O)\} ]^2 ) \ .
 \end{equation*}
 An estimate of $\sigma^2_0$ is naturally obtained by taking the empirical
-variance of $D_i = D^{\star}(\bar{Q}_n, g_n, Q_{W,n})(O_i)$, that is,
+variance of $D_i = D^{\star}(\overline{Q}_n, g_n, Q_{W,n})(O_i)$, that is,
 \begin{equation*}
@@
 \end{equation*}
-Thus, a natural estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n /
+Thus, an estimate of the standard error of $\psi_{n}(1)$ is $\sigma_n /
 n^{1/2}$. Conveniently, this idea generalizes naturally to estimating the
@@
 For TML estimators, it is standard practice to construct $D_i$ using the
-*targeted* version of the OR (denoted by $\bar{Q}_n^{\star}$ above). Recall that
+*targeted* version of the OR (denoted by $\overline{Q}_n^{\star}$ above). Recall that
 such estimates are obtained by modifying an initial OR estimate so as to ensure
@@
 described in the preceding subsection. To this end, The `drtmle()` function
-includes the `targeted_se` option.
+includes the `targeted_se` option, which specifies whether initial nuisance
+parameter estimates are to be in computing the standard error (i.e.,
+`targeted_se = FALSE`).
 
@@
 
-Simulation studies have revealed that, when employing machine learning to
+Simulation studies have revealed that when employing machine learning to
 estimate the OR and/or PS, the estimated standard errors are often *too small*,
@@
 Suppose we use $V$-fold cross-validation to obtain $V$ training-sample-specific
-estimates of the OR, $\bar{Q}_{n,v}, v = 1, \dots, V$ and the PS, $g_{n,v},
-v = 1, \dots, V$. For $i = 1, \dots, n$, let $\bar{Q}_{n,i}^{cv}(1)$ denote the
-estimate of $\bar{Q}_{0}(a, W_i)$ obtained when observation $i$ was in the
-validation fold. That is, $\bar{Q}_{n,i}^{cv}$ is an out-of-sample prediction of
-$Y_i$. We define $g_{n,i}^{cv}$ similarly. Using these estimates, we can define
-\begin{equation*}
- D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i - \bar{Q}_{n,i}^{cv}(1)\} +
- \bar{Q}_{n,i}^{cv}(1) - \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,i}^{cv}(1)
+estimates of the OR, $\overline{Q}_{n,v}, v = 1, \dots, V$ and the PS,
+$g_{n,v}, v = 1, \dots, V$. For $i = 1, \dots, n$, let
+$\overline{Q}_{n,i}^{cv}(1)$ denote the estimate of $\overline{Q}_{0}(a, W_i)$
+obtained when observation $i$ was in the validation fold. That is,
+$\overline{Q}_{n,i}^{cv}$ is an out-of-sample prediction of $Y_i$. We define
+$g_{n,i}^{cv}$ similarly. Using these estimates, we can define
+\begin{equation*}
+ D_i^{cv} = \frac{I(A_i = 1)}{g_{n,i}^{cv}} \{ Y_i -
+ \overline{Q}_{n,i}^{cv}(1)\} + \overline{Q}_{n,i}^{cv}(1) -
+ \frac{1}{n} \sum_{i=1}^n \overline{Q}_{n,i}^{cv}(1)
 \end{equation*}
@@
 In \texttt{drtmle}, cross-validated standard error estimates can be obtained by
-setting `se_cv = "full"` and `se_cvFolds` to a number greater than 1. For
+setting `se_cv = "full"` and `se_cvFolds` to a number greater than one. For
 example,
@@
 Obtaining fully cross-validated standard errors can be fairly time-intensive
-when super learner is used to estimate the OR and PS. In \texttt{drtmle}, we give
-an option to compute standard error estimates that are *partially
+when super learner is used to estimate the OR and PS. In \texttt{drtmle}, we
+give an option to compute standard error estimates that are *partially
 cross-validated* and do not require any additional fitting beyond a standard
@@
 the OR and PS (in practice, the number could be different for the OR and for the
-PS). Recall that a super learner estimate of, for example, $\bar{Q}_0(a, w)$ is
-obtained as a weighted combination of estimates from each candidate regression.
-Let $\bar{Q}_{n,k}$ denote the estimate of $\bar{Q}_{0}$ obtained by training
-algorithm $k$ using the full data set. The super learner estimate of
-$\bar{Q}_{0}(a,w)$ is
-\begin{equation*}
-  \bar{Q}_{n,SL}(a, w) = \sum_{k = 1}^K \alpha_{n,k} \bar{Q}_{n,k}(a, w) \ ,
+PS). Recall that a super learner estimate of, for example, $\overline{Q}_0(a,
+w)$ is obtained as a weighted combination of estimates from each candidate
+regression. Let $\overline{Q}_{n,k}$ denote the estimate of $\overline{Q}_{0}$
+obtained by training algorithm $k$ using the full data set. The super learner
+estimate of $\overline{Q}_{0}(a,w)$ is
+\begin{equation*}
+  \overline{Q}_{n,SL}(a, w) = \sum_{k = 1}^K \alpha_{n,k}
+  \overline{Q}_{n,k}(a, w) \ ,
 \end{equation*}
@@
 plus one obtained by fitting the algorithm using the full data). Let
-$\bar{Q}_{n,k,v}$ denote the $k$\textsuperscript{th} algorithm trained using the
-$v$\textsuperscript{th} training fold. For $i = 1, \dots, n$, let
-$\bar{Q}_{n,k,i}^{cv}(1)$ denote the estimate of $\bar{Q}_0(1, W_i)$ implied by
-the fitted algorithm $\bar{Q}_{n,k,v}$, which was obtained when observation $i$
-was in the validation sample. That is, for observations with $A_i = 1$,
-$\bar{Q}_{n,k,i}^{cv}(1)$ is an out-of-sample prediction of $Y_i$ based on the
-$k$\textsuperscript{th} algorithm. Now, we can construct a *partially
-cross-validated* super learner prediction as follows. For $i = 1, \dots, n$ let
-\begin{equation*}
-  \bar{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k}
-    \bar{Q}_{n,k,i}^{cv}(1)
+$\overline{Q}_{n,k,v}$ denote the $k$\textsuperscript{th} algorithm trained
+using the $v$\textsuperscript{th} training fold. For $i = 1, \dots, n$, let
+$\overline{Q}_{n,k,i}^{cv}(1)$ denote the estimate of $\overline{Q}_0(1, W_i)$
+implied by the fitted algorithm $\overline{Q}_{n,k,v}$, which was obtained when
+observation $i$ was in the validation set. That is, for observations with
+$A_i = 1$, $\overline{Q}_{n,k,i}^{cv}(1)$ is an out-of-sample prediction of
+$Y_i$ based on the $k$\textsuperscript{th} algorithm. Now, we can construct a
+*partially cross-validated* super learner prediction as follows. For
+$i = 1, \dots, n$ let
+\begin{equation*}
+  \overline{Q}_{n,SL,i}^{cv}(1) = \sum_{k = 1}^K \alpha_{n,k}
+    \overline{Q}_{n,k,i}^{cv}(1)
 \end{equation*}
@@
   D_i^{pcv} = \frac{I(A_i = 1)}{g_{n,SL,i}^{cv}} \{ Y_i -
-    \bar{Q}_{n,SL,i}^{cv}(1)\} + \bar{Q}_{n,SL,i}^{cv}(1) -
-    \frac{1}{n} \sum_{i=1}^n \bar{Q}_{n,SL,i}^{cv}(1),
+    \overline{Q}_{n,SL,i}^{cv}(1)\} + \overline{Q}_{n,SL,i}^{cv}(1) -
+    \frac{1}{n} \sum_{i=1}^n \overline{Q}_{n,SL,i}^{cv}(1),
 \end{equation*}
@@
 list(
-  `No-CV` = sl_fit$cov, # no cv
-  `Partial CV` = sl_fit_pcv$cov # partial cv
+  `No-CV` = sl_fit$drtmle$cov, # no cv
+  `Partial CV` = sl_fit_pcv$drtmle$cov # partial cv
 )
@@
 
-The \texttt{drtmle} package contains methods that compute doubly-robust confidence
-intervals. The `ci()` method computes confidence intervals about the estimated
-marginal means in a fitted object of S3 class `"drtmle"`, in addition to
-contrasts between those means. By default, the method gives confidence intervals
-about each mean.
+The \texttt{drtmle} package contains methods that compute doubly robust
+confidence intervals. The `ci()` method computes confidence intervals about the
+estimated marginal means in a fitted object of S3 class `"drtmle"`, in addition
+to contrasts between those means. By default, the method gives confidence
+intervals about each mean.
 
@@
 \end{equation*}
-where $\sigma^{\log}_n$ is the estimated standard error on the log-scale. By
-the delta method, an estimate of this standard error is given by
+where $\sigma^{\log}_n$ is the estimated standard error on the logarthmic scale.
+By the delta method, an estimate of this standard error is given by
 \begin{equation*}
@@
 \end{equation*}
-where $\Sigma_n$ is the doubly-robust covariance matrix estimate output from
+where $\Sigma_n$ is the doubly robust covariance matrix estimate output from
 `drtmle()` and $\nabla g(\psi)$ is the gradient of $\log\{\psi(1)/\psi(0)\}$.
@@
 
-Doubly-robust Wald-style hypothesis tests may be implemented using the
+Doubly robust Wald-style hypothesis tests may be implemented using the
 `wald_test()` method. As with confidence intervals, there are three options for
@@
 $\Delta_Y$ be indicators that $A$ and $Y$ are observed, respectively. We can
-redefine the OR to be $\bar{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
+redefine the OR to be $\overline{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
 A = a, \Delta_A = 1, \Delta_Y = 1, W = w)$ and similarly redefine the PS to be
@@
   g_0(a \mid w) &= Pr_0(\Delta_A = 1, \Delta_A A = a, \Delta_Y = 1 \mid W = w)
-  \notag \\
+  \nonumber \\
   &= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
   \label{modPS} \\
-  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w) \
-   \notag.
+  &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w)
+   \nonumber.
 \end{align}
 
-To demonstrate, we artificially introduce missing values to `A` and `Y` in our
-running example.
+To demonstrate, we artificially introduce missing values to `A` and `Y` into
+the data from our running example.
 
@@
 
-The only modification of the call to `drtmle()` is in how the PS estimation is
-performed. The options `glm_g` and `SL_g` are still used to fit generalized
-linear models or a super learner, respectively. However, the code allows for
-separate regression fits for each of the three components of the PS in (8). To
-perform separate generalized linear model fits for each of the three components,
-`glm_g` should be a named list with entries `"DeltaA"`, `"A"`, and `"DeltaY"`.
-Respectively, these should provide a formula expression for the regression of
-$\Delta_A$ on $W$, $A$ on $W$ in observations with $\Delta_A = 1$, and of
-$\Delta_Y$ on $W$ and $A$ in observations with $\Delta_A = \Delta_Y = 1$ (if
-`stratify = FALSE`) _or_ of $\Delta_Y$ on $W$ to be fit in observations with
-$\Delta_A = \Delta_Y = 1$ and $A = a$ for each $a$ specified in option `a_0` (if
-`stratify = TRUE`). If only a single formula is passed to `glm_g`, it is used
-for each of the three regressions. We illustrate a call to `drtmle()` with
-missing data and generalized linear model fits below.
+The only required modification of the call to `drtmle()` is in how the PS
+estimation is performed. The options `glm_g` and `SL_g` are still used to fit
+generalized linear models or a super learner, respectively. However, the code
+allows for separate regression fits for each of the three components of the PS
+in equation (\ref{modPS}). To perform separate generalized linear model fits for
+each of the three components, `glm_g` should be a named list with entries
+`"DeltaA"`, `"A"`, and `"DeltaY"`. Respectively, these should provide a formula
+expression for the regression of $\Delta_A$ on $W$, $A$ on $W$ in observations
+with $\Delta_A = 1$, and of $\Delta_Y$ on $W$ and $A$ in observations with
+$\Delta_A = \Delta_Y = 1$ (if `stratify = FALSE`) _or_ of $\Delta_Y$ on $W$ to
+be fit in observations with $\Delta_A = \Delta_Y = 1$ and $A = a$ for each $a$
+specified in option `a_0` (if `stratify = TRUE`). If only a single formula is
+passed to `glm_g`, it is used for each of the three regressions. We illustrate a
+call to `drtmle()` with missing data and generalized linear model fits below.
 
@@
 of the nuisance regressions. Theoretically, the use of cross-validated
-regression weakens the assumptions necessary to achieve asymptotic normality
-[@zheng:2010:working]. This is true of the regular TML and AIPTW estimators, as
-well as of the TML estimator with doubly-robust inference. The `drtmle()`
-function implements cross-validated estimation of the regression parameters
-through the `cvFolds` option, as shown in the code below.
+regression weakens entropy conditions on nuisance parameter estimators
+necessary to achieve asymptotic normality [@klaassen1987consistent;
+@bickel1993efficient; @zheng:2010:working]. This is true of the regular TML and
+AIPTW estimators, as well as of the TML estimator with doubly robust inference.
+The `drtmle()` function implements cross-validated estimation of the nuisance
+regression parameters through the `cvFolds` option, as shown in the code below.
 
@@
 setting the option `use_future = TRUE`. Since, internally, all computations are
-carried out using futures (with the API provided by the \texttt{future} R package),
-setting this option amounts only to parallelizing with futures [@futurePackage]
-via `future.apply::lapply()`; when `use_future = FALSE`, base `lapply()` is
-simply used instead.  Together, setting these options results in parallelization
-of the estimation of the OR, PS, and reduced-dimension regressions.
+carried out using futures (with the API provided by the \texttt{future}
+\texttt{R} package), setting this option amounts only to parallelizing with
+futures [@futurePackage; @bengtsson2021unifying] via `future.apply::lapply()`;
+when `use_future = FALSE`, base `lapply()` is simply used instead. Together,
+setting these options results in parallelization of the estimation of the OR,
+PS, and reduced-dimension regressions.
 
 If no adjustments are made prior to invoking the `drtmle()` function, futures
-are computed sequentially, as per the defaults of the \texttt{future} package. It
-is left to the user to specify appropriate system-specific future backends that
-may better suit their problem (e.g., using the \texttt{future.batchtools} package
-[@futurebatchtoolsPackage] or even simply setting a parallelized
-`future::plan()`). Below, we demonstrate registration using an ad-hoc cluster on
-the local system. The choice of backend is flexible, and submission of the job
-to a wide variety of schedulers (e.g., SLURM, TORQUE) is also permitted.
-Interested readers are invited to consult the documentation of the
-\texttt{future.batchtools} package for further information.
+are computed sequentially, as per the defaults of the \texttt{future} package.
+It is left to the user to specify appropriate system-specific future backends
+that may better suit both their analytical problem and computing environment
+(e.g., using the \texttt{future.batchtools} package [@futurebatchtoolsPackage]
+or even simply setting a parallelized `future::plan()`). Below, we demonstrate
+registration using an ad-hoc cluster on the local system. The choice of backend
+is flexible, and submission of the job to a variety of schedulers (e.g., SLURM,
+TORQUE) is also permitted. Interested readers are invited to consult the
+documentation of the \texttt{future.batchtools} package for further information.
 
 ```{r, message = FALSE, eval = FALSE}
+# set up parallel backend via future.batchtools
 library(future.batchtools)
@@
 clusterEvalQ(cl, library("SuperLearner"))
+
+# run drtmle() in parallel using the registered backend via use_future = TRUE
 pcv_sl_fit <- drtmle(W = W, A = A, Y = Y, family = binomial(),
@@
 
-The \texttt{drtmle} package was designed to provide a user-friendly implementation
-of the TMLE algorithm that facilitates doubly-robust inference about marginal
-means and average treatment effects. While the theory underlying doubly-robust
-inference is complex, the implementation of the methods only requires users to
-provide the data and specify how to estimate nuisance regressions. Though
-estimation of the OR and PS is familiar to those with backgrounds in causal
-inference-related estimation, the reduced-dimension regressions may appear
-strange. Indeed, it is difficult to provide intuition for these parameters and
-more difficult still to determine what scientific background knowledge might
-guide users in how to estimate these regression parameters. Therefore, we
-recommended estimating these quantities adaptively, e.g., with the super
-learner. The super learner library should contain several relatively smooth
-estimators such as `SL.mean`, `SL.glm`, and `SL.gam`. In particular, including
-`SL.mean` (an unadjusted generalized linear model) may be important due to the
-fact that, when the OR and PS are consistently estimated, the reduced-dimension
-regressions are identically equal to zero. Thus, this unadjusted regression
-estimator may do a good job estimating the reduced-dimension regressions in
-situations where the OR and PS are estimated well. In addition to relatively
-smooth estimators, we recommend additionally including several adaptive
-estimators such as `SL.npreg` or `SL.earth`.
+The \texttt{drtmle} package was designed to provide a user-friendly
+implementation of the TMLE algorithm that facilitates doubly robust inference
+about marginal means and average treatment effects. While the theory underlying
+doubly robust inference is complex, the implementation of the methods only
+requires users to provide the data and specify how to estimate nuisance
+regressions. Though estimation of the OR and PS is familiar to those with
+backgrounds in causal inference-related estimation, the reduced-dimension
+regressions may appear strange. Indeed, it is difficult to provide intuition for
+these parameters and more difficult still to determine what scientific
+background knowledge might guide users in how to estimate these regression
+parameters. Therefore, we recommended estimating these quantities adaptively,
+e.g., with the super learner algorithm. The super learner library should contain
+several relatively smooth estimators such as `SL.mean`, `SL.glm`, and `SL.gam`.
+In particular, including `SL.mean` (an unadjusted generalized linear model) may
+be important due to the fact that, when the OR and PS are consistently
+estimated, the reduced-dimension regressions are identically equal to zero.
+Thus, this unadjusted regression estimator may do a good job estimating the
+reduced-dimension regressions in situations where the OR and PS are estimated
+well. In addition to relatively smooth estimators, we recommend additionally
+including several adaptive estimators such as `SL.npreg` or `SL.earth`.
 
@@
 of the collaborative targeted maximum likelihood estimator (C-TMLE) of
-@vdl:2014:ijb that also facilitates collaborative doubly-robust inference.
+@vdl:2014:ijb that also facilitates collaborative doubly robust inference. The
+\texttt{drtmle} \texttt{R} package is publicly available on the Comprehensive
+\texttt{R} Archive Network (CRAN) at https://CRAN.R-project.org/package=drtmle,
+and the open source development repository is available on the GitHub platform
+for collaborative programming at https://github.com/benkeser/drtmle; the
+package's documentation (automatically built via the \texttt{pkgdown} \texttt{R}
+package) is hosted via GitHub Pages at https://benkeser.github.io/drtmle/.
 
@@
 
-The \texttt{drtmle} package also allows users to pass in their own estimated OR and
-PS via the `Qn` and `gn` options, respectively.  If `Qn` is specified by the
+The \texttt{drtmle} package also allows users to pass in their own estimated OR
+and PS via the `Qn` and `gn` options, respectively. If `Qn` is specified by the
 user, `drtmle()` will ignore the nuisance parameter estimation routines
@@
 
-## Additional methods included in drtmle
+## Additional methods included in \texttt{drtmle}
 
@@
 
-Doubly-robust estimators of marginal means are appealing in their robustness and
-efficiency properties. However, researchers may prefer the IPTW estimator for
+Doubly robust estimators of marginal means are appealing in their robustness and
+efficiency properties; however, researchers may prefer the IPTW estimator for
 its intuitive derivation and implementation. Heretofore, inference for IPTW
-estimators precluded the use of adaptive estimators of the PS. However,
-@vdl:2014:ijb proposed modified IPTW estimators that allow for adaptive
-PS estimation. Similarly as above, this estimator is based on an additional
-regression parameter, $\tilde{Q}_{r,0n}(a,w) \coloneqq E_0\{Y \mid A = a, g_n(W)
-= g_n(w)\}$, that is, the regression of the outcome $Y$ on the estimated
-propensity score amongst observations with $A=a$. The author showed that if
-$g_n^{\star}(a \mid w)$, an estimator of the propensity for treatment $A=a$,
-satisfied that
+estimators precluded the use of adaptive estimators of the PS, yet @vdl:2014:ijb
+proposed modified IPTW estimators that allow for adaptive PS estimation by
+incorporating an additional step to update the initial PS estimates. Like those
+presented above, this estimator is based on an additional regression parameter,
+$\tilde{Q}_{r,0n}(a,w) \coloneqq E_0\{Y \mid A = a, g_n(W) = g_n(w)\}$, that is,
+the regression of the outcome $Y$ on the estimated propensity score amongst
+observations with $A=a$. The author showed that if $g_n^{\star}(a \mid w)$, an
+estimator of the propensity for treatment $A=a$, satisfied that
 \begin{equation*}
@@
 \end{equation*}
-then the IPTW estimator based on $g_n^{\star}$ would have an asymptotic normal
-distribution with estimable variance. In fact, @vdl:2014:ijb proposed a
-TMLE algorithm that mapped an initial PS estimate into an estimate that
+then the IPTW estimator based on $g_n^{\star}$ would asymptotically attain a normal
+limiting distribution with estimable variance. In fact, @vdl:2014:ijb proposed a
+TMLE-type algorithm that mapped an initial PS estimate into an estimate that
 satisfies this key equation and variance estimators that can be used to generate
@@
 \begin{equation*}
-  \psi_{n,IPTW}(a) = \psi_{n,IPTW}(a) - \frac{1}{n} \sum\limits_{i=1}^n
-  \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)} \ \{ I(A_i = a) - g_n(a
-  \mid W = W_i)\} \ .
+  \psi_{n,\text{IPTW}}(a) = \psi_{n,\text{IPTW}}(a) - \frac{1}{n}
+  \sum\limits_{i=1}^n \frac{\tilde{Q}_{r,0}(a,W_i)}{g_n(a \mid W = W_i)}
+  \{ I(A_i = a) - g_n(a \mid W = W_i)\} \ .
 \end{equation*}
 Both this estimator and the TMLE-based IPTW estimator above are implemented in
-\texttt{drtmle}.
+\texttt{drtmle}. Alternative approaches for developing IPTW estimators that are
+asymptotically normal and attain the nonparametric efficiency bound have been
+investigated by @ertefaie2022nonparametric and @hejazi2022efficient, who relied
+on undersmoothing-based corrections rather than the TMLE-type corrections of
+@vdl:2014:ijb.
 
@@
 
-We have recently proposed estimators of the average treatment effect to be used
-in settings where this effect is weakly identifiable, as evidenced by estimated
-propensity scores of small magnitude.  In these settings, doubly robust
+@ju2020robust have recently proposed estimators of the average treatment effect
+to be used in settings where this effect is weakly identifiable, as evidenced by
+estimated propensity scores of small magnitude. In these settings, doubly robust
 estimators may exhibit non-robust behavior in small samples. The proposed
-solution was to target an alternative quantity in the propensity score
-estimation step. If we are interested in estimating $\psi_0(a)$ for $a = 0, 1$,
-we can estimate $Pr_0(A = a \mid \bar{Q}_0(1, W), \bar{Q}_0(0, W))$, which
+solution is to target an alternative quantity in the propensity score estimation
+step. If we are interested in estimating $\psi_0(a)$ for $a = 0, 1$, we can
+estimate $Pr_0(A = a \mid \overline{Q}_0(1, W), \overline{Q}_0(0, W))$, which
 describes the conditional probability of receiving treatment $A = a$ given the
@@
 *super-efficient*, meaning that it will generally have greater asymptotic
-efficiency than a standard TML estimator. Simulation experiments indicate that
-this approach also delivers more stable behavior in settings where estimated
-propensity scores become very small. However, the estimator is not doubly robust
--- the asymptotics rely on the outcome regression being consistently estimated
-at $n^{-1/4}$ rate. In this sense, this estimator trades off robustness against
+efficiency than a standard TML estimator (in fact, greater asymptotic efficiency
+than is ordinarily available to any regular asymptotically linear estimator).
+Simulation experiments indicate that this approach also delivers more stable
+behavior in settings where estimated propensity scores become very small.
+However, the estimator is not doubly robust -- the asymptotics rely on the
+outcome regression being consistently estimated at $n^{-1/4}$ rate. The highly
+adaptive lasso estimator [@benkeser2016highly; @vdl2017generally;
+@bibaut2019fast], implemented in the open source \texttt{hal9001} \texttt{R}
+package [@hejazi2020hal9001; @coyle2021hal9001], can attain such a convergence
+rate without sacrificing flexible estimation; nevertheless, as a consequence of
+this convergence rate requirement, this estimator trades off robustness against
 achieving greater stability and efficiency. See @ju2020robust for further
@@
 
-# Acknowledgements
-
-NSH is grateful for the support of the National Science Foundation (award no.
-DMS 2102840).
+# Acknowledgements {-}
+
+NSH grateful acknowledges the support of the National Science Foundation (award
+no. [DMS 2102840](https://www.nsf.gov/awardsearch/showAward?AWD_ID=2102840)).
 

### 2022-12-29 paper/reviews/response.md
updates to address reviewer comments

Reviewer 1

I have enjoyed reading about the authors’ software, which I believe has the potential to lead to more robust inferences in observational studies. The following are some relatively minor comments.

Sections 2.2, 2.3, and 2.4– GCOMP, IPTW and AIPTW are commonly used estimators in applications. To improve the readability of the manuscript, I would perhaps shorten and combine sections 2.2, 2.3 and 2.4 into just one section describing existing estimators.

_Response_: Thanks for the suggestion. We have shortened and combined these sections.

Section 3– Unless I am missing something, I would mention that, in addition to the specific moment conditions described by the authors, the estimators need to satisfy also additional and untestable small-bias requirements in order to achieve doubly-robust asymptotic linearity. For example, the second-order terms in Appendix B in Benkeser et al (2017) need to be asymptotically negligible.

_Response_: Thanks, we have added references to the exact regularity conditions as they appear in the original manuscripts. We have also added the sentence

	> The regularity conditions required to achieve asymptotic linearity include assumptions on the convergence rates of the OR, PS, R-OR, and R-PS to their true counterparts."

Page 6, line starting with “The key theory…”– Should the estimators of the OR and PS satisfy equation (5) rather than (4)?

Page 7, line starting with “Further, they showed…”– Should the estimators of the OR, PS, R-OR satisfy equations (8)-(9) instead of (5)-(6)?

_Response_: Thanks for pointing out. There was some Latex/bookdown referencing gore happening with the original submission. We appreciate the reviewers catching this. It has been corrected in the revision.

Because of the crucial role that the reduced univariate regressions play in obtaining doubly-robust inference, I wonder whether it could be useful to have a simple way of checking the model fit for their estimators. For instance, would it make sense to have functions that can plot the residuals and fitted values against the estimated covariate?

_Response_: The package includes a `plot` method that can be used towards this end. We refer to this method in Section 4.1.1.

Reviewer 2

In this software tutorial, the authors detail and implement state-of-the-art nonparametric estimators of counterfactual means, or the average treatment effect (ATE), from observational studies, together with valid doubly robust statistical inferential tools (confidence intervals and hypothesis tests). Standard inference for doubly robust estimators of the ATE (e.g., AIPW or TML estimators) that employ flexible nuisance estimators is not strictly doubly robust in the same sense as the estimators themselves, i.e., consistency of the estimated ATE only requires consistency of one of two nuisance functions, but this robustness does not apply to asymptotic linearity/normality. That said, valid doubly robust inferential procedures have been proposed by van der Laan (2014) and Benkeser et al. (2017), in which additional nuisance models must be fit and satisfy certain estimating equations. However, these procedures are challenging and involved, thus the drtmle software package fills an important gap by implementing these methods. Beyond these approaches, which extend the usual TML estimator (drtmle function), the authors also include extensions to cross-validated TMLE (cvFolds option, i.e., using sample-splitting), and implement an IPW estimator with valid inference when using a flexible estimator for the propensity score (adaptive_iptw function).

The article does a good job of motivating and describing the underlying methodology well for readers familiar with doubly robust estimators of the ATE, but not with the doubly robust inference literature. Overall, the manuscript is also very clearly written, and the package seems to have a nice design. Specific comments are detailed below.

The workhorse functions (drtmle and adaptive_iptw) allow for missingness in treatment (A) and outcome (Y) values (e.g., with DeltaA and DeltaY control parameters within ‘glm_g’). This bivariate missingness setting doesn’t appear (unless I’m missing something) to have been formally considered in either van der Laan (2014) or Benkeser et al. (2017). I am concerned whether such missingness may lead to non-identifiability of the counterfactual mean—or “treatment-specific marginal mean” as referred to by the authors—functionals, without further assumptions. Are the authors implicitly invoking concrete missingness assumptions under which the implemented approaches target the original functional of interest (e.g., something like Delta_A x Delta_Y independent of (A, Y) given W)? Or are the methods simply standard, or used for convenience? It would enhance the paper if this issue were clarified.

_Response_: Thanks for the suggestion. We have added the formal assumptions needed for identification to that section (copied here):

if we define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively, then we can still perform valid inference on $\psi_0$ under the assumptions that (i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
(iii) $Y(a) \perp \Delta_Y \mid \Delta_A = 1, A = a, W$.

Related to point 1., why is missingness in the covariates (W) not considered? I am aware that this would be a very challenging scenario methodologically (and this is a good enough reason to not deal with it), but I am just curious whether there is something fundamentally different about W in this setting which precludes consideration of it being partially missing. Any guidance on what a practitioner should do if W is partially missing?

_Response_: Thanks for the suggestion. We have added a brief paragraph on missingness in $W$.

In section 4.2 on standard error estimation, the discussion is focused on the AIPW and standard TMLE (i.e., not the DR-TMLE implemented in the package). They say “The ideas generalize immediately to the DR-TML estimator, but the formulas are more complex in this latter case”. Presumably, that means that inference within drtmle is based on the empirical variance of the influence function of the DR-TMLE, but authors should elaborate at least a little. If the formulas are too complex to be presented within the manuscript, readers should be referred to the relevant sections of van der Laan (2014) and/or Benkeser et al. (2017).

_Response_: Thanks for the suggestion. We have restructured this section a bit and refer to the specific location of the formula in Benkeser et al (2017).

Several times, the equations (6) and (7) are referred to as if they were estimating equations, but it seems like they are definitions of new nuisance functions (at the “population” level).

a. On page 7, “Further, they showed that if estimators of the OR, PS, and R-OR satisfy equations (5)–(6)”. Equation (5) is empirical, but 6 is not. Should this say equations (8)—(9)?

b. On page 8, “Now, we have argued that a TML estimator may achieve doubly robust limiting behavior if it additionally satisfies equations (6) and (7)”. Should this say equations (9) and (10)?

c. On Page 8, “iteratively modify initial estimates of the OR and PS via the fluctuateQ() and fluctuateG() functions until equations (5)–(7) are approximately solved”. Should this say equations (8)—(10)?

_Response_: All equation referencing has been corrected. Sorry for the confusion.

Other specific minor comments:

Page 3, “relying upon Neymon orthogonality” should say Neyman?

_Response_: Correct. Thanks for catching.

Page 4, “Under additional assumptions, these parameters have interpretations…”. Should at least give a reference here to the causal assumptions.

_Response_: We have added a few short sentences on causal assumptions.

Page 7, equation (9), I think the Q^bar term in the numerator and g_n term in the denominator should each have an asterisk.

_Response_: Yes, thanks. Good catch.

Page 9, “r round(EY0, 2)” looks like unrendered markdown. What is the numerical value?

_Response_: This has been corrected.

Page 18, “methods that compute doubly robust confidence intervals”. Can this be clarified to “Wald-style” confidence intervals? It seems from context that these are just based directly on the standard error estimates from the previous section, but would be good to clarify in opening paragraph.

_Response_: Done

### 2022-12-29 paper/using_drtmle_obstudies.Rmd
updates to address reviewer comments

@@
 effects within the double machine learning framework [@chernozhukov2017double;
-@chernozhukov2018double], reyling upon Neymon orthogonality and sample-splitting
+@chernozhukov2018double], reyling upon Neyman orthogonality and sample-splitting
 for robustness. From among this non-exhaustive list, \texttt{drtmle} stands out
@@
   \ , \ \mbox{for} \ a = 0,1 \ ,
-  \label{parameterDef}
+  (\#eq:parameterDef)
 \end{equation}
@@
 different treatment levels. For example, the average treatment effect is often
-defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$. Under additional
-assumptions, these parameters have interpretations as mean counterfactual
-outcomes under the different treatments, with contrasts between these means
-defining causal effects.
-
-## G-computation estimators
-
-An estimator of $\psi_0(a)$ may be motivated directly by (\ref{parameterDef}).
+defined as $\psi_0 \coloneqq \psi_0(1) - \psi_0(0)$.
+
+Under additional assumptions, these parameters identify the mean counterfactual
+outcomes under the different treatments. Specifically, we denote by $Y(a)$ a well-defined 
+counterfactual outcome that would be observed under treatment $a$.
+If the conditional randomization assumption ($Y(a) \perp A \mid W$) and positivity
+conditions ($Pr_0\{ Pr_0(A = a \mid W) > 0 \} = 1$) hold, then $\psi(a)$ identifies
+the mean of $Y(a)$.
+
+## Estimators
+
+The G-computation estimator of $\psi_0(a)$ may be motivated directly by \@ref(eq:parameterDef).
 Suppose we have available an estimate of the OR, $\overline{Q}_n$, and an estimate of
 the distribution function of baseline covariates, $Q_{W,n}$. An estimator of
-$\psi_0(a)$ may be obtained by plugging these into (\ref{parameterDef}),
+$\psi_0(a)$ may be obtained by plugging these into \@ref(eq:parameterDef),
 $\psi_n(a) = \int \overline{Q}_n(a,w) dQ_{W,n}(w)$. Typically, the empirical
@@
   \overline{Q}_n(a, W_i) \ .
-  \label{eqn:gcomp}
+  (\#eq:gcomp)
 \end{equation}
-This estimator is commonly referred to as the G-computation (GCOMP) estimator.
 Inference for GCOMP estimators based on parametric OR estimators may be
@@
 
-## Inverse probability of treatment estimators
-
 An alternative estimator of the treatment-specific population-level mean may be
-obtained by noting that (\ref{eqn:gcomp}) can be equivalently written as
+obtained by noting that \@ref(eq:gcomp) can be equivalently written as
 \begin{align}
@@
           &= E_0\biggl\{\frac{I(A=a)}{Pr_0(A = a \mid W)} Y \biggr\} \ .
-  \label{iptwID}
+  (\#eq:iptwID)
 \end{align}
@@
 score. Suppose we had available $g_n$, an estimate of the PS. Equation
-(\ref{iptwID}) motivates the estimator
+\@ref(eq:iptwID) motivates the estimator
 \begin{equation*}
@@
 estimator. Like GCOMP estimators, inference for IPTW estimators based on
-parametric PS estimators may be facilitated through the nonparametric bootstrap;
-however, unlike their GCOMP counterparts, IPTW estimators, being Z-estimators,
-are constructed as solutions to an IPTW estimating equation, the empirical
-variance of which may also be used for standard error estimation. If adaptive
-approaches are used to estimate the PS, inference may not be available without
+parametric PS estimators may be facilitated through the nonparametric bootstrap. 
+If adaptive approaches are used to estimate the PS, inference may not be available without
 further modification of the propensity score estimator, whether via targeting,
-as we discuss below, or undersmoothing, as explored by
-@ertefaie2022nonparametric and @hejazi2022efficient.
-
-## Doubly robust estimators
+as we discuss below, or undersmoothing (as in @ertefaie2022nonparametric and @hejazi2022efficient).
 
@@
 
-This motivates a new class of estimators that combine both the GCOMP and IPTW
-estimators. One prominent example is the augmented IPTW (AIPTW) estimator
+The augmented IPTW (AIPTW) estimator
 \begin{equation*}
@@
 \end{equation*}
-which adds an augmentation term to the IPTW estimator. Equivalently, the
+adds an augmentation term to the IPTW estimator. Equivalently, the
 estimator can be viewed as adding an augmentation to the GCOMP estimator
-\begin{equation} \label{oneStepEst}
+\begin{equation} 
   \psi_{n,\text{AIPTW}}(a) = \psi_{n,\text{GCOMP}}(a) + \frac{1}{n}
   \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
-  \overline{Q}_n(a,W_i)\} \ .
+  \overline{Q}_n(a,W_i)\} \ . (\#eq:oneStepEst)
 \end{equation}
 The AIPTW estimator overcomes the two limitations of the GCOMP and IPTW
-estimators: It exhibits _double robustness_ in that it is consistent for
+estimators. First, it exhibits _double robustness_ in that it is consistent for
 $\psi_0(a)$ if either $\overline{Q}_n$ is consistent for the true OR *or* $g_n$ is
@@
 an estimator of the OR $\overline{Q}_n^{\star}$ satisfies
-\begin{equation}\label{meanEIFZero}
+\begin{equation}
   \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n(a \mid W_i)} \{Y_i -
-  \overline{Q}_n^{\star}(a, W_i)\} = 0 \ ,
+  \overline{Q}_n^{\star}(a, W_i)\} = 0 \ , (\#eq:meanEIFZero)
 \end{equation}
@@
 TMLE provides a means of mapping an initial estimator of the OR into an
-estimator that satisfies (\ref{oneStepEst}). In addition to possessing the
+estimator that satisfies \@ref(eq:meanEIFZero). In addition to possessing the
 desirable asymptotic properties, TML estimators have been shown to outperform
@@
 The key to the theory underlying these results is finding estimators of the OR
-and PS that simultaneously satisfy equation (\ref{oneStepEst}) in addition to
+and PS that simultaneously satisfy equation \@ref(eq:meanEIFZero) in addition to
 two additional equations. These equations involve additional regression
@@
   \overline{Q}_{r,0n}(a,w) &\coloneqq E_{0}\bigl\{Y - \overline{Q}_n(W) \mid
-  A = a, g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} \label{eqn:qoverliner} \\
+  A = a, g_n(W)=g_n(w)\bigr\} \ , \ \mbox{and} (\#eq:qoverliner) \\
   g_{r,0n}(a \mid w) &\coloneqq Pr_0\left\{A = a \mid \overline{Q}_n(W) =
   \overline{Q}_n(w), g_n(W)=g_n(w)\right\} \ .
-  \label{eqn:gr_bivariate}
+  (\#eq:grbivariate)
 \end{align}
@@
 & \frac{1}{n} \sum\limits_{i=1}^n \frac{I(A_i = a)}{g_n^{\star}(a \mid W_i)}
-\{Y_i - \overline{Q}_n^{\star}(a,W_i)\} = 0 \ , \label{tmleEIFSolve} \\
-&\frac{1}{n} \sum\limits_{i=1}^n \frac{\overline{Q}_{r,n}(a, W_i)}
-{g_n(a \mid W_i)}\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
-\label{tmleAsolve} \\ &\frac{1}{n} \sum\limits_{i=1}^n
+\{Y_i - \overline{Q}_n^{\star}(a,W_i)\} = 0 \ , (\#eq:tmleEIFSolve) \\
+&\frac{1}{n} \sum\limits_{i=1}^n \frac{\overline{Q}_{r,n}^{\star}(a, W_i)}
+{g_n^{\star}(a \mid W_i)}\{I(A_i = a) - g_n^{\star}(a \mid W_i)\} = 0 \ , \ \mbox{and}
+(\#eq:tmleAsolve) \\ &\frac{1}{n} \sum\limits_{i=1}^n
 \frac{I(A_i = a)}{g_{r,n}^{\star}(a \mid W_i)}
@@
 = 0 \ ,
-\label{tmleQsolve}
+(\#eq:tmleQsolve)
 \end{align}
-then the GCOMP estimator based on $\overline{Q}_n^{\star}$ would be doubly
+then, under regularity conditions described in Theorem 3 of @vdl:2014:ijb, 
+the GCOMP estimator based on $\overline{Q}_n^{\star}$ would be doubly
 robust not only with respect to consistency, but also with respect to its
@@
 initial estimates of the OR, PS, R-OR, and R-PS into estimates that satisfy
-these key equations.
+these key equations. The regularity conditions required to achieve asymptotic
+linearity include assumptions on the convergence rates of the OR, PS, R-OR, and R-PS
+to their true counterparts.
 
@@
 estimators of the OR, PS, and R-OR satisfy equations
-(\ref{meanEIFZero})--(\ref{eqn:qoverliner}), and if, additionally, estimators of
+\@ref(eq:tmleEIFSolve)--\@ref(eq:tmleAsolve), and if, additionally, estimators of
 the OR, R-PS-1 $g_{r,n,1}^{\star}$, and R-PS-2 $g_{r,n,2}^{\star}$ satisfy
@@
 related and, practically, the two estimators are generally seen as competitors
-for estimating marginal means. In equation (\ref{oneStepEst}), we were able to
+for estimating marginal means. In equation \@ref(eq:oneStepEst), we were able to
 generate a doubly robust estimator by adding a term to the GCOMP estimator. This
-term is exactly the left-hand side of the key equation (\ref{meanEIFZero}), the
+term is exactly the left-hand side of the key equation \@ref(eq:meanEIFZero), the
 key equation
@@
 that a TML estimator achieves doubly-robust consistency by satisfying equation
-(\ref{meanEIFZero}), while the AIPTW achieves doubly robust consistency by
+\@ref(eq:meanEIFZero), while the AIPTW achieves doubly robust consistency by
 combining the GCOMP estimator with the left-hand side of this term. Now, we have
 argued that a TML estimator may achieve doubly robust limiting behavior if it
-additionally satisfies equations (\ref{eqn:qoverliner}) and
-(\ref{eqn:gr_bivariate}). It is thus sensible to wonder whether an AIPTW-style
+additionally satisfies equations \@ref(eq:tmleAsolve) and
+\@ref(eq:tmleQsolve). It is thus sensible to wonder whether an AIPTW-style
 estimator could be constructed by combining the left-hand side of these two
@@
 as follows:
+
   1. estimate the OR via the internal `estimateQ()` function;
@@
      `fluctuateQ()` and `fluctuateG()` functions until equations
-     (\ref{meanEIFZero})--(\ref{eqn:gr_bivariate}) are approximately solved;
+     \@ref(eq:tmleEIFSolve)--\@ref(eq:tmleQsolve) are approximately solved;
   5. compute the TML estimates and covariance estimates based on the final OR.
-We refer interested readers to @benkeser:2017:biometrika for a theoretical
+
+\noindent We refer interested readers to @benkeser:2017:biometrika for a theoretical
 derivation of how the iterative modification of initial OR and PS is performed.
@@
 each of these approaches in turn on a simulated data set in which the treatment
-is not randomized. The true parameter values of interest are $\psi_0(0)=$ `r
-round(EY0, 2)` and $\psi_0(1)=$ `r round(EY1, 2)`.
+is not randomized. The true parameter values of interest are $\psi_0(0)=$ `r round(EY0, 2)` 
+and $\psi_0(1)=$ `r round(EY1, 2)`.
 
@@
 To understand the variable names available for R-OR, recall from equation
-(\ref{eqn:qoverliner}) that the nuisance parameter involves a regression onto
+\@ref(eq:qoverliner) that the nuisance parameter involves a regression onto
 the estimated propensity score. Thus, the formula that is input via `glm_Qr`
@@
 if the estimators of @vdl:2014:ijb are preferred (`reduction = "bivariate"`),
-the `glm_gr` formula specifies a model for (\ref{eqn:gr_bivariate}) and assumes
+the `glm_gr` formula specifies a model for \@ref(eq:grbivariate) and assumes
 data with two columns, `"Qn"` and `"gn"`.
@@
                       glm_g = "W1 + W2", glm_Q = "W1 + W2*A",
-                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE)
+                      glm_gr = "Qn", glm_Qr = "gn", stratify = FALSE,
+                      returnModels = TRUE)
 glm_fit_uni
@@
                       glm_gr = "Qn + gn", glm_Qr = "gn", stratify = FALSE,
+                      returnModels = TRUE,
                       reduction = "bivariate")
@@
 ```
+
+The fit of the reduced-dimension regressions can be plotted using the 
+post-hoc `plot.drtmle` method.
 
@@
 skipped over by users interested only in implementation. For simplicity, we
-focus this discussion on the standard TML and AIPTW estimators. The ideas
-generalize immediately to the DR-TML estimator, but the formulas are more
-complex in this latter case.
+focus this discussion on the standard TML and AIPTW estimators.
 
@@
 asymptotic covariance matrix of a vector $n^{1/2} (\psi_{n}(a) : a)$.
+
+These ideas generalize immediately to the DR-TML estimator, where we can use
+the variance of the estimated influence function of that estimator. Unfortunately, 
+the formulas for the influence functions more complex, but are available in 
+Theorem 1 of @benkeser:2017:biometrika.
 
@@
 
-The \texttt{drtmle} package contains methods that compute doubly robust
+The \texttt{drtmle} package contains methods that compute Wald-style doubly robust
 confidence intervals. The `ci()` method computes confidence intervals about the
@@
 The `contrast` option also accepts as input a list of functions to be used to
-compute a confidence interval for a user-specified contrast. For example, we
+compute a Wald-style confidence interval for a user-specified contrast. For example, we
 might be interested in the risk ratio $\psi_0(1) / \psi_0(0)$. When
@@
 means of interest. The estimators previously discussed can be modified to allow
-such missingness processes to depend on observed covariates. Let $\Delta_A$ and
-$\Delta_Y$ be indicators that $A$ and $Y$ are observed, respectively. We can
+such missingness processes to depend on observed covariates. That is, if we 
+define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively,
+then we can still perform valid inference on $\psi_0$ under the assumptions that 
+(i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
+(iii) $Y(a) \perp \Delta_Y \mid \Delta_A = 1, A = a, W$. 
+In this case, we can
 redefine the OR to be $\overline{Q}_0(a,w) \coloneqq E_0(\Delta_Y Y \mid \Delta_A
@@
   &= Pr_0(\Delta_A = 1 \mid W) Pr_0(\Delta_A A = a \mid \Delta_A = 1, W = w)
-  \label{modPS} \\
+  (\#eq:modPS) \\
   &\hspace{1.5in} Pr_0(\Delta_Y = 1 \mid \Delta_A = 1, \Delta_A A = a, W = w)
@@
 allows for separate regression fits for each of the three components of the PS
-in equation (\ref{modPS}). To perform separate generalized linear model fits for
+in equation \@ref(eq:modPS). To perform separate generalized linear model fits for
 each of the three components, `glm_g` should be a named list with entries
@@
 ```
+
+\textbf{Remark: } In practice it is common to also have missingness in $W$. In this
+situation, there are varying assumptions under which one can still identify and 
+estimate $\psi_0$. The only native solution in `drtmle` currently is to augment $W$
+with columns of missingness indicators for each component of $W$. For example, if
+variable $W_1$ is observed, we could define a new indicator $\Delta_{W,1} = 1$ and 
+$\Delta_{W,1} = 0$ otherwise. For observations with $\Delta_{W,1} = 0$, we could use
+some (possibly very simple) imputation strategy to fill in a value for $W_1$. Candidate
+learners for the OR and PS may then include appropriate cross-products between $W_1$ 
+and $\Delta_{W,1}$. Other learners may opt to treat the imputed value of $W_1$ as 
+the true value and ignore $\Delta_{W,1}$. Super learner can be used to help determine
+the modeling strategy that leads to the optimal model fit conditional on $W_1$ and
+$\Delta_{W,1}$. It is important to note that this approach is only valid for drawing 
+causal inference under the assumption that $Y(a) \perp A \mid \Delta_W W$, which may 
+not be plausible depending on the form of missingness of $W$. Other approaches for handling
+missing data, e.g., multiple imputation, could in theory be coupled with `drtmle`, 
+though such applications have not been studied empirically and are not available natively
+in the `drtmle` package.
 

### 2023-01-04 paper/reviews/response.md
finalize revision

@@
 
-if we define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively, then we can still perform valid inference on $\psi_0$ under the assumptions that (i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
+if we define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively, then we can still perform valid inference on $E[Y(a)]$ under the assumptions that (i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
 (iii) $Y(a) \perp \Delta_Y \mid \Delta_A = 1, A = a, W$.
@@
 
-_Response_: Thanks for the suggestion. We have added a brief paragraph on missingness in $W$.
+_Response_: Thanks for the suggestion. We have added a paragraph on missingness in $W$, copied below.
+
+> In practice it is common to also have missingness in $W$. In this
+situation, there are varying assumptions under which one can still identify and 
+estimate $E[Y(a)]$. The only native solution in `drtmle` currently is to augment $W$
+with columns of missingness indicators for each component of $W$. For example, if
+variable $W_1$ is observed, we could define a new indicator $\Delta_{W,1} = 1$ and 
+$\Delta_{W,1} = 0$ otherwise. For observations with $\Delta_{W,1} = 0$, we could use
+some (possibly very simple) imputation strategy to fill in a value for $W_1$. Candidate
+learners for the OR and PS may then include appropriate cross-products between $W_1$ 
+and $\Delta_{W,1}$. Other learners may opt to treat the imputed value of $W_1$ as 
+the true value and ignore $\Delta_{W,1}$. Super learner can be used to help determine
+the modeling strategy that leads to the optimal model fit conditional on $W_1$ and
+$\Delta_{W,1}$. It is important to note that this approach is only valid for drawing 
+causal inference under the assumption that $Y(a)$ is independent of $A$ conditional on
+the partially observed variables in $W$ and their missingness indicators. This assumption
+may not be plausible depending on the form of missingness of $W$. Other approaches for 
+handling missing data, e.g., multiple imputation, could in theory be coupled with `drtmle`, 
+though such applications have not been studied empirically and are not available natively
+in the `drtmle` package.
 

### 2023-01-04 paper/using_drtmle_obstudies.Rmd
finalize revision

@@
 define $\Delta_A$ and $\Delta_Y$ as indicators that $A$ and $Y$ are observed, respectively,
-then we can still perform valid inference on $\psi_0$ under the assumptions that 
+then we can still perform valid inference on $E[Y(a)]$ under the assumptions that 
 (i) $Y(a) \perp \Delta_A \mid W$, (ii) $Y(a) \perp A \mid \Delta_A = 1, W$, and
@@
 situation, there are varying assumptions under which one can still identify and 
-estimate $\psi_0$. The only native solution in `drtmle` currently is to augment $W$
+estimate $E[Y(a)]$. The only native solution in `drtmle` currently is to augment $W$
 with columns of missingness indicators for each component of $W$. For example, if
@@
 $\Delta_{W,1}$. It is important to note that this approach is only valid for drawing 
-causal inference under the assumption that $Y(a) \perp A \mid \Delta_W W$, which may 
-not be plausible depending on the form of missingness of $W$. Other approaches for handling
-missing data, e.g., multiple imputation, could in theory be coupled with `drtmle`, 
+causal inference under the assumption that $Y(a)$ is independent of $A$ conditional on
+the partially observed variables in $W$ and their missingness indicators. This assumption
+may not be plausible depending on the form of missingness of $W$. Other approaches for 
+handling missing data, e.g., multiple imputation, could in theory be coupled with `drtmle`, 
 though such applications have not been studied empirically and are not available natively
